### 2017-05-24 GRB111117A.tex
Pushing overleaf paper

%                                                                 aa.dem
% AA vers. 8.2, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[rnote]{aa} % for the research notes
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
% according to the author-year natbib style

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
%\documentclass[referee]{aa}

\usepackage{graphicx,url,twoopt,natbib}
\usepackage[varg]{txfonts}           %% A&A font choice

\usepackage{pdfcomment}              %% for popup acronym meanings
\usepackage{acronym}                 %% for popup acronym meanings

\usepackage{url}
\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0.3,0.7}
\hypersetup{colorlinks=true,
	linkcolor=linkcolor, 
	citecolor=linkcolor,
	filecolor=linkcolor, 
	urlcolor=linkcolor}

\usepackage{natbib}

\bibpunct{(}{)}{,}{a}{}{,}    %% natbib cite format used by A&A and ApJ

% Manually specified definitions
\input{definitions.tex}

\usepackage{xcolor}
\newcommand\todo[1]{\textbf{(#1)}}

\begin{document}

\title{The host galaxy of the short GRB111117A at z = 2.211.
\thanks{Based on observations collected at the European Southern Observatory, Paranal, 
	Chile, Program ID: 088.A-0051 and 091.D-0904.}}

\titlerunning{GRB111117A}

\input{authorlist.tex}

\date{Received/ accepted}

\authorrunning{Selsing et al.}

\abstract{
Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to localize. We here present evidence of the, to date, highest redshift high-confidence short GRB, located at $z=2.211$. This value supersedes a previously estimated, lower redshift value. Spectroscopic observations along with a wealth of imaging data allow us to place tight constraints on the nature of the host galaxy at an unprecedented distance. 
The rest frame X-ray derived hydrogen column density is high compared to a complete sample of sGRBs and seems to follow the evolution with redshift as traced by the hosts of long GRBs (lGRBs). This could indicate that the part of the sGRB population hosted in late-type galaxies, live in similar environments to that of lGRB hosts. 
The host is in the brighter end of the expected host brightness distribution at $z = 2.211$, but only 30 per cent of the distribution should be missed due to host faintness at this redshift. This indicates that we are not missing redshifts for a dominant fraction of the sGRB hosts at this redshift. The redshift of GRB111117A is evidence against a lognormal delay-time model for sGRBs through the predicted redshift distribution of sGRBs, which is very sensitive to high-z sGRBs.
From the age of the universe at explosion time, an initial progenitor separation of a$_0 < 3.1 $R$_\odot$ is required for the case of a binary neutron star (NS) system. This puts constraints on the progenitor system evolution up to the time of explosion.
}

\keywords{Gamma-ray burst: individual: GRB~111117A --- }

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

There is now mounting evidence that most short-duration gamma-ray bursts (GRBs) come from the merger of neutron stars (NSs), either with another NS, or a black hole, due to their apparent association with kilonovae 
\citep{Barnes2013a, Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The
absence of associated supernovae in deep searches \citep[e.g.][]
{Hjorth2005a,Fox2005,Hjorth2005b} supports this idea and distinguishes the physical origin of sGRBs from their long-duration counterparts.
%\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}. 
The range of host galaxy types, in some cases apparently absent, and typically more massive 
and on average less actively star-forming than long GRB hosts  \citep{Fong2013b}, as well as their positions within
hosts \citep{Fong2013a}, suggest an origin for the progenitors that can be very 
long-lived, and associated with stellar mass rather than purely 
star-formation rate \citep{Berger2014}.

This progress on the progenitors of sGRBs has been enabled by
localization by the \emph{Swift} satellite \citep{Gehrels2009a}. The bulk of these localizations have associated galaxies at relatively low redshifts with a median redshift $z\sim0.5$ \citep{Berger2014}. Because the redshifts are measured from the hosts, measurements of the redshift distribution of sGRBs is biased towards low redshift due to cosmological dimming. 
%at least partly due to the faintness of the bursts and their afterglows, though this is not clear. 
So far, the electromagnetic signal from sGRBs are the only means to accurately localize NS mergers and holds the promise for detection of an associated gravitational wave signal \citep{Ghirlanda2016}. From the burst and following afterglow, modeling the properties and mapping the environments additionally allows insights into the phenomenon itself.

The total lifetimes of NS binaries is dependent on their initial separations and subsequent inspiral times, and impacts the timing and distribution of the enrichment of the ISM and subsequent stars and planets with heavy $r$-process elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some limits can be calculated based on models of star-formation histories of, and the spatial distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}. The most distant cosmological bursts, however, offer direct, hard limits.

In this \emph{Letter} we present the spectrum of the host galaxy of the short
GRB\,111117A, showing it to be at higher redshift than previously
estimated. We present the rest frame burst properties based on this new distance
compared to previous analyses \citep{Margutti2012,Sakamoto2013} and revisit the host properties derived from the new solution to the SED fit. 
%While no optical afterglow was detected, the excellent localization from a detection
%of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also allows
%us to discuss the positioning and environmental properties of this remarkably
%distant short GRB.

Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology provided by
\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$ km s$^{-1}$ and
$\Omega_m = 0.307$ and report all magnitudes in the AB system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observations and results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Spectroscopic observations and analysis}

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
	\caption{Imaging of the field with the X-shooter slit overlaid. Only one slit is shown, despite 4 epochs of spectroscopic observations because of the similarity in position angle. The image is the FORS2 R-band image, for which the photometry is shown in Fig. \ref{fig:SED}. The blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}}
	\label{fig:spec_setup}
\end{figure}

%(21000 \AA when the K-band blocking filter is used)
Spectroscopic observations were carried out using the cross-dispersed echelle spectrograph, VLT/X-shooter \citep{Vernet2011}, in four seperate epochs. The burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from 3000 \AA~to 24800 \AA~across three spectroscopic arms, simultaneously illuminated through the use of dichroics. The bias-correction, flat-fielding, order tracing, wavelength calibration, rectification, and flux calibration is then carried out using the VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} and managed by Reflex \citep{Freudling2013}. Because the spectra are curved across each detector, a rectification algorithm is employed which introduces correlations between neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the degree of correlation while conserving the maximal resolution. The observations are combined and extracted using scripts described in Selsing et al. 2017 (in prep.) and available online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The signal-to-noise of the continuum is the near-infrared arm is too low to use the optimal extraction algorithm \citep{Horne1986}, why an aperture containing the entire object has been chosen for which the flux is summed and the corresponding error is summed in quadrature.
An overview of the spectroscopic observations is given in Tab. \ref{tab:spec_overview}, and the position of the slit on the target is shown in Fig. \ref{fig:spec_setup}.

% \begin{figure}
% 	\centering
% 	\includegraphics[width=9cm]{figures/SEDspecphot.pdf}
% 	\caption{SED fit showning the best-fist SED to the derived photometry. The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is a version of the observed spectrum binned by 30 \AA. The reason for the spectral gap at 10000 \AA is from the merging of the visual and near-infrared arm.}
% 	\label{fig:SED}
% \end{figure}

We determine a redshift of $z = 2.211$ from from the simultaneous detection of emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with \hb~detected at low significance ($\sim 3 \sigma$). \ha is only visible in the first epoch, due to the K-band blocking filter used for the remainder observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and shows significant velocity structure along the slit. A drop in the continuum the the blue of the \lya line further supports the inferred redshift.

%The total extent of the lines in velocity space is $\sim$ 450 km/s. The line profiles shows an asymmetric "double-horned" profile, indicating that we are seeing a galaxy with a large degree of coherent rotational motion relative to the line-of-sight. If we assume that we are viewing a spiral galaxy exactly edge-on, this is a measure of the rotational velocity of the gas. If we assume that the spectral resolution and the turbulent width of the lines are negligible compared to the rotational velocity, we can based on the projected size of the source and the width of the lines, put a constraint on the dynamical mass of the galaxy. Based on the physical scale from extent and the velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we are viewing the host at an angle, this value is an upper limit, which should be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.

Using the integrated flux of \ha, we can infer the SFR \citep{Kennicutt1998, Moustakas2006}. From the SED-fit, see Sect. \ref{SED}, and the detection of \lya, the host is constrained to contain very little or no dust which is also confirmed by the shape of the spectral continuum, which is why we do not apply a dust-correction. Because of the redshift, \ha~falls on the far end of the NIR-arm where the night sky is very bright and additionally, several bright sky-lines intersect the line, making an accurate estimate of the \ha-flux difficult. %Because of the challenging measurement, two approaches are employed. First, 
A limit on the SFR is obtained by integrating the part of \ha~free of contamination and conservatively correcting for the missing fraction by assessing the fraction integrated. Converting to a \citet{Chabrier2003} initial mass function using \citet{Madau2014} we get a limit of SFR > 7 M$_\odot$ yr$^{-1}$. %Secondly, based on the width of \oiii, an aperture covering the line is integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have been masked and interpolated over. From the integrated \ha-line, we estimate SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/OIII_img.pdf}
	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:line}
\end{figure}

A measure of the optical-to-X-ray flux ratio is parametrized in terms of the "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time of the non-detection. We find $\beta_{OX} < 0.79$, consistent with what was reported in \citet{Sakamoto2013}.

Because the projected distance does not change much between redshift 1.3 and 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating to host offset are unaffected.

\subsection{Imaging observations and SED analysis} \label{SED}

We additionally imaged the field of GRB 111117A in multiple broad-band filters using the VLT equipped with FORS2 ($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst has faded. These new data are complemented by a re-analysis of some of the imagery used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us here (GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric observations and measured brightnesses is given in Tab.~\ref{tab:phot_overview}.

All data were reduced, analyzed and fitted in a similar manner as we have done previously and described in detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Very briefly, we use our own python and IRAF routines to perform a consistent standard reduction of the available imaging which includes bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and stacking of individual images for a given instrument in a given filter. The photometry of the host was tied against magnitudes of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. For our $R$ and $I$-band photometry we used the color transformations of Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. We subsequently convert all magnitudes into the AB system if necessary, and correct for the Galactic foreground of $E_{B-V}=0.027~\mathrm{mag}$.

The multi-color spectral energy distribution (SED) is then fit by \citet{Bruzual2003} stellar population synthesis models based on a \citet{Chabrier2003} initial mass function in \emph{LePhare} \citep{Ilbert2006}, where the redshift if fixed to the spectroscopic value of $z=2.211$. The best model is obtained with an unreddened galaxy template, and returns physical parameters of luminosity ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), single stellar population age ($\tau = 0.5_{-0.25}^{+0.5}$ Gyr) and star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). %We show the SED fit in Fig. \ref{fig:SED}.

Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the re-analysis of Gemini data to the $z$-band data of \citet{Margutti2012} and \citet{Sakamoto2013}. Both report $z$-band photometry brighter by 0.8-1.0 mag than what we derive here, whereas bluer data are in excellent agreement with previously published values.

In fact, the large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break driving the galaxy photometric redshift of both of these earlier works. It is this exactly this overestimate in $z$-band brightness that led the authors to erronously low estimates in their photo-$z$ for GRB~111117A. Using the revised photometry from Table~\ref{tab:phot_overview}, the photo-$z$ of the galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic value at the 1~$\sigma$ confidence level.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Redshift, line-fluxes, SFR, and Mass}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{XRT temporal and spectral analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901.} \todo{Broken link, Paolo?}} \citep{Evans2009}. The X-ray afterglow light curve can be fit with a single power-law decay with index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral analysis using data in photon counting (PC) mode in the widest time epoch where the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant (namely, from $t-T_0 = 205$ s to $t-T_0 = 203.5$ ks, for a total of 29.1 ks of data), in order to prevent spectral changes that can affect the X-ray column density determination. The obtained spectrum is well described by an absorbed power-law model. The best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and an intrinsic $N_{H}$ of $2.4_{-1.6}^{+2.4} \times 10^{22}$ cm$^{-2}$ ($z=2.211$), assuming the Galactic $N_{H}$ in the burst direction of $4.1 \times 10^{20}$ cm$^{-2}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reinterpretation of restframe properties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Classification}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB111117A very securely belongs the the short class of GRBs. Because the observed classification indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not change this designation significantly. The intrinsic luminosity goes up, as visible in the light curve in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to majority of long GRBs. 
\citet{Bromberg2013} investigated the degree to which the long and short population distributions overlap and quantified the certainty in the class membership. GRB111117A has $96_{-5}^{+3}$ percent probability of being short. This is compared to the other two highest-z short candidates, GRB060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent).

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
	\caption{Restframe XRT lightcurve of GRB111117A, compared to the general population of XRT lightcurves of GRBs. The grey shared region is a compilation of long GRB lightcurves where the color represents density and the light blue is other short GRB lightcurves for which redshifts has been determined. Despite the remarkably high redshift, the burst’s luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
	\label{fig:sxray_lightcurve}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restframe $N_H$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We plot the recalculated $N_H$ in Fig. \ref{fig:NH_z} where we compare with the distributions of complete samples of both long and short GRBs. The long sample is from \citet{Arcodia2016} and the short sample is from \citet{DAvanzo2014a}. 17 of the 99 long bursts do not have redshifts and likewise for 5 out of 16 for the short sample. Bursts without redshifts have been excluded for both groups. GRB111117A occupies a unique position in Fig. \ref{fig:NH_z} with a the highest $N_H$ of all short burst. The short sample, excluding GRB111117A, is located at low redshifts(z < 1) and are found to populate a similar column density environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred hydrogen column for GRB111117A seems to follow the trend with increasing $N_H$ as a function of redshift as found for the long GRB hosts \citep{Arcodia2016}.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/NH_z.pdf}
	\caption{Rest frame, X-ray derived hydrogen column densities for GRB111117A compared to complete samples of both long and short populations. In red is shown the long sample of \citet{Arcodia2016}. The detections are replaced with contours for clarity and the limits are shown with arrows. In blue is the sample from \citet{DAvanzo2014a}. GRB111117A is an outlier from the short sample, but seems to follow the distribution of long GRBs.}
	\label{fig:NH_z}
\end{figure}

% \begin{figure}
% 	\centering
% 	\includegraphics[width=9cm]{figures/Eiso_z.pdf}
% 	\caption{Isotropic equivalent energy for all the representative samples of both long and short GRBs. The sample is from \citet{Turpin2015} and the short is from \citet{DAvanzo2014a}.}
% 	\label{fig:NH_z}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Host galaxy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From the clear host association, GRB111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits clear emission lines indicative of a population of relatively young stars, as the majority of short GRBs, the host is a late-type galaxy \citep{Fong2013b}.
The host of GRB111117A is entirely consistent in terms of stellar mass and stellar age with the general population of short GRB hosts($\left\langle M _* \right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle = 0.3 $ Gyr) \citet{Leibler2010}. The SFR, on the other hand, is $\sim$1 order of magnitude higher than the typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to the SFRs found in the hosts of long GRBs at a corresponding redshift \citep{Kruhler2015}. Only two hosts in the sample of short GRBs compiled in \citet{Berger2014} have a more vigorous star formation, meaning that it is in the very upper end of the star formation distribution, while not a complete outlier. This is likely a selection effect, because a less star-forming galaxy would exhibit weaker emission lines, thus making the redshift harder to determine. Being a late-type host, both the stellar mass and sSFR are entirely consistent with the expected host of sGRBs\citep{Behroozi2014}.
That we detect \lya~is consistent with the SED-inferred absence of dust, despite the moderate stellar age and the high X-ray derived hydrogen column density which would suggest the opposite. The centroid of the Ly$\alpha$ emission is found to be redshifted by $\sim$ 220 km s$^{-1}$ with respect to systemic, which is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Comparison to long GRBs at z $\sim$ 2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implications for redshift distribution of short GRBs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A single sGRB at high redshift does little in terms of constraints on the redshift distribution and therethrough the progenitor channels, but other short hosts could be missed because they are intrinsically fainter and thus this high-z event is only detected due to the brightness of the host. \citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by the characteristic galaxy luminosity at their respective redshift, $L_B/L^{\star}_{B}$. 26 out of 39 hosts (66 per cent) in the sample has redshifts. To convert the SED-inferred $M_B$ of GRB111117A to $L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window $2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with measured $L_B/L^{\star}_{B}$. If we assume that we are able to get emission-line redshifts from hosts with $R < 25 $mags \citep{Kruhler2012}, then we would have missed around 30\% (8 out of 26 from the sample of \citealt{Berger2014} with measured $L_B/L^{\star}_{B}$), if they were at the redshift of GRB111117A. If we assume that the hosts missing redshift in \citet{Berger2014} follow the redshift distribution of the hosts \textit{with} redshift, this implies that we are \textit{not} missing a dominant fraction of sGRB redshift at $z \approx 2$, due to host faintness.

The theoretical redshift distributions of sGRBs depends on the type of delay-time function used to model the progenitor system. The redshift of GRB111117A puts constraints on the type of delay-time models suitable for modeling. The likelihood preferred lognormal time delay models investigated in \citet{Wanderman2015} predicts a rate of sGRBs at $z = 2.211$, $\sim$ 2 orders of magnitude lower compared to peak rate (z = 0.9). It is stated in \citet{Wanderman2015} that this preference depends critically on the absence of non-collapsar sGRB at $z \gtrsim 1.2$. The redshift of the burst, on the other hand, is close to the expected peak in sGRB rate calculated for the power law models \citep{Behroozi2014, Wanderman2015}.

%In terms of Fig. 9 in \citet{Fong2013a}, the host is indeed  brighter than the average galaxies using the evolving luminosity function of galaxies. \todo{What is the average brightness of a galaxy at z = 2?} 
%\todo{Reformulate paragraph}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Constraints on the progenitor separation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

At $z = 2.211$, the age of the universe is 2.97 Gyr. If the progenitor systems of sGRBs are the merger of two neutron stars, this sets a hard upper limit to the coalescence timescale for such a system. In the absence of other mechanisms, the timescale of the orbital decay of the system is set by the energy loss due to gravitational waves, which in turn is set by the mass of constituent compact objects and the separation of the two \citep{Postnov2014}. If we assume that the formation timescale of the first galaxies is short compared to the time since the big bang \citep{Richard2011} and if we assume a mass of 1.4 M$_\odot$ for each of the neutron stars at the time of system formation, this places a limit on the initial separation of the two neutronstars of a$_0 < 3.1 $R$_\odot$, where a$_0$ is the initial separation. Because we have the age of the stellar population, by using this as a limit to the orbital decay time we can put tighter constraints, a$_0 < 1.7 $R$_\odot$, but this does not account for the possibility there could be an underlying stellar population of older stars from a previous star-formation epoch, which is why we consider a$_0 < 3.1 $R$_\odot$ as a hard upper limit. The delay time between formation and explosion is well accomodated by the models of \citet{Belczynski2006}, although the longest formation channels are excluded. This is especially true given the late type nature of the host \citep{OShaughnessy2008}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We have in this letter provided a revised, spectroscopic redshift for GRB111117A based on emission lines discovered using an improved reduction scheme, setting it at $z = 2.211$. This value supersedes the previous photometric redshift of $z \sim 1.3$ \citep{Margutti2012, Sakamoto2013}. Part of the reason for the erroneous redshift estimate of previous authors is attributed to a discrepancy in the measured $z$-band magnitude.

The rest-frame parameters of the burst and the conditions of the burst environment have been recalculated using the new distance. The X-ray derived hydrogen column density towards GRB111117A is the highest within a complete sample of short hosts, but seems to trace the evolution with redshift as found for the hosts of long GRBs.

The SFR of the host is in the upper end of the sGRB host SFR distribution and this plays a role in allowing a redshift to be measured for this host. Despite the moderate age and high $N_H$, almost no dust is present.

Although a single burst carries little leverage in terms of constraining the redshift distribution of sGRB, the high redshift of GRB111117A needs to be accommodated. A lognormal delay time model predicts a very low volumetric density of bursts at $z = 2.211$, whereas a power law delay time model peaks near GRB111117A. If more sGRBs are at this redshift, but missed due to faintness of their hosts, a lognormal delay time model will be disfavored. Compared to a sample of short hosts, GRB111117A is more luminous than 70 per cent of the sample with measured luminosities. Of the host sample, for 30 per cent, we would be unable to determine a redshift, should they be at the redshift of GRB111117A. This implies that we are \textit{not} missing a dominant fraction of the sGRBs at $z \sim 2$.

Using the age of the universe at the time of explosion allows us to set constraints on the maximal separation between the engine constituents at the time of formation. We find that the maximal separation for two neutron stars at formation time is a$_0 < 3.1 $R$_\odot$, which excludes some of the formation channels with the longest timescales.

\begin{acknowledgements}
TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
\end{acknowledgements}

\bibliographystyle{mnras}
\bibliography{GRB111117A}

\newpage

\input{tables/spec_overview.tex}

\input{tables/phot_overview.tex}

\end{document}

### 2017-05-24 authorlist.tex
Pushing overleaf paper

\author{J.~Selsing\inst{1}
\and T. Kr\"{u}hler\inst{2}
\and D.~Malesani\inst{1,3}
\and P. D'Avanzo\inst{4}
\and J.~Palmerio\inst{5}
\and S.~D.~Vergani\inst{6}
\and J.~Japelj\inst{6}
\and B.~Milvang-Jensen\inst{1}
\and D.~Watson\inst{1}
\and P.~Jakobsson\inst{7}
%
\and Z. Cano\inst{8}
\and J.~P.~U.~Fynbo\inst{1}
\and A. Gomboc\inst{9}
\and K.~E.~Heintz\inst{7,1}
\and M. Sparre\inst{1,10}
\and XSGRB collaboration
}

\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
%2
\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
%3
\and DTU Space, National Space Institute, Technical University of Denmark, Elektrovej 327, DK-2800 Lyngby, Denmark
%4
\and INAF - Osservatorio Astronomico di Brera, via E. Bianchi 46, I-23807, Merate (LC), Italy
%5
\and Sorbonne Universités, UPMC Univ. Paris 6 et CNRS, UMR 7095, Institut d’Astrophysique de Paris, 98 bis bd Arago, 75014 Paris, France
%6
\and GEPI, Observatoire de Paris, PSL Research University, CNRS, Univ. Paris Diderot, Sorbonne Paris Cité, 5 Place Jules Janssen, 92195 Meudon, France
%7
\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
%8
\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain.
%9
\and University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
%10
\and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
%\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
}

### 2017-05-24 tables/phot_overview.tex
Pushing overleaf paper

\begin{table*}[!ht]

	\centering
	\caption{Overview of the photometric observations. \label{tab:phot_overview}}
	\begin{tabular}{ccccccc}
		\hline\hline
{Obs. Date} &  Exptime & Telescope/Instrument & Filter & Airmass & Image Quality & Host Brightness\tablefootmark{a}  \\ [1.5pt]
        \hline
{} & {ks} &    & {} & & (arcsec)  & (mag$_{\mathrm{AB}}$)  \\ [1.5pt]
		\hline
2013-08-30T07:43 & 1.45 & VLT/FORS2 & $g$ & 1.55 & 0.99 & $24.08\pm 0.09$ \\
2011-11-17T20:07 & 0.80 & GTC/OSIRIS & $g$ & 1.15 & 1.67 & $24.13\pm 0.09$ \\
2011-11-17T20:07 & 1.20 & GTC/OSIRIS & $r$ & 1.11 & 1.50 & $23.93\pm 0.08$ \\        
2013-07-17T08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
2011-11-28T21:10 & 3.60 & TNG/LRS & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
2011-11-17T20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   
2013-08-03T09:23 & 1.35 & VLT/FORS2 & $I$ & 1.54 & 0.93 & $24.22\pm 0.15$ \\           
2011-11-28T06:14 & 1.80 & Gemini/GMOS-N & $z$ & 1.01 & 0.84 & $24.24\pm 0.47$ \\  
2013-07-13T09:33 & 1.08 & VLT/FORS2 & $z$ & 1.49 & 0.63 & $23.76\pm 0.21$ \\             
2013-06-24T09:14 & 1.98 & VLT/HAWK-I & $J$ & 1.70 & 0.63 & $23.13\pm 0.18$ \\        
2013-06-27T09:21 & 1.68 & VLT/HAWK-I & $H$ & 1.63 & 0.91 & $22.94\pm 0.29$ \\   
2013-06-28T09:14 & 1.92 & VLT/HAWK-I & $K_\mathrm{s}$ & 1.65 & 0.76 & $23.07\pm 0.32$ \\   
\hline\noalign{\smallskip}

\end{tabular}

\tablefoot{
\tablefoottext{a}{All magnitudes are given in the AB system and are not corrected for the expected Galactic foreground extinction corresponding to a reddening of $E_{B-V}=0.027$\,mag. \todo{Thomas, Can you help with the reference for the data which are not ours?}}}
\end{table*}

### 2017-05-24 py/BPT.py

@@ -1,0 +1,123 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import matplotlib as mpl
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+mpl.rcParams.update(params)
+
+def propagate_uncertainties(func, parameters, parameter_errors, samples):
+    """
+    For a given function, evaluates the function and computes the 1 \sigma errors.
+    """
+
+    for i, k in enumerate(parameter_errors):
+        parameter_errors[i] = k * (np.random.randn(samples))
+
+    for i, k in enumerate(parameters):
+        parameters[i] = k + parameter_errors[i]
+
+    func_eval = func (*parameters)
+    # print(func_eval)
+    # print('Discarded iterations: ', len(func_eval[np.isnan(func_eval)]))
+    if len(func_eval[np.isnan(func_eval)])/samples  > 0.3:
+        print('High number of dicards. Potentially biased result.')
+    return np.mean(func_eval[~np.isnan(func_eval)]), np.std(func_eval[~np.isnan(func_eval)])
+
+def fraction(x, y):
+    return np.log10(x / y)
+
+
+def main():
+    """
+    # Script to produce N_H vs. z
+    """
+    # Read in SDSS line-fluxes
+    sdss = fits.open('/Users/jselsing/Work/spectral_libraries/sdss_catalogs/portsmouth_emlinekin_full-DR12.fits')
+    # mask = np.logical_and((sdss[1].data.field('Z') >= 0.5),(sdss[1].data.field('Z') <= 2.0))
+
+
+    mask = (sdss[1].data.field('Z') <= 0.1)
+
+    Hb_sdss = sdss[1].data.field('FLUX')[:, 15]
[…]

### 2017-05-24 py/Eiso_z.py

@@ -1,0 +1,130 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import matplotlib as mpl
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+mpl.rcParams.update(params)
+from astropy.cosmology import Planck15 as cosmo
+
+
+
+def iqr(a):
+    """Calculate the IQR for an array of numbers."""
+    a = np.asarray(a)
+    q1 = stats.scoreatpercentile(a, 25)
+    q3 = stats.scoreatpercentile(a, 75)
+    return q3 - q1
+
+
+def _freedman_diaconis_bins(a):
+    """Calculate number of hist bins using Freedman-Diaconis rule."""
+    # From http://stats.stackexchange.com/questions/798/
+    a = np.asarray(a)
+    h = 2 * iqr(a) / (len(a) ** (1 / 3))
+    # fall back to sqrt(a) bins if iqr is 0
+    if h == 0:
+        return int(np.sqrt(a.size))
+    else:
+        return int(np.ceil((a.max() - a.min()) / h))
+
+
+def filt_nan(input_array, fill_value=np.nan):
+    """
+    Small functino to remove values in array, not convertible to float. Return the filtered array
+    """
+    holder = []
+    for ii in input_array:
+        try:
+            holder.append(float(ii))
+        except ValueError:
+            holder.append(fill_value)
+    return np.array(holder)
+
+
+def main():
+    """
+    # Script to produce Eiso vs. z
+    """
+
+    Eiso_111117A = 3.38 * (cosmo.luminosity_distance(2.211)**2/cosmo.luminosity_distance(1.3)**2)*(2.3 / 3.211)*1e51
+
+
+    # Read in long GRBs from Turpin et al. 2015, https://arxiv.org/abs/1503.02760
[…]

### 2017-05-24 py/NH_z.py

@@ -1,0 +1,166 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats
+import matplotlib as mpl
+
+params = {
+   'axes.labelsize': 16,
+   'font.size': 12,
+   'legend.fontsize': 12,
+   'xtick.labelsize': 16,
+   'ytick.labelsize': 16,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+mpl.rcParams.update(params)
+
+def iqr(a):
+    """Calculate the IQR for an array of numbers."""
+    a = np.asarray(a)
+    q1 = stats.scoreatpercentile(a, 25)
+    q3 = stats.scoreatpercentile(a, 75)
+    return q3 - q1
+
+
+def _freedman_diaconis_bins(a):
+    """Calculate number of hist bins using Freedman-Diaconis rule."""
+    # From http://stats.stackexchange.com/questions/798/
+    a = np.asarray(a)
+    h = 2 * iqr(a) / (len(a) ** (1 / 3))
+    # fall back to sqrt(a) bins if iqr is 0
+    if h == 0:
+        return int(np.sqrt(a.size))
+    else:
+        return int(np.ceil((a.max() - a.min()) / h))
+
+
+def filt_nan(input_array, fill_value=np.nan):
+    """
+    Small functino to remove values in array, not convertible to float. Return the filtered array
+    """
+    holder = []
+    for ii in input_array:
+        try:
+            holder.append(float(ii))
+        except ValueError:
+            holder.append(fill_value)
+    return np.array(holder)
+
+
+def main():
+    """
+    # Script to produce N_H vs. z
+    """
+    # Read in long GRBs
+    burst_table = pd.read_csv("../data/Comparison sample - Long NH.csv")
+    name, z_long, NH_long = burst_table["GRB"].values, burst_table["z"].values, burst_table["Nnew(z)"].values
+
[…]

### 2017-05-24 py/SED_plot.py

@@ -1,0 +1,121 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import numpy as np
+from scipy.interpolate import interp1d
+from scipy.signal import medfilt
+
+# Plotting
+import matplotlib.pyplot as pl
+import seaborn; seaborn.set_style('ticks')
+
+import matplotlib as mpl
+from matplotlib.ticker import FormatStrFormatter
+params = {
+   'axes.labelsize': 16,
+   'font.size': 16,
+   'legend.fontsize': 16,
+   'xtick.labelsize': 16,
+   'ytick.labelsize': 16,
+   'text.usetex': True,
+   'figure.figsize': [7.281, 4.5]
+   }
+mpl.rcParams.update(params)
+
+
+def jansky_to_AB(jansky):
+    return 23.9 - 2.5*np.log10(jansky)
+
+def AB_to_jansky(AB):
+    return 10**((23.9 - AB)/2.5)
+
+
+def main():
+    # Small script to plot SED, photometry and spectrum
+
+    # Load data
+    spectrum = np.genfromtxt("../data/stitched_spectrum_bin50.dat")
+    wl_spec = spectrum[:, 0]
+    flux_spec = spectrum[:, 1]
+    error_spec = spectrum[:, 2]
+    # pl.plot(wl_spec, flux_spec)
+    # pl.show()
+    SED = np.genfromtxt("../data/SED.dat")
+    wl_SED = SED[:, 0]
+    AB_SED = SED[:, 1]
+    phot = np.genfromtxt("../data/photometry.dat")
+    mag = phot[:, 0]
+    magmask = (mag != -99)
+    mag = phot[:, 0][magmask]
+    magerr = phot[:, 1][magmask]
+    wl_phot = phot[:, 2][magmask]
+    wl_plot_wid = phot[:, 3][magmask]/2.35
+    # pl.scatter(wl_phot, mag)
+    # print(phot[:, 2])
+    # print(phot[:, 3])
+    # exit()
+
+
+    nu = 3e5/(wl_spec*1e-13) # nu
+    f_nu = (flux_spec*(1e-13*(wl_spec**2.))/3e5)/1e-29
+    f_nu[f_nu <= 0] = 0.1
[…]

### 2017-05-24 py/combine_arms_and_bin.py

@@ -1,0 +1,227 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+import numpy as np
+from scipy.interpolate import interp1d
+from util import *
+from scipy import optimize
+from scipy.signal import medfilt
+import lmfit
+# Plotting
+import matplotlib.pyplot as pl
+import seaborn; seaborn.set_style('ticks')
+from numpy.polynomial import chebyshev
+import matplotlib as mpl
+from matplotlib.ticker import FormatStrFormatter
+
+def pow(x, y, z):
+    return y * x ** z
+
+
+def load_array(data_array, tell = None):
+
+    # Check is reposnse has been written based on value in sixt column
+    has_resp = False
+    # if np.nanmean(data_array[:, 6]) < 1e-10:
+    #     has_resp = True
+
+    if has_resp:
+        try:
+            tellcorr = data_array[:, 8]
+            tellcorr[np.isnan(tellcorr)] = 1
+            slitcorr = 1#np.median(data_array[:, 7])
+            return data_array[:, 1], data_array[:, 2]*tellcorr*slitcorr, data_array[:, 3]*tellcorr*slitcorr, data_array[:, 4]
+        except:
+            slitcorr = 1#np.median(data_array[:, 7])
+            return data_array[:, 1], data_array[:, 2]*slitcorr, data_array[:, 3]*slitcorr, data_array[:, 4]
+    elif not has_resp:
+        try:
+            tellcorr = tell#data_array[:, 7]
+            # print(tellcorr)
+            tellcorr[np.isnan(tellcorr)] = 1
+            slitcorr = 1#np.median(data_array[:, 6])
+            return data_array[:, 1], data_array[:, 2]*tellcorr*slitcorr, data_array[:, 3]*tellcorr*slitcorr, data_array[:, 4]
+        except:
+            slitcorr = 1#np.median(data_array[:, 6])
[…]

### 2017-05-24 py/line_flux_ha.py

@@ -1,0 +1,111 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats, interpolate
+import matplotlib as mpl
+from astropy.convolution import Gaussian1DKernel, Gaussian2DKernel, convolve
+from astropy.cosmology import Planck15 as cosmo
+from astropy import units as u
+from astropy import constants as c
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+
+def convert_air_to_vacuum(air_wave) :
+    # convert air to vacuum. Based onhttp://idlastro.gsfc.nasa.gov/ftp/pro/astro/airtovac.pro
+    # taken from https://github.com/desihub/specex/blob/master/python/specex_air_to_vacuum.py
+
+    sigma2 = (1e4/air_wave)**2
+    fact = 1. +  5.792105e-2/(238.0185 - sigma2) +  1.67917e-3/( 57.362 - sigma2)
+    vacuum_wave = air_wave*fact
+
+    # comparison with http://www.sdss.org/dr7/products/spectra/vacwavelength.html
+    # where : AIR = VAC / (1.0 + 2.735182E-4 + 131.4182 / VAC^2 + 2.76249E8 / VAC^4)
+    # air_wave=numpy.array([4861.363,4958.911,5006.843,6548.05,6562.801,6583.45,6716.44,6730.82])
+    # expected_vacuum_wave=numpy.array([4862.721,4960.295,5008.239,6549.86,6564.614,6585.27,6718.29,6732.68])
+    return vacuum_wave
+
+
+
+def main():
+    """
+    # Script to get lineflux
+    """
+    # Get extraction
+    data = np.genfromtxt("../data/NIRext.dat", dtype=None)
+    # data = np.genfromtxt("../data/NIROB4skysubstdext.dat", dtype=None)
[…]

### 2017-05-24 py/line_flux_hb.py

@@ -1,0 +1,99 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats, interpolate
+import matplotlib as mpl
+from astropy.convolution import Gaussian1DKernel, Gaussian2DKernel, convolve
+from astropy.cosmology import Planck15 as cosmo
+from astropy import units as u
+from astropy import constants as c
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+
+def convert_air_to_vacuum(air_wave) :
+    # convert air to vacuum. Based onhttp://idlastro.gsfc.nasa.gov/ftp/pro/astro/airtovac.pro
+    # taken from https://github.com/desihub/specex/blob/master/python/specex_air_to_vacuum.py
+
+    sigma2 = (1e4/air_wave)**2
+    fact = 1. +  5.792105e-2/(238.0185 - sigma2) +  1.67917e-3/( 57.362 - sigma2)
+    vacuum_wave = air_wave*fact
+
+    # comparison with http://www.sdss.org/dr7/products/spectra/vacwavelength.html
+    # where : AIR = VAC / (1.0 + 2.735182E-4 + 131.4182 / VAC^2 + 2.76249E8 / VAC^4)
+    # air_wave=numpy.array([4861.363,4958.911,5006.843,6548.05,6562.801,6583.45,6716.44,6730.82])
+    # expected_vacuum_wave=numpy.array([4862.721,4960.295,5008.239,6549.86,6564.614,6585.27,6718.29,6732.68])
+    return vacuum_wave
+
+
+
+
+
+
+def main():
+    """
+    # Script to get lineflux
+    """
+    # Get extraction
+    data = np.genfromtxt("../data/NIRext.dat", dtype=None)
+    # data = np.genfromtxt("../data/NIROB4skysubstdext.dat", dtype=None)
+    # print(np.std([18, 31, 38, 20]))
+    # exit()
[…]

### 2017-05-24 py/line_flux_nii.py

@@ -1,0 +1,93 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats, interpolate
+import matplotlib as mpl
+from astropy.convolution import Gaussian1DKernel, Gaussian2DKernel, convolve
+from astropy.cosmology import Planck15 as cosmo
+from astropy import units as u
+from astropy import constants as c
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+
+def convert_air_to_vacuum(air_wave) :
+    # convert air to vacuum. Based onhttp://idlastro.gsfc.nasa.gov/ftp/pro/astro/airtovac.pro
+    # taken from https://github.com/desihub/specex/blob/master/python/specex_air_to_vacuum.py
+
+    sigma2 = (1e4/air_wave)**2
+    fact = 1. +  5.792105e-2/(238.0185 - sigma2) +  1.67917e-3/( 57.362 - sigma2)
+    vacuum_wave = air_wave*fact
+
+    # comparison with http://www.sdss.org/dr7/products/spectra/vacwavelength.html
+    # where : AIR = VAC / (1.0 + 2.735182E-4 + 131.4182 / VAC^2 + 2.76249E8 / VAC^4)
+    # air_wave=numpy.array([4861.363,4958.911,5006.843,6548.05,6562.801,6583.45,6716.44,6730.82])
+    # expected_vacuum_wave=numpy.array([4862.721,4960.295,5008.239,6549.86,6564.614,6585.27,6718.29,6732.68])
+    return vacuum_wave
+
+def main():
+    """
+    # Script to get lineflux
+    """
+    # Get extraction
+    data = np.genfromtxt("../data/NIRext.dat", dtype=None)
+    # data = np.genfromtxt("../data/NIROB4skysubstdext.dat", dtype=None)
+    # print(np.std([18, 31, 38, 20]))
+    # exit()
+    wl = data[:, 1]
[…]

### 2017-05-24 py/line_flux_oiii.py

@@ -1,0 +1,107 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+from astropy.io import fits
+import pandas as pd
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import numpy as np
+from scipy import stats, interpolate
+import matplotlib as mpl
+from astropy.convolution import Gaussian1DKernel, Gaussian2DKernel, convolve
+from astropy.cosmology import Planck15 as cosmo
+from astropy import units as u
+from astropy import constants as c
+
+params = {
+   'axes.labelsize': 10,
+   'font.size': 10,
+   'legend.fontsize': 10,
+   'xtick.labelsize': 10,
+   'ytick.labelsize': 10,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+
+def convert_air_to_vacuum(air_wave) :
+    # convert air to vacuum. Based onhttp://idlastro.gsfc.nasa.gov/ftp/pro/astro/airtovac.pro
+    # taken from https://github.com/desihub/specex/blob/master/python/specex_air_to_vacuum.py
+
+    sigma2 = (1e4/air_wave)**2
+    fact = 1. +  5.792105e-2/(238.0185 - sigma2) +  1.67917e-3/( 57.362 - sigma2)
+    vacuum_wave = air_wave*fact
+
+    # comparison with http://www.sdss.org/dr7/products/spectra/vacwavelength.html
+    # where : AIR = VAC / (1.0 + 2.735182E-4 + 131.4182 / VAC^2 + 2.76249E8 / VAC^4)
+    # air_wave=numpy.array([4861.363,4958.911,5006.843,6548.05,6562.801,6583.45,6716.44,6730.82])
+    # expected_vacuum_wave=numpy.array([4862.721,4960.295,5008.239,6549.86,6564.614,6585.27,6718.29,6732.68])
+    return vacuum_wave
+
+
+
+def main():
+    """
+    # Script to get lineflux
+    """
+    # Get extraction
+    data = np.genfromtxt("../data/NIRext.dat", dtype=None)
+    # data = np.genfromtxt("../data/NIROB4skysubstdext.dat", dtype=None)
[…]

### 2017-05-24 py/schecter_extrapolate.py

@@ -1,0 +1,51 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+# Imports
+import numpy as np
+
+# Plotting
+import matplotlib.pyplot as pl
+import seaborn; seaborn.set_style('ticks')
+from scipy import interpolate
+
+def main():
+
+    # z = [0.30, 0.50, 0.70, 0.90, 1.10, 1.30]
+    # z_arr = np.arange(0, 3, 0.01)
+    # MB = [-20.36, -20.72, -21.15, -21.21, -21.38, -21.86]
+    # MB_up = [0.13, 0.05, 0.07, 0.00, 0.04, 0.07]
+    # MB_low = [0.11, 0.07, 0.07, 0.03, 0.05, 0.08]
+    # f = interpolate.interp1d(z, MB, kind="slinear", fill_value="extrapolate")
+    # pl.errorbar(z, MB, yerr=[MB_up, MB_low])
+    # pl.plot(z_arr, f(z_arr))
+    # print(f(2.211))
+    # pl.show()
+
+    # LB from Berger 2014
+    L_B = np.array([0.1, 1.0, 0.3, np.nan, np.nan, 0.1, 0.1, 0.6, np.nan, 0.1, 1.4, 1.9, 1.2, np.nan, np.nan, np.nan, np.nan, np.nan, 0.3, 1.0, np.nan, 0.5, np.nan, np.nan, 0.6, 0.8, 1.0, 5.0, np.nan, np.nan, 1.6, 0.6, 0.9, 0.4, 1.0, 1.2, 1.0, 0.2, 1.3])
+    print(len(L_B))
+    print(26/39)
+    print(len(L_B[L_B > 1.17]))
+    print(np.sum((~np.isnan(L_B)).astype("int")))
+    print(1 - 7/26)
+    L_B_nonan = L_B[~np.isnan(L_B)]
+
+    MB_star = -21.83#f(2.211)
+    L_B_star = 10 ** (-1*(-22 - MB_star)/2.5)
+    # seaborn.distplot(L_B_nonan, rug=True)
+    # pl.show()
+    print(L_B_star)
+    print(L_B_star*10**(-0.4))
+    print(len(L_B[L_B > L_B_star*10**(-0.4)]))
+    print(1 - 18/26)
+    print(np.percentile(L_B[~np.isnan(L_B)], [14, 50, 85]))
+
+if __name__ == '__main__':
+    main()
+
+
+
+

### 2017-05-24 py/util.py

@@ -1,0 +1,229 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+import numpy as np
+from astropy.stats import sigma_clip
+from scipy import signal
+import matplotlib.pyplot as pl
+from scipy.special import wofz, erf
+
+__all__ = ["correct_for_dust", "bin_image", "avg", "convert_air_to_vacuum", "convert_vacuum_to_air", "bin_spectrum"]
+
+
+
+
+def avg(flux, error, mask=None, axis=2, weight=False, weight_map=None):
+
+    """Calculate the weighted average with errors
+    ----------
+    flux : array-like
+        Values to take average of
+    error : array-like
+        Errors associated with values, assumed to be standard deviations.
+    mask : array-like
+        Array of bools, where true means a masked value.
+    axis : int, default 0
+        axis argument passed to numpy
+
+    Returns
+    -------
+    average, error : tuple
+
+    Notes
+    -----
+    """
+    try:
+        if not mask:
+            mask = np.zeros_like(flux).astype("bool")
+    except:
+        pass
+        # print("All values are masked... Returning nan")
+        # if np.sum(mask.astype("int")) == 0:
+        #     return np.nan, np.nan, np.nan
+
+
+    # Normalize to avoid numerical issues in flux-calibrated data
+    norm = abs(np.median(flux[flux > 0]))
+    if norm == np.nan or norm == np.inf or norm == 0:
+        print("Nomalization factor in avg has got a bad value. It's "+str(norm)+" ... Replacing with 1")
+
+    flux_func = flux.copy() / norm
+    error_func = error.copy() / norm
+
+    # Calculate average based on supplied weight map
+    if weight_map is not None:
+        # Remove non-contributing pixels
+        flux_func[mask] = 0
+        error_func[mask] = 0
+        average = np.sum(weight_map * flux_func, axis = axis)
+        variance = np.sum(weight_map ** 2 * error_func ** 2.0, axis = axis)
+
+    # Inverse variance weighted average
+    elif weight:
+        ma_flux_func = np.ma.array(flux_func, mask=mask)
+        ma_error_func = np.ma.array(error_func, mask=mask)
+        w = 1.0 / (ma_error_func ** 2.0)
[…]

### 2017-05-24 py/Bergersample.py
Rename and update redshift dist calc

@@ -1,0 +1,70 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+# Imports
+import numpy as np
+
+# Plotting
+import matplotlib.pyplot as pl
+import seaborn; seaborn.set_style('ticks')
+from scipy import interpolate
+
+def main():
+
+    # GRB111117A L_B
+    MB_star = -21.83 # From Marchesini et al. 2007
+    L_B_star = 10 ** (-1*(-22 - MB_star)/2.5)
+
+    #L_B for GRB111117A
+    print("L_B:")
+    print(L_B_star)
+    print()
+    # L_B for a galaxy, 1 mag fainter
+    dM = 1
+    print("L_B_limit:")
+    print(L_B_star*10**(-0.4*(dM)))
+    print()
+
+    # LB from Berger 2014
+    L_B = np.array([0.1, 1.0, 0.3, np.nan, np.nan, 0.1, 0.1, 0.6, np.nan, 0.1, 1.4, 1.9, 1.2, np.nan, np.nan, np.nan, np.nan, np.nan, 0.3, 1.0, np.nan, 0.5, np.nan, np.nan, 0.6, 0.8, 1.0, 5.0, np.nan, np.nan, 1.6, 0.6, 0.9, 0.4, 1.0, 1.2, 1.0, 0.2, 1.3])
+    #Exclude nans
+    L_B_nonan = L_B[~np.isnan(L_B)]
+    # Length of list
+    ln = len(L_B)
+    # Nr with redshifts
+    lnredshifts = len(L_B_nonan)
+    # Brigter than GRB111117A
+    lnbrigter = len(L_B[L_B > L_B_star])
+    #Redshift completeness
+    print("Redshift completeness:")
+    print(lnredshifts/ln)
+    print()
+    #Brighter than:
+    print("Brighter than:")
+    print(1 - lnbrigter/lnredshifts)
+    print()
+    # print(len(L_B[L_B > 1.17]))
+    # print(np.sum((~np.isnan(L_B)).astype("int")))
+    # print(1 - 7/26)
+
+
+    MB_star = -21.83#f(2.211)
+    L_B_star = 10 ** (-1*(-22 - MB_star)/2.5)
+
+
+    lnunobs = len(L_B[L_B > L_B_star*10**(-0.4)])
+    # Fraction unobservable
+    print("Fraction unobservable:")
[…]

### 2017-05-24 py/schecter_extrapolate.py

@@ -1,51 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-from __future__ import division, print_function
-
-# Imports
-import numpy as np
-
-# Plotting
-import matplotlib.pyplot as pl
-import seaborn; seaborn.set_style('ticks')
-from scipy import interpolate
-
-def main():
-
-    # z = [0.30, 0.50, 0.70, 0.90, 1.10, 1.30]
-    # z_arr = np.arange(0, 3, 0.01)
-    # MB = [-20.36, -20.72, -21.15, -21.21, -21.38, -21.86]
-    # MB_up = [0.13, 0.05, 0.07, 0.00, 0.04, 0.07]
-    # MB_low = [0.11, 0.07, 0.07, 0.03, 0.05, 0.08]
-    # f = interpolate.interp1d(z, MB, kind="slinear", fill_value="extrapolate")
-    # pl.errorbar(z, MB, yerr=[MB_up, MB_low])
-    # pl.plot(z_arr, f(z_arr))
-    # print(f(2.211))
-    # pl.show()
-
-    # LB from Berger 2014
-    L_B = np.array([0.1, 1.0, 0.3, np.nan, np.nan, 0.1, 0.1, 0.6, np.nan, 0.1, 1.4, 1.9, 1.2, np.nan, np.nan, np.nan, np.nan, np.nan, 0.3, 1.0, np.nan, 0.5, np.nan, np.nan, 0.6, 0.8, 1.0, 5.0, np.nan, np.nan, 1.6, 0.6, 0.9, 0.4, 1.0, 1.2, 1.0, 0.2, 1.3])
-    print(len(L_B))
-    print(26/39)
-    print(len(L_B[L_B > 1.17]))
-    print(np.sum((~np.isnan(L_B)).astype("int")))
-    print(1 - 7/26)
-    L_B_nonan = L_B[~np.isnan(L_B)]
-
-    MB_star = -21.83#f(2.211)
-    L_B_star = 10 ** (-1*(-22 - MB_star)/2.5)
-    # seaborn.distplot(L_B_nonan, rug=True)
-    # pl.show()
-    print(L_B_star)
-    print(L_B_star*10**(-0.4))
-    print(len(L_B[L_B > L_B_star*10**(-0.4)]))
-    print(1 - 18/26)
-    print(np.percentile(L_B[~np.isnan(L_B)], [14, 50, 85]))
-
-if __name__ == '__main__':
-    main()
-
-
-
-

### 2017-05-24 GRB111117A.tex
Test tex update
Confirm progenitor seperation and push calc code

@@
 This puts constraints on the progenitor system evolution up to the time of explosion.
+
 }
@@
 In the absence of other mechanisms, the timescale of the orbital decay of the system is set by the energy loss due to gravitational waves, which in turn is set by the mass of constituent compact objects and the separation of the two \citep{Postnov2014}.
-If we assume that the formation timescale of the first galaxies is short compared to the time since the big bang \citep{Richard2011} and if we assume a mass of 1.4 M$_\odot$ for each of the neutron stars at the time of system formation, this places a limit on the initial separation of the two neutronstars of a$_0 < 3.1 $R$_\odot$, where a$_0$ is the initial separation.
-Because we have the age of the stellar population, by using this as a limit to the orbital decay time we can put tighter constraints, a$_0 < 1.7 $R$_\odot$, but this does not account for the possibility there could be an underlying stellar population of older stars from a previous star-formation epoch, which is why we consider a$_0 < 3.1 $R$_\odot$ as a hard upper limit.
+If we assume that the formation timescale of the first galaxies is short compared to the time since the big bang \citep{Richard2011} and if we assume a mass of 1.4 M$_\odot$ for each of the neutron stars at the time of system formation, this places a limit on the initial separation of the two neutronstars of a$_0 < 3.2 $R$_\odot$, where a$_0$ is the initial separation.
+Because we have the age of the stellar population, by using this as a limit to the orbital decay time we can put tighter constraints, a$_0 < 2.1 $R$_\odot$, but this does not account for the possibility there could be an underlying stellar population of older stars from a previous star-formation epoch, which is why we consider a$_0 < 3.2 $R$_\odot$ as a hard upper limit.
 The delay time between formation and explosion is well accomodated by the models of \citet{Belczynski2006}, although the longest formation channels are excluded.
@@
 Using the age of the universe at the time of explosion allows us to set constraints on the maximal separation between the engine constituents at the time of formation.
-We find that the maximal separation for two neutron stars at formation time is a$_0 < 3.1 $R$_\odot$, which excludes some of the formation channels with the longest timescales.
+We find that the maximal separation for two neutron stars at formation time is a$_0 < 3.2 $R$_\odot$, which excludes some of the formation channels with the longest timescales.
+
+We make all the data, code and calculation related to the paper along with the paper itself available at \url{https://github.com/jselsing/GRB111117A}.
 

### 2017-05-24 py/progenitor_seperation.py

@@ -1,0 +1,54 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+from __future__ import division, print_function
+
+# Imports
+from astropy import units as u
+from astropy import constants as c
+
+
+def calc_sep(age, m1, m2):
+    """
+    Small calculation of progenitor seperation.
+    age in Myr, progenitor masses in Msun
+    returns seperation in units of Rsun
+    """
+    # Convert to seconds
+    t0 = (age*1e6 * u.yr).to(u.s)
+    # Convert to kg
+    M_ns1 = (1.4 * u.Msun).to(u.kg)
+    M_ns2 = (1.4 * u.Msun).to(u.kg)
+
+    # Total system mass
+    M = M_ns1 + M_ns2
+    # Reduced mass
+    mu = M_ns1*M_ns2/M
+
+    # Calculate initial progenitor seperation from eq 29 from Postnov and Yungelson 2014
+    a0 = ((t0 * ((c.G**3)/(c.c**5)) * ((256 * M**2 * mu)/5))**(1/4))
+
+    return a0.value/u.Rsun.to(u.m) * u.Rsun
+
+
+def main():
+
+    # Calculate initial progenitor seperation from eq 29 from Postnov and Yungelson 2014
+    # Time since big bang
+    t_constraint = 2970 #Myr
+    # t_constraint = 500 #Myr
+    # Neutron star masses
+    m1 = 1.4 # Msun
+    m2 = 1.4 # Msun
+    # Get seperation
+    a0 = calc_sep(t_constraint, m1, m2)
+
+    print(a0)
+
+
+if __name__ == '__main__':
+    main()
+
+
+
+

### 2017-05-24 GRB111117A.tex
Circulated paper

@@
 The redshift of GRB111117A is evidence against a lognormal delay-time model for sGRBs through the predicted redshift distribution of sGRBs, which is very sensitive to high-z sGRBs.
-From the age of the universe at explosion time, an initial progenitor separation of a$_0 < 3.1 $R$_\odot$ is required for the case of a binary neutron star (NS) system.
+From the age of the universe at explosion time, an initial progenitor separation of a$_0 < 3.2 $R$_\odot$ is required for the case of a binary neutron star (NS) system.
 This puts constraints on the progenitor system evolution up to the time of explosion.
@@
 The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and shows significant velocity structure along the slit.
-A drop in the continuum the the blue of the \lya line further supports the inferred redshift.
+A drop in the continuum to the blue of the \lya~line further supports the inferred redshift.
 
@@
 
-A measure of the optical-to-X-ray flux ratio is parametrized in terms of the "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}.
+%A measure of the optical-to-X-ray flux ratio is parametrized in terms of the "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}.
 Using the optical afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time of the non-detection.
@@
 
-Because the projected distance does not change much between redshift 1.3 and 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating to host offset are unaffected.
+%Because the projected distance does not change much between redshift 1.3 and 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating to host offset are unaffected.
 

### 2017-05-28 GRB111117A.tex
Implement Jures comments
Closes #1

@@
 
-\title{The host galaxy of the short GRB111117A at z = 2.211.
-\thanks{Based on observations collected at the European Southern Observatory, Paranal, 
-	Chile, Program ID: 088.A-0051 and 091.D-0904.}}
+\title{The host galaxy of the short GRB~111117A at z = 2.211\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}}
 
-\titlerunning{GRB111117A}
+\titlerunning{GRB~111117A}
 
@@
 Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to localize.
-We here present evidence of the, to date, highest redshift high-confidence short GRB, located at $z=2.211$.
+We here present a spectroscopic observation of the host galaxy of GRB 111117A and measure its redshift to be $z = 2.211$, making it the most distant high-confidence short GRB with a secure redshift measurement.
 This value supersedes a previously estimated, lower redshift value.
@@
 The rest frame X-ray derived hydrogen column density is high compared to a complete sample of sGRBs and seems to follow the evolution with redshift as traced by the hosts of long GRBs (lGRBs).
-This could indicate that the part of the sGRB population hosted in late-type galaxies, live in similar environments to that of lGRB hosts.
-The host is in the brighter end of the expected host brightness distribution at $z = 2.211$, but only 30 per cent of the distribution should be missed due to host faintness at this redshift.
+This could indicate that the part of the sGRB population hosted in late-type galaxies live in similar environments to that of lGRB hosts.
+The host lies in the brighter end of the expected host brightness distribution at $z = 2.211$, and is actively star forming.
+Using the host as a benchmark for redshift determination, we assess that only 30 per cent of sGRB redshifts should be missed due to host faintness at this redshift.
 This indicates that we are not missing redshifts for a dominant fraction of the sGRB hosts at this redshift.
-The redshift of GRB111117A is evidence against a lognormal delay-time model for sGRBs through the predicted redshift distribution of sGRBs, which is very sensitive to high-z sGRBs.
-From the age of the universe at explosion time, an initial progenitor separation of a$_0 < 3.2 $R$_\odot$ is required for the case of a binary neutron star (NS) system.
+The redshift of GRB~111117A is evidence against a lognormal delay-time model for sGRBs through the predicted redshift distribution of sGRBs, which is very sensitive to high-z sGRBs.
+%From the age of the universe at explosion time, an initial progenitor separation of a$_0 < 3.2 $R$_\odot$ is required for the case of a binary neutron star (NS) system.
 This puts constraints on the progenitor system evolution up to the time of explosion.
-
 }
@@
 
-There is now mounting evidence that most short-duration gamma-ray bursts (GRBs) come from the merger of neutron stars (NSs), either with another NS, or a black hole, due to their apparent association with kilonovae
-\citep{Barnes2013a, Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The
-absence of associated supernovae in deep searches \citep[e.g.][]
-{Hjorth2005a,Fox2005,Hjorth2005b} supports this idea and distinguishes the physical origin of sGRBs from their long-duration counterparts.
+There is now mounting evidence that most short-duration gamma-ray bursts (GRBs) come from the merger of neutron stars (NSs), either with another NS, or a black hole, due to their apparent association with kilonovae \citep{Barnes2013a, Tanvir2013b, Yang2015, Jin2016, Rosswog2016}.
+The absence of associated supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b} supports this idea and distinguishes the physical origin of sGRBs from their long-duration counterparts.
 %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}. 
 The range of host galaxy types, in some cases apparently absent, and typically more massive 
-and on average less actively star-forming than long GRB hosts  \citep{Fong2013b}, as well as their positions within
-hosts \citep{Fong2013a}, suggest an origin for the progenitors that can be very 
-long-lived, and associated with stellar mass rather than purely 
-star-formation rate \citep{Berger2014}.
+and on average less actively star-forming than long GRB hosts  \citep{Fong2013b}, as well as their positions within hosts \citep{Fong2013a}, suggest an origin for the progenitors that can be very long-lived, and associated with stellar mass rather than purely star-formation rate \citep{Berger2014}.
+The bulk of sGRBs have associated galaxies at relatively low redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts are measured from the hosts, this value is biased towards low $z$.
+%This progress on the progenitors of sGRBs has been enabled by the afterglow detection and burst localization by the \emph{Swift} satellite \citep{Gehrels2009a}.
+The bulk of these localizations have associated galaxies at relatively low redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts are measured from the hosts, this value is biased towards low $z$.
 
-This progress on the progenitors of sGRBs has been enabled by
-localization by the \emph{Swift} satellite \citep{Gehrels2009a}.
-The bulk of these localizations have associated galaxies at relatively low redshifts with a median redshift $z\sim0.5$ \citep{Berger2014}.
-Because the redshifts are measured from the hosts, measurements of the redshift distribution of sGRBs is biased towards low redshift due to cosmological dimming.
-%at least partly due to the faintness of the bursts and their afterglows, though this is not clear. 
-So far, the electromagnetic signal from sGRBs are the only means to accurately localize NS mergers and holds the promise for detection of an associated gravitational wave signal \citep{Ghirlanda2016}.
+%So far, the electromagnetic signal from sGRBs are the only means to accurately localize NS mergers and holds the promise for detection of an associated gravitational wave signal \citep{Ghirlanda2016}.
 From the burst and following afterglow, modeling the properties and mapping the environments additionally allows insights into the phenomenon itself.
@@
 
-In this \emph{Letter} we present the spectrum of the host galaxy of the short
-GRB\,111117A, showing it to be at higher redshift than previously
-estimated. We present the rest frame burst properties based on this new distance
+In this \emph{Letter} we present the spectrum of the host galaxy of the short GRB 111117A at z=2.211, a significantly higher redshift than the previously estimated redshift based on photometric studies.
+We present the rest frame burst properties based on this new distance
 compared to previous analyses \citep{Margutti2012,Sakamoto2013} and revisit the host properties derived from the new solution to the SED fit. 
@@
 Only one slit is shown, despite 4 epochs of spectroscopic observations because of the similarity in position angle.
-The image is the FORS2 R-band image, for which the photometry is shown in Fig.
+The image is the FORS2 R-band image and%, for which the photometry is shown in Fig.
 \ref{fig:SED}.
-The blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}}
+	the blue asterisk i ndicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}.
+The errorcircle of the \emph{Chandra}-position is too small to resolve in the image.}
 	\label{fig:spec_setup}
@@
 2017 (in prep.) and available online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}.
-The signal-to-noise of the continuum is the near-infrared arm is too low to use the optimal extraction algorithm \citep{Horne1986}, why an aperture containing the entire object has been chosen for which the flux is summed and the corresponding error is summed in quadrature.
+The signal-to-noise of the continuum in the near-infrared arm is too low to use the optimal extraction algorithm \citep{Horne1986} and therefore the extraction is carried out with a simple aperture.
 An overview of the spectroscopic observations is given in Tab. \ref{tab:spec_overview}, and the position of the slit on the target is shown in Fig. \ref{fig:spec_setup}.
@@
 
-We determine a redshift of $z = 2.211$ from from the simultaneous detection of emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with \hb~detected at low significance ($\sim 3 \sigma$).
-\ha is only visible in the first epoch, due to the K-band blocking filter used for the remainder observations.
-The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and shows significant velocity structure along the slit.
+We determine a redshift of $z = 2.211$ from the simultaneous detection of emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with \hb~detected at low significance ($\sim 3 \sigma$).
+\ha~is only visible in the first epoch, due to the K-band blocking filter used for the remainder observations.
+The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show significant velocity structure along the slit.
 A drop in the continuum to the blue of the \lya~line further supports the inferred redshift.
@@
 
-The multi-color spectral energy distribution (SED) is then fit by \citet{Bruzual2003} stellar population synthesis models based on a \citet{Chabrier2003} initial mass function in \emph{LePhare} \citep{Ilbert2006}, where the redshift if fixed to the spectroscopic value of $z=2.211$.
+The multi-color spectral energy distribution (SED) is then fit by \citet{Bruzual2003} stellar population synthesis models based on a \citet{Chabrier2003} initial mass function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the spectroscopic value of $z=2.211$.
 The best model is obtained with an unreddened galaxy template, and returns physical parameters of luminosity ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), single stellar population age ($\tau = 0.5_{-0.25}^{+0.5}$ Gyr) and star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$).
@@
 In fact, the large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break driving the galaxy photometric redshift of both of these earlier works.
-It is this exactly this overestimate in $z$-band brightness that led the authors to erronously low estimates in their photo-$z$ for GRB~111117A.
 Using the revised photometry from Table~\ref{tab:phot_overview}, the photo-$z$ of the galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic value at the 1~$\sigma$ confidence level.
@@
 
-As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB111117A very securely belongs the the short class of GRBs.
+As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A very securely belongs the short class of GRBs.
 Because the observed classification indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not change this designation significantly.
-The intrinsic luminosity goes up, as visible in the light curve in Fig.
+The intrinsic luminosity increases, as visible in the light curve in Fig.
 \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to majority of long GRBs.
 \citet{Bromberg2013} investigated the degree to which the long and short population distributions overlap and quantified the certainty in the class membership.
-GRB111117A has $96_{-5}^{+3}$ percent probability of being short.
-This is compared to the other two highest-z short candidates, GRB060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent).
+GRB~111117A has $96_{-5}^{+3}$ percent probability of being short.
+Compared to the other two highest-z short candidates, GRB060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class membership for GRB~111117A is much higher.
 
@@
 	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
-	\caption{Restframe XRT lightcurve of GRB111117A, compared to the general population of XRT lightcurves of GRBs.
+	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
 The grey shared region is a compilation of long GRB lightcurves where the color represents density and the light blue is other short GRB lightcurves for which redshifts has been determined.
@@
 Bursts without redshifts have been excluded for both groups.
-GRB111117A occupies a unique position in Fig.
-\ref{fig:NH_z} with a the highest $N_H$ of all short burst.
-The short sample, excluding GRB111117A, is located at low redshifts(z < 1) and are found to populate a similar column density environment to long GRBs at similar redshifts \citep{DAvanzo2014a}.
-The inferred hydrogen column for GRB111117A seems to follow the trend with increasing $N_H$ as a function of redshift as found for the long GRB hosts \citep{Arcodia2016}.
+GRB~111117A occupies a unique position in Fig.
+\ref{fig:NH_z} with the highest $N_H$ of all short burst.
+The short sample, excluding GRB~111117A, is located at low redshifts(z < 1) and are found to populate a similar column density environment to long GRBs at similar redshifts \citep{DAvanzo2014a}.
+The inferred hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$ as a function of redshift as found for the long GRB hosts \citep{Arcodia2016}.
 
@@
 	\includegraphics[width=9cm]{figures/NH_z.pdf}
-	\caption{Rest frame, X-ray derived hydrogen column densities for GRB111117A compared to complete samples of both long and short populations.
+	\caption{Rest frame, X-ray derived hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations.
 In red is shown the long sample of \citet{Arcodia2016}.
@@
 In blue is the sample from \citet{DAvanzo2014a}.
-GRB111117A is an outlier from the short sample, but seems to follow the distribution of long GRBs.}
+GRB~111117A is an outlier from the short sample, but seems to follow the distribution of long GRBs.}
 	\label{fig:NH_z}
@@
 
-From the clear host association, GRB111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits clear emission lines indicative of a population of relatively young stars, as the majority of short GRBs, the host is a late-type galaxy \citep{Fong2013b}.
-The host of GRB111117A is entirely consistent in terms of stellar mass and stellar age with the general population of short GRB hosts($\left\langle M _* \right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle = 0.3 $ Gyr) \citet{Leibler2010}.
+%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits clear emission lines indicative of a population of relatively young stars,
+As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a late-type galaxy and is entirely consistent in terms of stellar mass and stellar age with the general population of short GRB hosts($\left\langle M _* \right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle = 0.3 $ Gyr) \citet{Leibler2010}.
 The SFR, on the other hand, is $\sim$1 order of magnitude higher than the typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to the SFRs found in the hosts of long GRBs at a corresponding redshift \citep{Kruhler2015}.
@@
 26 out of 39 hosts (66 per cent) in the sample has redshifts.
-To convert the SED-inferred $M_B$ of GRB111117A to $L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window $2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with measured $L_B/L^{\star}_{B}$.
-If we assume that we are able to get emission-line redshifts from hosts with $R < 25 $mags \citep{Kruhler2012}, then we would have missed around 30\% (8 out of 26 from the sample of \citealt{Berger2014} with measured $L_B/L^{\star}_{B}$), if they were at the redshift of GRB111117A.
+To convert the SED-inferred $M_B$ of GRB~111117A to $L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window $2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with measured $L_B/L^{\star}_{B}$.
+If we assume that we are able to get emission-line redshifts from hosts with $R < 25$ mags \citep{Kruhler2012}, then we would have missed around 30\% (8 out of 26 from the sample of \citealt{Berger2014} with measured $L_B/L^{\star}_{B}$), if they were at the redshift of GRB~111117A.
 If we assume that the hosts missing redshift in \citet{Berger2014} follow the redshift distribution of the hosts \textit{with} redshift, this implies that we are \textit{not} missing a dominant fraction of sGRB redshift at $z \approx 2$, due to host faintness.
 
-The theoretical redshift distributions of sGRBs depends on the type of delay-time function used to model the progenitor system.
-The redshift of GRB111117A puts constraints on the type of delay-time models suitable for modeling.
+The theoretical redshift distribution of sGRBs depends on the type of delay-time function used to model the progenitor system.
+The redshift of GRB~111117A puts constraints on the type of delay-time models suitable for modeling.
 The likelihood preferred lognormal time delay models investigated in \citet{Wanderman2015} predicts a rate of sGRBs at $z = 2.211$, $\sim$ 2 orders of magnitude lower compared to peak rate (z = 0.9).
@@
 In the absence of other mechanisms, the timescale of the orbital decay of the system is set by the energy loss due to gravitational waves, which in turn is set by the mass of constituent compact objects and the separation of the two \citep{Postnov2014}.
-If we assume that the formation timescale of the first galaxies is short compared to the time since the big bang \citep{Richard2011} and if we assume a mass of 1.4 M$_\odot$ for each of the neutron stars at the time of system formation, this places a limit on the initial separation of the two neutronstars of a$_0 < 3.2 $R$_\odot$, where a$_0$ is the initial separation.
-Because we have the age of the stellar population, by using this as a limit to the orbital decay time we can put tighter constraints, a$_0 < 2.1 $R$_\odot$, but this does not account for the possibility there could be an underlying stellar population of older stars from a previous star-formation epoch, which is why we consider a$_0 < 3.2 $R$_\odot$ as a hard upper limit.
+If we assume that the formation timescale of the first galaxies is short compared to the time since the big bang \citep{Richard2011} and if we assume a mass of 1.4 M$_\odot$ for each of the neutron stars at the time of system formation, this places a limit on the initial separation of the two neutron stars of a$_0 < 3.2 $R$_\odot$, where a$_0$ is the initial separation.
+Assuming our measured age of the stellar population ($\tau = 0.5$ Gyr) as a limit for the orbital decay time we can put a tighter constraint of a$_0 < 2.1 $R$_\odot$.
+However, this does not account for the possibility there could be an underlying stellar population of older stars from a previous star-formation epoch, which is why we consider a$_0 < 3.2 $R$_\odot$ as a hard upper limit.
 The delay time between formation and explosion is well accomodated by the models of \citet{Belczynski2006}, although the longest formation channels are excluded.
@@
 
-We have in this letter provided a revised, spectroscopic redshift for GRB111117A based on emission lines discovered using an improved reduction scheme, setting it at $z = 2.211$.
+We have in this letter provided a revised, spectroscopic redshift for GRB~111117A based on emission lines discovered using an improved reduction scheme, setting it at $z = 2.211$.
 This value supersedes the previous photometric redshift of $z \sim 1.3$ \citep{Margutti2012, Sakamoto2013}.
@@
 The rest-frame parameters of the burst and the conditions of the burst environment have been recalculated using the new distance.
-The X-ray derived hydrogen column density towards GRB111117A is the highest within a complete sample of short hosts, but seems to trace the evolution with redshift as found for the hosts of long GRBs.
+The X-ray derived hydrogen column density towards GRB~111117A is the highest within a complete sample of short hosts, but seems to trace the evolution with redshift as found for the hosts of long GRBs.
 
@@
 
-Although a single burst carries little leverage in terms of constraining the redshift distribution of sGRB, the high redshift of GRB111117A needs to be accommodated.
-A lognormal delay time model predicts a very low volumetric density of bursts at $z = 2.211$, whereas a power law delay time model peaks near GRB111117A.
+Although a single burst carries little leverage in terms of constraining the redshift distribution of sGRB, the high redshift of GRB~111117A needs to be accommodated.
+A lognormal delay time model predicts a very low volumetric density of bursts at $z = 2.211$, whereas a power law delay time model peaks near GRB~111117A.
 If more sGRBs are at this redshift, but missed due to faintness of their hosts, a lognormal delay time model will be disfavored.
-Compared to a sample of short hosts, GRB111117A is more luminous than 70 per cent of the sample with measured luminosities.
-Of the host sample, for 30 per cent, we would be unable to determine a redshift, should they be at the redshift of GRB111117A.
+Compared to a sample of short hosts, GRB~111117A is more luminous than 70 per cent of the sample with measured luminosities.
+Of the host sample, for 30 per cent, we would be unable to determine a redshift, should they be at the redshift of GRB~111117A.
 This implies that we are \textit{not} missing a dominant fraction of the sGRBs at $z \sim 2$.
@@
 \begin{acknowledgements}
-TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
+TK acknowledges support through the Sofja Kovalevskaja Award to P.
+Schady.
+CT acknowledges support from a Spanish National Research Grant of Excellence under project AYA 2014-58381-P and funding associated to a Ramón y Cajál fellowship under grant number RyC-2012-09984.
 \end{acknowledgements}

### 2017-05-28 GRB111117A.tex
Add co-authors and add line-breaks to better track changes

@@
 \abstract{
-Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to localize.
-We here present a spectroscopic observation of the host galaxy of GRB 111117A and measure its redshift to be $z = 2.211$, making it the most distant high-confidence short GRB with a secure redshift measurement.
-This value supersedes a previously estimated, lower redshift value.
-Spectroscopic observations along with a wealth of imaging data allow us to place tight constraints on the nature of the host galaxy at an unprecedented distance.
-The rest frame X-ray derived hydrogen column density is high compared to a complete sample of sGRBs and seems to follow the evolution with redshift as traced by the hosts of long GRBs (lGRBs).
-This could indicate that the part of the sGRB population hosted in late-type galaxies live in similar environments to that of lGRB hosts.
-The host lies in the brighter end of the expected host brightness distribution at $z = 2.211$, and is actively star forming.
-Using the host as a benchmark for redshift determination, we assess that only 30 per cent of sGRB redshifts should be missed due to host faintness at this redshift.
-This indicates that we are not missing redshifts for a dominant fraction of the sGRB hosts at this redshift.
-The redshift of GRB~111117A is evidence against a lognormal delay-time model for sGRBs through the predicted redshift distribution of sGRBs, which is very sensitive to high-z sGRBs.
-%From the age of the universe at explosion time, an initial progenitor separation of a$_0 < 3.2 $R$_\odot$ is required for the case of a binary neutron star (NS) system.
-This puts constraints on the progenitor system evolution up to the time of explosion.
+Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to
+localize. We here present a spectroscopic observation of the host galaxy of GRB
+111117A and measure its redshift to be $z = 2.211$, making it the most distant
+high-confidence short GRB with a secure redshift measurement. This value
+supersedes a previously estimated, lower redshift value. Spectroscopic
+observations along with a wealth of imaging data allow us to place tight
+constraints on the nature of the host galaxy at an unprecedented distance.
+
+The rest frame X-ray derived hydrogen column density is high compared to a
+complete sample of sGRBs and seems to follow the evolution with redshift as
+traced by the hosts of long GRBs (lGRBs). This could indicate that the part of
+the sGRB population hosted in late-type galaxies live in similar environments to
+that of lGRB hosts.
+
+The host lies in the brighter end of the expected host brightness distribution
+at $z = 2.211$, and is actively star forming. Using the host as a benchmark for
+redshift determination, we assess that only 30 per cent of sGRB redshifts should
+be missed due to host faintness at this redshift. This indicates that we are not
+missing redshifts for a dominant fraction of the sGRB hosts at this redshift.
+The redshift of GRB~111117A is evidence against a lognormal delay-time model for
+sGRBs through the predicted redshift distribution of sGRBs, which is very
+sensitive to high-z sGRBs.
+
+From the age of the universe at explosion time, an initial progenitor separation
+of a$_0 < 3.2 $R$_\odot$ is required for the case of a binary neutron star (NS)
+system. This puts constraints on the progenitor system evolution up to the time
+of explosion.
 }
@@
 
-There is now mounting evidence that most short-duration gamma-ray bursts (GRBs) come from the merger of neutron stars (NSs), either with another NS, or a black hole, due to their apparent association with kilonovae \citep{Barnes2013a, Tanvir2013b, Yang2015, Jin2016, Rosswog2016}.
-The absence of associated supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b} supports this idea and distinguishes the physical origin of sGRBs from their long-duration counterparts.
-%\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}. 
-The range of host galaxy types, in some cases apparently absent, and typically more massive 
-and on average less actively star-forming than long GRB hosts  \citep{Fong2013b}, as well as their positions within hosts \citep{Fong2013a}, suggest an origin for the progenitors that can be very long-lived, and associated with stellar mass rather than purely star-formation rate \citep{Berger2014}.
-The bulk of sGRBs have associated galaxies at relatively low redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts are measured from the hosts, this value is biased towards low $z$.
-%This progress on the progenitors of sGRBs has been enabled by the afterglow detection and burst localization by the \emph{Swift} satellite \citep{Gehrels2009a}.
-The bulk of these localizations have associated galaxies at relatively low redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts are measured from the hosts, this value is biased towards low $z$.
+There is now mounting evidence that most short-duration gamma-ray bursts (GRBs)
+come from the merger of neutron stars (NSs), either with another NS, or a black
+hole, due to their apparent association with kilonovae \citep{Barnes2013a,
+	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
+supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b}
+supports this idea and distinguishes the physical origin of sGRBs from their
+long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006,
+% Gal-Yam2006}.
+The range of host galaxy types, in some cases apparently absent, and typically
+more massive and on average less actively star-forming than long GRB hosts 
+\citep{Fong2013b}, as well as their positions within hosts \citep{Fong2013a},
+suggest an origin for the progenitors that can be very long-lived, and
+associated with stellar mass rather than purely star-formation rate
+\citep{Berger2014}. The bulk of sGRBs have associated galaxies at relatively low
+redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts
+are measured from the hosts, this value is biased towards low $z$.
 
-%So far, the electromagnetic signal from sGRBs are the only means to accurately localize NS mergers and holds the promise for detection of an associated gravitational wave signal \citep{Ghirlanda2016}.
-From the burst and following afterglow, modeling the properties and mapping the environments additionally allows insights into the phenomenon itself.
+%This progress on the progenitors of sGRBs has been enabled by the afterglow
+% detection and burst localization by the \emph{Swift} satellite
+% \citep{Gehrels2009a}. The bulk of these localizations have associated galaxies
+% at relatively low redshifts with a median $z\sim0.5$ \citep{Berger2014} and
+% because most redshifts are measured from the hosts, this value is biased
+% towards low $z$.
 
-The total lifetimes of NS binaries is dependent on their initial separations and subsequent inspiral times, and impacts the timing and distribution of the enrichment of the ISM and subsequent stars and planets with heavy $r$-process elements \citep{VandeVoort2015, Wallner2015,  Ji2016}.
-Some limits can be calculated based on models of star-formation histories of, and the spatial distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}.
-The most distant cosmological bursts, however, offer direct, hard limits.
+%So far, the electromagnetic signal from sGRBs are the only means to accurately
+% localize NS mergers and holds the promise for detection of an associated
+% gravitational wave signal \citep{Ghirlanda2016}. From the burst and following
+% afterglow, modeling the properties and mapping the environments additionally
+% allows insights into the phenomenon itself.
 
-In this \emph{Letter} we present the spectrum of the host galaxy of the short GRB 111117A at z=2.211, a significantly higher redshift than the previously estimated redshift based on photometric studies.
-We present the rest frame burst properties based on this new distance
-compared to previous analyses \citep{Margutti2012,Sakamoto2013} and revisit the host properties derived from the new solution to the SED fit. 
-%While no optical afterglow was detected, the excellent localization from a detection
+The total lifetimes of NS binaries is dependent on their initial separations and
+subsequent inspiral times, and impacts the timing and distribution of the
+enrichment of the ISM and subsequent stars and planets with heavy $r$-process
+elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some limits can be
+calculated based on models of star-formation histories of, and the spatial
+distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}. The most
+distant cosmological bursts, however, offer direct, hard limits.
+
+In this \emph{Letter} we present the spectrum of the host galaxy of the short
+GRB 111117A at z=2.211, a significantly higher redshift than the previously
+estimated redshift based on photometric studies. We present the rest frame burst
+properties based on this new distance compared to previous analyses
+\citep{Margutti2012,Sakamoto2013} and revisit the host properties derived from
+the new solution to the SED fit. %While no optical afterglow was detected, the
+% excellent localization from a detection
 %of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also allows
@@
 Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology provided by
-\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$ km s$^{-1}$ and
-$\Omega_m = 0.307$ and report all magnitudes in the AB system.
+\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$ km s$^{-1}$
+and $\Omega_m = 0.307$ and report all magnitudes in the AB system.
 
@@
 
-%(21000 \AA when the K-band blocking filter is used)
-Spectroscopic observations were carried out using the cross-dispersed echelle spectrograph, VLT/X-shooter \citep{Vernet2011}, in four seperate epochs.
-The burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO program 091.D-0904 (PI: Hjorth).
-X-shooter can cover the wavelength range from 3000 \AA~to 24800 \AA~across three spectroscopic arms, simultaneously illuminated through the use of dichroics.
-The bias-correction, flat-fielding, order tracing, wavelength calibration, rectification, and flux calibration is then carried out using the VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} and managed by Reflex \citep{Freudling2013}.
-Because the spectra are curved across each detector, a rectification algorithm is employed which introduces correlations between neighboring pixels.
-We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the degree of correlation while conserving the maximal resolution.
-The observations are combined and extracted using scripts described in Selsing et al.
-2017 (in prep.) and available online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}.
-The signal-to-noise of the continuum in the near-infrared arm is too low to use the optimal extraction algorithm \citep{Horne1986} and therefore the extraction is carried out with a simple aperture.
+Spectroscopic observations were carried out using the cross-dispersed echelle
+spectrograph, VLT/X-shooter \citep{Vernet2011}, in four seperate epochs. The
+burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger
+under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO
+program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
+3000 \AA~to 24800 \AA~(21000 \AA~when the K-band blocking filter is used)~across
+three spectroscopic arms, simultaneously illuminated through the use of
+dichroics. The bias-correction, flat-fielding, order tracing, wavelength
+calibration, rectification, and flux calibration is then carried out using the
+VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}. 
+%Because the spectra are curved across each detector, a rectification algorithm
+%is employed which introduces correlations between neighboring pixels. We select
+%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
+%degree of correlation while conserving the maximal resolution.
+The observations are combined and extracted using scripts described in Selsing
+et al. 2017 (in prep.) and available
+online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
+signal-to-noise of the continuum in the near-infrared arm is too low to use the
+optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
+carried out with a simple aperture.
+
 An overview of the spectroscopic observations is given in Tab. \ref{tab:spec_overview}, and the position of the slit on the target is shown in Fig. \ref{fig:spec_setup}.
@@
 
-We determine a redshift of $z = 2.211$ from the simultaneous detection of emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with \hb~detected at low significance ($\sim 3 \sigma$).
-\ha~is only visible in the first epoch, due to the K-band blocking filter used for the remainder observations.
-The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show significant velocity structure along the slit.
-A drop in the continuum to the blue of the \lya~line further supports the inferred redshift.
+We determine a redshift of $z = 2.211$ from the simultaneous detection of
+emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with
+\hb~detected at low significance ($\sim 3 \sigma$). \ha~is only visible in the
+first epoch, due to the K-band blocking filter used for the remainder
+observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and
+show significant velocity structure along the slit. A drop in the continuum to
+the blue of the \lya~line further supports the inferred redshift.
 
-%The total extent of the lines in velocity space is $\sim$ 450 km/s.
-The line profiles shows an asymmetric "double-horned" profile, indicating that we are seeing a galaxy with a large degree of coherent rotational motion relative to the line-of-sight.
-If we assume that we are viewing a spiral galaxy exactly edge-on, this is a measure of the rotational velocity of the gas.
-If we assume that the spectral resolution and the turbulent width of the lines are negligible compared to the rotational velocity, we can based on the projected size of the source and the width of the lines, put a constraint on the dynamical mass of the galaxy.
-Based on the physical scale from extent and the velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$.
-Because we are viewing the host at an angle, this value is an upper limit, which should be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.
+%The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
+% profiles shows an asymmetric "double-horned" profile, indicating that we are
+% seeing a galaxy with a large degree of coherent rotational motion relative to
+% the line-of-sight. If we assume that we are viewing a spiral galaxy exactly
+% edge-on, this is a measure of the rotational velocity of the gas. If we assume
+% that the spectral resolution and the turbulent width of the lines are
+% negligible compared to the rotational velocity, we can based on the projected
+% size of the source and the width of the lines, put a constraint on the
+% dynamical mass of the galaxy. Based on the physical scale from extent and the
+% velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we
+% are viewing the host at an angle, this value is an upper limit, which should
+% be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.
 
-Using the integrated flux of \ha, we can infer the SFR \citep{Kennicutt1998, Moustakas2006}.
-From the SED-fit, see Sect.
-\ref{SED}, and the detection of \lya, the host is constrained to contain very little or no dust which is also confirmed by the shape of the spectral continuum, which is why we do not apply a dust-correction.
-Because of the redshift, \ha~falls on the far end of the NIR-arm where the night sky is very bright and additionally, several bright sky-lines intersect the line, making an accurate estimate of the \ha-flux difficult.
-%Because of the challenging measurement, two approaches are employed.
-First,
-A limit on the SFR is obtained by integrating the part of \ha~free of contamination and conservatively correcting for the missing fraction by assessing the fraction integrated.
-Converting to a \citet{Chabrier2003} initial mass function using \citet{Madau2014} we get a limit of SFR > 7 M$_\odot$ yr$^{-1}$.
-%Secondly, based on the width of \oiii, an aperture covering the line is integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have been masked and interpolated over.
-From the integrated \ha-line, we estimate SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.
+Using the integrated flux of \ha, we can infer the SFR \citep{Kennicutt1998}.
+From the SED-fit, see Sect. \ref{SED}, and the detection of \lya, the host is
+constrained to contain very little or no dust which is also confirmed by the
+shape of the spectral continuum, which is why we do not apply a dust-correction.
+Because of the redshift, \ha~falls on the far end of the NIR-arm where the night
+sky is very bright and additionally, several bright sky-lines intersect the
+line, making an accurate estimate of the \ha-flux difficult. A limit on the SFR
+is obtained by integrating the part of \ha~free of contamination and
+conservatively correcting for the missing fraction by assessing the fraction
+integrated. Converting to a \citet{Chabrier2003} initial mass function using
+\citet{Madau2014} we get a limit of SFR > 7 M$_\odot$ yr$^{-1}$.
+
+%Secondly, based on the width of \oiii, an aperture covering the line
+% is integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have
+% been masked and interpolated over. From the integrated \ha-line, we estimate
+% SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.
 
@@
 
-%A measure of the optical-to-X-ray flux ratio is parametrized in terms of the "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}.
-Using the optical afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time of the non-detection.
-We find $\beta_{OX} < 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
+%A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
+% "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical
+% afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
+% interpolated and evaluated at the time of the non-detection. We find
+% $\beta_{OX} < 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
 
-%Because the projected distance does not change much between redshift 1.3 and 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating to host offset are unaffected.
+%Because the projected distance does not change much between redshift 1.3 and
+% 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating
+% to host offset are unaffected.
 
@@
 
-We additionally imaged the field of GRB 111117A in multiple broad-band filters using the VLT equipped with FORS2 ($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst has faded.
-These new data are complemented by a re-analysis of some of the imagery used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us here (GTC $gri$-band, TNG $R$-band, and Gemini $z$-band).
-A log of the photometric observations and measured brightnesses is given in Tab.~\ref{tab:phot_overview}.
+We additionally imaged the field of GRB 111117A in multiple broad-band filters
+using the VLT equipped with FORS2 ($gRIz$ filters) and HAWK-I
+($JHK_{\mathrm{s}}$ filters), long after the burst has faded. These new data are
+complemented by a re-analysis of some of the imagery used in
+\citet{Margutti2012} and \citet{Sakamoto2013} that are available to us here (GTC
+$gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
+observations and measured brightnesses is given in Tab.~\ref{tab:phot_overview}.
 
-All data were reduced, analyzed and fitted in a similar manner as we have done previously and described in detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
-Very briefly, we use our own python and IRAF routines to perform a consistent standard reduction of the available imaging which includes bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and stacking of individual images for a given instrument in a given filter.
-The photometry of the host was tied against magnitudes of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters.
-For our $R$ and $I$-band photometry we used the color transformations of Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}.
+All data were reduced, analyzed and fitted in a similar manner as we have done
+previously and described in detail in \citet{Kruhler2011a} and, more recently,
+in \citet{Schulze2016}. Very briefly, we use our own python and IRAF routines to
+perform a consistent standard reduction of the available imaging which includes
+bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
+stacking of individual images for a given instrument in a given filter. The
+photometry of the host was tied against magnitudes of field stars from the SDSS
+and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. For our $R$
+and $I$-band photometry we used the color transformations of
+Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}.
 We subsequently convert all magnitudes into the AB system if necessary, and correct for the Galactic foreground of $E_{B-V}=0.027~\mathrm{mag}$.
 
-The multi-color spectral energy distribution (SED) is then fit by \citet{Bruzual2003} stellar population synthesis models based on a \citet{Chabrier2003} initial mass function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the spectroscopic value of $z=2.211$.
-The best model is obtained with an unreddened galaxy template, and returns physical parameters of luminosity ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), single stellar population age ($\tau = 0.5_{-0.25}^{+0.5}$ Gyr) and star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$).
-%We show the SED fit in Fig.
-\ref{fig:SED}.
+The multi-color spectral energy distribution (SED) is then fit by
+\citet{Bruzual2003} stellar population synthesis models based on a
+\citet{Chabrier2003} initial mass function in \emph{LePhare} \citep{Ilbert2006},
+where the redshift is fixed to the spectroscopic value of $z=2.211$. The best
+model is obtained with an unreddened galaxy template, and returns physical
+parameters of luminosity ($M_B=-22.0\pm0.1$\,mag), stellar mass
+($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), single stellar population age ($\tau =
+0.5_{-0.25}^{+0.5}$ Gyr) and star-formation rate
+($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). %We show the SED
+% fit in Fig. \ref{fig:SED}.
 
-Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the re-analysis of Gemini data to the $z$-band data of \citet{Margutti2012} and \citet{Sakamoto2013}.
-Both report $z$-band photometry brighter by 0.8-1.0 mag than what we derive here, whereas bluer data are in excellent agreement with previously published values.
+Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the
+re-analysis of Gemini data to the $z$-band data of \citet{Margutti2012} and
+\citet{Sakamoto2013}. Both report $z$-band photometry brighter by 0.8-1.0 mag
+than what we derive here, whereas bluer data are in excellent agreement with
+previously published values.
 
-In fact, the large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break driving the galaxy photometric redshift of both of these earlier works.
-Using the revised photometry from Table~\ref{tab:phot_overview}, the photo-$z$ of the galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic value at the 1~$\sigma$ confidence level.
+In fact, the large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
+driving the galaxy photometric redshift of both of these earlier works. Using
+the revised photometry from Table~\ref{tab:phot_overview}, the photo-$z$ of the
+galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the
+spectroscopic value at the 1~$\sigma$ confidence level.
 
@@
 
-We retrieved the automated data products provided by the \textit{Swift}-XRT GRB repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901.} \todo{Broken link, Paolo?}} \citep{Evans2009}.
-The X-ray afterglow light curve can be fit with a single power-law decay with index $\alpha=1.27_{-0.10}^{+0.12}$.
-We performed a time-integrated spectral analysis using data in photon counting (PC) mode in the widest time epoch where the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant (namely, from $t-T_0 = 205$ s to $t-T_0 = 203.5$ ks, for a total of 29.1 ks of data), in order to prevent spectral changes that can affect the X-ray column density determination.
-The obtained spectrum is well described by an absorbed power-law model.
-The best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and an intrinsic $N_{H}$ of $2.4_{-1.6}^{+2.4} \times 10^{22}$ cm$^{-2}$ ($z=2.211$), assuming the Galactic $N_{H}$ in the burst direction of $4.1 \times 10^{20}$ cm$^{-2}$.
+We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
+repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901.}
+	\todo{Broken link, Paolo?}} \citep{Evans2009}. The X-ray afterglow light curve
+can be fit with a single power-law decay with index
+$\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral analysis
+using data in photon counting (PC) mode in the widest time epoch where the
+$0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
+(namely, from $t-T_0 = 205$ s to $t-T_0 = 203.5$ ks, for a total of 29.1 ks of
+data), in order to prevent spectral changes that can affect the X-ray column
+density determination. The obtained spectrum is well described by an absorbed
+power-law model. The best-fit spectral parameters are a photon index of $2.1 \pm
+0.4$ and an intrinsic $N_{H}$ of $2.4_{-1.6}^{+2.4} \times 10^{22}$ cm$^{-2}$
+($z=2.211$), assuming the Galactic $N_{H}$ in the burst direction of $4.1 \times
+10^{20}$ cm$^{-2}$.
 
@@
 
-As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A very securely belongs the short class of GRBs.
-Because the observed classification indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not change this designation significantly.
-The intrinsic luminosity increases, as visible in the light curve in Fig.
-\ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to majority of long GRBs.
-\citet{Bromberg2013} investigated the degree to which the long and short population distributions overlap and quantified the certainty in the class membership.
-GRB~111117A has $96_{-5}^{+3}$ percent probability of being short.
-Compared to the other two highest-z short candidates, GRB060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class membership for GRB~111117A is much higher.
+As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A very
+securely belongs the short class of GRBs. Because the observed classification
+indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
+\citep{Littlejohns2013a}, the updated redshift does not change this designation
+significantly. The intrinsic luminosity increases, as visible in the light curve
+in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
+majority of long GRBs. \citet{Bromberg2013} investigated the degree to which the
+long and short population distributions overlap and quantified the certainty in
+the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
+being short. Compared to the other two highest-z short candidates, GRB060121
+\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
+($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010,
+	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
+membership for GRB~111117A is much higher.
 
@@
 
-We plot the recalculated $N_H$ in Fig.
-\ref{fig:NH_z} where we compare with the distributions of complete samples of both long and short GRBs.
-The long sample is from \citet{Arcodia2016} and the short sample is from \citet{DAvanzo2014a}.
-17 of the 99 long bursts do not have redshifts and likewise for 5 out of 16 for the short sample.
-Bursts without redshifts have been excluded for both groups.
-GRB~111117A occupies a unique position in Fig.
-\ref{fig:NH_z} with the highest $N_H$ of all short burst.
-The short sample, excluding GRB~111117A, is located at low redshifts(z < 1) and are found to populate a similar column density environment to long GRBs at similar redshifts \citep{DAvanzo2014a}.
-The inferred hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$ as a function of redshift as found for the long GRB hosts \citep{Arcodia2016}.
+We plot the recalculated $N_H$ in Fig. \ref{fig:NH_z} where we compare with the
+distributions of complete samples of both long and short GRBs. The long sample
+is from \citet{Arcodia2016} and the short sample is from \citet{DAvanzo2014a}.
+17 of the 99 long bursts do not have redshifts and likewise for 5 out of 16 for
+the short sample. Bursts without redshifts have been excluded for both groups.
+GRB~111117A occupies a unique position in Fig. \ref{fig:NH_z} with the highest
+$N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
+low redshifts(z < 1) and are found to populate a similar column density
+environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
+hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
+as a function of redshift as found for the long GRB hosts \citep{Arcodia2016}.
 
@@
 %From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits clear emission lines indicative of a population of relatively young stars,
-As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a late-type galaxy and is entirely consistent in terms of stellar mass and stellar age with the general population of short GRB hosts($\left\langle M _* \right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle = 0.3 $ Gyr) \citet{Leibler2010}.
-The SFR, on the other hand, is $\sim$1 order of magnitude higher than the typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to the SFRs found in the hosts of long GRBs at a corresponding redshift \citep{Kruhler2015}.
-Only two hosts in the sample of short GRBs compiled in \citet{Berger2014} have a more vigorous star formation, meaning that it is in the very upper end of the star formation distribution, while not a complete outlier.
-This is likely a selection effect, because a less star-forming galaxy would exhibit weaker emission lines, thus making the redshift harder to determine.
-Being a late-type host, both the stellar mass and sSFR are entirely consistent with the expected host of sGRBs\citep{Behroozi2014}.
-That we detect \lya~is consistent with the SED-inferred absence of dust, despite the moderate stellar age and the high X-ray derived hydrogen column density which would suggest the opposite.
-The centroid of the Ly$\alpha$ emission is found to be redshifted by $\sim$ 220 km s$^{-1}$ with respect to systemic, which is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.
+As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a
+late-type galaxy and is entirely consistent in terms of stellar mass and stellar
+age with the general population of short GRB hosts($\left\langle M _*
+\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
+0.3 $ Gyr) \citet{Leibler2010}. The SFR, on the other hand, is $\sim$1 order of
+magnitude higher than the typical SFR for short GRB hosts galaxies
+\citep{Berger2014} and more similar to the SFRs found in the hosts of long GRBs
+at a corresponding redshift \citep{Kruhler2015}. Only two hosts in the sample of
+short GRBs compiled in \citet{Berger2014} have a more vigorous star formation,
+meaning that it is in the very upper end of the star formation distribution,
+while not a complete outlier. This is likely a selection effect, because a less
+star-forming galaxy would exhibit weaker emission lines, thus making the
+redshift harder to determine. Being a late-type host, both the stellar mass and
+sSFR are entirely consistent with the expected host of
+sGRBs\citep{Behroozi2014}.
+
+That we detect \lya~is consistent with the SED-inferred absence of dust, despite
+the moderate stellar age and the high X-ray derived hydrogen column density
+which would suggest the opposite. The centroid of the Ly$\alpha$ emission is
+found to be redshifted by $\sim$ 220 km s$^{-1}$ with respect to systemic, which
+is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where
+the outflow is attributed to star formation.
 
@@
 
-A single sGRB at high redshift does little in terms of constraints on the redshift distribution and therethrough the progenitor channels, but other short hosts could be missed because they are intrinsically fainter and thus this high-z event is only detected due to the brightness of the host.
-\citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by the characteristic galaxy luminosity at their respective redshift, $L_B/L^{\star}_{B}$.
-26 out of 39 hosts (66 per cent) in the sample has redshifts.
-To convert the SED-inferred $M_B$ of GRB~111117A to $L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window $2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with measured $L_B/L^{\star}_{B}$.
-If we assume that we are able to get emission-line redshifts from hosts with $R < 25$ mags \citep{Kruhler2012}, then we would have missed around 30\% (8 out of 26 from the sample of \citealt{Berger2014} with measured $L_B/L^{\star}_{B}$), if they were at the redshift of GRB~111117A.
-If we assume that the hosts missing redshift in \citet{Berger2014} follow the redshift distribution of the hosts \textit{with} redshift, this implies that we are \textit{not} missing a dominant fraction of sGRB redshift at $z \approx 2$, due to host faintness.
+A single sGRB at high redshift does little in terms of constraints on the
+redshift distribution and therethrough the progenitor channels, but other short
+hosts could be missed because they are intrinsically fainter and thus this
+high-z event is only detected due to the brightness of the host.
+\citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by
+the characteristic galaxy luminosity at their respective redshift,
+$L_B/L^{\star}_{B}$. 26 out of 39 hosts (66 per cent) in the sample has
+redshifts. To convert the SED-inferred $M_B$ of GRB~111117A to
+$L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the
+Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window
+$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} =
+1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with
+measured $L_B/L^{\star}_{B}$. If we assume that we are able to get emission-line
+redshifts from hosts with $R < 25$ mags \citep{Kruhler2012}, then we would have
+missed around 30\% (8 out of 26 from the sample of \citealt{Berger2014} with
+measured $L_B/L^{\star}_{B}$), if they were at the redshift of GRB~111117A. If
+we assume that the hosts missing redshift in \citet{Berger2014} follow the
+redshift distribution of the hosts \textit{with} redshift, this implies that we
+are \textit{not} missing a dominant fraction of sGRB redshift at $z \approx 2$,
+due to host faintness.
 
-The theoretical redshift distribution of sGRBs depends on the type of delay-time function used to model the progenitor system.
-The redshift of GRB~111117A puts constraints on the type of delay-time models suitable for modeling.
-The likelihood preferred lognormal time delay models investigated in \citet{Wanderman2015} predicts a rate of sGRBs at $z = 2.211$, $\sim$ 2 orders of magnitude lower compared to peak rate (z = 0.9).
-It is stated in \citet{Wanderman2015} that this preference depends critically on the absence of non-collapsar sGRB at $z \gtrsim 1.2$.
-The redshift of the burst, on the other hand, is close to the expected peak in sGRB rate calculated for the power law models \citep{Behroozi2014, Wanderman2015}.
+The theoretical redshift distribution of sGRBs depends on the type of delay-time
+function used to model the progenitor system. The redshift of GRB~111117A puts
+constraints on the type of delay-time models suitable for modeling. The
+likelihood preferred lognormal time delay models investigated in
+\citet{Wanderman2015} predicts a rate of sGRBs at $z = 2.211$, $\sim$ 2 orders
+of magnitude lower compared to peak rate (z = 0.9). It is stated in
+\citet{Wanderman2015} that this preference depends critically on the absence of
+non-collapsar sGRB at $z \gtrsim 1.2$. The redshift of the burst, on the other
+hand, is close to the expected peak in sGRB rate calculated for the power law
+models \citep{Behroozi2014, Wanderman2015}.
 
@@
 
-At $z = 2.211$, the age of the universe is 2.97 Gyr.
-If the progenitor systems of sGRBs are the merger of two neutron stars, this sets a hard upper limit to the coalescence timescale for such a system.
-In the absence of other mechanisms, the timescale of the orbital decay of the system is set by the energy loss due to gravitational waves, which in turn is set by the mass of constituent compact objects and the separation of the two \citep{Postnov2014}.
-If we assume that the formation timescale of the first galaxies is short compared to the time since the big bang \citep{Richard2011} and if we assume a mass of 1.4 M$_\odot$ for each of the neutron stars at the time of system formation, this places a limit on the initial separation of the two neutron stars of a$_0 < 3.2 $R$_\odot$, where a$_0$ is the initial separation.
-Assuming our measured age of the stellar population ($\tau = 0.5$ Gyr) as a limit for the orbital decay time we can put a tighter constraint of a$_0 < 2.1 $R$_\odot$.
-However, this does not account for the possibility there could be an underlying stellar population of older stars from a previous star-formation epoch, which is why we consider a$_0 < 3.2 $R$_\odot$ as a hard upper limit.
-The delay time between formation and explosion is well accomodated by the models of \citet{Belczynski2006}, although the longest formation channels are excluded.
-This is especially true given the late type nature of the host \citep{OShaughnessy2008}.
+At $z = 2.211$, the age of the universe is 2.97 Gyr. If the progenitor systems
+of sGRBs are the merger of two neutron stars, this sets a hard upper limit to
+the coalescence timescale for such a system. In the absence of other mechanisms,
+the timescale of the orbital decay of the system is set by the energy loss due
+to gravitational waves, which in turn is set by the mass of constituent compact
+objects and the separation of the two \citep{Postnov2014}. If we assume that the
+formation timescale of the first galaxies is short compared to the time since
+the big bang \citep{Richard2011} and if we assume a mass of 1.4 M$_\odot$ for
+each of the neutron stars at the time of system formation, this places a limit
+on the initial separation of the two neutron stars of a$_0 < 3.2 $R$_\odot$,
+where a$_0$ is the initial separation.
+
+Assuming our measured age of the stellar population ($\tau = 0.5$ Gyr) as a
+limit for the orbital decay time we can put a tighter constraint of a$_0 < 2.1
+$R$_\odot$. However, this does not account for the possibility there could be an
+underlying stellar population of older stars from a previous star-formation
+epoch, which is why we consider a$_0 < 3.2 $R$_\odot$ as a hard upper limit. The
+delay time between formation and explosion is well accomodated by the models of
+\citet{Belczynski2006}, although the longest formation channels are excluded.
+This is especially true given the late type nature of the host
+\citep{OShaughnessy2008}.
 
@@
 
-We have in this letter provided a revised, spectroscopic redshift for GRB~111117A based on emission lines discovered using an improved reduction scheme, setting it at $z = 2.211$.
-This value supersedes the previous photometric redshift of $z \sim 1.3$ \citep{Margutti2012, Sakamoto2013}.
-Part of the reason for the erroneous redshift estimate of previous authors is attributed to a discrepancy in the measured $z$-band magnitude.
+We have in this letter provided a revised, spectroscopic redshift for
+GRB~111117A based on emission lines discovered using an improved reduction
+scheme, setting it at $z = 2.211$. This value supersedes the previous
+photometric redshift of $z \sim 1.3$ \citep{Margutti2012, Sakamoto2013}. Part of
+the reason for the erroneous redshift estimate of previous authors is attributed
+to a discrepancy in the measured $z$-band magnitude.
 
-The rest-frame parameters of the burst and the conditions of the burst environment have been recalculated using the new distance.
-The X-ray derived hydrogen column density towards GRB~111117A is the highest within a complete sample of short hosts, but seems to trace the evolution with redshift as found for the hosts of long GRBs.
+The rest-frame parameters of the burst and the conditions of the burst
+environment have been recalculated using the new distance. The X-ray derived
+hydrogen column density towards GRB~111117A is the highest within a complete
+sample of short hosts, but seems to trace the evolution with redshift as found
+for the hosts of long GRBs.
 
-The SFR of the host is in the upper end of the sGRB host SFR distribution and this plays a role in allowing a redshift to be measured for this host.
-Despite the moderate age and high $N_H$, almost no dust is present.
+The SFR of the host is in the upper end of the sGRB host SFR distribution and
+this plays a role in allowing a redshift to be measured for this host. Despite
+the moderate age and high $N_H$, almost no dust is present.
 
-Although a single burst carries little leverage in terms of constraining the redshift distribution of sGRB, the high redshift of GRB~111117A needs to be accommodated.
-A lognormal delay time model predicts a very low volumetric density of bursts at $z = 2.211$, whereas a power law delay time model peaks near GRB~111117A.
-If more sGRBs are at this redshift, but missed due to faintness of their hosts, a lognormal delay time model will be disfavored.
-Compared to a sample of short hosts, GRB~111117A is more luminous than 70 per cent of the sample with measured luminosities.
-Of the host sample, for 30 per cent, we would be unable to determine a redshift, should they be at the redshift of GRB~111117A.
-This implies that we are \textit{not} missing a dominant fraction of the sGRBs at $z \sim 2$.
+Although a single burst carries little leverage in terms of constraining the
+redshift distribution of sGRB, the high redshift of GRB~111117A needs to be
+accommodated. A lognormal delay time model predicts a very low volumetric
+density of bursts at $z = 2.211$, whereas a power law delay time model peaks
+near GRB~111117A. If more sGRBs are at this redshift, but missed due to
+faintness of their hosts, a lognormal delay time model will be disfavored.
+Compared to a sample of short hosts, GRB~111117A is more luminous than 70 per
+cent of the sample with measured luminosities. Of the host sample, for 30 per
+cent, we would be unable to determine a redshift, should they be at the redshift
+of GRB~111117A. This implies that we are \textit{not} missing a dominant
+fraction of the sGRBs at $z \sim 2$.
 
-Using the age of the universe at the time of explosion allows us to set constraints on the maximal separation between the engine constituents at the time of formation.
-We find that the maximal separation for two neutron stars at formation time is a$_0 < 3.2 $R$_\odot$, which excludes some of the formation channels with the longest timescales.
+Using the age of the universe at the time of explosion allows us to set
+constraints on the maximal separation between the engine constituents at the
+time of formation. We find that the maximal separation for two neutron stars at
+formation time is a$_0 < 3.2 $R$_\odot$, which excludes some of the formation
+channels with the longest timescales.
 
-We make all the data, code and calculation related to the paper along with the paper itself available at \url{https://github.com/jselsing/GRB111117A}.
+We make all the data, code and calculation related to the paper along with the
+paper itself available at \url{https://github.com/jselsing/GRB111117A}.
 

### 2017-05-28 authorlist.tex
Add co-authors and add line-breaks to better track changes

@@
 \author{J.~Selsing\inst{1}
-\and T. Kr\"{u}hler\inst{2}
+\and T.~Kr\"{u}hler\inst{2}
 \and D.~Malesani\inst{1,3}
-\and P. D'Avanzo\inst{4}
+\and P.~D'Avanzo\inst{4}
 \and J.~Palmerio\inst{5}
 \and S.~D.~Vergani\inst{6}
-\and J.~Japelj\inst{6}
+\and J.~Japelj\inst{7}
 \and B.~Milvang-Jensen\inst{1}
 \and D.~Watson\inst{1}
-\and P.~Jakobsson\inst{7}
+\and P.~Jakobsson\inst{8}
 %
-\and Z. Cano\inst{8}
+\and Z.~Cano\inst{9}
+%\and L.~B.~Christensen\inst{1}
+\and V.~D'Elia\inst{10}
+\and A.~de~Ugarte~Postigo\inst{9, 1}
 \and J.~P.~U.~Fynbo\inst{1}
-\and A. Gomboc\inst{9}
-\and K.~E.~Heintz\inst{7,1}
-\and M. Sparre\inst{1,10}
-\and XSGRB collaboration
+\and A.~Gomboc\inst{11}
+\and K.~E.~Heintz\inst{8,1}
+\and A.~J.~Levan \inst{12}
+\and M.~Sparre\inst{1,13}
+\and N.~R.~Tanvir\inst{14}
+\and C.~C.~Th\"one\inst{9}
 }
@@
 %7
+\and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands
+%8
 \and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
-%8
+%9
 \and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain.
-%9
+%10
+\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy;
+ASI-Science Data Centre, Via del Politecnico snc, I-00133 Rome, Italy
+%11
 \and University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
-%10
+%12
+\and Department of Physics, University of Warwick, Coventry CV4 7AL, UK
+%13
 \and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
+%14
+\and Department of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK
 %\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile

### 2017-05-31 GRB111117A.tex
TK: Typographical changes, nothing on the science

@@
 
-\title{The host galaxy of the short GRB~111117A at z = 2.211\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}}
+\title{The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}}
 
@@
 
+% TK: I made the following typographical changes:
+% All physical constants are in italics, all mathematical constants are normal, so SFR
+% used in formulas is italics, while it is not italics when used in text as an acronym
+% Super- and subscripts are usually normal
+% Write out all numbers < 12
+% Added fixed spaces in GRB, e.g., GRB~111117A and units (this prevents a line break)
+% Thin spaces \, between thousands when > 9999
+% Table is never abbreviated
+% Tried to make the text a bit easier to read, e.g., things like the following:
+% Conclusions: Part of the reason is attributed ... -> The reason is ...
+
 \abstract{
-Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to
-localize. We here present a spectroscopic observation of the host galaxy of GRB
-111117A and measure its redshift to be $z = 2.211$, making it the most distant
-high-confidence short GRB with a secure redshift measurement. This value
-supersedes a previously estimated, lower redshift value. Spectroscopic
-observations along with a wealth of imaging data allow us to place tight
-constraints on the nature of the host galaxy at an unprecedented distance.
-
-The rest frame X-ray derived hydrogen column density is high compared to a
-complete sample of sGRBs and seems to follow the evolution with redshift as
-traced by the hosts of long GRBs (lGRBs). This could indicate that the part of
-the sGRB population hosted in late-type galaxies live in similar environments to
+% TK I rephased the abstract a bit, check if you like it or not. The original version is commented out
+It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and their hosts to measure their redshifts. These measurements, however, are critical to constrain sGRB progenitors 
+and delay time models. Here, 
+%Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to
+%localize. 
+we present spectroscopy of the host galaxy of GRB~111117A 
+and measure its redshift to be $z = 2.211$. This makes GRB~111117A the most distant
+high-confidence short GRB detected to date.
+%with a secure redshift measurement. 
+Our spectroscopic redshift supersedes a lower redshift previously estimated from photometry. %Spectroscopic observations along with a wealth of imaging data allow us to place tight
+%constraints on the nature of the host galaxy at an unprecedented distance.
+We use the spectroscopic redshift, as well as new imaging data to constrain the nature of the host galaxy as well as physical parameters of the GRB.
+The rest frame X-ray derived hydrogen column density, for example, is high compared to a complete sample of sGRBs and seems to follow the evolution with redshift as traced by the hosts of long GRBs (lGRBs).
+This could indicate that the part of the sGRB population hosted in late-type galaxies live in similar environments to
 that of lGRB hosts.
@@
 The host lies in the brighter end of the expected host brightness distribution
-at $z = 2.211$, and is actively star forming. Using the host as a benchmark for
+at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
 redshift determination, we assess that only 30 per cent of sGRB redshifts should
-be missed due to host faintness at this redshift. This indicates that we are not
-missing redshifts for a dominant fraction of the sGRB hosts at this redshift.
-The redshift of GRB~111117A is evidence against a lognormal delay-time model for
+be missed due to host faintness at $z\sim2$. This indicates that we are not
+missing redshifts for a dominant fraction of the sGRB hosts in this redshift range.
+The high redshift of GRB~111117A is evidence against a lognormal delay-time model for
 sGRBs through the predicted redshift distribution of sGRBs, which is very
-sensitive to high-z sGRBs.
+sensitive to high-$z$ sGRBs.
 
 From the age of the universe at explosion time, an initial progenitor separation
-of a$_0 < 3.2 $R$_\odot$ is required for the case of a binary neutron star (NS)
-system. This puts constraints on the progenitor system evolution up to the time
+of $a_0 < 3.2~R_\odot$ is required in the case of a binary neutron star system. 
+This puts constraints on the progenitor system evolution up to the time
 of explosion.
@@
 % Gal-Yam2006}.
-The range of host galaxy types, in some cases apparently absent, and typically
-more massive and on average less actively star-forming than long GRB hosts 
-\citep{Fong2013b}, as well as their positions within hosts \citep{Fong2013a},
-suggest an origin for the progenitors that can be very long-lived, and
-associated with stellar mass rather than purely star-formation rate
-\citep{Berger2014}. The bulk of sGRBs have associated galaxies at relatively low
-redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts
-are measured from the hosts, this value is biased towards low $z$.
+
+% TK I rephrased this paragraph, i found the sentences too long and split . Original
+% version is commented out.
+
+The host galaxies of short GRBs are diverse. They are more massive and less actively 
+star-forming on average than long GRB hosts \citep{Fong2013b}, while in some cases, no host 
+galaxy can be identified. Together with their position 
+within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
+long-lived, and is associated with stellar mass rather than star-formation rate
+\citep{Berger2014}. The median redshift for sGRBs is $z\sim0.5$ \citep{Berger2014}, and
+because most of these measurements come from the associated hosts, it is arguably biased
+towards lower redshifts.
+
+%The range of host galaxy types, in some cases apparently absent, and typically
+%more massive and on average less actively star-forming than long GRB hosts 
+%\citep{Fong2013b}, as well as their positions within hosts \citep{Fong2013a},
+%suggest an origin for the progenitors that can be very long-lived, and
+%associated with stellar mass rather than purely star-formation rate
+%\citep{Berger2014}. The bulk of sGRBs have associated galaxies at relatively low
+%redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts
+%are measured from the hosts, this value is biased towards low $z$.
 
@@
 
-The total lifetimes of NS binaries is dependent on their initial separations and
+The total lifetimes of NS binaries depends on their initial separations and
 subsequent inspiral times, and impacts the timing and distribution of the
@@
 
-In this \emph{Letter} we present the spectrum of the host galaxy of the short
-GRB 111117A at z=2.211, a significantly higher redshift than the previously
-estimated redshift based on photometric studies. We present the rest frame burst
+In this \emph{Letter} we present the spectrum of the host galaxy of the short GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is significantly higher 
+than the previously estimated redshift based on photometric studies. 
+We present the GRB's rest frame 
 properties based on this new distance compared to previous analyses
@@
 Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology provided by
-\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$ km s$^{-1}$
+\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
 and $\Omega_m = 0.307$ and report all magnitudes in the AB system.
@@
 	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
-	\caption{Imaging of the field with the X-shooter slit overlaid.
-Only one slit is shown, despite 4 epochs of spectroscopic observations because of the similarity in position angle.
-The image is the FORS2 R-band image and%, for which the photometry is shown in Fig.
+	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid.
+Only one slit is shown, despite four epochs of spectroscopic observations because of the similarity in position angle.
+The image is the FORS2 $R$-band image and%, for which the photometry is shown in Fig.
 \ref{fig:SED}.
-	the blue asterisk i ndicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}.
-The errorcircle of the \emph{Chandra}-position is too small to resolve in the image.}
+	the blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}.
+The errorcircle of the \emph{Chandra}-position is too small to be visible in the image.}
 	\label{fig:spec_setup}
@@
 program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
-3000 \AA~to 24800 \AA~(21000 \AA~when the K-band blocking filter is used)~across
+3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across
 three spectroscopic arms, simultaneously illuminated through the use of
@@
 
-An overview of the spectroscopic observations is given in Tab. \ref{tab:spec_overview}, and the position of the slit on the target is shown in Fig. \ref{fig:spec_setup}.
+An overview of the spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the position of the slit on the target is shown in Fig.~\ref{fig:spec_setup}.
 
@@
 \hb~detected at low significance ($\sim 3 \sigma$). \ha~is only visible in the
-first epoch, due to the K-band blocking filter used for the remainder
+first epoch, due to the $K$-band blocking filter used for the remainder
 observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and
@@
 
-Using the integrated flux of \ha, we can infer the SFR \citep{Kennicutt1998}.
-From the SED-fit, see Sect. \ref{SED}, and the detection of \lya, the host is
-constrained to contain very little or no dust which is also confirmed by the
-shape of the spectral continuum, which is why we do not apply a dust-correction.
-Because of the redshift, \ha~falls on the far end of the NIR-arm where the night
-sky is very bright and additionally, several bright sky-lines intersect the
-line, making an accurate estimate of the \ha-flux difficult. A limit on the SFR
-is obtained by integrating the part of \ha~free of contamination and
-conservatively correcting for the missing fraction by assessing the fraction
-integrated. Converting to a \citet{Chabrier2003} initial mass function using
-\citet{Madau2014} we get a limit of SFR > 7 M$_\odot$ yr$^{-1}$.
+% TK: Some rephrasing below.
+
+Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
+galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
+21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
+intersect the line, making an accurate estimate of the \ha-flux difficult. 
+We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
+correcting for the missing flux using the line shape. After converting the 
+\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
+\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
+From the SED-fit (Sect. \ref{SED}), and the detection of \lya, the host is
+constrained to contain very little or no dust, which is why we do not apply 
+a dust-correction to the measured \ha~flux here.
 
@@
 
-We additionally imaged the field of GRB 111117A in multiple broad-band filters
-using the VLT equipped with FORS2 ($gRIz$ filters) and HAWK-I
+In addition to the spectroscopy presented above, we imaged the field of GRB~111117A 
+in multiple broad-band filters using the VLT equipped with FORS2 ($gRIz$ filters) and HAWK-I
 ($JHK_{\mathrm{s}}$ filters), long after the burst has faded. These new data are
@@
 $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
-observations and measured brightnesses is given in Tab.~\ref{tab:phot_overview}.
+observations and measured brightnesses is given in Table~\ref{tab:phot_overview}.
 
-All data were reduced, analyzed and fitted in a similar manner as we have done
-previously and described in detail in \citet{Kruhler2011a} and, more recently,
-in \citet{Schulze2016}. Very briefly, we use our own python and IRAF routines to
-perform a consistent standard reduction of the available imaging which includes
+% TK: I shortened a bit here
+All data were reduced, analyzed and fitted in a similar manner as described in detail in
+\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Very briefly, we use our own
+\texttt{python} and IRAF routines to perform a standard reduction which includes
 bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
-stacking of individual images for a given instrument in a given filter. The
-photometry of the host was tied against magnitudes of field stars from the SDSS
-and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. For our $R$
-and $I$-band photometry we used the color transformations of
+stacking of individual images. The photometry of the host was tied against magnitudes 
+of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. 
+For the $R$ and $I$-band photometry we used the color transformations of
 Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
-We subsequently convert all magnitudes into the AB system if necessary, and correct for the Galactic foreground of $E_{B-V}=0.027~\mathrm{mag}$.
+We convert all magnitudes into the AB system if necessary, and correct for the Galactic 
+foreground of $E_{B-V}=0.027~\mathrm{mag}$.
 
-The multi-color spectral energy distribution (SED) is then fit by
-\citet{Bruzual2003} stellar population synthesis models based on a
+The multi-color spectral energy distribution (SED) is fit by
+\citet{Bruzual2003} single stellar population models based on a
 \citet{Chabrier2003} initial mass function in \emph{LePhare} \citep{Ilbert2006},
@@
 model is obtained with an unreddened galaxy template, and returns physical
-parameters of luminosity ($M_B=-22.0\pm0.1$\,mag), stellar mass
-($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), single stellar population age ($\tau =
-0.5_{-0.25}^{+0.5}$ Gyr) and star-formation rate
+parameters of absolute magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass
+($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), stellar population age ($\tau =
+0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
 ($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). %We show the SED
@@
 
-Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the
-re-analysis of Gemini data to the $z$-band data of \citet{Margutti2012} and
-\citet{Sakamoto2013}. Both report $z$-band photometry brighter by 0.8-1.0 mag
-than what we derive here, whereas bluer data are in excellent agreement with
-previously published values.
-
-In fact, the large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
-driving the galaxy photometric redshift of both of these earlier works. Using
-the revised photometry from Table~\ref{tab:phot_overview}, the photo-$z$ of the
+Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
+re-analysis of Gemini data to the $z$-band measurements of \citet{Margutti2012} and
+\citet{Sakamoto2013}. Both of these authors report $z$-band photometry that is brighter 
+by 0.8~mag to 1.0 mag than our value, whereas data taken in bluer filters are in excellent agreement. In fact, the large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
+driving the galaxy photometric redshift of the earlier works. Using
+the revised photometry from Table~\ref{tab:phot_overview}, the photometric redshift of the
 galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{XRT temporal and spectral analysis}
+\subsection{X-ray temporal and spectral analysis}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
-repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901.}
-	\todo{Broken link, Paolo?}} \citep{Evans2009}. The X-ray afterglow light curve
+repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}} \citep{Evans2009}. 
+% TK: Fixed the link
+The X-ray afterglow light curve
 can be fit with a single power-law decay with index
@@
 $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
-(namely, from $t-T_0 = 205$ s to $t-T_0 = 203.5$ ks, for a total of 29.1 ks of
-data), in order to prevent spectral changes that can affect the X-ray column
+(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
+data) to prevent spectral changes that can affect the X-ray column
 density determination. The obtained spectrum is well described by an absorbed
 power-law model. The best-fit spectral parameters are a photon index of $2.1 \pm
-0.4$ and an intrinsic $N_{H}$ of $2.4_{-1.6}^{+2.4} \times 10^{22}$ cm$^{-2}$
-($z=2.211$), assuming the Galactic $N_{H}$ in the burst direction of $4.1 \times
-10^{20}$ cm$^{-2}$.
+0.4$ and an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4} \times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$.
 
@@
 
-As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A very
-securely belongs the short class of GRBs. Because the observed classification
+As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A securely 
+belongs the short class of GRBs. Because the observed classification
 indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
 \citep{Littlejohns2013a}, the updated redshift does not change this designation
-significantly. The intrinsic luminosity increases, as visible in the light curve
+significantly. The intrinsic luminosity increases, as visible in the X-ray light curve
 in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
-majority of long GRBs. \citet{Bromberg2013} investigated the degree to which the
+the majority of long GRBs. \citet{Bromberg2013} investigated the degree to which the
 long and short population distributions overlap and quantified the certainty in
 the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
-being short. Compared to the other two highest-z short candidates, GRB060121
+being short. Compared to the other two short GRB candidates at high redshift, GRB060121
 \citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
@@
 The grey shared region is a compilation of long GRB lightcurves where the color represents density and the light blue is other short GRB lightcurves for which redshifts has been determined.
-Despite the remarkably high redshift, the burst’s luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
+Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
 	\label{fig:sxray_lightcurve}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restframe $N_H$}
+\subsection{Restframe $N_\mathrm{H}$}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We plot the recalculated $N_H$ in Fig. \ref{fig:NH_z} where we compare with the
+We plot the recalculated $N_\mathrm{H}$ in Fig. \ref{fig:NH_z} where we compare with the
 distributions of complete samples of both long and short GRBs. The long sample
 is from \citet{Arcodia2016} and the short sample is from \citet{DAvanzo2014a}.
-17 of the 99 long bursts do not have redshifts and likewise for 5 out of 16 for
+17 of the 99 long bursts do not have redshifts and likewise 5 out of 16 for
 the short sample. Bursts without redshifts have been excluded for both groups.
@@
 $N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
-low redshifts(z < 1) and are found to populate a similar column density
+low redshifts ($z < 1$) and is found to populate a similar column density
 environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
 hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
-as a function of redshift as found for the long GRB hosts \citep{Arcodia2016}.
+as a function of redshift as found for the long GRB afterglows \citep{Arcodia2016}.
 
@@
 	\caption{Rest frame, X-ray derived hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations.
-In red is shown the long sample of \citet{Arcodia2016}.
+In red we show the sample of lGRBs from \citet{Arcodia2016}.
 The detections are replaced with contours for clarity and the limits are shown with arrows.
-In blue is the sample from \citet{DAvanzo2014a}.
+In blue is the sample of sGRBs from \citet{DAvanzo2014a}.
 GRB~111117A is an outlier from the short sample, but seems to follow the distribution of long GRBs.}
@@
 late-type galaxy and is entirely consistent in terms of stellar mass and stellar
-age with the general population of short GRB hosts($\left\langle M _*
+age with the general population of short GRB hosts ($\left\langle M _*
 \right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
-0.3 $ Gyr) \citet{Leibler2010}. The SFR, on the other hand, is $\sim$1 order of
+0.3 $~Gyr) \citet{Leibler2010}. The SFR, on the other hand, is $\sim$1 order of
 magnitude higher than the typical SFR for short GRB hosts galaxies
@@
 short GRBs compiled in \citet{Berger2014} have a more vigorous star formation,
-meaning that it is in the very upper end of the star formation distribution,
+meaning that it is in the upper end of the star formation distribution,
 while not a complete outlier. This is likely a selection effect, because a less
@@
 sSFR are entirely consistent with the expected host of
-sGRBs\citep{Behroozi2014}.
+sGRBs \citep{Behroozi2014}.
 
-That we detect \lya~is consistent with the SED-inferred absence of dust, despite
+The detection of \lya~is consistent with the SED-inferred absence of dust, despite
 the moderate stellar age and the high X-ray derived hydrogen column density
 which would suggest the opposite. The centroid of the Ly$\alpha$ emission is
-found to be redshifted by $\sim$ 220 km s$^{-1}$ with respect to systemic, which
+found to be redshifted by $\sim$~220~km~s$^{-1}$ with respect to systemic, which
 is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where
@@
 hosts could be missed because they are intrinsically fainter and thus this
-high-z event is only detected due to the brightness of the host.
+high-$z$ event is only detected due to the brightness of the host.
 \citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by
 the characteristic galaxy luminosity at their respective redshift,
-$L_B/L^{\star}_{B}$. 26 out of 39 hosts (66 per cent) in the sample has
+$L_B/L^{\star}_{B}$. 26 out of 39 hosts (66 per cent) in the sample have
 redshifts. To convert the SED-inferred $M_B$ of GRB~111117A to
@@
 1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with
-measured $L_B/L^{\star}_{B}$. If we assume that we are able to get emission-line
-redshifts from hosts with $R < 25$ mags \citep{Kruhler2012}, then we would have
+measured $L_B/L^{\star}_{B}$. If we assume that we are able to obtain emission-line
+redshifts from hosts with $R < 25$~mag \citep{Kruhler2012}, then we would have
 missed around 30\% (8 out of 26 from the sample of \citealt{Berger2014} with
@@
 likelihood preferred lognormal time delay models investigated in
-\citet{Wanderman2015} predicts a rate of sGRBs at $z = 2.211$, $\sim$ 2 orders
-of magnitude lower compared to peak rate (z = 0.9). It is stated in
+\citet{Wanderman2015} predicts a rate of sGRBs at $z = 2.211$, $\sim$ two orders
+of magnitude lower compared to peak rate ($z = 0.9$). It is stated in
 \citet{Wanderman2015} that this preference depends critically on the absence of
@@
 
-At $z = 2.211$, the age of the universe is 2.97 Gyr. If the progenitor systems
+At $z = 2.211$, the age of the universe is 3 Gyr. If the progenitor systems
 of sGRBs are the merger of two neutron stars, this sets a hard upper limit to
@@
 formation timescale of the first galaxies is short compared to the time since
-the big bang \citep{Richard2011} and if we assume a mass of 1.4 M$_\odot$ for
+the big bang \citep{Richard2011} and if we assume a mass of 1.4~$M_\odot$ for
 each of the neutron stars at the time of system formation, this places a limit
-on the initial separation of the two neutron stars of a$_0 < 3.2 $R$_\odot$,
-where a$_0$ is the initial separation.
+on the initial separation of the two neutron stars of $a_0 < 3.2~R_\odot$,
+where $a_0$ is the initial separation.
 
-Assuming our measured age of the stellar population ($\tau = 0.5$ Gyr) as a
-limit for the orbital decay time we can put a tighter constraint of a$_0 < 2.1
-$R$_\odot$. However, this does not account for the possibility there could be an
+Assuming our measured age of the stellar population ($\tau = 0.5$~Gyr) as a
+limit for the orbital decay time we can put a tighter constraint of $a_0 < 2.1~R_\odot$. 
+However, this does not account for the possibility there could be an
 underlying stellar population of older stars from a previous star-formation
-epoch, which is why we consider a$_0 < 3.2 $R$_\odot$ as a hard upper limit. The
-delay time between formation and explosion is well accomodated by the models of
+epoch, which is why we consider $a_0 < 3.2~R_\odot$ as a hard upper limit. The
+delay time between formation and explosion is well accommodated by the models of
 \citet{Belczynski2006}, although the longest formation channels are excluded.
@@
 
-We have in this letter provided a revised, spectroscopic redshift for
-GRB~111117A based on emission lines discovered using an improved reduction
-scheme, setting it at $z = 2.211$. This value supersedes the previous
-photometric redshift of $z \sim 1.3$ \citep{Margutti2012, Sakamoto2013}. Part of
-the reason for the erroneous redshift estimate of previous authors is attributed
-to a discrepancy in the measured $z$-band magnitude.
+In this \emph{Letter}, we have provided a revised, spectroscopic redshift for the short
+GRB~111117A based on emission lines %discovered using an improved reduction scheme, 
+setting it at $z = 2.211$. This value supersedes and corrects the previous
+% TK: No need to repeat a wrong redshift here, so slightly rephrased
+photometric redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. The erroneous redshift estimate of previous authors is attributed to a discrepancy in the measured $z$-band magnitude.
 
@@
 hydrogen column density towards GRB~111117A is the highest within a complete
-sample of short hosts, but seems to trace the evolution with redshift as found
+sample of short hosts, but is indicative of an evolution with redshift as found
 for the hosts of long GRBs.
@@
 The SFR of the host is in the upper end of the sGRB host SFR distribution and
-this plays a role in allowing a redshift to be measured for this host. Despite
-the moderate age and high $N_H$, almost no dust is present.
+this helped us to measure the redshift for this galaxy. Despite
+the moderate age and high $N_\mathrm{H}$, almost no dust is present.
 
@@
 redshift distribution of sGRB, the high redshift of GRB~111117A needs to be
-accommodated. A lognormal delay time model predicts a very low volumetric
+accommodated in progenitor models. A lognormal delay time model predicts a very low volumetric
 density of bursts at $z = 2.211$, whereas a power law delay time model peaks
-near GRB~111117A. If more sGRBs are at this redshift, but missed due to
-faintness of their hosts, a lognormal delay time model will be disfavored.
-Compared to a sample of short hosts, GRB~111117A is more luminous than 70 per
-cent of the sample with measured luminosities. Of the host sample, for 30 per
-cent, we would be unable to determine a redshift, should they be at the redshift
-of GRB~111117A. This implies that we are \textit{not} missing a dominant
+near GRB~111117A. If more sGRBs are at similarly high redshifts, but are missed due to
+the faintness of their hosts, a lognormal delay time model will be disfavored.
+Compared to a sample of short GRB hosts, GRB~111117A is more luminous than 70 per
+cent of the sample with measured luminosities. For 30 per
+cent of the host sample, we would be unable to determine a redshift should they be at 
+a similar redshift as GRB~111117A. This implies that we are \textit{not} missing a dominant
 fraction of the sGRBs at $z \sim 2$.
@@
 time of formation. We find that the maximal separation for two neutron stars at
-formation time is a$_0 < 3.2 $R$_\odot$, which excludes some of the formation
+formation time is $a_0 < 3.2~R_\odot$, which excludes some of the formation
 channels with the longest timescales.
 
-We make all the data, code and calculation related to the paper along with the
-paper itself available at \url{https://github.com/jselsing/GRB111117A}.
+All data, code and calculation related to the paper along with the
+paper itself are available at \url{https://github.com/jselsing/GRB111117A}.
 

### 2017-06-01 GRB111117A.tex
Nials comments
Closes #3

@@
 
-% TK: I made the following typographical changes:
-% All physical constants are in italics, all mathematical constants are normal, so SFR
-% used in formulas is italics, while it is not italics when used in text as an acronym
-% Super- and subscripts are usually normal
-% Write out all numbers < 12
-% Added fixed spaces in GRB, e.g., GRB~111117A and units (this prevents a line break)
-% Thin spaces \, between thousands when > 9999
-% Table is never abbreviated
-% Tried to make the text a bit easier to read, e.g., things like the following:
-% Conclusions: Part of the reason is attributed ... -> The reason is ...
+\abstract{
+It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and
+their hosts to measure their redshifts. These measurements, however, are
+critical to constrain sGRB progenitors and delay time models. Here, we present
+spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
+= 2.211$. This makes GRB~111117A the most distant high-confidence short GRB
+detected to date. Our spectroscopic redshift supersedes a lower redshift value
+for this burst previously estimated from photometry.
 
-\abstract{
-% TK I rephased the abstract a bit, check if you like it or not. The original version is commented out
-It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and their hosts to measure their redshifts. These measurements, however, are critical to constrain sGRB progenitors 
-and delay time models. Here, 
-%Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to
-%localize. 
-we present spectroscopy of the host galaxy of GRB~111117A 
-and measure its redshift to be $z = 2.211$. This makes GRB~111117A the most distant
-high-confidence short GRB detected to date.
-%with a secure redshift measurement. 
-Our spectroscopic redshift supersedes a lower redshift previously estimated from photometry. %Spectroscopic observations along with a wealth of imaging data allow us to place tight
-%constraints on the nature of the host galaxy at an unprecedented distance.
-We use the spectroscopic redshift, as well as new imaging data to constrain the nature of the host galaxy as well as physical parameters of the GRB.
-The rest frame X-ray derived hydrogen column density, for example, is high compared to a complete sample of sGRBs and seems to follow the evolution with redshift as traced by the hosts of long GRBs (lGRBs).
-This could indicate that the part of the sGRB population hosted in late-type galaxies live in similar environments to
-that of lGRB hosts.
+We use the spectroscopic redshift, as well as new imaging data to constrain the
+nature of the host galaxy as well as physical parameters of the GRB. The rest
+frame X-ray derived hydrogen column density, for example, is high compared to a
+complete sample of sGRBs and seems to follow the evolution with redshift as
+traced by the hosts of long GRBs (lGRBs). This is consistent with the proportion
+of the sGRB population originating in late-type galaxies live in similar
+environments to that of lGRB hosts.
 
@@
 at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
-redshift determination, we assess that only 30 per cent of sGRB redshifts should
-be missed due to host faintness at $z\sim2$. This indicates that we are not
-missing redshifts for a dominant fraction of the sGRB hosts in this redshift range.
-The high redshift of GRB~111117A is evidence against a lognormal delay-time model for
-sGRBs through the predicted redshift distribution of sGRBs, which is very
-sensitive to high-$z$ sGRBs.
+redshift determination, we find that less than 55 per cent of all sGRB redshifts
+should be missed due to host faintness at $z\sim2$. The high redshift of
+GRB~111117A is evidence against a lognormal delay-time model for sGRBs through
+the predicted redshift distribution of sGRBs, which is very sensitive to
+high-$z$ sGRBs.
 
@@
 supports this idea and distinguishes the physical origin of sGRBs from their
-long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006,
-% Gal-Yam2006}.
-
-% TK I rephrased this paragraph, i found the sentences too long and split . Original
-% version is commented out.
+long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.
 
@@
 star-forming on average than long GRB hosts \citep{Fong2013b}, while in some cases, no host 
-galaxy can be identified. Together with their position 
+galaxy can be identified \citep{Berger2010a, Tunnicliffe2014}. Together with their position 
 within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
@@
 towards lower redshifts.
-
-%The range of host galaxy types, in some cases apparently absent, and typically
-%more massive and on average less actively star-forming than long GRB hosts 
-%\citep{Fong2013b}, as well as their positions within hosts \citep{Fong2013a},
-%suggest an origin for the progenitors that can be very long-lived, and
-%associated with stellar mass rather than purely star-formation rate
-%\citep{Berger2014}. The bulk of sGRBs have associated galaxies at relatively low
-%redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts
-%are measured from the hosts, this value is biased towards low $z$.
-
-%This progress on the progenitors of sGRBs has been enabled by the afterglow
-% detection and burst localization by the \emph{Swift} satellite
-% \citep{Gehrels2009a}. The bulk of these localizations have associated galaxies
-% at relatively low redshifts with a median $z\sim0.5$ \citep{Berger2014} and
-% because most redshifts are measured from the hosts, this value is biased
-% towards low $z$.
 
@@
 Spectroscopic observations were carried out using the cross-dispersed echelle
-spectrograph, VLT/X-shooter \citep{Vernet2011}, in four seperate epochs. The
+spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
 burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger
@@
 
-% TK: Some rephrasing below.
-
 Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
@@
 
-% TK: I shortened a bit here
 All data were reduced, analyzed and fitted in a similar manner as described in detail in
@@
 We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
-repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}} \citep{Evans2009}. 
-% TK: Fixed the link
-The X-ray afterglow light curve
-can be fit with a single power-law decay with index
-$\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral analysis
-using data in photon counting (PC) mode in the widest time epoch where the
-$0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
+repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
+\citep{Evans2009}. The X-ray afterglow light curve can be fit with a single power-law decay with
+index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
+analysis using data in photon counting (PC) mode in the widest time epoch where
+the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
 (namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
-data) to prevent spectral changes that can affect the X-ray column
-density determination. The obtained spectrum is well described by an absorbed
-power-law model. The best-fit spectral parameters are a photon index of $2.1 \pm
-0.4$ and an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4} \times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$.
+data) to prevent spectral changes that can affect the X-ray column density
+determination. The obtained spectrum is well described by an absorbed power-law
+model. The best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
+an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
+\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
+the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \cite{Willingale2013}.
 
@@
 the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
-being short. Compared to the other two short GRB candidates at high redshift, GRB060121
+being short. Compared to the other two sGRB candidates at high redshift, GRB060121
 \citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restframe $N_\mathrm{H}$}
+\subsection{Restframe $N_\mathrm{H}$} \label{restnH}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We plot the recalculated $N_\mathrm{H}$ in Fig. \ref{fig:NH_z} where we compare with the
-distributions of complete samples of both long and short GRBs. The long sample
-is from \citet{Arcodia2016} and the short sample is from \citet{DAvanzo2014a}.
+distributions of complete samples of both long and short GRBs. The lGRB sample
+is from \citet{Arcodia2016} and the sGRB sample is from \citet{DAvanzo2014a}.
 17 of the 99 long bursts do not have redshifts and likewise 5 out of 16 for
@@
 
-%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits clear emission lines indicative of a population of relatively young stars,
+%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars,
 As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a
@@
 \right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
-0.3 $~Gyr) \citet{Leibler2010}. The SFR, on the other hand, is $\sim$1 order of
-magnitude higher than the typical SFR for short GRB hosts galaxies
-\citep{Berger2014} and more similar to the SFRs found in the hosts of long GRBs
-at a corresponding redshift \citep{Kruhler2015}. Only two hosts in the sample of
-short GRBs compiled in \citet{Berger2014} have a more vigorous star formation,
-meaning that it is in the upper end of the star formation distribution,
-while not a complete outlier. This is likely a selection effect, because a less
+0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
+and sSFR are entirely within the range expected for the hosts of sGRBs
+\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
+typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to
+the SFRs found in the hosts of long GRBs at a corresponding redshift
+\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
 star-forming galaxy would exhibit weaker emission lines, thus making the
-redshift harder to determine. Being a late-type host, both the stellar mass and
-sSFR are entirely consistent with the expected host of
-sGRBs \citep{Behroozi2014}.
+redshift harder to determine. Additionally, it is natural to expect some
+evolution in the hosts of sGRBs with redshift as illustrated in Sect.
+\ref{restnH}.
 
-The detection of \lya~is consistent with the SED-inferred absence of dust, despite
-the moderate stellar age and the high X-ray derived hydrogen column density
-which would suggest the opposite. The centroid of the Ly$\alpha$ emission is
-found to be redshifted by $\sim$~220~km~s$^{-1}$ with respect to systemic, which
-is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where
-the outflow is attributed to star formation.
+%The detection of \lya~is consistent with the SED-inferred absence of dust, despite
+%the moderate stellar age and the high X-ray derived hydrogen column density
+%which would suggest the opposite. The centroid of the Ly$\alpha$ emission is
+%found to be redshifted by $\sim$~220~km~s$^{-1}$ with respect to systemic, which
+%is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where
+%the outflow is attributed to star formation.
 
@@
 
-A single sGRB at high redshift does little in terms of constraints on the
-redshift distribution and therethrough the progenitor channels, but other short
-hosts could be missed because they are intrinsically fainter and thus this
-high-$z$ event is only detected due to the brightness of the host.
+A single sGRB at high redshift does little in terms of constraining the redshift
+distribution and therethrough the progenitor channels. In particular, other
+short-GRB hosts could be missed because they are intrinsically fainter and thus
+this high-$z$ event is only detected due to the brightness of the host.
 \citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by
@@
 1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with
-measured $L_B/L^{\star}_{B}$. If we assume that we are able to obtain emission-line
-redshifts from hosts with $R < 25$~mag \citep{Kruhler2012}, then we would have
-missed around 30\% (8 out of 26 from the sample of \citealt{Berger2014} with
-measured $L_B/L^{\star}_{B}$), if they were at the redshift of GRB~111117A. If
-we assume that the hosts missing redshift in \citet{Berger2014} follow the
-redshift distribution of the hosts \textit{with} redshift, this implies that we
-are \textit{not} missing a dominant fraction of sGRB redshift at $z \approx 2$,
-due to host faintness.
+measured $L_B/L^{\star}_{B}$.
+
+If we assume that we are able to obtain emission-line redshifts from hosts with
+$R < 25$~mag \citep{Kruhler2012}, then we would have missed around 30 per cent
+(8 out of 26 from the sample of \citealt{Berger2014} with measured
+$L_B/L^{\star}_{B}$), if they were at the redshift of GRB~111117A. Because the
+average SFR of galaxies hosting lGRBs is higher than for galaxies hosting sGRBs,
+the fraction of missed burst redshifts is likely higher although the cosmic SFR
+evolution could play a role in improving redshift determinability.
+
+%If we additionally assume that the hosts missing redshift in
+%\citet{Berger2014} follow the redshift distribution of the hosts \textit{with}
+%redshift, this implies that we are \textit{not} missing a dominant fraction of
+%sGRB redshift at $z \approx 2$, due to host faintness.
+
+A fraction of the bursts missing redshift are host-less and is therefore likely
+at moderate redshifts \citep{Tunnicliffe2014}, but should some of the remainder
+be at high redshift, the missed fraction will increase. If we assume that
+\textit{all} the bursts that are missing redshifts in \citet{Berger2014} are at
+high-$z$ and missed due to host faintness, 22 out of 39 hosts (55 per cent)
+would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
+missed burst at high-$z$ and illustrates that we are likely \textit{not} missing a
+large fraction of sGRBs redshift at $z \approx 2$, due to host faintness.
 
@@
 
-At $z = 2.211$, the age of the universe is 3 Gyr. If the progenitor systems
-of sGRBs are the merger of two neutron stars, this sets a hard upper limit to
-the coalescence timescale for such a system. In the absence of other mechanisms,
-the timescale of the orbital decay of the system is set by the energy loss due
-to gravitational waves, which in turn is set by the mass of constituent compact
+At $z = 2.211$, the age of the universe is 3 Gyr. If the progenitor systems of
+sGRBs are the merger of two neutron stars, this sets a hard upper limit to the
+coalescence timescale for such a system. In the absence of other mechanisms, the
+timescale of the orbital decay of the system is set by the energy loss due to
+gravitational waves, which in turn is set by the mass of constituent compact
 objects and the separation of the two \citep{Postnov2014}. If we assume that the
@@
 the big bang \citep{Richard2011} and if we assume a mass of 1.4~$M_\odot$ for
-each of the neutron stars at the time of system formation, this places a limit
-on the initial separation of the two neutron stars of $a_0 < 3.2~R_\odot$,
-where $a_0$ is the initial separation.
+each of the neutron stars at the time of system formation, this places a hard
+upper limit on the initial separation, $a_0$, of the two neutron stars of $a_0 <
+3.2~R_\odot$.
 
-Assuming our measured age of the stellar population ($\tau = 0.5$~Gyr) as a
-limit for the orbital decay time we can put a tighter constraint of $a_0 < 2.1~R_\odot$. 
-However, this does not account for the possibility there could be an
-underlying stellar population of older stars from a previous star-formation
-epoch, which is why we consider $a_0 < 3.2~R_\odot$ as a hard upper limit. The
-delay time between formation and explosion is well accommodated by the models of
+If we use the stellar population age from our SED fit, then we obtain a
+(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
+does not account for the possibility there could be an underlying stellar
+population of older stars from a previous star-formation epoch. The delay time
+between formation and explosion is well accommodated by the models of
 \citet{Belczynski2006}, although the longest formation channels are excluded.
@@
 setting it at $z = 2.211$. This value supersedes and corrects the previous
-% TK: No need to repeat a wrong redshift here, so slightly rephrased
 photometric redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. The erroneous redshift estimate of previous authors is attributed to a discrepancy in the measured $z$-band magnitude.
@@
 Compared to a sample of short GRB hosts, GRB~111117A is more luminous than 70 per
-cent of the sample with measured luminosities. For 30 per
-cent of the host sample, we would be unable to determine a redshift should they be at 
-a similar redshift as GRB~111117A. This implies that we are \textit{not} missing a dominant
+cent of the sample with measured luminosities. Conservatively, for 55 per
+cent of the sGRB hosts, we would be unable to determine a redshift should they be at 
+a similar redshift as GRB~111117A. This implies that we are \textit{not} missing a large
 fraction of the sGRBs at $z \sim 2$.

### 2017-06-01 py/Bergersample.py

@@ -22,7 +22,7 @@ def main():
     print(L_B_star)
     print()
     # L_B for a galaxy, 1 mag fainter
-    dM = 1
+    dM = 0.5
     print("L_B_limit:")
     print(L_B_star*10**(-0.4*(dM)))
     print()
@@ -54,12 +54,13 @@ def main():
     L_B_star = 10 ** (-1*(-22 - MB_star)/2.5)
 
 
-    lnunobs = len(L_B[L_B > L_B_star*10**(-0.4)])
+    lnunobs = len(L_B[L_B > L_B_star*10**(-0.4*(dM))])
     # Fraction unobservable
     print("Fraction unobservable:")
+    print(lnredshifts - lnunobs, lnredshifts)
     print(1 - lnunobs/lnredshifts)
 
-    
+
     # print(np.percentile(L_B[~np.isnan(L_B)], [14, 50, 85]))
 
 if __name__ == '__main__':

### 2017-06-01 GRB111117A.tex
Antonios comments
Closes #4

@@
 	\centering
-	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
+	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs2.pdf}
 	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid.
 Only one slit is shown, despite four epochs of spectroscopic observations because of the similarity in position angle.
-The image is the FORS2 $R$-band image and%, for which the photometry is shown in Fig.
-\ref{fig:SED}.
-	the blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}.
-The errorcircle of the \emph{Chandra}-position is too small to be visible in the image.}
+The image is the FORS2 $R$-band image for which the photometry is shown in Fig.~\ref{fig:SED}.
+	The blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}.
+The insert shows the 2D-image of the \oiii$\lambda$5007 emission line.
+The location of a bright skyline is marked by the blue box.
+The location of the emission line is indicated with the red ellipse.
+Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
 	\label{fig:spec_setup}
@@
 program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
-3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across
-three spectroscopic arms, simultaneously illuminated through the use of
-dichroics. The bias-correction, flat-fielding, order tracing, wavelength
-calibration, rectification, and flux calibration is then carried out using the
-VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}. 
-%Because the spectra are curved across each detector, a rectification algorithm
+3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used)
+across three spectroscopic arms. The bias-correction, flat-fielding, order
+tracing, wavelength calibration, rectification, and flux calibration is then
+carried out using the VLT/X-shooter pipeline version 2.8.4
+\citep{Modigliani2010}. %Because the spectra are curved across each detector, a
+% rectification algorithm
 %is employed which introduces correlations between neighboring pixels. We select
@@
 optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
-carried out with a simple aperture.
-
-An overview of the spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the position of the slit on the target is shown in Fig.~\ref{fig:spec_setup}.
-
-% \begin{figure}
-% 	\centering
-% 	\includegraphics[width=9cm]{figures/SEDspecphot.pdf}
-% 	\caption{SED fit showning the best-fist SED to the derived photometry.
-The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
-Overplotted in grey is a version of the observed spectrum binned by 30 \AA.
-The reason for the spectral gap at 10000 \AA is from the merging of the visual and near-infrared arm.}
-% 	\label{fig:SED}
-% \end{figure}
+carried out with a simple aperture. An overview of the spectroscopic
+observations is given in Table~\ref{tab:spec_overview}, and the position of the
+slit on the target is shown in Fig.~\ref{fig:spec_setup}.We show the extracted
+spectrum in Fig.~\ref{fig:SED}.
 
@@
 emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with
-\hb~detected at low significance ($\sim 3 \sigma$). \ha~is only visible in the
-first epoch, due to the $K$-band blocking filter used for the remainder
-observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and
-show significant velocity structure along the slit. A drop in the continuum to
-the blue of the \lya~line further supports the inferred redshift.
+\hb~detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
+in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
+epoch, due to the $K$-band blocking filter used for the remainder observations.
+The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
+significant velocity structure along the slit. A drop in the continuum to the
+blue of the \lya~line further supports the inferred redshift.
 
@@
 \citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
-From the SED-fit (Sect. \ref{SED}), and the detection of \lya, the host is
+From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
 constrained to contain very little or no dust, which is why we do not apply 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=9cm]{figures/OIII_img.pdf}
-	\caption{2D-image of the \oiii$\lambda$5007 emission line.
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
+%	\caption{2D-image of the \oiii$\lambda$5007 emission line.
 The location of a bright skyline is marked by the blue box.
@@
 Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
-	\label{fig:line}
-\end{figure}
+%	\label{fig:line}
+%\end{figure}
 
@@
 
-In addition to the spectroscopy presented above, we imaged the field of GRB~111117A 
-in multiple broad-band filters using the VLT equipped with FORS2 ($gRIz$ filters) and HAWK-I
-($JHK_{\mathrm{s}}$ filters), long after the burst has faded. These new data are
-complemented by a re-analysis of some of the imagery used in
-\citet{Margutti2012} and \citet{Sakamoto2013} that are available to us here (GTC
-$gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
-observations and measured brightnesses is given in Table~\ref{tab:phot_overview}.
+In addition to the spectroscopy presented above, we imaged the field of
+GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
+($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
+has faded. These new data are complemented by a re-analysis of some of the
+imagery used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available
+to us here (GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the
+photometric observations and measured brightnesses is given in
+Table~\ref{tab:phot_overview}.
 
@@
 
-The multi-color spectral energy distribution (SED) is fit by
-\citet{Bruzual2003} single stellar population models based on a
-\citet{Chabrier2003} initial mass function in \emph{LePhare} \citep{Ilbert2006},
-where the redshift is fixed to the spectroscopic value of $z=2.211$. The best
-model is obtained with an unreddened galaxy template, and returns physical
-parameters of absolute magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass
-($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), stellar population age ($\tau =
-0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
-($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). %We show the SED
-% fit in Fig. \ref{fig:SED}.
+The multi-color spectral energy distribution (SED) is fit by \citet{Bruzual2003}
+single stellar population models based on a \citet{Chabrier2003} initial mass
+function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
+the spectroscopic value of $z=2.211$. The best model is obtained with an
+unreddened galaxy template, and returns physical parameters of absolute
+magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) =
+9.9\pm0.2$), stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and
+star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9}
+M_\odot\,\mathrm{yr}^{-1}$). We show the SED fit in Fig.~\ref{fig:SED}.
 
@@
 \times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
-the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \cite{Willingale2013}.
+the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.
 
@@
 \citep{Littlejohns2013a}, the updated redshift does not change this designation
-significantly. The intrinsic luminosity increases, as visible in the X-ray light curve
-in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
-the majority of long GRBs. \citet{Bromberg2013} investigated the degree to which the
+significantly. 
+%The intrinsic luminosity increases, as visible in the X-ray light curve
+%in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
+%the majority of long GRBs. 
+\citet{Bromberg2013} investigated the degree to which the
 long and short population distributions overlap and quantified the certainty in
 the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
-being short. Compared to the other two sGRB candidates at high redshift, GRB060121
+being short. Compared to the other two sGRB candidates at high redshift, GRB~060121
 \citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
-($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010,
+($17_{-15}^{+14}$ per cent) and GRB~090426 \citep{Antonelli2009, Levesque2010,
 	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
-	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
+Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
+intermediate duration bursts. This comes from both events having very soft
+spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
+bursts are very clearly related in their properties to long GRBs
+\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
+mergers.
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
+%	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
 The grey shared region is a compilation of long GRB lightcurves where the color represents density and the light blue is other short GRB lightcurves for which redshifts has been determined.
 Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
-	\label{fig:sxray_lightcurve}
-\end{figure}
+%	\label{fig:sxray_lightcurve}
+%\end{figure}
 
@@
 
-We plot the recalculated $N_\mathrm{H}$ in Fig. \ref{fig:NH_z} where we compare with the
+We plot the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare with the
 distributions of complete samples of both long and short GRBs. The lGRB sample
@@
 the short sample. Bursts without redshifts have been excluded for both groups.
-GRB~111117A occupies a unique position in Fig. \ref{fig:NH_z} with the highest
+GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
 $N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
@@
 	\caption{Rest frame, X-ray derived hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations.
-In red we show the sample of lGRBs from \citet{Arcodia2016}.
-The detections are replaced with contours for clarity and the limits are shown with arrows.
-In blue is the sample of sGRBs from \citet{DAvanzo2014a}.
-GRB~111117A is an outlier from the short sample, but seems to follow the distribution of long GRBs.}
+The detections are replaced with contours for clarity and the limits are shown with arrows.}
 	\label{fig:NH_z}
@@
 redshift harder to determine. Additionally, it is natural to expect some
-evolution in the hosts of sGRBs with redshift as illustrated in Sect.
-\ref{restnH}.
+evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.
 
@@
 \begin{acknowledgements}
-TK acknowledges support through the Sofja Kovalevskaja Award to P.
-Schady.
+TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady. 
 CT acknowledges support from a Spanish National Research Grant of Excellence under project AYA 2014-58381-P and funding associated to a Ramón y Cajál fellowship under grant number RyC-2012-09984.
+AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation Grant for Researchers and Cultural Creators,and the Spanish Ministry of Economy and Competitiveness through project AYA2014-58381-P.
+Partly based on observations made with the Gran Telescopio Canarias (GTC).
+
 \end{acknowledgements}
@@
 \newpage
+%\appendix
 
@@
 
+ \begin{figure*}
+ 	\centering
+ 	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
+ 	\caption{SED fit showning the best-fist SED to the derived photometry.
+The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
+Overplotted in grey is the observed spectrum.
+The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
+ 	\label{fig:SED}
+ \end{figure*}
+
 \end{document}

### 2017-06-01 tables/phot_overview.tex
Antonios comments
Closes #4

@@
 2013-07-17T08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
-2011-11-28T21:10 & 3.60 & TNG/LRS & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
+2011-11-28T21:10 & 3.60 & TNG/DOLORES & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
 2011-11-17T20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   

### 2017-06-01 GRB111117A.tex
Christinas comments
Contained partly in Nials comments. Closes #5

@@
 
-In this \emph{Letter}, we have provided a revised, spectroscopic redshift for the short
-GRB~111117A based on emission lines %discovered using an improved reduction scheme, 
-setting it at $z = 2.211$. This value supersedes and corrects the previous
-photometric redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. The erroneous redshift estimate of previous authors is attributed to a discrepancy in the measured $z$-band magnitude.
+In this \emph{Letter}, we have provided a revised, spectroscopic redshift for
+the short GRB~111117A based on emission lines setting it at $z = 2.211$. This
+value supersedes the previous photometric redshift of \citet{Margutti2012} and
+\citet{Sakamoto2013}. %The erroneous redshift estimate of previous authors is
+attributed to a discrepancy in the measured $z$-band magnitude.
 
@@
 hydrogen column density towards GRB~111117A is the highest within a complete
-sample of short hosts, but is indicative of an evolution with redshift as found
+sample of short hosts and is indicative of an evolution with redshift as found
 for the hosts of long GRBs.
@@
 The SFR of the host is in the upper end of the sGRB host SFR distribution and
-this helped us to measure the redshift for this galaxy. Despite
-the moderate age and high $N_\mathrm{H}$, almost no dust is present.
+this helped us to measure the redshift for this galaxy. Despite the moderate age
+and high $N_\mathrm{H}$, almost no dust is present.
 

### 2017-06-01 py/SED_plot.py

@@ -14,13 +14,13 @@ import seaborn; seaborn.set_style('ticks')
 import matplotlib as mpl
 from matplotlib.ticker import FormatStrFormatter
 params = {
-   'axes.labelsize': 16,
-   'font.size': 16,
-   'legend.fontsize': 16,
-   'xtick.labelsize': 16,
-   'ytick.labelsize': 16,
+   'axes.labelsize': 24,
+   'font.size': 24,
+   'legend.fontsize': 24,
+   'xtick.labelsize': 24,
+   'ytick.labelsize': 24,
    'text.usetex': True,
-   'figure.figsize': [7.281, 4.5]
+   'figure.figsize': [16, 16/1.618]
    }
 mpl.rcParams.update(params)
 
@@ -36,7 +36,7 @@ def main():
     # Small script to plot SED, photometry and spectrum
 
     # Load data
-    spectrum = np.genfromtxt("../data/stitched_spectrum_bin50.dat")
+    spectrum = np.genfromtxt("../data/stitched_spectrum_bin1.dat")
     wl_spec = spectrum[:, 0]
     flux_spec = spectrum[:, 1]
     error_spec = spectrum[:, 2]
@@ -71,11 +71,11 @@ def main():
     print(AB - AB_err_hi)
     AB_err_lo = -2.5*np.log10(flux_spec - error_spec) - 5*np.log10(wl_spec) - 2.406
 
-    pl.errorbar(wl_spec, AB, yerr=[AB - AB_err_lo, AB_err_hi - AB], fmt=".k", capsize=0, elinewidth=0.5, ms=3, alpha=0.3, label="X-shooter spectrum", zorder=1)
-    pl.errorbar(wl_phot, mag, xerr=wl_plot_wid, yerr=magerr, fmt=".", capsize=0, ms=10, color = "#4C72B0", label = "Photometric points", zorder=3, mec="black", mew = 1)
+    # pl.errorbar(wl_spec, AB, yerr=[AB - AB_err_lo, AB_err_hi - AB], fmt=".k", capsize=0, elinewidth=0.5, ms=3, alpha=0.3, label="X-shooter spectrum", zorder=1)
+    pl.errorbar(wl_phot, mag, xerr=wl_plot_wid, yerr=magerr, fmt=".", capsize=0, ms=20, color = "#4C72B0", label = "Photometric points", zorder=3, mec="black", mew = 1)
[…]

### 2017-06-01 py/combine_arms_and_bin.py

@@ -177,9 +177,9 @@ def main():
     VIS_wl, VIS_flux, VIS_error, VIS_bp = bin_spectrum(VIS_wl, VIS_flux, VIS_error, VIS_bp.astype("bool"), 3, weight=True)
 
     # Construct lists
-    UVB_mask = (UVB_wl > 3300) & (UVB_wl < 5600)
-    VIS_mask = (VIS_wl < 9800)
-    NIR_mask = (NIR_wl > 13000) & (NIR_wl < 20000) & (NIR_flux > -1e-19)
+    UVB_mask = (UVB_wl > 3300) & (UVB_wl < 5500)
+    VIS_mask = (VIS_wl > 5700) & (VIS_wl < 9600)
+    NIR_mask = (NIR_wl > 12000) & (NIR_wl < 21000) & (NIR_flux > -1e-19)
     waves = [UVB_wl[UVB_mask], VIS_wl[VIS_mask], NIR_wl[NIR_mask]]
     flux = [UVB_flux[UVB_mask], VIS_flux[VIS_mask], NIR_flux[NIR_mask]]
     error = [UVB_error[UVB_mask], VIS_error[VIS_mask], NIR_error[NIR_mask]]
@@ -190,7 +190,7 @@ def main():
 
     np.savetxt(data_dir + "stitched_spectrum.dat", zip(wl, flux, error, bpmap), fmt = ['%10.6e', '%10.6e', '%10.6e', '%10.6f'], header=" wl flux error bpmap")
 
-    hbin = 10
+    hbin = 1
     wl_bin, flux_bin, error_bin, bp_bin = bin_spectrum(wl, flux, error, bpmap.astype("bool"), hbin, weight=True)
 
 

### 2017-06-01 py/SED_plot.py
small edit

@@ -74,7 +74,7 @@ def main():
     # pl.errorbar(wl_spec, AB, yerr=[AB - AB_err_lo, AB_err_hi - AB], fmt=".k", capsize=0, elinewidth=0.5, ms=3, alpha=0.3, label="X-shooter spectrum", zorder=1)
     pl.errorbar(wl_phot, mag, xerr=wl_plot_wid, yerr=magerr, fmt=".", capsize=0, ms=20, color = "#4C72B0", label = "Photometric points", zorder=3, mec="black", mew = 1)
     # pl.errorbar(wl_spec, AB, fmt=".k", capsize=0, elinewidth=0.5, ms=3, alpha=0.3)
-    # pl.plot(wl_spec[::3], AB[::3], color="black", linestyle="steps-mid", lw=0.3, alpha=0.3, label="X-shooter spectrum", zorder=1)
+    pl.plot(wl_spec[::3], AB[::3], color="black", linestyle="steps-mid", lw=0.3, alpha=0.3, label="X-shooter spectrum", zorder=1)
     pl.plot(wl_SED, AB_SED, linestyle="steps-mid", lw=2.0, color = "#C44E52", label = "SED fit", zorder=2)
     # pl.axhline(0, linestyle="dashed", color = "black", lw = 0.4)
     pl.legend(loc=2)

### 2017-06-02 GRB111117A.tex
Address Zachs comments
Phew. Closes #6

@@
 We use the spectroscopic redshift, as well as new imaging data to constrain the
-nature of the host galaxy as well as physical parameters of the GRB. The rest
-frame X-ray derived hydrogen column density, for example, is high compared to a
+nature of the host galaxy as well as physical parameters of the GRB. The rest-frame X-ray derived hydrogen column density, for example, is high compared to a
 complete sample of sGRBs and seems to follow the evolution with redshift as
 traced by the hosts of long GRBs (lGRBs). This is consistent with the proportion
-of the sGRB population originating in late-type galaxies live in similar
+of the sGRB population originating in late-type galaxies, live in similar
 environments to that of lGRB hosts.
@@
 
-From the age of the universe at explosion time, an initial progenitor separation
-of $a_0 < 3.2~R_\odot$ is required in the case of a binary neutron star system. 
-This puts constraints on the progenitor system evolution up to the time
-of explosion.
+From the age of the universe at the time of GRB explosion, an initial neutron
+star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
+progenitor is a pair of inspiralling NSs. This constraint excludes some of the
+longest formation channels for this burst.
+
 }
@@
 
-There is now mounting evidence that most short-duration gamma-ray bursts (GRBs)
-come from the merger of neutron stars (NSs), either with another NS, or a black
+There is now mounting evidence that most short-duration $\gamma$-ray bursts
+come from the merger of NSs, either with another NS, or a black
 hole, due to their apparent association with kilonovae \citep{Barnes2013a,
@@
 
-The host galaxies of short GRBs are diverse. They are more massive and less actively 
-star-forming on average than long GRB hosts \citep{Fong2013b}, while in some cases, no host 
+The host galaxies of sGRBs are diverse. They are more massive and less actively 
+star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases, no host 
 galaxy can be identified \citep{Berger2010a, Tunnicliffe2014}. Together with their position 
 within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
-long-lived, and is associated with stellar mass rather than star-formation rate
+long-lived, and is associated with stellar mass rather than star-formation rate (SFR)
 \citep{Berger2014}. The median redshift for sGRBs is $z\sim0.5$ \citep{Berger2014}, and
@@
 The total lifetimes of NS binaries depends on their initial separations and
-subsequent inspiral times, and impacts the timing and distribution of the
-enrichment of the ISM and subsequent stars and planets with heavy $r$-process
+subsequent inspiral times. The delay time from formation to explosion impacts
+the timing and distribution of the enrichment of the ISM with heavy $r$-process
 elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some limits can be
@@
 
-In this \emph{Letter} we present the spectrum of the host galaxy of the short GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is significantly higher 
+In this \emph{Letter} we present a spectrum of the host galaxy of the short GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is significantly higher 
 than the previously estimated redshift based on photometric studies. 
@@
 %distant short GRB.
-
-Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology provided by
+Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology parameters provided by
 \citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
-and $\Omega_m = 0.307$ and report all magnitudes in the AB system.
+and $\Omega_m = 0.307$. All magnitudes are given in the AB system.
 
@@
 spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
-burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger
-under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO
-program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
-3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used)
-across three spectroscopic arms. The bias-correction, flat-fielding, order
-tracing, wavelength calibration, rectification, and flux calibration is then
-carried out using the VLT/X-shooter pipeline version 2.8.4
-\citep{Modigliani2010}. %Because the spectra are curved across each detector, a
+burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
+ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
+091.D-0904 (PI: Hjorth). X-shooter covers the wavelength range from 3000~\AA~to
+24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
+spectroscopic arms. We carried out the bias-correction, flat-fielding, order
+tracing, wavelength calibration, rectification, and flux calibration using the
+VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
+%Because the spectra are curved across each detector, a
 % rectification algorithm
@@
 et al. 2017 (in prep.) and available
-online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
-signal-to-noise of the continuum in the near-infrared arm is too low to use the
-optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
-carried out with a simple aperture. An overview of the spectroscopic
-observations is given in Table~\ref{tab:spec_overview}, and the position of the
-slit on the target is shown in Fig.~\ref{fig:spec_setup}.We show the extracted
-spectrum in Fig.~\ref{fig:SED}.
+online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
+%The
+%signal-to-noise of the continuum in the near-infrared arm is too low to use the
+%optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
+%carried out with a simple aperture.
+An overview of the spectroscopic observations is given in
+Table~\ref{tab:spec_overview}, and the slit position is shown in
+Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in Fig.~\ref{fig:SED}.
 
 We determine a redshift of $z = 2.211$ from the simultaneous detection of
-emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with
-\hb~detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
+emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
+\hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
 in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
-epoch, due to the $K$-band blocking filter used for the remainder observations.
+epoch due to the $K$-band blocking filter used for the remainder observations.
 The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
@@
 21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
-intersect the line, making an accurate estimate of the \ha-flux difficult. 
+are superposed on the line, making an accurate estimate of the \ha-flux difficult. 
 We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
@@
 
-%Because the projected distance does not change much between redshift 1.3 and
-% 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating
-% to host offset are unaffected.
-
 \subsection{Imaging observations and SED analysis} \label{SED}
@@
 ($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
-has faded. These new data are complemented by a re-analysis of some of the
-imagery used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available
-to us here (GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the
-photometric observations and measured brightnesses is given in
+faded. These new data are complemented by a re-analysis of some of the imaging
+used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
+(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
+observations and measured brightnesses is given in
 Table~\ref{tab:phot_overview}.
@@
 All data were reduced, analyzed and fitted in a similar manner as described in detail in
-\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Very briefly, we use our own
-\texttt{python} and IRAF routines to perform a standard reduction which includes
+\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Briefly, we use our own
+\texttt{Python} and IRAF routines to perform a standard reduction which includes
 bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
@@
 of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. 
-For the $R$ and $I$-band photometry we used the color transformations of
-Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
-We convert all magnitudes into the AB system if necessary, and correct for the Galactic 
+%For the $R$ and $I$-band photometry, we used the color transformations of
+%Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
+We convert all magnitudes into the AB system, and correct for a Galactic 
 foreground of $E_{B-V}=0.027~\mathrm{mag}$.
@@
 function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
-the spectroscopic value of $z=2.211$. The best model is obtained with an
+the spectroscopic value of $z=2.211$. The best fit model is an
 unreddened galaxy template, and returns physical parameters of absolute
@@
 Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
-re-analysis of Gemini data to the $z$-band measurements of \citet{Margutti2012} and
+re-analysis of the Gemini data to the $z$-band measurements of \citet{Margutti2012} and
 \citet{Sakamoto2013}. Both of these authors report $z$-band photometry that is brighter 
-by 0.8~mag to 1.0 mag than our value, whereas data taken in bluer filters are in excellent agreement. In fact, the large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
+by 0.8~mag to 1.0 mag compared to our value, where data taken in bluer filters are in excellent agreement. The large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
 driving the galaxy photometric redshift of the earlier works. Using
@@
 repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
-\citep{Evans2009}. The X-ray afterglow light curve can be fit with a single power-law decay with
-index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
-analysis using data in photon counting (PC) mode in the widest time epoch where
-the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
-(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
-data) to prevent spectral changes that can affect the X-ray column density
-determination. The obtained spectrum is well described by an absorbed power-law
-model. The best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
+\citep{Evans2009}. 
+%The X-ray afterglow light curve can be fit with a single power-law decay with an
+%index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
+%analysis using data obtained in photon counting (PC) mode in the widest time epoch where
+%the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
+%(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
+%data) to prevent spectral changes that can affect the X-ray column density
+%determination. 
+The obtained spectrum is well described by an absorbed power-law
+model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
 an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Reinterpretation of restframe properties}
+\section{Reinterpretation of the restframe properties}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Because the projected distance does not change much between $z = 1.3$ and $z =
+2.211$, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013}
+relating to host offset are unaffected.
 
@@
 
-As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A securely 
-belongs the short class of GRBs. Because the observed classification
+As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is securely 
+classified as belonging to the short class of GRBs. Because the observed classification
 indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
@@
 the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
-being short. Compared to the other two sGRB candidates at high redshift, GRB~060121
+being a sGRB. Compared to the other two sGRB candidates at high redshift, GRB~060121
 \citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
@@
 
-Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
-intermediate duration bursts. This comes from both events having very soft
-spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
-bursts are very clearly related in their properties to long GRBs
-\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
-mergers.
+%Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
+%intermediate duration bursts. This comes from both events having very soft
+%spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
+%bursts are very clearly related in their properties to lGRBs
+%\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
+%mergers.
 
@@
 
-We plot the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare with the
-distributions of complete samples of both long and short GRBs. The lGRB sample
-is from \citet{Arcodia2016} and the sGRB sample is from \citet{DAvanzo2014a}.
-17 of the 99 long bursts do not have redshifts and likewise 5 out of 16 for
-the short sample. Bursts without redshifts have been excluded for both groups.
-GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
-$N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
-low redshifts ($z < 1$) and is found to populate a similar column density
-environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
+We show the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare
+with the distributions of complete samples of both long and short GRBs. The lGRB
+sample is from \citet{Arcodia2016} and the sGRB sample is from
+\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have redshifts, and
+likewise for 5 out of 16 in the short sample. Bursts without redshifts have been
+excluded for both groups. GRB~111117A occupies a unique position in
+Fig.~\ref{fig:NH_z} with the highest $N_H$ of all short bursts. Additionally the
+upper limit at $z = 2.609$ belongs to GRB~090426 and does not belong in the
+short sample. The short sample, excluding GRB~111117A, is located at low
+redshifts ($z < 1$) and is found to populate a similar column density
+environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
 hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
-as a function of redshift as found for the long GRB afterglows \citep{Arcodia2016}.
+as a function of redshift as found for the lGRB afterglows \citep{Arcodia2016}.
 
@@
 %From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars,
-As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a
+As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
 late-type galaxy and is entirely consistent in terms of stellar mass and stellar
-age with the general population of short GRB hosts ($\left\langle M _*
+age with the general population of sGRB hosts ($\left\langle M _*
 \right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
@@
 \citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
-typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to
-the SFRs found in the hosts of long GRBs at a corresponding redshift
+typical SFR for sGRB host galaxies \citep{Berger2014} and more similar to
+the SFR found in the hosts of lGRBs at a corresponding redshift
 \citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
-star-forming galaxy would exhibit weaker emission lines, thus making the
+star forming galaxy would exhibit weaker emission lines, thus making the
 redshift harder to determine. Additionally, it is natural to expect some
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implications for redshift distribution of short GRBs}
+\section{Implications for the redshift distribution of sGRBs}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A single sGRB at high redshift does little in terms of constraining the redshift
-distribution and therethrough the progenitor channels. In particular, other
-short-GRB hosts could be missed because they are intrinsically fainter and thus
-this high-$z$ event is only detected due to the brightness of the host.
-\citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by
-the characteristic galaxy luminosity at their respective redshift,
-$L_B/L^{\star}_{B}$. 26 out of 39 hosts (66 per cent) in the sample have
-redshifts. To convert the SED-inferred $M_B$ of GRB~111117A to
-$L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the
-Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window
-$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} =
-1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with
-measured $L_B/L^{\star}_{B}$.
+distribution of sGRBs. In particular, other sGRB hosts could be missed because
+they are intrinsically fainter and thus this high-$z$ event is only detected due
+to the brightness of it's host. \citet{Berger2014} compiled a sample of sGRB
+host luminosities, normalized by the characteristic galaxy luminosity at their
+respective redshift, $L_B/L^{\star}_{B}$. Out of the 39 hosts in the sample, 26
+(66 per cent) have redshifts. To convert the SED-inferred $M_B$ of GRB~111117A
+to $L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band magnitude of
+the Schechter function for blue galaxies ($U - V < 0.25$) in the redshift window
+$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007}. For GRB~111117A, we obtain
+$L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in
+\citet{Berger2014} with measured $L_B/L^{\star}_{B}$.
 
@@
 missed burst at high-$z$ and illustrates that we are likely \textit{not} missing a
-large fraction of sGRBs redshift at $z \approx 2$, due to host faintness.
+large fraction of sGRBs redshift at $z \approx 2$ due to host faintness.
 
 The theoretical redshift distribution of sGRBs depends on the type of delay-time
-function used to model the progenitor system. The redshift of GRB~111117A puts
-constraints on the type of delay-time models suitable for modeling. The
-likelihood preferred lognormal time delay models investigated in
-\citet{Wanderman2015} predicts a rate of sGRBs at $z = 2.211$, $\sim$ two orders
-of magnitude lower compared to peak rate ($z = 0.9$). It is stated in
-\citet{Wanderman2015} that this preference depends critically on the absence of
-non-collapsar sGRB at $z \gtrsim 1.2$. The redshift of the burst, on the other
-hand, is close to the expected peak in sGRB rate calculated for the power law
-models \citep{Behroozi2014, Wanderman2015}.
+function used to model the progenitor system. The likelihood preferred lognormal
+time delay models investigated by \citet{Wanderman2015} predict a sGRB rate at
+$z = 2.211$, $\sim$ two orders of magnitude lower compared to the peak rate at
+$z = 0.9$. According to \citet{Wanderman2015}, this preference depends
+critically on the absence of non-collapsar sGRBs at $z \gtrsim 1.2$. The
+redshift of GRB~111117A, on the other hand, is close to the expected peak in
+sGRB rate calculated using the power law delay time models \citep{Behroozi2014,
+	Wanderman2015}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Constraints on the progenitor separation}
+\section{Constraints on progenitor separation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 At $z = 2.211$, the age of the universe is 3 Gyr. If the progenitor systems of
-sGRBs are the merger of two neutron stars, this sets a hard upper limit to the
-coalescence timescale for such a system. In the absence of other mechanisms, the
-timescale of the orbital decay of the system is set by the energy loss due to
-gravitational waves, which in turn is set by the mass of constituent compact
+sGRBs are the merger of two NSs, this sets a hard upper limit to the coalescence
+timescale for such a system. In the absence of other mechanisms, the timescale
+of the orbital decay of the system is set by the energy loss due to
+gravitational waves, which in turn is set by the mass of the constituent compact
 objects and the separation of the two \citep{Postnov2014}. If we assume that the
 formation timescale of the first galaxies is short compared to the time since
-the big bang \citep{Richard2011} and if we assume a mass of 1.4~$M_\odot$ for
-each of the neutron stars at the time of system formation, this places a hard
-upper limit on the initial separation, $a_0$, of the two neutron stars of $a_0 <
-3.2~R_\odot$.
+the Big Bang \citep{Richard2011}, and if we assume a mass of 1.4~$M_\odot$ for
+each of the NSs at the time of system formation, this places a hard upper limit
+on the initial separation, of $a_0 < 3.2~R_\odot$.
 
@@
 between formation and explosion is well accommodated by the models of
-\citet{Belczynski2006}, although the longest formation channels are excluded.
+\citet{Belczynski2006}, although the longest delay times are excluded.
 This is especially true given the late type nature of the host
@@
 \citet{Sakamoto2013}. %The erroneous redshift estimate of previous authors is
-attributed to a discrepancy in the measured $z$-band magnitude.
+%attributed to a discrepancy in the measured $z$-band magnitude.
 
-The rest-frame parameters of the burst and the conditions of the burst
-environment have been recalculated using the new distance. The X-ray derived
-hydrogen column density towards GRB~111117A is the highest within a complete
-sample of short hosts and is indicative of an evolution with redshift as found
-for the hosts of long GRBs.
-
-The SFR of the host is in the upper end of the sGRB host SFR distribution and
-this helped us to measure the redshift for this galaxy. Despite the moderate age
-and high $N_\mathrm{H}$, almost no dust is present.
+Using the new distance, the X-ray derived $n_H$ towards GRB~111117A is the
+highest within a complete sample of sGRB hosts and is consistent with the
+$n_H-z$ evolution traced by the hosts of lGRBs.The SFR of the host is in the
+upper end of the sGRB host SFR distribution and despite the moderate age and
+high $N_\mathrm{H}$, almost no dust is present.
 
@@
 the faintness of their hosts, a lognormal delay time model will be disfavored.
-Compared to a sample of short GRB hosts, GRB~111117A is more luminous than 70 per
+Compared to a sample of sGRB hosts, GRB~111117A is more luminous than 70 per
 cent of the sample with measured luminosities. Conservatively, for 55 per
@@
 constraints on the maximal separation between the engine constituents at the
-time of formation. We find that the maximal separation for two neutron stars at
+time of formation. We find that the maximal separation of two NSs at
 formation time is $a_0 < 3.2~R_\odot$, which excludes some of the formation
@@
 Partly based on observations made with the Gran Telescopio Canarias (GTC).
+ZC acknowledges support from the Spanish research project AYA 2014-58381-P and support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2014-21669.
 

### 2017-06-09 GRB111117A.tex
Address Andrews comments
Closes #7

@@
 We use the spectroscopic redshift, as well as new imaging data to constrain the
-nature of the host galaxy as well as physical parameters of the GRB. The rest-frame X-ray derived hydrogen column density, for example, is high compared to a
+nature of the host galaxy and the physical parameters of the GRB. The rest-frame X-ray derived hydrogen column density, for example, is high compared to a
 complete sample of sGRBs and seems to follow the evolution with redshift as
-traced by the hosts of long GRBs (lGRBs). This is consistent with the proportion
-of the sGRB population originating in late-type galaxies, live in similar
+traced by the hosts of long GRBs (lGRBs). This is consistent with a proportion
+of the sGRB population originating in late-type galaxies, living in similar
 environments to that of lGRB hosts.
@@
 	\centering
-	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs2.pdf}
+	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs4.pdf}
 	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid.
@@
 model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
-an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
-\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
+an intrinsic equivalent hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
+\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a solar abundance and a Galactic $N_{\mathrm{H}}$ in
 the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.
@@
 
-Because the projected distance does not change much between $z = 1.3$ and $z =
+Because the projected distance does not change significantly between $z = 1.3$ and $z =
 2.211$, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013}
@@
 
-As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is securely 
-classified as belonging to the short class of GRBs. Because the observed classification
-indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
-\citep{Littlejohns2013a}, the updated redshift does not change this designation
-significantly. 
-%The intrinsic luminosity increases, as visible in the X-ray light curve
+As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is
+securely classified as belonging to the short class of GRBs. Because the
+observed classification indicators, $T_{90}$ and hardness ratio, do not depend
+strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not
+change this designation. %The intrinsic luminosity increases, as
+% visible in the X-ray light curve
 %in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
 %the majority of long GRBs.
-\citet{Bromberg2013} investigated the degree to which the
-long and short population distributions overlap and quantified the certainty in
-the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
-being a sGRB. Compared to the other two sGRB candidates at high redshift, GRB~060121
-\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
-($17_{-15}^{+14}$ per cent) and GRB~090426 \citep{Antonelli2009, Levesque2010,
-	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
-membership for GRB~111117A is much higher.
+\citet{Bromberg2013} investigated the degree to which the long and short
+population distributions overlap and quantified the certainty in the class
+membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
+percent probability of being a sGRB. Compared to the other two sGRB candidates
+at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
+\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
+\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
+per cent), the certainty in class membership for GRB~111117A is much higher.
+This means that GRB~111117A is likely older by 1.8 Gyr than any other
+non-collapsar GRB ever detected.
 
@@
 
+For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
+\citep{Lyman2017}, which is a challenge to reconcile  with the relatively large
+offset from the host center derived in \citet{Margutti2012} and
+\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
+burst position that we don’t see. Along with the absence of dust, the large
+offset from the host center indicates that the high $N_H$ arises because the
+density in the GRB surrounding is high, or because the light from the afterglow
+traverses a region of star formation.
+
 \begin{figure}
@@
 	\includegraphics[width=9cm]{figures/NH_z.pdf}
-	\caption{Rest frame, X-ray derived hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations.
+	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations.
 The detections are replaced with contours for clarity and the limits are shown with arrows.}
@@
 
-%If we additionally assume that the hosts missing redshift in
-%\citet{Berger2014} follow the redshift distribution of the hosts \textit{with}
-%redshift, this implies that we are \textit{not} missing a dominant fraction of
-%sGRB redshift at $z \approx 2$, due to host faintness.
-
 A fraction of the bursts missing redshift are host-less and is therefore likely
@@
 
-%In terms of Fig.
-9 in \citet{Fong2013a}, the host is indeed  brighter than the average galaxies using the evolving luminosity function of galaxies.
-\todo{What is the average brightness of a galaxy at z = 2?}
-%\todo{Reformulate paragraph}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 on the initial separation, of $a_0 < 3.2~R_\odot$.
+
+In practice most NS-NS binaries will be eccentric at formation because of the SN
+natal kicks, and this can be quite a big effect on merger times, although
+eccentric systems normally circularise first. It could also be a BH-NS system,
+and as the merger time scales roughly as $M^3$ the mass here almost as big an
+effect as the separation ($a^4$).
 
@@
 upper end of the sGRB host SFR distribution and despite the moderate age and
-high $N_\mathrm{H}$, almost no dust is present.
+high $N_\mathrm{H}$, almost no dust is present. The high $N_H$ is difficult to
+reconcile with the large projected host offset and the absence of dust. One
+possible explanation could be, that GRB~111117A is formed through the prompt
+channel of sGRBs \todo{ref} and originates in a star forming region located in
+the outskirts of the host.
 

### 2017-06-11 GRB111117A.tex
Addresses Lises comments
Closes #8

@@
 spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
-= 2.211$. This makes GRB~111117A the most distant high-confidence short GRB
+= 2.211$. This makes GRB~111117A the most distant high-confidence short duration GRB
 detected to date. Our spectroscopic redshift supersedes a lower redshift value
@@
 star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
-progenitor is a pair of inspiralling NSs. This constraint excludes some of the
+progenitor is a circular pair of inspiralling NSs. This constraint excludes some of the
 longest formation channels for this burst.
@@
 
-There is now mounting evidence that most short-duration $\gamma$-ray bursts
+There is mounting evidence that short-duration $\gamma$-ray bursts
 come from the merger of NSs, either with another NS, or a black
@@
 The host galaxies of sGRBs are diverse. They are more massive and less actively
-star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases, no host 
-galaxy can be identified \citep{Berger2010a, Tunnicliffe2014}. Together with their position 
-within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
-long-lived, and is associated with stellar mass rather than star-formation rate (SFR)
-\citep{Berger2014}. The median redshift for sGRBs is $z\sim0.5$ \citep{Berger2014}, and
-because most of these measurements come from the associated hosts, it is arguably biased
-towards lower redshifts.
+star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases,
+no host galaxy can be identified above detection threshold of deep follow-up
+observations \citep{Berger2010a, Tunnicliffe2014}. Together with their position
+within their hosts \citep{Fong2013a}, this suggests a progenitor system that can
+be very long-lived, and is associated with stellar mass rather than
+star-formation rate (SFR) \citep{Berger2014}. The median redshift for sGRBs is
+$z\sim0.5$ \citep{Berger2014}, and because most of these measurements come from
+the associated hosts, it is arguably biased towards lower redshifts.
 
@@
 
-The total lifetimes of NS binaries depends on their initial separations and
-subsequent inspiral times. The delay time from formation to explosion impacts
-the timing and distribution of the enrichment of the ISM with heavy $r$-process
-elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some limits can be
-calculated based on models of star-formation histories of, and the spatial
-distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}. The most
-distant cosmological bursts, however, offer direct, hard limits.
+The total lifetimes of NS binaries depends on their orbit, mass, spin, initial
+separations and subsequent inspiral times. The delay time from formation to
+explosion impacts the timing and distribution of the enrichment of the ISM with
+heavy $r$-process elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some
+limits can be calculated based on models of star-formation histories of, and the
+spatial distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}.
+The most distant cosmological bursts, however, offer direct, hard limits on the
+coalescence time scales.
 
-In this \emph{Letter} we present a spectrum of the host galaxy of the short GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is significantly higher 
-than the previously estimated redshift based on photometric studies. 
-We present the GRB's rest frame 
-properties based on this new distance compared to previous analyses
-\citep{Margutti2012,Sakamoto2013} and revisit the host properties derived from
-the new solution to the SED fit. %While no optical afterglow was detected, the
-% excellent localization from a detection
-%of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also allows
-%us to discuss the positioning and environmental properties of this remarkably
-%distant short GRB.
+In this \emph{Letter} we present a spectrum of the host galaxy of the short
+GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This
+value is significantly higher than the previously estimated redshift based on
+photometric studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest
+frame properties based on this new distance compared to previous analyses and
+revisit the host properties derived from the new solution to the spectral energy distribution (SED) fit.
+%While no optical afterglow was detected, the excellent localization from a
+%detection of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also
+%allows us to discuss the positioning and environmental properties of this
+%remarkably distant short GRB.
 Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology parameters provided by
@@
 	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs4.pdf}
-	\caption{Imaging of the field of GRB~111117A with the X-shooter slit overlaid.
-Only one slit is shown, despite four epochs of spectroscopic observations because of the similarity in position angle.
-The image is the FORS2 $R$-band image for which the photometry is shown in Fig.~\ref{fig:SED}.
-	The blue asterisk indicates the GRB position as derived from the \emph{Chandra} observations in \citet{Sakamoto2013}.
-The insert shows the 2D-image of the \oiii$\lambda$5007 emission line.
-The location of a bright skyline is marked by the blue box.
-The location of the emission line is indicated with the red ellipse.
-Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
+	\caption{
+	Imaging of the field of GRB~111117A with the X-shooter slit overlaid. The slit
+	position represents 4 epochs of spectroscopic observations taken at similar
+	position angles. The image is the FORS2 $R$-band image for which the photometry
+	is shown in Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as
+	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The
+	insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The location
+	of a bright skyline is marked by the blue box. The location of the emission
+	line is indicated with the red ellipse. Because the host is observed in
+	nodding-mode, negative images of the emission line appear on both sides in the
+	spatial direction.
+	}
 	\label{fig:spec_setup}
@@
 %The
-%signal-to-noise of the continuum in the near-infrared arm is too low to use the
+%signal-to-noise ratio of the continuum in the near-infrared arm is too low to use the
 %optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
@@
 We determine a redshift of $z = 2.211$ from the simultaneous detection of
-emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
+emission lines belonging to \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
 \hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
@@
 From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
-constrained to contain very little or no dust, which is why we do not apply 
-a dust-correction to the measured \ha~flux here.
+constrained to contain very little or no dust, although the presence of Lya does
+not exclude dust. Therefore we do not apply a dust-correction to the measured
+\ha~flux here.
 
@@
 observations and measured brightnesses is given in
-Table~\ref{tab:phot_overview}.
+Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, there
+is likely no afterglow contamination on the measured photometry.
 
-All data were reduced, analyzed and fitted in a similar manner as described in detail in
-\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Briefly, we use our own
-\texttt{Python} and IRAF routines to perform a standard reduction which includes
-bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
-stacking of individual images. The photometry of the host was tied against magnitudes 
-of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. 
+All data were reduced, analyzed and fitted in a similar manner as described in
+detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
+Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
+reduction which includes bias/flat-field correction, de-fringing (if necessary),
+sky-subtraction, and stacking of individual images. The photometry of the host
+was calibrated relative to field stars from the SDSS and 2MASS catalogs in the
+case of $grizJHK_{\mathrm{s}}$ filters.
 %For the $R$ and $I$-band photometry, we used the color transformations of
@@
 
-The multi-color spectral energy distribution (SED) is fit by \citet{Bruzual2003}
-single stellar population models based on a \citet{Chabrier2003} initial mass
-function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
-the spectroscopic value of $z=2.211$. The best fit model is an
-unreddened galaxy template, and returns physical parameters of absolute
-magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) =
-9.9\pm0.2$), stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and
-star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9}
-M_\odot\,\mathrm{yr}^{-1}$). We show the SED fit in Fig.~\ref{fig:SED}.
+The multi-color SED is fit by \citet{Bruzual2003} single stellar population
+models based on a \citet{Chabrier2003} with initial mass function in
+\emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the
+spectroscopic value of $z=2.211$. The best fit model is an unreddened galaxy
+template, and returns physical parameters of absolute magnitude
+($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$),
+stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
+($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). We show the SED
+fit in Fig.~\ref{fig:SED}.
 
 Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
-re-analysis of the Gemini data to the $z$-band measurements of \citet{Margutti2012} and
-\citet{Sakamoto2013}. Both of these authors report $z$-band photometry that is brighter 
-by 0.8~mag to 1.0 mag compared to our value, where data taken in bluer filters are in excellent agreement. The large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
-driving the galaxy photometric redshift of the earlier works. Using
-the revised photometry from Table~\ref{tab:phot_overview}, the photometric redshift of the
-galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the
-spectroscopic value at the 1~$\sigma$ confidence level.
+re-analysis of the Gemini data to the $z$-band measurements of
+\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
+$z$-band photometry that is brighter by 0.8~mag to 1.0 mag compared to our
+value, where data taken in bluer filters are in excellent agreement. The large
+$i-z$ color was mistakenly interpreted as a 4000\,\AA\,break driving the galaxy
+photometric redshift of the earlier works. Using the revised photometry from
+Table~\ref{tab:phot_overview}, the photometric redshift of the galaxy is
+$z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic
+value at the 1~$\sigma$ confidence level.
 
@@
 strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not
-change this designation. %The intrinsic luminosity increases, as
-% visible in the X-ray light curve
-%in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
+change this designation.
+%The intrinsic luminosity is higher as demonstrated in the X-ray light curve
+%(Fig. \ref{fig:sxray_lightcurve}), but it is still is sub-luminous compared to
 %the majority of long GRBs.
@@
 per cent), the certainty in class membership for GRB~111117A is much higher.
-This means that GRB~111117A is likely older by 1.8 Gyr than any other
-non-collapsar GRB ever detected.
+This means that GRB~111117A happened when the universe was younger by 1.8 Gyr
+compared to any other non-collapsar GRB ever detected. This number is 3.2 Gyr
+for the next-highest spectroscopic redshift. This has implications for the
+$r$-process element enrichment timescale of the universe.
 
@@
 %	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
-The grey shared region is a compilation of long GRB lightcurves where the color represents density and the light blue is other short GRB lightcurves for which redshifts has been determined.
+The grey shaded region is a compilation of long GRB lightcurves from \todo{reference} where the color represents density and the light blue lines are other short GRB lightcurves from \todo{reference} for which redshifts has been determined.
 Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
@@
 
-For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
-\citep{Lyman2017}, which is a challenge to reconcile  with the relatively large
-offset from the host center derived in \citet{Margutti2012} and
-\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
-burst position that we don’t see. Along with the absence of dust, the large
-offset from the host center indicates that the high $N_H$ arises because the
-density in the GRB surrounding is high, or because the light from the afterglow
-traverses a region of star formation.
+%For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
+%\citep{Lyman2017}, which is a challenge to reconcile  with the relatively large
+%offset from the host center derived in \citet{Margutti2012} and
+%\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
+%burst position that we don’t see. Along with the absence of dust, the large
+%offset from the host center indicates that the high $N_H$ arises because the
+%density in the GRB surrounding is high, or because the light from the afterglow
+%traverses a region of star formation.
 
@@
 
-%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars,
+%From the clear host association, GRB~111117A does not belong to the hostless
+%class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
+%this is indicative of a population of relatively young stars.
 As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
@@
 
-%The detection of \lya~is consistent with the SED-inferred absence of dust, despite
-%the moderate stellar age and the high X-ray derived hydrogen column density
-%which would suggest the opposite. The centroid of the Ly$\alpha$ emission is
-%found to be redshifted by $\sim$~220~km~s$^{-1}$ with respect to systemic, which
-%is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where
-%the outflow is attributed to star formation.
+%The detection of \lya~is consistent with the SED-inferred absence of dust,
+%despite the moderate stellar age which would suggest the opposite. The centroid
+%of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
+%with respect to systemic, which is similar to what is found for long GRB hosts
+%\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.
 
@@
 
-If we use the stellar population age from our SED fit, then we obtain a
+Using the inferred stellar population age from our SED fit, then we obtain a
 (softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
@@
 $n_H-z$ evolution traced by the hosts of lGRBs.The SFR of the host is in the
-upper end of the sGRB host SFR distribution and despite the moderate age and
-high $N_\mathrm{H}$, almost no dust is present. The high $N_H$ is difficult to
-reconcile with the large projected host offset and the absence of dust. One
-possible explanation could be, that GRB~111117A is formed through the prompt
-channel of sGRBs \todo{ref} and originates in a star forming region located in
-the outskirts of the host.
+upper end of the sGRB host SFR distribution and no dust is present. The high
+$N_H$ is difficult to reconcile with the large projected host offset and the
+absence of dust. One possible explanation could be, that GRB~111117A is formed
+through the prompt channel of sGRBs \todo{ref} and originates in a star forming
+region located in the outskirts of the host.
 

### 2017-06-11 GRB111117A.tex
Addresses Johans comments
Closes #14

@@
 spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
-= 2.211$. This makes GRB~111117A the most distant high-confidence short duration GRB
-detected to date. Our spectroscopic redshift supersedes a lower redshift value
-for this burst previously estimated from photometry.
+= 2.211$. This makes GRB~111117A the most distant high-confidence short duration
+GRB detected to date. Our spectroscopic redshift supersedes a lower photometric
+redshift value for this burst previously estimated.
 
@@
 sample is from \citet{Arcodia2016} and the sGRB sample is from
-\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have redshifts, and
-likewise for 5 out of 16 in the short sample. Bursts without redshifts have been
+\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have measured redshifts, and
+likewise for 5 out of 16 in the short sample. Bursts without measured redshifts have been
 excluded for both groups. GRB~111117A occupies a unique position in
@@
 environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
-hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
+hydrogen column density for GRB~111117A seems to follow the trend with increasing $N_H$
 as a function of redshift as found for the lGRB afterglows \citep{Arcodia2016}.
@@
 respective redshift, $L_B/L^{\star}_{B}$. Out of the 39 hosts in the sample, 26
-(66 per cent) have redshifts. To convert the SED-inferred $M_B$ of GRB~111117A
+(66 per cent) have measured redshifts. To convert the SED-inferred $M_B$ of GRB~111117A
 to $L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band magnitude of
@@
 
-In this \emph{Letter}, we have provided a revised, spectroscopic redshift for
-the short GRB~111117A based on emission lines setting it at $z = 2.211$. This
-value supersedes the previous photometric redshift of \citet{Margutti2012} and
-\citet{Sakamoto2013}. %The erroneous redshift estimate of previous authors is
-%attributed to a discrepancy in the measured $z$-band magnitude.
+In this \emph{Letter}, we have provided a revised, spectroscopic redshift
+measurement for the short GRB~111117A based on host galaxy emission lines
+setting it at $z = 2.211$. This value supersedes the previous photometric
+redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. 
+%The erroneous redshift estimate of previous authors is attributed to a
+%discrepancy in the measured $z$-band magnitude.
 

### 2017-06-11 GRB111117A.tex
Addressing Steves comments
Changing comparison sample to D’Avanzo for completeness. Closes #15

@@
 at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
-redshift determination, we find that less than 55 per cent of all sGRB redshifts
+redshift determination, we find that less than 64 per cent of all sGRB redshifts
 should be missed due to host faintness at $z\sim2$. The high redshift of
@@
 	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
-supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b}
+supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
 supports this idea and distinguishes the physical origin of sGRBs from their
@@
 VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
-%Because the spectra are curved across each detector, a
-% rectification algorithm
+%Because the spectra are curved across each detector, a rectification algorithm
 %is employed which introduces correlations between neighboring pixels. We select
@@
 online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
-%The
-%signal-to-noise ratio of the continuum in the near-infrared arm is too low to use the
-%optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
-%carried out with a simple aperture.
+%The signal-to-noise ratio of the continuum in the near-infrared arm is too low
+%to use the optimal extraction algorithm \citep{Horne1986} and therefore the
+%extraction is carried out with a simple aperture.
 An overview of the spectroscopic observations is given in
@@
 host luminosities, normalized by the characteristic galaxy luminosity at their
-respective redshift, $L_B/L^{\star}_{B}$. Out of the 39 hosts in the sample, 26
-(66 per cent) have measured redshifts. To convert the SED-inferred $M_B$ of GRB~111117A
-to $L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band magnitude of
-the Schechter function for blue galaxies ($U - V < 0.25$) in the redshift window
-$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007}. For GRB~111117A, we obtain
-$L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in
-\citet{Berger2014} with measured $L_B/L^{\star}_{B}$.
+respective redshift, $L_B/L^{\star}_{B}$. To convert the SED-inferred $M_B$ of
+GRB~111117A to $L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band
+magnitude of the Schechter function for blue galaxies ($U - V < 0.25$) in the
+redshift window $2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and obtain
+$L_B/L^{\star}_{B} = 1.2$.
+
+Using the complete, flux-limited selection of bursts from \citet{DAvanzo2014a},
+excluding GRB~111117A and the non-sGRB GRB~090426, we have a statistically
+homogeneous sample from which we can address the implications of the redshift of
+GRB~111117A. Out of the 14 hosts in the sample, 10 (71 per cent) have measured
+redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the host of
+GRB~111117A  is brighter than 80 per cent of the hosts with measured
+$L_B/L^{\star}_{B}$. If we conservatively assume that \textit{all} the host
+missing $L_B/L^{\star}_{B}$ are brighter than the host of GRB~111117A, the host
+still brighter than > 60 per cent of sGRB hosts. For all 26 hosts with
+$L_B/L^{\star}_{B}$ from \citet{Berger2014}, the host of GRB~111117A is brighter
+than 73 per cent.
 
 If we assume that we are able to obtain emission-line redshifts from hosts with
-$R < 25$~mag \citep{Kruhler2012}, then we would have missed around 30 per cent
-(8 out of 26 from the sample of \citealt{Berger2014} with measured
-$L_B/L^{\star}_{B}$), if they were at the redshift of GRB~111117A. Because the
-average SFR of galaxies hosting lGRBs is higher than for galaxies hosting sGRBs,
-the fraction of missed burst redshifts is likely higher although the cosmic SFR
+$R < 25$~mag \citep{Kruhler2012}, then we would have missed around 50 per cent
+of the redshifts (5 out of 10 hosts), if they were at the redshift of
+GRB~111117A due to the hosts being too faint. The corresponding number is around
+30 per cent (8 out of 26) from the full sample of \citet{Berger2014}, reflecting
+the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because the average
+SFR of galaxies hosting lGRBs is higher than for galaxies hosting sGRBs, the
+fraction of missed burst redshifts is likely higher although the cosmic SFR
 evolution could play a role in improving redshift determinability.
@@
 be at high redshift, the missed fraction will increase. If we assume that
-\textit{all} the bursts that are missing redshifts in \citet{Berger2014} are at
-high-$z$ and missed due to host faintness, 22 out of 39 hosts (55 per cent)
+\textit{all} the bursts that are missing redshifts are at high-$z$ and missed
+due to host faintness, 9 out of 14 hosts in the complete sample (64 per cent)
 would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
-missed burst at high-$z$ and illustrates that we are likely \textit{not} missing a
-large fraction of sGRBs redshift at $z \approx 2$ due to host faintness.
+missed burst redshifts at high-$z$. The corresponding fraction of missed bursts
+is 0.55 (22 out of 39 hosts) using the full sample from \citet{Berger2014}.
 
@@
 gravitational waves, which in turn is set by the mass of the constituent compact
-objects and the separation of the two \citep{Postnov2014}. If we assume that the
+objects, the eccentricity of the orbit and the separation of the two \citep{Postnov2014}. If we assume that the
 formation timescale of the first galaxies is short compared to the time since
 the Big Bang \citep{Richard2011}, and if we assume a mass of 1.4~$M_\odot$ for
-each of the NSs at the time of system formation, this places a hard upper limit
-on the initial separation, of $a_0 < 3.2~R_\odot$.
+each of the NSs in a circular orbit at the time of system formation, this places
+a hard upper limit on the initial separation, of $a_0 < 3.2~R_\odot$.
 
 In practice most NS-NS binaries will be eccentric at formation because of the SN
-natal kicks, and this can be quite a big effect on merger times, although
-eccentric systems normally circularise first. It could also be a BH-NS system,
-and as the merger time scales roughly as $M^3$ the mass here almost as big an
-effect as the separation ($a^4$).
+natal kicks, although eccentric systems normally circularise first. For more
+eccentric orbits, the coalescence timescale decreases, leading to a decrease in
+the constraint on the initial separation and larger initial separations. As
+noted by \citet{Postnov2014}, it takes eccentricities > 0.6 to significantly
+shorten the merger time. 
+%It could also be a BH-NS system, and as the merger time scales roughly as $M^3$
+%the mass here almost as big an effect as the separation ($a^4$).
 
@@
 redshift distribution of sGRB, the high redshift of GRB~111117A needs to be
-accommodated in progenitor models. A lognormal delay time model predicts a very low volumetric
-density of bursts at $z = 2.211$, whereas a power law delay time model peaks
-near GRB~111117A. If more sGRBs are at similarly high redshifts, but are missed due to
-the faintness of their hosts, a lognormal delay time model will be disfavored.
-Compared to a sample of sGRB hosts, GRB~111117A is more luminous than 70 per
-cent of the sample with measured luminosities. Conservatively, for 55 per
-cent of the sGRB hosts, we would be unable to determine a redshift should they be at 
-a similar redshift as GRB~111117A. This implies that we are \textit{not} missing a large
-fraction of the sGRBs at $z \sim 2$.
+accommodated in progenitor models. A lognormal delay time model predicts a very
+low volumetric density of bursts at $z = 2.211$, whereas a power law delay time
+model peaks near GRB~111117A. If more sGRBs are at similarly high redshifts, but
+are missed due to the faintness of their hosts, a lognormal delay time model
+will be disfavored. Compared to a sample of sGRB hosts, GRB~111117A is more
+luminous than 80 per cent of the sample with measured luminosities.
+Conservatively, for 64 per cent of sGRB hosts, we would be unable to determine a
+redshift should they be at a similar redshift as GRB~111117A.
 

### 2017-06-11 py/Bergersample.py

@@ -10,6 +10,7 @@ import numpy as np
 import matplotlib.pyplot as pl
 import seaborn; seaborn.set_style('ticks')
 from scipy import interpolate
+import pandas as pd
 
 def main():
 
@@ -28,7 +29,19 @@ def main():
     print()
 
     # LB from Berger 2014
-    L_B = np.array([0.1, 1.0, 0.3, np.nan, np.nan, 0.1, 0.1, 0.6, np.nan, 0.1, 1.4, 1.9, 1.2, np.nan, np.nan, np.nan, np.nan, np.nan, 0.3, 1.0, np.nan, 0.5, np.nan, np.nan, 0.6, 0.8, 1.0, 5.0, np.nan, np.nan, 1.6, 0.6, 0.9, 0.4, 1.0, 1.2, 1.0, 0.2, 1.3])
+    burst_table = pd.read_csv("../data/Berger 2014 - Table 2.csv")
+    name, L_B = burst_table["GRB"].values, burst_table["LB"].values
+    #Flux complete from D'Avanzo 2014 #130515A not in Berger sample, 090426c, 111117Ac excluded
+    flux_complete = ["051221A", "060313", "061201", "070714B", "080123", "080503", "080905A", "090426c", "090510", "090515", "100117A", "100625A", "101219A", "111117Ac", "130515A", "130603B"]
+
+    # Filter for names
+    complete_L_B = [ii for ii, kk in enumerate(name) if kk in flux_complete]
+    print(len(complete_L_B))
+    complete = True
+    if complete == True:
+        L_B = L_B[complete_L_B]
+    # print(np.nanmean(L_B))
+    # exit()
     #Exclude nans
     L_B_nonan = L_B[~np.isnan(L_B)]
     # Length of list
@@ -39,10 +52,12 @@ def main():
     lnbrigter = len(L_B[L_B > L_B_star])
     #Redshift completeness
     print("Redshift completeness:")
+    print(lnredshifts, ln)
     print(lnredshifts/ln)
     print()
     #Brighter than:
     print("Brighter than:")
+    print(lnbrigter, lnredshifts)
     print(1 - lnbrigter/lnredshifts)
     print()
     # print(len(L_B[L_B > 1.17]))

### 2017-06-11 py/progenitor_seperation.py

@@ -17,8 +17,8 @@ def calc_sep(age, m1, m2):
     # Convert to seconds
     t0 = (age*1e6 * u.yr).to(u.s)
     # Convert to kg
-    M_ns1 = (1.4 * u.Msun).to(u.kg)
-    M_ns2 = (1.4 * u.Msun).to(u.kg)
+    M_ns1 = (m1 * u.Msun).to(u.kg)
+    M_ns2 = (m2 * u.Msun).to(u.kg)
 
     # Total system mass
     M = M_ns1 + M_ns2

### 2017-06-12 GRB111117A.tex
Address Andrejas comments and drop letter format
Closes #17

@@
 supports this idea and distinguishes the physical origin of sGRBs from their
-long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.
+long-duration counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.
 
@@
 
-%So far, the electromagnetic signal from sGRBs are the only means to accurately
-% localize NS mergers and holds the promise for detection of an associated
-% gravitational wave signal \citep{Ghirlanda2016}. From the burst and following
-% afterglow, modeling the properties and mapping the environments additionally
-% allows insights into the phenomenon itself.
+So far, the electromagnetic signals from sGRBs are the only means to accurately
+localize NS mergers and hold the promise for detection of an associated
+gravitational wave signal \citep{Ghirlanda2016}. From the burst and following
+afterglow, modeling the properties and mapping the environments additionally
+allows insights into the phenomenon itself.
 
@@
 
-In this \emph{Letter} we present a spectrum of the host galaxy of the short
+We here present a spectrum of the host galaxy of the short
 GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This
@@
 %remarkably distant short GRB.
-Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology parameters provided by
+We use the $\Lambda$CDM cosmology parameters provided by
 \citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
@@
 	\centering
-	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs4.pdf}
+	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
 	\caption{
@@
 	is shown in Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as
-	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. The
-	insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The location
-	of a bright skyline is marked by the blue box. The location of the emission
-	line is indicated with the red ellipse. Because the host is observed in
-	nodding-mode, negative images of the emission line appear on both sides in the
-	spatial direction.
+	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. 
+%	The insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The
+%	location of a bright skyline is marked by the blue box. The location of the
+%	emission line is indicated with the red ellipse. Because the host is observed
+%	in nodding-mode, negative images of the emission line appear on both sides in
+%	the spatial direction.
 	}
 	\label{fig:spec_setup}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=9cm]{figures/OIII_img.pdf}
+	\caption{2D-image of the \oiii$\lambda$5007 emission line.
+The location of a bright skyline is marked by the blue box.
+The location of the emission line is indicated with the red ellipse.
+Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
+	\label{fig:line}
 \end{figure}
@@
 VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
-%Because the spectra are curved across each detector, a rectification algorithm
-%is employed which introduces correlations between neighboring pixels. We select
-%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
-%degree of correlation while conserving the maximal resolution.
+Because the spectra are curved across each detector, a rectification algorithm
+is employed which introduces correlations between neighboring pixels. We select
+a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
+degree of correlation while conserving the maximal resolution.
 The observations are combined and extracted using scripts described in Selsing
@@
 online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
-%The signal-to-noise ratio of the continuum in the near-infrared arm is too low
-%to use the optimal extraction algorithm \citep{Horne1986} and therefore the
-%extraction is carried out with a simple aperture.
+The signal-to-noise ratio of the continuum in the near-infrared arm is too low
+to use the optimal extraction algorithm \citep{Horne1986} and therefore the
+extraction is carried out with a simple aperture covering the entire host.
 An overview of the spectroscopic observations is given in
@@
 
+\input{tables/spec_overview.tex}
+
 We determine a redshift of $z = 2.211$ from the simultaneous detection of
@@
 \hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
-in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
-epoch due to the $K$-band blocking filter used for the remainder observations.
+in the insert in Fig.~\ref{fig:line}.  \ha~is only visible in the first
+epoch due to the $K$-band blocking filter used for the remaining observations.
 The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
@@
 
-%The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
-% profiles shows an asymmetric "double-horned" profile, indicating that we are
-% seeing a galaxy with a large degree of coherent rotational motion relative to
-% the line-of-sight. If we assume that we are viewing a spiral galaxy exactly
-% edge-on, this is a measure of the rotational velocity of the gas. If we assume
-% that the spectral resolution and the turbulent width of the lines are
-% negligible compared to the rotational velocity, we can based on the projected
-% size of the source and the width of the lines, put a constraint on the
-% dynamical mass of the galaxy. Based on the physical scale from extent and the
-% velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we
-% are viewing the host at an angle, this value is an upper limit, which should
-% be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.
+The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
+ profiles shows an asymmetric "double-horned" profile, indicating that we are
+ seeing a galaxy with a large degree of coherent rotational motion relative to
+ the line-of-sight. If we assume that we are viewing a spiral galaxy exactly
+ edge-on, this is a measure of the rotational velocity of the gas. If we assume
+ that the spectral resolution and the turbulent width of the lines are
+ negligible compared to the rotational velocity, we can based on the projected
+ size of the source and the width of the lines, put a constraint on the
+ dynamical mass of the galaxy. Based on the physical scale from extent and the
+ velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we
+ are viewing the host at an angle, this value is an upper limit, which should
+ be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.
 
 Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a
-galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
-21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
-are superposed on the line, making an accurate estimate of the \ha-flux difficult. 
-We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
-correcting for the missing flux using the line shape. After converting the 
-\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
-\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
-From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
-constrained to contain very little or no dust, although the presence of Lya does
-not exclude dust. Therefore we do not apply a dust-correction to the measured
-\ha~flux here.
+galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
+at 21\,000~\AA~where the night sky is very bright. In addition, several bright
+sky-lines are superposed on the line, making an accurate estimate of the
+\ha-flux difficult. We obtain a limit on the SFR by integrating the part of
+\ha~free of contamination and correcting for the missing flux using the line
+shape. After converting the \citet{Kennicutt1998} relation to a
+\citet{Chabrier2003} initial mass function using \citet{Madau2014}, we derive a
+limit of $SFR > 7 M_\odot$~yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and
+the detection of \lya, the host is constrained to contain very little or no
+dust, although the presence of Lya does not exclude dust. Therefore we do not
+apply a dust-correction to the measured \ha~flux here. Secondly, based on the
+width of \oiii, an aperture covering the line is integrated over, where
+synthetic sky lines \citep{Noll2012, Jones2013} have been masked and
+interpolated over. From the integrated \ha-line, we estimate SFR = $18 \pm 3$
+M$_\odot$ yr$^{-1}$.
 
-%Secondly, based on the width of \oiii, an aperture covering the line
-% is integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have
-% been masked and interpolated over. From the integrated \ha-line, we estimate
-% SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
-%	\caption{2D-image of the \oiii$\lambda$5007 emission line.
-The location of a bright skyline is marked by the blue box.
-The location of the emission line is indicated with the red ellipse.
-Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
-%	\label{fig:line}
-%\end{figure}
-
-%A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
-% "darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical
-% afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
-% interpolated and evaluated at the time of the non-detection. We find
-% $\beta_{OX} < 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
+A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
+"darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical
+afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
+interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
+< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
 
@@
 case of $grizJHK_{\mathrm{s}}$ filters.
-%For the $R$ and $I$-band photometry, we used the color transformations of
-%Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
+For the $R$ and $I$-band photometry, we used the color transformations of
+Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
 We convert all magnitudes into the AB system, and correct for a Galactic 
 foreground of $E_{B-V}=0.027~\mathrm{mag}$.
+
+\input{tables/phot_overview.tex}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Redshift, line-fluxes, SFR, and Mass}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ \begin{figure*}
+	\centering
+	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
+	\caption{SED fit showning the best-fist SED to the derived photometry.
+The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
+Overplotted in grey is the observed spectrum.
+The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
+	\label{fig:SED}
+\end{figure*}
 
@@
 \citep{Evans2009}. 
-%The X-ray afterglow light curve can be fit with a single power-law decay with an
-%index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
-%analysis using data obtained in photon counting (PC) mode in the widest time epoch where
-%the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
-%(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
-%data) to prevent spectral changes that can affect the X-ray column density
-%determination. 
+The X-ray afterglow light curve can be fit with a single power-law decay with an
+index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
+analysis using data obtained in photon counting (PC) mode in the widest time epoch where
+the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
+(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
+data) to prevent spectral changes that can affect the X-ray column density
+determination. 
 The obtained spectrum is well described by an absorbed power-law
@@
 change this designation.
-%The intrinsic luminosity is higher as demonstrated in the X-ray light curve
-%(Fig. \ref{fig:sxray_lightcurve}), but it is still is sub-luminous compared to
-%the majority of long GRBs.
+The intrinsic luminosity is higher as demonstrated in the X-ray light curve
+(Fig. \ref{fig:sxray_lightcurve}), but it is still is sub-luminous compared to
+the majority of long GRBs.
 \citet{Bromberg2013} investigated the degree to which the long and short
@@
 
-%Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
-%intermediate duration bursts. This comes from both events having very soft
-%spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
-%bursts are very clearly related in their properties to lGRBs
-%\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
-%mergers.
+Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
+intermediate duration bursts. This comes from both events having very soft
+spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
+bursts are very clearly related in their properties to lGRBs
+\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
+mergers.
 
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
-%	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
+\begin{figure}
+	\centering
+	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
+	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
 The grey shaded region is a compilation of long GRB lightcurves from \todo{reference} where the color represents density and the light blue lines are other short GRB lightcurves from \todo{reference} for which redshifts has been determined.
+The thin green line is the light-curve using $z = 1.3$ and the think green line is for $z = 2.211$.
 Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
-%	\label{fig:sxray_lightcurve}
-%\end{figure}
+	\label{fig:sxray_lightcurve}
+\end{figure}
 
@@
 redshifts ($z < 1$) and is found to populate a similar column density
-environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
-hydrogen column density for GRB~111117A seems to follow the trend with increasing $N_H$
-as a function of redshift as found for the lGRB afterglows \citep{Arcodia2016}.
+environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. 
+The inferred hydrogen column density for GRB~111117A seems to follow the trend
+with increasing $N_H$ as a function of redshift as found for the lGRB afterglows
+\citep{Arcodia2016}. This is similar to what is found by \citet{Kopac2012,
+	Margutti2013}, that $N_H$ seems to be comparable for long and short GRBs when
+compared at similar redshifts.
 
-%For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
-%\citep{Lyman2017}, which is a challenge to reconcile with the relatively large
-%offset from the host center derived in \citet{Margutti2012} and
-%\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
-%burst position that we don’t see. Along with the absence of dust, the large
-%offset from the host center indicates that the high $N_H$ arises because the
-%density in the GRB surrounding is high, or because the light from the afterglow
-%traverses a region of star formation.
+For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
+\citep{Lyman2017}, which is a challenge to reconcile with the relatively large
+offset from the host center derived in \citet{Margutti2012} and
+\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
+burst position that we don’t see. Along with the absence of dust, the large
+offset from the host center indicates that the high $N_H$ arises because the
+density in the GRB surrounding is high, or because the light from the afterglow
+traverses a region of star formation.
 
@@
 
-%From the clear host association, GRB~111117A does not belong to the hostless
-%class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
-%this is indicative of a population of relatively young stars.
+From the clear host association, GRB~111117A does not belong to the hostless
+class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
+this is indicative of a population of relatively young stars.
 As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
@@
 
-%The detection of \lya~is consistent with the SED-inferred absence of dust,
-%despite the moderate stellar age which would suggest the opposite. The centroid
-%of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
-%with respect to systemic, which is similar to what is found for long GRB hosts
-%\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.
+The detection of \lya~is consistent with the SED-inferred absence of dust,
+despite the moderate stellar age which would suggest the opposite. The centroid
+of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
+with respect to systemic, which is similar to what is found for long GRB hosts
+\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.
 
@@
 missed burst redshifts at high-$z$. The corresponding fraction of missed bursts
-is 0.55 (22 out of 39 hosts) using the full sample from \citet{Berger2014}.
+is 0.55 (22 out of 39 hosts) using the full sample from \citet{Berger2014}. This
+indicates that we could be missing a significant high-$z$ sGRBs.
 
@@
 eccentric orbits, the coalescence timescale decreases, leading to a decrease in
-the constraint on the initial separation and larger initial separations. As
+the constraint on the initial separation and larger initial separation constraints. As
 noted by \citet{Postnov2014}, it takes eccentricities > 0.6 to significantly
@@
 
-In this \emph{Letter}, we have provided a revised, spectroscopic redshift
-measurement for the short GRB~111117A based on host galaxy emission lines
-setting it at $z = 2.211$. This value supersedes the previous photometric
-redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. 
-%The erroneous redshift estimate of previous authors is attributed to a
-%discrepancy in the measured $z$-band magnitude.
+We have here provided a revised, spectroscopic redshift measurement for the
+short GRB~111117A based on host galaxy emission lines setting it at $z = 2.211$.
+This value supersedes the previous photometric redshift of \citet{Margutti2012}
+and \citet{Sakamoto2013}. %The erroneous redshift estimate of previous authors is
+attributed to a discrepancy in the measured $z$-band magnitude.
 
@@
 will be disfavored. Compared to a sample of sGRB hosts, GRB~111117A is more
-luminous than 80 per cent of the sample with measured luminosities.
-Conservatively, for 64 per cent of sGRB hosts, we would be unable to determine a
-redshift should they be at a similar redshift as GRB~111117A.
+luminous than 80 per cent of a complete sample of sGRB hosts with measured
+luminosities. Conservatively, for 64 per cent of sGRB hosts, we would be unable
+to determine a redshift should they be at a similar redshift as GRB~111117A.
+This could indicate that a significant fraction of sGRBs are at high $z$.
 
@@
 
-\newpage
-%\appendix
-
-\input{tables/spec_overview.tex}
-
-\input{tables/phot_overview.tex}
-
- \begin{figure*}
- 	\centering
- 	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
- 	\caption{SED fit showning the best-fist SED to the derived photometry.
-The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
-Overplotted in grey is the observed spectrum.
-The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
- 	\label{fig:SED}
- \end{figure*}
-
 \end{document}

### 2017-06-12 authorlist.tex
Adding Steve
Address Andrejas comments and drop letter format
Closes #17

@@
 \and A.~J.~Levan \inst{12}
-\and M.~Sparre\inst{1,13}
-\and N.~R.~Tanvir\inst{14}
+\and S.~Schulze \inst{13, 14}
+\and M.~Sparre\inst{1,15}
+\and N.~R.~Tanvir\inst{16}
 \and C.~C.~Th\"one\inst{9}
@@
 %11
-\and University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
+\and Centre for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
 %12
@@
 %13
+\and Instituto de Astrof\'isica, Facultad de F\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+%14
+\and Millennium Institute of Astrophysics, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+%15
 \and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
-%14
+%16
 \and Department of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK

### 2017-06-13 GRB111117A.tex
Address Valerios comments
Closes #2

@@
 
-The host lies in the brighter end of the expected host brightness distribution
+The host lies in the brighter end of the expected sGRB host brightness distribution
 at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
 redshift determination, we find that less than 64 per cent of all sGRB redshifts
-should be missed due to host faintness at $z\sim2$. The high redshift of
+should be missed due to host faintness for hosts at $z\sim2$. The high redshift of
 GRB~111117A is evidence against a lognormal delay-time model for sGRBs through
@@
 
+The \textit{Swift} satellite \citep{Gehrels2004} greatly improved the
+understanding of sGRB progenitors thanks to its quick localization capability.
+The bulk of these localizations have associated galaxies at relatively low
+redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, and because
+most of these measurements come from the associated hosts, it is arguably biased
+towards lower redshifts.
+
 The host galaxies of sGRBs are diverse. They are more massive and less actively
@@
 within their hosts \citep{Fong2013a}, this suggests a progenitor system that can
-be very long-lived, and is associated with stellar mass rather than
-star-formation rate (SFR) \citep{Berger2014}. The median redshift for sGRBs is
-$z\sim0.5$ \citep{Berger2014}, and because most of these measurements come from
-the associated hosts, it is arguably biased towards lower redshifts.
+be very long-lived, and is associated with host stellar mass rather than
+star-formation rate (SFR).
 
-So far, the electromagnetic signals from sGRBs are the only means to accurately
-localize NS mergers and hold the promise for detection of an associated
-gravitational wave signal \citep{Ghirlanda2016}. From the burst and following
-afterglow, modeling the properties and mapping the environments additionally
-allows insights into the phenomenon itself.
+The electromagnetic signals from sGRBs are a promising channel to accurately
+localize NS mergers, which holds the promise for a detection of an associated
+gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
+of the sGRB and the GW will provide new a promising ways to constrain the binary
+inclination angle \citep{Arun2014} and measure cosmological distances
+\citep{Nissanke2010}.
 
@@
 heavy $r$-process elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some
-limits can be calculated based on models of star-formation histories of, and the
-spatial distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}.
-The most distant cosmological bursts, however, offer direct, hard limits on the
+limits can be calculated using host galaxy star-formation history models and
+spatial distribution of sGRBs within their hosts \citep[][]{Berger2014}. The
+most distant cosmological bursts, however, offer direct, hard limits on the
 coalescence time scales.
@@
 	\caption{
-	Imaging of the field of GRB~111117A with the X-shooter slit overlaid. The slit
+	FORS2 R-band imaging of the field of GRB~111117A with the X-shooter slit overlaid. The slit
 	position represents 4 epochs of spectroscopic observations taken at similar
-	position angles. The image is the FORS2 $R$-band image for which the photometry
+	position angles. The corresponding photometry
 	is shown in Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as
@@
 We determine a redshift of $z = 2.211$ from the simultaneous detection of
-emission lines belonging to \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
-\hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
-in the insert in Fig.~\ref{fig:line}.  \ha~is only visible in the first
-epoch due to the $K$-band blocking filter used for the remaining observations.
-The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
-significant velocity structure along the slit. A drop in the continuum to the
-blue of the \lya~line further supports the inferred redshift.
+emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$5007,
+and \ha. \hb~is detected at low significance ($\sim 3 \sigma$). We show
+\oiii$\lambda$5007 in the insert in Fig.~\ref{fig:line}.  \ha~is only visible in
+the first epoch due to the $K$-band blocking filter used for the remaining
+observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and
+show significant velocity structure along the slit. A drop in the continuum
+bluewards of the \lya~line further supports the inferred redshift.
 
-The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
- profiles shows an asymmetric "double-horned" profile, indicating that we are
- seeing a galaxy with a large degree of coherent rotational motion relative to
- the line-of-sight. If we assume that we are viewing a spiral galaxy exactly
- edge-on, this is a measure of the rotational velocity of the gas. If we assume
- that the spectral resolution and the turbulent width of the lines are
- negligible compared to the rotational velocity, we can based on the projected
- size of the source and the width of the lines, put a constraint on the
- dynamical mass of the galaxy. Based on the physical scale from extent and the
- velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we
- are viewing the host at an angle, this value is an upper limit, which should
- be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.
+%The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
+% profiles shows an asymmetric "double-horned" profile, indicating that we are
+% seeing a galaxy with a large degree of coherent rotational motion relative to
+% the line-of-sight. If we assume that we are viewing a spiral galaxy exactly
+% edge-on, this is a measure of the rotational velocity of the gas. If we assume
+% that the spectral resolution and the turbulent width of the lines are
+% negligible compared to the rotational velocity, we can based on the projected
+% size of the source and the width of the lines, put a constraint on the
+% dynamical mass of the galaxy. Based on the physical scale from extent and the
+% velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we
+% are viewing the host at an angle, this value is an upper limit, which should
+% be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.
 
@@
 sky-lines are superposed on the line, making an accurate estimate of the
-\ha-flux difficult. We obtain a limit on the SFR by integrating the part of
+\ha-flux difficult. We obtain a limit on the SFR by numerically integrating the part of
 \ha~free of contamination and correcting for the missing flux using the line
-shape. After converting the \citet{Kennicutt1998} relation to a
+profile. After converting the \citet{Kennicutt1998} relation to a
 \citet{Chabrier2003} initial mass function using \citet{Madau2014}, we derive a
-limit of $SFR > 7 M_\odot$~yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and
-the detection of \lya, the host is constrained to contain very little or no
-dust, although the presence of Lya does not exclude dust. Therefore we do not
-apply a dust-correction to the measured \ha~flux here. Secondly, based on the
+limit of $SFR > 7 M_\odot$~yr$^{-1}$. Additionally, based on the
 width of \oiii, an aperture covering the line is integrated over, where
@@
 interpolated over. From the integrated \ha-line, we estimate SFR = $18 \pm 3$
-M$_\odot$ yr$^{-1}$.
+M$_\odot$ yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and
+the detection of \lya, the host is constrained to contain very little or no
+dust, although the presence of Lya does not exclude dust. Therefore we do not
+apply a dust-correction to the measured \ha~flux here.
 
@@
 
-The multi-color SED is fit by \citet{Bruzual2003} single stellar population
+The multi-color SED is fit using the \citet{Bruzual2003} single stellar population
 models based on a \citet{Chabrier2003} with initial mass function in
@@
 
-As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is
-securely classified as belonging to the short class of GRBs. Because the
-observed classification indicators, $T_{90}$ and hardness ratio, do not depend
-strongly on redshift \citep{Littlejohns2013a}, the updated redshift does not
-change this designation.
+As pointed out by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is
+securely classified as a sGRBs. Because the observed classification indicators,
+$T_{90}$ and hardness ratio, do not depend strongly on redshift
+\citep{Littlejohns2013a}, the updated redshift does not change this designation.
 The intrinsic luminosity is higher as demonstrated in the X-ray light curve
 (Fig. \ref{fig:sxray_lightcurve}), but it is still is sub-luminous compared to
-the majority of long GRBs.
+the majority of long GRBs. The intrinsic spectral lag shortens, further pushing
+the classification in the short regime.
+
 \citet{Bromberg2013} investigated the degree to which the long and short
@@
 sample is from \citet{Arcodia2016} and the sGRB sample is from
-\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have measured redshifts, and
-likewise for 5 out of 16 in the short sample. Bursts without measured redshifts have been
-excluded for both groups. GRB~111117A occupies a unique position in
-Fig.~\ref{fig:NH_z} with the highest $N_H$ of all short bursts. Additionally the
-upper limit at $z = 2.609$ belongs to GRB~090426 and does not belong in the
-short sample. The short sample, excluding GRB~111117A, is located at low
-redshifts ($z < 1$) and is found to populate a similar column density
-environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. 
-The inferred hydrogen column density for GRB~111117A seems to follow the trend
-with increasing $N_H$ as a function of redshift as found for the lGRB afterglows
-\citep{Arcodia2016}. This is similar to what is found by \citet{Kopac2012,
-	Margutti2013}, that $N_H$ seems to be comparable for long and short GRBs when
-compared at similar redshifts.
+\citet{DAvanzo2014a}. Both comparison samples consider the largest temporal
+interval of constant hardness ratio to prevent spectral changes that can affect
+the X-ray derived column density. The 17 (5) of the 99 (16) long (short) GRBs
+which do not have redshift have been excluded from our analysis.
+
+GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
+$N_H$ of all short bursts. Additionally the upper limit at $z = 2.609$ belongs
+to GRB~090426 and does not belong in the short sample. The short sample,
+excluding GRB~111117A, is located at low redshifts ($z < 1$) and is found to
+populate a column density environment similar to that of lGRBs at comparable
+redshifts \citep{DAvanzo2014a}. The inferred hydrogen column density for
+GRB~111117A seems to follow the trend with increasing $N_H$ as a function of
+redshift as found for the lGRB afterglows \citep{Arcodia2016}. This is similar
+to what is found by \citet{Kopac2012, Margutti2013}, that $N_H$ seems to be
+comparable for long and short GRBs when compared at similar redshifts.
 

### 2017-06-14 GRB111117A.tex
Addressing Susannas comments
Closes #11

@@
 
-%The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
-% profiles shows an asymmetric "double-horned" profile, indicating that we are
-% seeing a galaxy with a large degree of coherent rotational motion relative to
-% the line-of-sight. If we assume that we are viewing a spiral galaxy exactly
-% edge-on, this is a measure of the rotational velocity of the gas. If we assume
-% that the spectral resolution and the turbulent width of the lines are
-% negligible compared to the rotational velocity, we can based on the projected
-% size of the source and the width of the lines, put a constraint on the
-% dynamical mass of the galaxy. Based on the physical scale from extent and the
-% velocity with, we infer M$_\text{dyn} \gtrsim 10^{10.8}$ M$_\odot$. Because we
-% are viewing the host at an angle, this value is an upper limit, which should
-% be divided by $\sin(i)$, where $i$ is the inclination relative to edge-on.
+The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
+profiles shows an asymmetric "double-horned" profile, indicating that we are
+seeing a galaxy with a large degree of coherent rotational motion relative to
+the line-of-sight. If we assume that we are viewing a spiral galaxy edge-on,
+this is a measure of the rotational velocity of the gas. If we assume that the
+spectral resolution and the turbulent width of the lines are negligible compared
+to the rotational velocity, we can, based on the projected size of the source
+and the width of the lines, put a constraint on the dynamical mass of the galaxy
+\citep{DeBlok2014}. Based on the physical size along the slit and the velocity
+width of \oiii$\lambda5007$, we infer M$_\text{dyn} \gtrsim 10^{10.8}$
+M$_\odot$. Because we are viewing the host at an angle, this value is an upper
+limit, which should be divided by $\sin(i)$, where $i$ is the inclination
+relative to edge-on.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Classification}
+\subsection{Classification} \label{classification}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 sample is from \citet{Arcodia2016} and the sGRB sample is from
-\citet{DAvanzo2014a}. Both comparison samples consider the largest temporal
-interval of constant hardness ratio to prevent spectral changes that can affect
-the X-ray derived column density. The 17 (5) of the 99 (16) long (short) GRBs
-which do not have redshift have been excluded from our analysis.
+\citet{DAvanzo2014a}. From the sGRB sample of \citet{DAvanzo2014a} we have
+excluded GRB~090426 which does not belong in a short sample as highlighted in
+Sect. \ref{classification}. Both comparison samples consider the largest
+temporal interval of constant hardness ratio to prevent spectral changes that
+can affect the X-ray derived column density. The 17 (5) of the 99 (15) long
+(short) GRBs which do not have redshift have been excluded from our analysis.
 
 GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
-$N_H$ of all short bursts. Additionally the upper limit at $z = 2.609$ belongs
-to GRB~090426 and does not belong in the short sample. The short sample,
-excluding GRB~111117A, is located at low redshifts ($z < 1$) and is found to
-populate a column density environment similar to that of lGRBs at comparable
-redshifts \citep{DAvanzo2014a}. The inferred hydrogen column density for
-GRB~111117A seems to follow the trend with increasing $N_H$ as a function of
-redshift as found for the lGRB afterglows \citep{Arcodia2016}. This is similar
-to what is found by \citet{Kopac2012, Margutti2013}, that $N_H$ seems to be
-comparable for long and short GRBs when compared at similar redshifts.
+$N_H$ and highest redshift of all sGRBs. The short sample, excluding
+GRB~111117A, is located at low redshifts ($z < 1$) and is found to populate a
+column density environment similar to that of lGRBs at comparable redshifts
+\citep{DAvanzo2014a}. The inferred hydrogen column density for GRB~111117A seems
+to follow the trend with increasing $N_H$ as a function of redshift as found for
+the lGRB afterglows \citep{Arcodia2016}. This is similar to what is found by
+\citet{Kopac2012, Margutti2013}, that $N_H$ seems to be comparable for long and
+short GRBs when compared at similar redshifts.
 
@@
 % 	\caption{Isotropic equivalent energy for all the representative samples of both long and short GRBs. The sample is from \citet{Turpin2015} and the short is from \citet{DAvanzo2014a}.}
-% 	\label{fig:NH_z}
+% 	\label{fig:Eiso_z}
 % \end{figure}
@@
 class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
-this is indicative of a population of relatively young stars.
-As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
-late-type galaxy and is entirely consistent in terms of stellar mass and stellar
-age with the general population of sGRB hosts ($\left\langle M _*
-\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
-0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
-and sSFR are entirely within the range expected for the hosts of sGRBs
-\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
-typical SFR for sGRB host galaxies \citep{Berger2014} and more similar to
-the SFR found in the hosts of lGRBs at a corresponding redshift
-\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
-star forming galaxy would exhibit weaker emission lines, thus making the
-redshift harder to determine. Additionally, it is natural to expect some
-evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.
+this is indicative of a population of relatively young stars. As the majority of
+sGRBs \citep{Fong2013b}, the host of GRB~111117A is a late-type galaxy and is
+entirely consistent in terms of stellar mass and stellar age with the general
+population of sGRB hosts ($\left\langle M _* \right\rangle = 10^{10.1}
+M_{\odot}$ and $\left\langle \tau _* \right\rangle = 0.3 $~Gyr)
+\citep{Leibler2010}. Being a late-type host, both the stellar mass and sSFR are
+entirely within the range expected for the hosts of sGRBs \citep{Behroozi2014}.
+
+The SFR is $\sim$1 order of magnitude higher than the typical SFR for sGRB host
+galaxies \citep{Berger2014} and more similar to the SFR found in the hosts of
+lGRBs at a corresponding redshift \citep{Kruhler2015}. Only two hosts in the
+sample of short GRBs compiled in \citet{Berger2014} have a more vigorous star
+formation, meaning that it is in the very upper end of the star formation
+distribution. The cosmic SFR evolution of the universe likely plays a role due
+to the proximity of GRB~111117A to the peak of cosmic SFR \citep{Madau2014}.
+The high SFR is partly a selection effect, because a less star forming galaxy
+would exhibit weaker emission lines, thus making the redshift harder to
+determine. Additionally, it is natural to expect some evolution in the hosts of
+sGRBs with redshift as illustrated for $N_H$ in Sect.~\ref{restnH}.
 
@@
 \begin{acknowledgements}
+
 TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady. 
@@
 ZC acknowledges support from the Spanish research project AYA 2014-58381-P and support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2014-21669. 
+RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural Creators
 

### 2017-06-14 py/Eiso_z.py

@@ -12,11 +12,11 @@ from scipy import stats
 import matplotlib as mpl
 
 params = {
-   'axes.labelsize': 10,
-   'font.size': 10,
-   'legend.fontsize': 10,
-   'xtick.labelsize': 10,
-   'ytick.labelsize': 10,
+   'axes.labelsize': 16,
+   'font.size': 12,
+   'legend.fontsize': 12,
+   'xtick.labelsize': 16,
+   'ytick.labelsize': 16,
    'text.usetex': False,
    'figure.figsize': [7.281, 4.5]
    }
@@ -77,6 +77,10 @@ def main():
     burst_table = pd.read_csv("../data/Comparison sample - Short.csv")
     name, z_short, Eiso_short = burst_table["GRB"].values, burst_table["z"].values, burst_table["Eiso"].values
 
+    # Remove GRB090426
+    idx = np.where(name == "090426")[0]
+    name, z_short, Eiso_short = np.delete(name, idx), np.delete(z_short, idx), np.delete(Eiso_short, idx)
+
     Eiso_short = np.log10(filt_nan(Eiso_short, fill_value=np.nan)*1e51)
 
 
@@ -122,7 +126,7 @@ def main():
 
 
     # Save figure for tex
-    pl.legend()
+    pl.legend(loc=4)
     pl.savefig("../figures/Eiso_z.pdf", dpi="figure")
     pl.show()
 

### 2017-06-14 py/NH_z.py

@@ -87,6 +87,10 @@ def main():
     burst_table = pd.read_csv("../data/Comparison sample - Short.csv")
     name, z_short, NH_short = burst_table["GRB"].values, burst_table["z"].values, burst_table["NH(z)"].values
 
+    # Remove GRB090426
+    idx = np.where(name == "090426")[0]
+    name, z_short, NH_short = np.delete(name, idx), np.delete(z_short, idx), np.delete(NH_short, idx)
+
     NH_short_dec = np.log10(1e21*filt_nan(NH_short, fill_value=np.nan))
     NH_short_lim = np.log10(1e21*np.array([float(ii[1:]) for ii in NH_short[np.isnan(NH_short_dec)]]))
 
@@ -133,7 +137,6 @@ def main():
     color = sns.color_palette()[0]
     g.x = z_short_dec
     g.y = NH_short_dec
-    print(g.y)
     g = g.plot_marginals(sns.distplot, hist=False, rug=True, kde=False, color=color, rug_kws={"height": 0.7, "lw": 2.0})
     g = g.plot_joint(pl.scatter, color=color, label="D'Avanzo et al. 2014")
     g.x = z_short_lim
@@ -145,17 +148,9 @@ def main():
     g.y = 1
     g = g.plot_joint(pl.plot, color=sns.color_palette()[2], label="Arcodia et al. 2016")
 
-
-
-
-
     ax = pl.gca()
-    # ax.axhline(0.5, color="black", linestyle="dashed", alpha=0.5)
-    # ax.annotate(r"$\beta_{OX} = 0.5$", (19.6, 0.45))
     g.set_axis_labels(r"z", r"log(N$_H$) [cm$^{-2}$]")
     pl.tight_layout()
-
-
 
     # Save figure for tex
     pl.legend(loc=1)

### 2017-06-20 GRB111117A.tex
Address Klaas' comments
Closes #19

@@
 
+The classification of GRBs in two groups, initially comes from the bimodal
+distribution of burst duration and spectral hardness \citet{Kouveliotou1993},
+where the duration T$_{90} < 2$ has been the standard dividing line between long
+and short GRBs. Additionally, it has been found for lGRBs that there is a
+spectral lag in the arrival-time of photons, with the most energetic ones
+arriving first, where this lag is consistent with zero for sGRBs
+\citep{Norris2006}. Because both populations have continuous, overlapping
+distributions in their observables and because telescopes observe in differing
+bands, it is difficult to impose a single demarcation criterion between the two
+classes. For this reason, the distinction between long and short GRBs is
+preferably based on a combination of high-energy properties \citep{Zhang2009,
+	Bromberg2012a, Bromberg2013}.
+
 The \textit{Swift} satellite \citep{Gehrels2004} greatly improved the
@@
 within their hosts \citep{Fong2013a}, this suggests a progenitor system that can
-be very long-lived, and is associated with host stellar mass rather than
+be very long lived in comparison to lGRBs, and is tracing with host stellar mass rather than
 star-formation rate (SFR).
@@
 ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
-091.D-0904 (PI: Hjorth). X-shooter covers the wavelength range from 3000~\AA~to
+091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
+arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA~to
 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
@@
 tracing, wavelength calibration, rectification, and flux calibration using the
-VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
-Because the spectra are curved across each detector, a rectification algorithm
-is employed which introduces correlations between neighboring pixels. We select
-a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
-degree of correlation while conserving the maximal resolution.
-The observations are combined and extracted using scripts described in Selsing
-et al. 2017 (in prep.) and available
-online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
-The signal-to-noise ratio of the continuum in the near-infrared arm is too low
-to use the optimal extraction algorithm \citep{Horne1986} and therefore the
-extraction is carried out with a simple aperture covering the entire host.
-An overview of the spectroscopic observations is given in
+VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010} run in physical
+mode. Because the echelle orders are curved across each detector, a rectification
+algorithm is employed which introduces correlations between neighboring pixels.
+We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to
+minimize the degree of correlation while conserving the maximal resolution. The
+observations are combined and extracted using scripts described in Selsing et
+al. 2017 (in prep.) and available
+online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
+signal-to-noise ratio of the continuum in the near-infrared arm is too low to
+use the optimal extraction algorithm \citep{Horne1986} and therefore the
+extraction is carried out with a simple aperture covering the entire host. An
+overview of the spectroscopic observations is given in
 Table~\ref{tab:spec_overview}, and the slit position is shown in
@@
 
+Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a
+galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
+at 21\,000~\AA~where the night sky is very bright. In addition, several bright
+sky-lines are superposed on the line, making an accurate estimate of the
+\ha-flux difficult. We obtain a limit on the SFR by numerically integrating the part of
+\ha~free of contamination and correcting for the missing flux using the line
+profile and obtain $F_{H \alpha} > 4.1 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting the \citet{Kennicutt1998} relation to a
+\citet{Chabrier2003} initial mass function using \citet{Madau2014}, we derive a
+limit of $SFR > 7 M_\odot$~yr$^{-1}$. Additionally, based on the
+width of \oiii, an aperture covering the line is integrated over, where
+synthetic sky lines \citep{Noll2012, Jones2013} have been masked and
+interpolated over. From the integrated \ha-line, we estimate SFR = $18 \pm 3$
+M$_\odot$ yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and
+the detection of \lya, the host is constrained to contain very little or no
+dust, although the presence of Lya does not exclude dust. Therefore we do not
+apply a dust-correction to the measured \ha~flux here.
+
 The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
@@
 width of \oiii$\lambda5007$, we infer M$_\text{dyn} \gtrsim 10^{10.8}$
-M$_\odot$. Because we are viewing the host at an angle, this value is an upper
-limit, which should be divided by $\sin(i)$, where $i$ is the inclination
-relative to edge-on.
-
-Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a
-galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
-at 21\,000~\AA~where the night sky is very bright. In addition, several bright
-sky-lines are superposed on the line, making an accurate estimate of the
-\ha-flux difficult. We obtain a limit on the SFR by numerically integrating the part of
-\ha~free of contamination and correcting for the missing flux using the line
-profile. After converting the \citet{Kennicutt1998} relation to a
-\citet{Chabrier2003} initial mass function using \citet{Madau2014}, we derive a
-limit of $SFR > 7 M_\odot$~yr$^{-1}$. Additionally, based on the
-width of \oiii, an aperture covering the line is integrated over, where
-synthetic sky lines \citep{Noll2012, Jones2013} have been masked and
-interpolated over. From the integrated \ha-line, we estimate SFR = $18 \pm 3$
-M$_\odot$ yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and
-the detection of \lya, the host is constrained to contain very little or no
-dust, although the presence of Lya does not exclude dust. Therefore we do not
-apply a dust-correction to the measured \ha~flux here.
+M$_\odot$. Because we are viewing the host inclined at an angle relative to
+edge-on and because the slit is not aligned along the long axis of the host,
+this value is an upper limit.
 
@@
 We convert all magnitudes into the AB system, and correct for a Galactic 
-foreground of $E_{B-V}=0.027~\mathrm{mag}$.
+foreground of $E_{B-V}=0.027~\mathrm{mag}$ \citep{Schlegel1998, Schlafly2011}.
 
@@
 data) to prevent spectral changes that can affect the X-ray column density
-determination. 
+determination \citep{Kopac2012}. 
 The obtained spectrum is well described by an absorbed power-law
@@
 to follow the trend with increasing $N_H$ as a function of redshift as found for
-the lGRB afterglows \citep{Arcodia2016}. This is similar to what is found by
+the lGRB afterglows \citep{Campana2010, Starling2013, Arcodia2016}. This is similar to what is found by
 \citet{Kopac2012, Margutti2013}, that $N_H$ seems to be comparable for long and
@@
 This value supersedes the previous photometric redshift of \citet{Margutti2012}
-and \citet{Sakamoto2013}. %The erroneous redshift estimate of previous authors is
-attributed to a discrepancy in the measured $z$-band magnitude.
+and \citet{Sakamoto2013}. The erroneous redshift estimate of previous authors is
+attributed to a discrepancy in the measured $z$-band magnitude, and highlights
+the importance of deep spectroscopic studies of sGRB hosts at medium resolution.
 

### 2017-06-21 GRB111117A.tex
Address Jans comments
Closes #21

@@
 to follow the trend with increasing $N_H$ as a function of redshift as found for
-the lGRB afterglows \citep{Campana2010, Starling2013, Arcodia2016}. This is similar to what is found by
-\citet{Kopac2012, Margutti2013}, that $N_H$ seems to be comparable for long and
-short GRBs when compared at similar redshifts.
+the lGRB afterglows \citep{Campana2010, Starling2013, Arcodia2016}. This is
+related to what is found by \citet{Kopac2012, Margutti2013}, that $N_H$ seems to
+be comparable for long and short GRBs when compared at similar redshifts.
 
-For lGRBs, $N_H$ correlates with the surface luminosity at the explosion site
-\citep{Lyman2017}, which is a challenge to reconcile with the relatively large
-offset from the host center derived in \citet{Margutti2012} and
-\citet{Sakamoto2013}. This could suggest that there is some galaxy under the
-burst position that we don’t see. Along with the absence of dust, the large
-offset from the host center indicates that the high $N_H$ arises because the
-density in the GRB surrounding is high, or because the light from the afterglow
-traverses a region of star formation.
+The redshift evolution of $N_H$ in the hosts of lGRBs is not reproduced by
+\citet{Buchner2017}, using a different $N_H$ inference methodology. Instead a
+correlation between $N_H$ and host stellar mass is suggested. Assuming that the
+different $N_H$-fitting methodologies yield comparable results, GRB~111117A is
+an outlier from the relation suggested by \citet{Buchner2017} by more than the
+intrinsic scatter, although some lGRB hosts populate a similar region in the
+$N_H$-$M_\star$ relation. Additionally, for lGRBs, $N_H$ correlates with the
+surface luminosity at the explosion site \citep{Lyman2017}. The large offset of
+GRB~111117A relative to the host center derived in \citet{Margutti2012} and
+\citet{Sakamoto2013} is difficult to reconcile with galaxy-scale gas providing
+the X-ray absorption. Along with the absence of dust, the large offset from the
+host center indicates that the high $N_H$ arises because the density in the GRB
+surrounding is high, or because the light from the afterglow traverses a region
+of star formation.
 

### 2017-06-26 GRB111117A.tex
Small edit to include limit on missed fraction

@@
 their hosts to measure their redshifts. These measurements, however, are
-critical to constrain sGRB progenitors and delay time models. Here, we present
-spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
-= 2.211$. This makes GRB~111117A the most distant high-confidence short duration
-GRB detected to date. Our spectroscopic redshift supersedes a lower photometric
-redshift value for this burst previously estimated.
+critical to constrain the nature of sGRB progenitors, the redshift distribution
+of these events and the $r$-process element enrichment history of
+the universe. Here, we present spectroscopy of the host galaxy of GRB~111117A
+and measure its redshift to be $z = 2.211$. This makes GRB~111117A the most
+distant high-confidence short duration GRB detected to date. Our spectroscopic
+redshift supersedes a lower, previously estimated photometric redshift value for
+this burst.
 
 We use the spectroscopic redshift, as well as new imaging data to constrain the
-nature of the host galaxy and the physical parameters of the GRB. The rest-frame X-ray derived hydrogen column density, for example, is high compared to a
+nature of the host galaxy and the physical parameters of the GRB. The rest-frame
+X-ray derived hydrogen column density, for example, is the highest compared to a
 complete sample of sGRBs and seems to follow the evolution with redshift as
-traced by the hosts of long GRBs (lGRBs). This is consistent with a proportion
-of the sGRB population originating in late-type galaxies, living in similar
-environments to that of lGRB hosts.
+traced by the hosts of long GRBs (lGRBs).
 
@@
 at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
-redshift determination, we find that less than 64 per cent of all sGRB redshifts
+redshift determination, we find that between 43 and 71 per cent of all sGRB redshifts
 should be missed due to host faintness for hosts at $z\sim2$. The high redshift of
@@
 star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
-progenitor is a circular pair of inspiralling NSs. This constraint excludes some of the
-longest formation channels for this burst.
-
+progenitor system is a circular pair of inspiralling NSs. This constraint
+excludes some of the longest sGRB formation channels for this burst.
 }
@@
 distribution of burst duration and spectral hardness \citet{Kouveliotou1993},
-where the duration T$_{90} < 2$ has been the standard dividing line between long
+where the duration T$_{90} < 2$ has been regarded as the dividing line between long
 and short GRBs. Additionally, it has been found for lGRBs that there is a
 spectral lag in the arrival-time of photons, with the most energetic ones
-arriving first, where this lag is consistent with zero for sGRBs
+arriving first. This lag is consistent with zero for sGRBs
 \citep{Norris2006}. Because both populations have continuous, overlapping
@@
 revisit the host properties derived from the new solution to the spectral energy distribution (SED) fit.
-%While no optical afterglow was detected, the excellent localization from a
-%detection of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also
-%allows us to discuss the positioning and environmental properties of this
-%remarkably distant short GRB.
+While no optical afterglow was detected, the excellent localization from a
+detection of the X-ray afterglow by the \emph{Chandra X-ray Observatory}
+allows us to discuss the positioning and environmental properties of this
+remarkably distant sGRB.
 We use the $\Lambda$CDM cosmology parameters provided by
@@
 	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. 
-%	The insert shows the 2D-image of the \oiii$\lambda$5007 emission line. The
-%	location of a bright skyline is marked by the blue box. The location of the
-%	emission line is indicated with the red ellipse. Because the host is observed
-%	in nodding-mode, negative images of the emission line appear on both sides in
-%	the spatial direction.
 	}
@@
 sky-lines are superposed on the line, making an accurate estimate of the
-\ha-flux difficult. We obtain a limit on the SFR by numerically integrating the part of
-\ha~free of contamination and correcting for the missing flux using the line
-profile and obtain $F_{H \alpha} > 4.1 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting the \citet{Kennicutt1998} relation to a
-\citet{Chabrier2003} initial mass function using \citet{Madau2014}, we derive a
-limit of $SFR > 7 M_\odot$~yr$^{-1}$. Additionally, based on the
-width of \oiii, an aperture covering the line is integrated over, where
-synthetic sky lines \citep{Noll2012, Jones2013} have been masked and
-interpolated over. From the integrated \ha-line, we estimate SFR = $18 \pm 3$
-M$_\odot$ yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and
-the detection of \lya, the host is constrained to contain very little or no
-dust, although the presence of Lya does not exclude dust. Therefore we do not
-apply a dust-correction to the measured \ha~flux here.
+\ha-flux difficult. We obtain a limit on the SFR by numerically integrating the
+part of \ha~free of contamination and correcting for the missing flux using the
+line profile and obtain $F_{H \alpha} > 4.1 \times 10^{-17} \mathrm{erg}
+\mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting the \citet{Kennicutt1998}
+relation to a \citet{Chabrier2003} initial mass function using
+\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$.
+Additionally, based on the width of \oiii, an aperture covering the line is
+integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have been
+masked and interpolated over. From the integrated \ha-line, we estimate SFR =
+$18 \pm 3$ M$_\odot$ yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and the
+detection of \lya, the host is constrained to contain very little or no dust,
+although the presence of Lya does not exclude dust. Therefore we do not apply a
+dust-correction to the measured \ha~flux here. \oii is close to a region of
+strong telluric absorption, which is why no SFR is inferred from this line.
 
@@
 A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
-"darkness"-parameter $\beta_{OX}$\citep{Jakobsson2004}. Using the optical
+"darkness"-parameter $\beta_{OX} $\citep{Jakobsson2004}. Using the optical
 afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
@@
 \citep{Littlejohns2013a}, the updated redshift does not change this designation.
-The intrinsic luminosity is higher as demonstrated in the X-ray light curve
-(Fig. \ref{fig:sxray_lightcurve}), but it is still is sub-luminous compared to
-the majority of long GRBs. The intrinsic spectral lag shortens, further pushing
-the classification in the short regime.
+The intrinsic spectral lag shortens, but since it is already consistent with
+zero, this does not affect the classification.
+
+The intrinsic luminosity is shown in the X-ray light curve (Fig.
+\ref{fig:sxray_lightcurve}) and it is sub-luminous compared to the majority of
+long GRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
+distribution at 10 ks. The sub-samples comprise of 333 long, 19 short GRBs, and
+GRB~111117A. The mean and the 1-$\sigma$ dispersions of the samples are
+L$_{lGRB} = 46.59 \pm 0.87$ and L$_{sGRB} = 44.96 \pm 0.94$. GRB~111117A had a
+luminosity of 44.95 at 10 ks which is very close to the peak of the sGRB
+luminosity distribution at 10 ks, but an outlier from the lGRG distribution,
+further supporting the short classification.
 
@@
 per cent), the certainty in class membership for GRB~111117A is much higher.
-This means that GRB~111117A happened when the universe was younger by 1.8 Gyr
-compared to any other non-collapsar GRB ever detected. This number is 3.2 Gyr
-for the next-highest spectroscopic redshift. This has implications for the
-$r$-process element enrichment timescale of the universe.
 
@@
 
+The redshift and secure classification of GRB~111117A means that it occurred
+when the universe was younger by 1.8 Gyr compared to any other non-collapsar GRB
+ever detected. This number is 3.2 Gyr for the next-highest spectroscopic
+redshift. If the merger of NSs is the primary agent for the $r$-process element
+enrichment of the universe \citep{Goriely2011, Ji2016, Komiya2016}, GRB~111117A
+marks the earliest detection of this process.
+
 \begin{figure}
@@
 	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
-The grey shaded region is a compilation of long GRB lightcurves from \todo{reference} where the color represents density and the light blue lines are other short GRB lightcurves from \todo{reference} for which redshifts has been determined.
-The thin green line is the light-curve using $z = 1.3$ and the think green line is for $z = 2.211$.
-Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and still subluminous compared to the lGRB population.}
+The grey shaded region is a compilation of long GRB lightcurves \citep{Evans2007, Evans2009} where the color represents density.
+The light blue lines are sGRB lightcurves from bursts with duration of $T_{90} \lesssim 2$ s and those that were classified as short in \citet{Kann2011, Berger2014, DAvanzo2014a}.
+The thick green line is GRB~111117A.
+Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and subluminous compared to the lGRB population.}
 	\label{fig:sxray_lightcurve}
@@
 surrounding is high, or because the light from the afterglow traverses a region
-of star formation.
+of dense gas.
 
@@
 	\includegraphics[width=9cm]{figures/NH_z.pdf}
-	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations.
-The detections are replaced with contours for clarity and the limits are shown with arrows.}
+	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations of GRBs.
+The detections are replaced with contours and the limits are shown with arrows.}
 	\label{fig:NH_z}
@@
 If we assume that we are able to obtain emission-line redshifts from hosts with
-$R < 25$~mag \citep{Kruhler2012}, then we would have missed around 50 per cent
-of the redshifts (5 out of 10 hosts), if they were at the redshift of
+$R < 24.5$~mag \citep{Kruhler2012}, then we would have missed 60 per cent
+of the redshifts (6 out of 10 hosts), if they were at the redshift of
 GRB~111117A due to the hosts being too faint. The corresponding number is around
-30 per cent (8 out of 26) from the full sample of \citet{Berger2014}, reflecting
+45 per cent (12 out of 26) from the full sample of \citet{Berger2014}, reflecting
 the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because the average
@@
 \textit{all} the bursts that are missing redshifts are at high-$z$ and missed
-due to host faintness, 9 out of 14 hosts in the complete sample (64 per cent)
+due to host faintness, 10 out of 14 hosts in the complete sample (71 per cent)
 would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
-missed burst redshifts at high-$z$. The corresponding fraction of missed bursts
-is 0.55 (22 out of 39 hosts) using the full sample from \citet{Berger2014}. This
-indicates that we could be missing a significant high-$z$ sGRBs.
+missed burst redshifts at high-$z$. Conversely, if all bursts missing redshift
+are at low redshift and missed for other reasons, 6 out of 14 hosts (43 per
+cent) would be missed at $z = 2.211$. The two limits indicate that we are
+missing between 43 and 71 per cent of sGRB hosts at $z sim 2$ due to host
+faintness. For the 13 hosts missing redshift in \citet{Berger2014}, this means
+that between 6 and 9 of these should be at high redshift, but missed due to host
+faintness.
 
@@
 $z = 0.9$. According to \citet{Wanderman2015}, this preference depends
-critically on the absence of non-collapsar sGRBs at $z \gtrsim 1.2$. The
-redshift of GRB~111117A, on the other hand, is close to the expected peak in
-sGRB rate calculated using the power law delay time models \citep{Behroozi2014,
-	Wanderman2015}.
+critically on the absence of non-collapsar sGRBs at $z \gtrsim 1.2$. The higher
+determined redshift of GRB~111117A, and the likely number of additional high-z
+sGRB could change the preferred time delay models. The redshift of GRB~111117A,
+on the other hand, is close to the expected peak in sGRB rate calculated using
+the power law delay time models \citep{Behroozi2014, Wanderman2015}.
 
@@
 In practice most NS-NS binaries will be eccentric at formation because of the SN
-natal kicks, although eccentric systems normally circularise first. For more
-eccentric orbits, the coalescence timescale decreases, leading to a decrease in
-the constraint on the initial separation and larger initial separation constraints. As
-noted by \citet{Postnov2014}, it takes eccentricities > 0.6 to significantly
-shorten the merger time. 
-%It could also be a BH-NS system, and as the merger time scales roughly as $M^3$
-%the mass here almost as big an effect as the separation ($a^4$).
+natal kicks. For more eccentric orbits, the coalescence timescale decreases,
+leading to a decrease in the constraint on the initial separation and larger
+initial separation constraints. As noted by \citet{Postnov2014}, it takes
+eccentricities > 0.6 to significantly shorten the merger time.
 
@@
 highest within a complete sample of sGRB hosts and is consistent with the
-$n_H-z$ evolution traced by the hosts of lGRBs.The SFR of the host is in the
+$n_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
 upper end of the sGRB host SFR distribution and no dust is present. The high
@@
 absence of dust. One possible explanation could be, that GRB~111117A is formed
-through the prompt channel of sGRBs \todo{ref} and originates in a star forming
+through a prompt channel of sGRB formation and originates in a star forming
 region located in the outskirts of the host.
@@
 luminous than 80 per cent of a complete sample of sGRB hosts with measured
-luminosities. Conservatively, for 64 per cent of sGRB hosts, we would be unable
+luminosities. For between 43 and 71 per cent of sGRB hosts, we would be unable
 to determine a redshift should they be at a similar redshift as GRB~111117A.
-This could indicate that a significant fraction of sGRBs are at high $z$.
+This could indicate that potentially, a significant fraction of sGRBs are at
+high $z$ but missed due to host faintness.
 

### 2017-06-26 tables/phot_overview.tex
Small edit to include limit on missed fraction

@@
 \tablefoot{
-\tablefoottext{a}{All magnitudes are given in the AB system and are not corrected for the expected Galactic foreground extinction corresponding to a reddening of $E_{B-V}=0.027$\,mag.
-\todo{Thomas, Can you help with the reference for the data which are not ours?}}}
+\tablefoottext{a}{All magnitudes are given in the AB system and are not corrected for the expected Galactic foreground extinction corresponding to a reddening of $E_{B-V}=0.027$\,mag.}}
 \end{table*}

### 2017-06-27 GRB111117A.tex
Fixing affil, adding ack

@@
 
-\keywords{Gamma-ray burst: individual: GRB~111117A --- }
+\keywords{Gamma-ray burst: individual: GRB~111117A -- Gamma-ray burst: general -- Galaxies: high-redshift -- binaries: general -- X-rays: bursts -- Techniques: imaging spectroscopy}
 
@@
 luminous than 80 per cent of a complete sample of sGRB hosts with measured
-luminosities. For between 43 and 71 per cent of sGRB hosts, we would be unable
-to determine a redshift should they be at a similar redshift as GRB~111117A.
-This could indicate that potentially, a significant fraction of sGRBs are at
-high $z$ but missed due to host faintness.
+luminosities. Assuming a host brightness redshift determination threshold, for
+between 43 and 71 per cent of sGRB hosts, we would be unable to determine a
+redshift should they be at a similar redshift as GRB~111117A. This could
+indicate that potentially, a significant fraction of sGRBs are at high $z$ but
+missed due to host faintness.
 
@@
 TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
-CT acknowledges support from a Spanish National Research Grant of Excellence under project AYA 2014-58381-P and funding associated to a Ramón y Cajál fellowship under grant number RyC-2012-09984.
-AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation Grant for Researchers and Cultural Creators,and the Spanish Ministry of Economy and Competitiveness through project AYA2014-58381-P.
-Partly based on observations made with the Gran Telescopio Canarias (GTC).
-ZC acknowledges support from the Spanish research project AYA 2014-58381-P and support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2014-21669. 
-RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural Creators
+% 
+JJ acknowledges support from NOVA and NWO-FAPESP grant for advanced
+instrumentation in astronomy.
+%
+CT acknowledges support from a Spanish National Research Grant of Excellence
+under project AYA 2014-58381-P and funding associated to a Ramón y Cajál
+fellowship under grant number RyC-2012-09984.
+%
+AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation
+Grant for Researchers and Cultural Creators,and the Spanish Ministry of Economy
+and Competitiveness through project AYA2014-58381-P. Partly based on
+observations made with the Gran Telescopio Canarias (GTC).
+%
+ZC acknowledges support from the Spanish research project AYA 2014-58381-P and
+support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2014-21669.
+%
+RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural
+Creators and support from ASI (Italian Space Agency) through the Contract n. 2015-046-R.0 and from European Union Horizon 2020 Programme under the AHEAD project (grant agreement n. 654215).
+%
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{TheAstropyCollaboration2013}.
+The analysis and plotting has been achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy and Scipy \citep{VanderWalt2011} along with other community-developed packages.
 

### 2017-06-27 authorlist.tex
Addressing Susannas comments
Closes #11
Small edit to include limit on missed fraction
Fix affil
Fixing affil, adding ack

@@
 \and P.~D'Avanzo\inst{4}
-\and J.~Palmerio\inst{5}
-\and S.~D.~Vergani\inst{6}
-\and J.~Japelj\inst{7}
+\and S.~Schulze \inst{5, 6, 7}
+\and J.~Palmerio\inst{8}
+\and S.~D.~Vergani\inst{9}
+\and J.~Japelj\inst{10}
 \and B.~Milvang-Jensen\inst{1}
 \and D.~Watson\inst{1}
-\and P.~Jakobsson\inst{8}
+\and P.~Jakobsson\inst{11}
 %
-\and Z.~Cano\inst{9}
+\and J.~Bolmer \inst{2}
+\and Z.~Cano\inst{12}
+\and S.~Covino \inst{13}
 %\and L.~B.~Christensen\inst{1}
-\and V.~D'Elia\inst{10}
-\and A.~de~Ugarte~Postigo\inst{9, 1}
+\and V.~D'Elia\inst{13, 14}
+\and A.~de~Ugarte~Postigo\inst{12, 1}
 \and J.~P.~U.~Fynbo\inst{1}
-\and A.~Gomboc\inst{11}
-\and K.~E.~Heintz\inst{8,1}
-\and A.~J.~Levan \inst{12}
-\and S.~Schulze \inst{13, 14}
-\and M.~Sparre\inst{1,15}
-\and N.~R.~Tanvir\inst{16}
-\and C.~C.~Th\"one\inst{9}
+\and A.~Gomboc\inst{15}
+\and K.~E.~Heintz\inst{11,1}
+\and A.~J.~Levan \inst{16}
+\and S.~Piranomonte \inst{17}
+\and G.~Pugliese \inst{10}
+\and R.~S\'{a}nchez-Ram\'{\i}rez \inst{12, 18}
+\and M.~Sparre\inst{1,19}
+\and N.~R.~Tanvir\inst{20}
+\and C.~C.~Th\"one\inst{12}
+\and K.~Wiersema \inst{20}
 }
@@
 %5
+\and Instituto de Astrof\'isica, Facultad de F\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+%6
+\and Millennium Institute of Astrophysics, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+%7
+\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
+%8
 \and Sorbonne Universités, UPMC Univ. Paris 6 et CNRS, UMR 7095, Institut d’Astrophysique de Paris, 98 bis bd Arago, 75014 Paris, France
-%6
+%9
 \and GEPI, Observatoire de Paris, PSL Research University, CNRS, Univ. Paris Diderot, Sorbonne Paris Cité, 5 Place Jules Janssen, 92195 Meudon, France
-%7
+%10
 \and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands
-%8
+%11
 \and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
-%9
-\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain.
-%10
-\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy;
-ASI-Science Data Centre, Via del Politecnico snc, I-00133 Rome, Italy
-%11
+%12
+\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain
+%13
+\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
+%14
+\and ASI-Science Data Centre, Via del Politecnico snc, I-00133 Rome, Italy
+%15
 \and Centre for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
-%12
+%16
 \and Department of Physics, University of Warwick, Coventry CV4 7AL, UK
-%13
-\and Instituto de Astrof\'isica, Facultad de F\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
-%14
-\and Millennium Institute of Astrophysics, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
-%15
+%17
+\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate (LC), Italy
+%18
+\and Istituto de Astrofisica e Planetologia Spaziali, INAF, Via Fosso del Cavaliere 100, I-00133 Roma, Italy
+%19
 \and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
-%16
+%20
 \and Department of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK

### 2017-06-29 GRB111117A.tex
Final edits

@@
 
-\title{The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}}
+\title{The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}}
 
@@
 their hosts to measure their redshifts. These measurements, however, are
-critical to constrain the nature of sGRB progenitors, the redshift distribution
-of these events and the $r$-process element enrichment history of
-the universe. Here, we present spectroscopy of the host galaxy of GRB~111117A
-and measure its redshift to be $z = 2.211$. This makes GRB~111117A the most
-distant high-confidence short duration GRB detected to date. Our spectroscopic
-redshift supersedes a lower, previously estimated photometric redshift value for
-this burst.
+critical to constrain the nature of sGRB progenitors, their redshift
+distribution and the $r$-process element enrichment history of the universe.
+Here, we present spectroscopy of the host galaxy of GRB~111117A and measure its
+redshift to be $z = 2.211$. This makes GRB~111117A the most distant
+high-confidence short duration GRB detected to date. Our spectroscopic redshift
+supersedes a lower, previously estimated photometric redshift value for this
+burst.
 
@@
 
-\keywords{Gamma-ray burst: individual: GRB~111117A -- Gamma-ray burst: general -- Galaxies: high-redshift -- binaries: general -- X-rays: bursts -- Techniques: imaging spectroscopy}
+\keywords{gamma-ray burst: individual: GRB~111117A -- gamma-ray burst: general -- galaxies: high-redshift -- binaries: general -- X-rays: bursts -- techniques: imaging spectroscopy}
 
@@
 most of these measurements come from the associated hosts, it is arguably biased
-towards lower redshifts.
-
-The host galaxies of sGRBs are diverse. They are more massive and less actively
-star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases,
-no host galaxy can be identified above detection threshold of deep follow-up
-observations \citep{Berger2010a, Tunnicliffe2014}. Together with their position
-within their hosts \citep{Fong2013a}, this suggests a progenitor system that can
-be very long lived in comparison to lGRBs, and is tracing with host stellar mass rather than
+towards lower redshifts. The host galaxies of sGRBs are diverse. They are more
+massive and less actively star-forming on average than lGRB hosts
+\citep{Fong2013b}, while in some cases, no host galaxy can be identified above
+detection threshold of deep follow-up observations \citep{Berger2010a,
+	Tunnicliffe2014}. Together with their position within their hosts
+\citep{Fong2013a}, this suggests a progenitor system that can be very long lived
+in comparison to lGRBs, and is tracing with host stellar mass rather than
 star-formation rate (SFR).
@@
 gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
-of the sGRB and the GW will provide new a promising ways to constrain the binary
+of a sGRB and a GW will provide new a promising ways to constrain the binary
 inclination angle \citep{Arun2014} and measure cosmological distances
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
+	\caption{SED fit showning the best-fist SED to the derived photometry.
+The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
+Overplotted in grey is the observed spectrum.
+The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
+	\label{fig:SED}
+\end{figure*}
+
 Spectroscopic observations were carried out using the cross-dispersed echelle
@@
 VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010} run in physical
-mode. Because the echelle orders are curved across each detector, a rectification
-algorithm is employed which introduces correlations between neighboring pixels.
-We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to
-minimize the degree of correlation while conserving the maximal resolution. The
-observations are combined and extracted using scripts described in Selsing et
-al. 2017 (in prep.) and available
+mode. Because the echelle orders are curved across each detector, a
+rectification algorithm is employed which introduces correlations between
+neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
+UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
+maximal resolution. The observations are combined and extracted using scripts
+described in Selsing et al. 2017 (in prep.) and available
 online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
@@
 use the optimal extraction algorithm \citep{Horne1986} and therefore the
-extraction is carried out with a simple aperture covering the entire host. An
-overview of the spectroscopic observations is given in
-Table~\ref{tab:spec_overview}, and the slit position is shown in
-Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in Fig.~\ref{fig:SED}.
+extraction is carried out with a simple aperture covering the entire host, where
+we sum the flux in the aperture and derive the error by summing the
+corresponding pixel errors in quadrature. An overview of the spectroscopic
+observations is given in Table~\ref{tab:spec_overview}, and the slit position is
+shown in Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in
+Fig.~\ref{fig:SED}.
 
@@
 
-Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a
-galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
+Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
+host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
 at 21\,000~\AA~where the night sky is very bright. In addition, several bright
@@
 although the presence of Lya does not exclude dust. Therefore we do not apply a
-dust-correction to the measured \ha~flux here. \oii is close to a region of
+dust-correction to the measured \ha~flux here. \oii~is close to a region of
 strong telluric absorption, which is why no SFR is inferred from this line.
@@
 A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
-"darkness"-parameter $\beta_{OX} $\citep{Jakobsson2004}. Using the optical
-afterglow limits\citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
+"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
+afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
 interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
@@
 
- \begin{figure*}
-	\centering
-	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
-	\caption{SED fit showning the best-fist SED to the derived photometry.
-The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
-Overplotted in grey is the observed spectrum.
-The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
-	\label{fig:SED}
-\end{figure*}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 As pointed out by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is
-securely classified as a sGRBs. Because the observed classification indicators,
+securely classified as a sGRB. Because the observed classification indicators,
 $T_{90}$ and hardness ratio, do not depend strongly on redshift
@@
 \citet{Bromberg2013} investigated the degree to which the long and short
-population distributions overlap and quantified the certainty in the class
+population distributions overlap and quantified the certainty in class
 membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
@@
 
-The redshift and secure classification of GRB~111117A means that it occurred
+The redshift and secure classification of GRB~111117A mean that it occurred
 when the universe was younger by 1.8 Gyr compared to any other non-collapsar GRB
@@
 redshift. If the merger of NSs is the primary agent for the $r$-process element
-enrichment of the universe \citep{Goriely2011, Ji2016, Komiya2016}, GRB~111117A
+enrichment of the universe \citep{Goriely2011, Ji2016, Komiya2016}, this
 marks the earliest detection of this process.
@@
 \citet{DAvanzo2014a}. From the sGRB sample of \citet{DAvanzo2014a} we have
-excluded GRB~090426 which does not belong in a short sample as highlighted in
+excluded GRB~090426 which does likely not belong in a short sample as highlighted in
 Sect. \ref{classification}. Both comparison samples consider the largest
@@
 this is indicative of a population of relatively young stars. As the majority of
-sGRBs \citep{Fong2013b}, the host of GRB~111117A is a late-type galaxy and is
-entirely consistent in terms of stellar mass and stellar age with the general
-population of sGRB hosts ($\left\langle M _* \right\rangle = 10^{10.1}
+sGRBs \citep{Fong2013b}, the host of GRB~111117A is therefore a late-type galaxy
+and is entirely consistent in terms of stellar mass and stellar age with the
+general population of sGRB hosts ($\left\langle M _* \right\rangle = 10^{10.1}
 M_{\odot}$ and $\left\langle \tau _* \right\rangle = 0.3 $~Gyr)
@@
 entirely within the range expected for the hosts of sGRBs \citep{Behroozi2014}.
+Our constraints on the dynamical mass is also well accommodated by the expected sGRB
+host halo mass \citep{Behroozi2014}.
 
@@
 distribution of sGRBs. In particular, other sGRB hosts could be missed because
-they are intrinsically fainter and thus this high-$z$ event is only detected due
-to the brightness of it's host. \citet{Berger2014} compiled a sample of sGRB
-host luminosities, normalized by the characteristic galaxy luminosity at their
-respective redshift, $L_B/L^{\star}_{B}$. To convert the SED-inferred $M_B$ of
-GRB~111117A to $L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band
-magnitude of the Schechter function for blue galaxies ($U - V < 0.25$) in the
-redshift window $2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and obtain
-$L_B/L^{\star}_{B} = 1.2$.
+they are intrinsically fainter and thus the high redshift of GRB~111117A is only
+measured due to the brightness of its host. \citet{Berger2014} compiled a
+sample of sGRB host luminosities, normalized by the characteristic galaxy
+luminosity at their respective redshift, $L_B/L^{\star}_{B}$. To convert the
+SED-inferred $M_B$ of GRB~111117A to $L_B/L^{\star}_{B}$, we use the
+characteristic absolute $B$-band magnitude of the Schechter function for blue
+galaxies ($U - V < 0.25$) in the redshift window $2.0 \leq z \leq 2.5$ from
+\citet{Marchesini2007} and obtain $L_B/L^{\star}_{B} = 1.2$.
 
 Using the complete, flux-limited selection of bursts from \citet{DAvanzo2014a},
-excluding GRB~111117A and the non-sGRB GRB~090426, we have a statistically
-homogeneous sample from which we can address the implications of the redshift of
-GRB~111117A. Out of the 14 hosts in the sample, 10 (71 per cent) have measured
-redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the host of
-GRB~111117A  is brighter than 80 per cent of the hosts with measured
-$L_B/L^{\star}_{B}$. If we conservatively assume that \textit{all} the host
-missing $L_B/L^{\star}_{B}$ are brighter than the host of GRB~111117A, the host
-still brighter than > 60 per cent of sGRB hosts. For all 26 hosts with
+excluding GRB~111117A and the likely non-sGRB GRB~090426, we have a
+statistically homogeneous sample from which we can address the implications of
+the redshift of GRB~111117A. Out of the 14 hosts in the sample, 10 (71 per cent)
+have measured redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the
+host of GRB~111117A  is brighter than 80 per cent of the hosts with measured
+$L_B/L^{\star}_{B}$. Even if we conservatively assume that \textit{all} the
+hosts missing $L_B/L^{\star}_{B}$ are brighter than the host of GRB~111117A, the
+host is still brighter than > 60 per cent of sGRB hosts. For all 26 hosts with
 $L_B/L^{\star}_{B}$ from \citet{Berger2014}, the host of GRB~111117A is brighter
@@
 
-If we assume that we are able to obtain emission-line redshifts from hosts with
-$R < 24.5$~mag \citep{Kruhler2012}, then we would have missed 60 per cent
-of the redshifts (6 out of 10 hosts), if they were at the redshift of
+If we assume that we are able to obtain emission-line redshifts from hosts, 0.5
+mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed 60
+per cent of the redshifts (6 out of 10 hosts), were they at the redshift of
 GRB~111117A due to the hosts being too faint. The corresponding number is around
-45 per cent (12 out of 26) from the full sample of \citet{Berger2014}, reflecting
-the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because the average
-SFR of galaxies hosting lGRBs is higher than for galaxies hosting sGRBs, the
-fraction of missed burst redshifts is likely higher although the cosmic SFR
-evolution could play a role in improving redshift determinability.
+45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
+reflecting the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because
+the average SFR of galaxies hosting lGRBs is higher than for galaxies hosting
+sGRBs, the fraction of missed burst redshifts is likely higher although the
+cosmic SFR evolution could play a role in improving redshift determinability.
 
-A fraction of the bursts missing redshift are host-less and is therefore likely
-at moderate redshifts \citep{Tunnicliffe2014}, but should some of the remainder
-be at high redshift, the missed fraction will increase. If we assume that
-\textit{all} the bursts that are missing redshifts are at high-$z$ and missed
-due to host faintness, 10 out of 14 hosts in the complete sample (71 per cent)
-would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
-missed burst redshifts at high-$z$. Conversely, if all bursts missing redshift
-are at low redshift and missed for other reasons, 6 out of 14 hosts (43 per
-cent) would be missed at $z = 2.211$. The two limits indicate that we are
-missing between 43 and 71 per cent of sGRB hosts at $z sim 2$ due to host
-faintness. For the 13 hosts missing redshift in \citet{Berger2014}, this means
-that between 6 and 9 of these should be at high redshift, but missed due to host
-faintness.
+A fraction of the bursts missing redshift are host-less but appear to be
+spatially correlated with galaxies that are likely at moderate redshifts
+\citep{Tunnicliffe2014}, but should some of the remainder be at high redshift,
+the missed fraction will increase. If we assume that \textit{all} the bursts
+that are missing redshifts are at high-$z$ and missed due to host faintness, 10
+out of 14 hosts in the complete sample (71 per cent) would be missed at $z =
+2.211$. This serves as an upper limit on the fraction of missed burst redshifts
+at high-$z$. Conversely, if all bursts missing redshift are at low redshift and
+missed for other reasons, 6 out of 14 hosts (43 per cent) would be missed at $z
+= 2.211$. The two limits indicate that we would miss between 43 and 71 per cent
+of sGRB hosts at $z \sim 2$ due to host faintness.
 
@@
 
-Using the new distance, the X-ray derived $n_H$ towards GRB~111117A is the
+Using the new distance, the X-ray derived $N_H$ towards GRB~111117A is the
 highest within a complete sample of sGRB hosts and is consistent with the
-$n_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
+$N_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
 upper end of the sGRB host SFR distribution and no dust is present. The high
@@
 constraints on the maximal separation between the engine constituents at the
-time of formation. We find that the maximal separation of two NSs at
+time of formation. We find that the maximal separation of two NSs at system
 formation time is $a_0 < 3.2~R_\odot$, which excludes some of the formation
@@
 \begin{acknowledgements}
-
+We thank Jens Hjorth and Lise Christensen for useful discussions regarding the interpretation of this event. 
+%	
 TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
@@
 instrumentation in astronomy.
+%
+NRT and KW acknowledge support from STFC Consolidated
+Grant ST/N000757/1
 %
@@
 The analysis and plotting has been achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy and Scipy \citep{VanderWalt2011} along with other community-developed packages.
+%
+This work made use of observations obtained with the Italian 3.6 m Telescopio Nazionale Galileo (TNG) operated on the island of La Palma by the Fundaci\'on Galileo Galilei of the INAF (Istituto Nazionale di Astrofisica) at the Spanish Observatorio del Roque de los Muchachos of the Instituto de Astrof\'isica de Canarias.
+%
+Based on data from the GTC Archive at CAB (INTA-CSIC) and on observations obtained at the Gemini Observatory, which is operated by the Association of Universities for Research in Astronomy, Inc., under a cooperative agreement with the NSF on behalf of the Gemini partnership: the National Science Foundation (United States), the National Research Council (Canada), CONICYT (Chile), Ministerio de Ciencia, Tecnología e Innovación Productiva (Argentina), and Ministério da Ciência, Tecnologia e Inovação (Brazil).
 

### 2017-06-29 py/LB_observability.py
rename

@@ -37,7 +37,7 @@ def main():
     # Filter for names
     complete_L_B = [ii for ii, kk in enumerate(name) if kk in flux_complete]
     print(len(complete_L_B))
-    complete = True
+    complete = False
     if complete == True:
         L_B = L_B[complete_L_B]
     # print(np.nanmean(L_B))

### 2017-07-01 GRB111117A.tex
Adding data
Astropy magnitude conversion and Susannas edits

@@
 
-\title{The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}}
+\title{The host galaxy of the short GRB~111117A at $z = 2.211$: impact on the short GRB redshift distribution and progenitor channels\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}}
 
@@
 The classification of GRBs in two groups, initially comes from the bimodal
-distribution of burst duration and spectral hardness \citet{Kouveliotou1993},
+distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
 where the duration T$_{90} < 2$ has been regarded as the dividing line between long
@@
 gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
-of a sGRB and a GW will provide new a promising ways to constrain the binary
+of a sGRB and a GW will provide new promising ways to constrain the binary
 inclination angle \citep{Arun2014} and measure cosmological distances
@@
 The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
-Overplotted in grey is the observed spectrum.
+Overplotted in grey is the observed spectrum, binned by 6 \AA~for presentation purposes.
+Slit losses are responsible for discrepancies between the measured photometry and the X-shooter spectrum.
 The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
@@
 tracing, wavelength calibration, rectification, and flux calibration using the
-VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010} run in physical
+VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run in physical
 mode. Because the echelle orders are curved across each detector, a
@@
 described in Selsing et al. 2017 (in prep.) and available
-online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
-signal-to-noise ratio of the continuum in the near-infrared arm is too low to
-use the optimal extraction algorithm \citep{Horne1986} and therefore the
-extraction is carried out with a simple aperture covering the entire host, where
-we sum the flux in the aperture and derive the error by summing the
-corresponding pixel errors in quadrature. An overview of the spectroscopic
-observations is given in Table~\ref{tab:spec_overview}, and the slit position is
-shown in Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in
-Fig.~\ref{fig:SED}.
+online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
+where the full spectral point spread function is modeled across each arm and
+used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
+spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
+slit position is shown in Fig.~\ref{fig:spec_setup}. We show the extracted
+spectrum in Fig.~\ref{fig:SED}.
 
@@
 edge-on and because the slit is not aligned along the long axis of the host,
-this value is an upper limit.
-
-A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
-"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
-afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
-interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
-< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
+this value is a lower limit.
 
@@
 observations and measured brightnesses is given in
-Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, there
-is likely no afterglow contamination on the measured photometry.
+Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, see
+Sect. \ref{xray}, there is likely no afterglow contamination on the measured
+photometry.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{X-ray temporal and spectral analysis}
+\subsection{X-ray temporal and spectral analysis}\label{xray}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
+"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
+afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
+interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
+< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 surrounding is high, or because the light from the afterglow traverses a region
-of dense gas.
+of dense gas. The optical darkness of the GRB is an additional hint of the high
+density in the GRB surroundings or along the line of sight.
 
@@
 	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations of GRBs.
-The detections are replaced with contours and the limits are shown with arrows.}
+The detections are replaced with contours and the limits are shown with arrows.
+Marginalizations over both axes are shown where the limits are shown as transparent bars and detections as solid ones.}
 	\label{fig:NH_z}
@@
 If we assume that we are able to obtain emission-line redshifts from hosts, 0.5
-mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed 60
-per cent of the redshifts (6 out of 10 hosts), were they at the redshift of
-GRB~111117A due to the hosts being too faint. The corresponding number is around
-45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
+mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed
+60 per cent of the redshifts (6 out of 10 hosts), due too the host being too
+faint, were they at the redshift of GRB~11117A. The corresponding number is
+around 45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
 reflecting the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because
@@
 on the other hand, is close to the expected peak in sGRB rate calculated using
-the power law delay time models \citep{Behroozi2014, Wanderman2015}.
+the power law delay time models \citep{Behroozi2014, Wanderman2015,
+	Ghirlanda2016}, meaning we could be missing a large fraction of sGRBs.
 
@@
 does not account for the possibility there could be an underlying stellar
-population of older stars from a previous star-formation epoch. The delay time
-between formation and explosion is well accommodated by the models of
-\citet{Belczynski2006}, although the longest delay times are excluded.
-This is especially true given the late type nature of the host
-\citep{OShaughnessy2008}.
+population of older stars from a previous star-formation epoch. To investigate
+the possible impact of the presence of an old stellar population, we followed
+\citet{Papovich2001} and re-fitted the observed SED with the best-fit template
+to which an additional stellar population of old stars was added. For each
+galaxy, this old population was set as the SPSs with same parameters that the
+best-fit SED excepts the age, which was set to the age of the Universe at the
+observed redshift. In principle, this can constrain the maximum contribution of
+old populations within the photometric error bars \citep[see][ for
+details]{Papovich2001}. We find a negligible contribution to the stellar mass
+(i.e. variations much smaller than the statistical uncertainty associated with
+the best-fit template). The delay time between formation and explosion is well
+accommodated by the models of \citet{Belczynski2006}, although the longest delay
+times are excluded. This is especially true given the late type nature of the
+host \citep{OShaughnessy2008}.
 
@@
 $N_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
-upper end of the sGRB host SFR distribution and no dust is present. The high
-$N_H$ is difficult to reconcile with the large projected host offset and the
-absence of dust. One possible explanation could be, that GRB~111117A is formed
-through a prompt channel of sGRB formation and originates in a star forming
-region located in the outskirts of the host.
+upper end of the sGRB host SFR distribution and no significant amount of dust is
+present. The high $N_H$ is difficult to reconcile with the large projected host
+offset and the absence of dust. One possible explanation could be, that
+GRB~111117A is formed through a prompt channel of sGRB formation and originates
+in a star forming region located in the outskirts of the host.
 
@@
 
-All data, code and calculation related to the paper along with the
+All data, code and calculations related to the paper along with the
 paper itself are available at \url{https://github.com/jselsing/GRB111117A}.
@@
 We thank Jens Hjorth and Lise Christensen for useful discussions regarding the interpretation of this event.
+We thank Mathieu Puech for testing the possible contribution from an older stellar population in the SED.
 %	
 TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
+%
+SDV is supported by the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro 
 %

### 2017-07-01 py/SED_plot.py

@@ -6,6 +6,8 @@ from __future__ import division, print_function
 import numpy as np
 from scipy.interpolate import interp1d
 from scipy.signal import medfilt
+import astropy.units as u
+import astropy.constants as c
 
 # Plotting
 import matplotlib.pyplot as pl
@@ -24,24 +26,15 @@ params = {
    }
 mpl.rcParams.update(params)
 
-
-def jansky_to_AB(jansky):
-    return 23.9 - 2.5*np.log10(jansky)
-
-def AB_to_jansky(AB):
-    return 10**((23.9 - AB)/2.5)
-
-
 def main():
     # Small script to plot SED, photometry and spectrum
 
     # Load data
-    spectrum = np.genfromtxt("../data/stitched_spectrum_bin1.dat")
-    wl_spec = spectrum[:, 0]
-    flux_spec = spectrum[:, 1]
-    error_spec = spectrum[:, 2]
-    # pl.plot(wl_spec, flux_spec)
-    # pl.show()
+    spectrum = np.genfromtxt("../data/spectroscopy/stitched_spectrum_bin5.dat")
+    wl_spec = spectrum[:, 0] * u.AA
+    flux_spec = spectrum[:, 1] * (u.erg/(u.s * u.cm**2 * u.AA))
+    error_spec = spectrum[:, 2] * (u.erg/(u.s * u.cm**2 * u.AA))
+
     SED = np.genfromtxt("../data/SED.dat")
     wl_SED = SED[:, 0]
     AB_SED = SED[:, 1]
@@ -52,29 +45,14 @@ def main():
     magerr = phot[:, 1][magmask]
     wl_phot = phot[:, 2][magmask]
     wl_plot_wid = phot[:, 3][magmask]/2.35
-    # pl.scatter(wl_phot, mag)
-    # print(phot[:, 2])
-    # print(phot[:, 3])
-    # exit()
 
+    AB = flux_spec.to(u.ABmag, u.spectral_density(wl_spec))
 
-    nu = 3e5/(wl_spec*1e-13) # nu
-    f_nu = (flux_spec*(1e-13*(wl_spec**2.))/3e5)/1e-29
-    f_nu[f_nu <= 0] = 0.1
-    f_nu_err = (error_spec*(1e-13*(wl_spec**2.))/3e5)/1e-29
-    f_nu_err[f_nu_err <= 0] = 1
-
-
[…]

### 2017-07-01 py/combine_arms_and_bin.py

@@ -29,7 +29,7 @@ def load_array(data_array, tell = None):
 
     if has_resp:
         try:
-            tellcorr = data_array[:, 8]
+            tellcorr = 1#data_array[:, 8]
             tellcorr[np.isnan(tellcorr)] = 1
             slitcorr = 1#np.median(data_array[:, 7])
             return data_array[:, 1], data_array[:, 2]*tellcorr*slitcorr, data_array[:, 3]*tellcorr*slitcorr, data_array[:, 4]
@@ -38,7 +38,7 @@ def load_array(data_array, tell = None):
             return data_array[:, 1], data_array[:, 2]*slitcorr, data_array[:, 3]*slitcorr, data_array[:, 4]
     elif not has_resp:
         try:
-            tellcorr = tell#data_array[:, 7]
+            tellcorr = 1#tell#data_array[:, 7]
             # print(tellcorr)
             tellcorr[np.isnan(tellcorr)] = 1
             slitcorr = 1#np.median(data_array[:, 6])
@@ -159,7 +159,7 @@ def main():
     # Small script to stitch X-shooter arms. Inspired by https://github.com/skylergrammer/Astro-Python/blob/master/stitch_spec.py
 
     # Load data from individual files
-    data_dir = "../data/"
+    data_dir = "../data/spectroscopy/"
 
     scale = False # True, False
 
@@ -179,7 +179,7 @@ def main():
     # Construct lists
     UVB_mask = (UVB_wl > 3300) & (UVB_wl < 5500)
     VIS_mask = (VIS_wl > 5700) & (VIS_wl < 9600)
-    NIR_mask = (NIR_wl > 12000) & (NIR_wl < 21000) & (NIR_flux > -1e-19)
+    NIR_mask = (NIR_wl > 10500) & (NIR_wl < 20500) #& (NIR_flux > -1e-19)
     waves = [UVB_wl[UVB_mask], VIS_wl[VIS_mask], NIR_wl[NIR_mask]]
     flux = [UVB_flux[UVB_mask], VIS_flux[VIS_mask], NIR_flux[NIR_mask]]
     error = [UVB_error[UVB_mask], VIS_error[VIS_mask], NIR_error[NIR_mask]]
@@ -190,7 +190,7 @@ def main():
 
[…]

### 2017-08-16 GRB111117A.tex
Addressing referee part 1

@@
 Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
-host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
-at 21\,000~\AA~where the night sky is very bright. In addition, several bright
+host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at
+21\,000~\AA~where the night sky is very bright. In addition, several bright
 sky-lines are superposed on the line, making an accurate estimate of the
-\ha-flux difficult. We obtain a limit on the SFR by numerically integrating the
-part of \ha~free of contamination and correcting for the missing flux using the
-line profile and obtain $F_{H \alpha} > 4.1 \times 10^{-17} \mathrm{erg}
-\mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting the \citet{Kennicutt1998}
-relation to a \citet{Chabrier2003} initial mass function using
-\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$.
-Additionally, based on the width of \oiii, an aperture covering the line is
-integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have been
-masked and interpolated over. From the integrated \ha-line, we estimate SFR =
-$18 \pm 3$ M$_\odot$ yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and the
-detection of \lya, the host is constrained to contain very little or no dust,
-although the presence of Lya does not exclude dust. Therefore we do not apply a
-dust-correction to the measured \ha~flux here. \oii~is close to a region of
-strong telluric absorption, which is why no SFR is inferred from this line.
+\ha-flux difficult. Due to their velocity structure, the lines exhibit clear
+deviations from a Gaussian. Given the low S/N of the spectra we do not attempt
+any parametric fits. We thus obtain a limit on the SFR by numerically
+integrating the part of \ha~free of contamination and obtain $F_{H \alpha} > 4.1
+\times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting
+the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass
+function using \citet{Madau2014}, we derive a limit of $SFR > 7
+M_\odot$~yr$^{-1}$. 
+%Additionally, based on the width of \oiii, an aperture
+%covering \ha is integrated over, where sky lines have been masked and
+%interpolated over, using a synthetic sky spectrum \citep{Noll2012, Jones2013}.
+%From the integrated \ha-line, we estimate SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.
+%The error on the SFR is derived from the integrated variance spectrum over the
+%\ha-line and excluding the intrinsic scatter of the relation. 
+From the SED-fit
+(Sect.~\ref{SED}), and the detection of \lya, the host is constrained to contain
+very little or no dust, although the presence of Lya does not exclude dust.
+Therefore we do not apply a dust-correction to the measured \ha~flux here.
+\oii~is close to a region of strong telluric absorption, which is why no SFR is
+inferred from this line.
 
@@
 observations and measured brightnesses is given in
-Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, see
-Sect. \ref{xray}, there is likely no afterglow contamination on the measured
-photometry.
+Table~\ref{tab:phot_overview}. Most data were taken long after the GRB when no
+afterglow contribution was present. Given the faintness of the afterglow (see
+Sect. \ref{xray}, and \citealt{Cucchiara2011, Cenko2011}), we also expect
+negligible contribution to the earliest epochs, which is confirmed by the
+consistency between the two $g$-band measurements.
 
@@
 Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
-reduction which includes bias/flat-field correction, de-fringing (if necessary),
-sky-subtraction, and stacking of individual images. The photometry of the host
-was calibrated relative to field stars from the SDSS and 2MASS catalogs in the
-case of $grizJHK_{\mathrm{s}}$ filters.
-For the $R$ and $I$-band photometry, we used the color transformations of
-Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
-We convert all magnitudes into the AB system, and correct for a Galactic 
-foreground of $E_{B-V}=0.027~\mathrm{mag}$ \citep{Schlegel1998, Schlafly2011}.
+reduction which includes bias/flat-field correction, de-fringing (where
+necessary), sky-subtraction, and stacking of individual images. On the final
+reduced products, seeing-matched aperture photometry was carried out using
+\texttt{Source Extractor} \citep{Bertin1996} where the size of the aperture was
+chosen to be twice the seeing FWHM. (THOMAS! WAS THE APERTURE FIXED BETWEEN
+FILTERS AND IS THIS CORRECT?) Photometric calibration was fixed relative to
+field stars from the SDSS and 2MASS catalogs in the case of $griz$ and
+$JHK_{\mathrm{s}}$ filters, respectively. For the $R$- and $I$-band photometry,
+we used the color transformations of Lupton%
+\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. We
+convert all magnitudes into the AB system, and correct for a Galactic foreground
+of $E_{B-V}=0.027~\mathrm{mag}$ \citep{Schlegel1998, Schlafly2011}.
+
+Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the
+re-analysis of the Gemini data compared to the $z$-band measurements of
+\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
+$z$-band magnitude that is brighter by 0.8-1.0 mag compared to our value, where
+data taken in other filters are consistent within the errors. Visual inspection
+of the Gemini image shows only a marginal detection of the host, which we report
+here as a $\sim$ 2-$\sigma$ measurement. More conservatively, the 3-$\sigma$
+upper limit for the Gemini image is $z > 24.06$. Objects of magnitude $z \sim
+23.2$ are clearly seen and are significantly brighter than the GRB host galaxy.
+The consistency between the deeper FORS2 $z$-band image and the upper limit
+derived for the Gemini image lends credence to the inferred magnitudes, see
+Fig. \ref{photplot}.
 
@@
 
-The multi-color SED is fit using the \citet{Bruzual2003} single stellar population
-models based on a \citet{Chabrier2003} with initial mass function in
+The multi-color SED is fit using the \citet{Bruzual2003} single stellar
+population models based on a \citet{Chabrier2003} with initial mass function in
 \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the
-spectroscopic value of $z=2.211$. The best fit model is an unreddened galaxy
+spectroscopic value of $z=2.211$. For the SED fitting, we create a grid of
+templates consisting of $\sim 10^6$  different galaxy templates with four
+metallicities 0.02, 0.2, 0.4, 1.0 $Z_{\odot}$, different ages, star-formation
+histories and degrees of extinction. For every model, we calculate the
+likelihood, and create a probability density function (PDF) for a given
+parameter by marginalizing over the other parameters. We quote the median of the
+PDF as the best fit parameters and the errors are the 16th and 84th percentiles
+of the PDFs \citep[e.g. see][for details on the SED fitting
+procedure]{Schulze2016}.
+
+The best fit model is an unreddened galaxy
 template, and returns physical parameters of absolute magnitude
@@
 
-Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
-re-analysis of the Gemini data to the $z$-band measurements of
-\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
-$z$-band photometry that is brighter by 0.8~mag to 1.0 mag compared to our
-value, where data taken in bluer filters are in excellent agreement. The large
-$i-z$ color was mistakenly interpreted as a 4000\,\AA\,break driving the galaxy
-photometric redshift of the earlier works. Using the revised photometry from
+The large $i-z$ color was mistakenly interpreted as a 4000\,\AA~ break driving
+the galaxy photometric redshift of the earlier works. Even without fixing the
+redshift to the spectroscopic value, using the revised photometry from
 Table~\ref{tab:phot_overview}, the photometric redshift of the galaxy is

### 2017-08-16 GRB111117A.tex
Referee comments part 2

@@
 	\caption{SED fit showning the best-fist SED to the derived photometry.
-The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
-Overplotted in grey is the observed spectrum, binned by 6 \AA~for presentation purposes.
+The detection of \lya{}  is predicted from the SED fit and confirmed by the spectroscopic observations.
+Overplotted in grey is the observed spectrum, binned by 6 \AA{} for presentation purposes.
 Slit losses are responsible for discrepancies between the measured photometry and the X-shooter spectrum.
-The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
+The reason for the spectral gaps at 5500 \AA{} and 10000 \AA{} is from the merging of the arms.}
 	\label{fig:SED}
@@
 091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
-arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA~to
-24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
+arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
+24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across three
 spectroscopic arms. We carried out the bias-correction, flat-fielding, order
@@
 emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$5007,
-and \ha. \hb~is detected at low significance ($\sim 3 \sigma$). We show
-\oiii$\lambda$5007 in the insert in Fig.~\ref{fig:line}.  \ha~is only visible in
+and \ha. \hb{} is detected at low significance ($\sim$ 3-$\sigma$). We show
+\oiii$\lambda$5007 in the insert in Fig.~\ref{fig:line}.  \ha{} is only visible in
 the first epoch due to the $K$-band blocking filter used for the remaining
@@
 show significant velocity structure along the slit. A drop in the continuum
-bluewards of the \lya~line further supports the inferred redshift.
+bluewards of the \lya{} line further supports the inferred redshift.
 
 Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
-host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at
-21\,000~\AA~where the night sky is very bright. In addition, several bright
+host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed at
+21\,000~\AA{} where the night sky is very bright. In addition, several bright
 sky-lines are superposed on the line, making an accurate estimate of the
 \ha-flux difficult. Due to their velocity structure, the lines exhibit clear
-deviations from a Gaussian. Given the low S/N of the spectra we do not attempt
-any parametric fits. We thus obtain a limit on the SFR by numerically
-integrating the part of \ha~free of contamination and obtain $F_{H \alpha} > 4.1
+deviations from a Gaussian and given the low S/N of the spectra we do not attempt
+any parametric fits. We instead obtain a limit on the SFR by numerically
+integrating the part of \ha{} free of contamination and obtain $F_{H \alpha} > 4.1
 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting
@@
 very little or no dust, although the presence of Lya does not exclude dust.
-Therefore we do not apply a dust-correction to the measured \ha~flux here.
-\oii~is close to a region of strong telluric absorption, which is why no SFR is
+Therefore we do not apply a dust-correction to the measured \ha{} flux here.
+\oii{} is close to a region of strong telluric absorption, which is why no SFR is
 inferred from this line.
@@
 upper limit for the Gemini image is $z > 24.06$. Objects of magnitude $z \sim
-23.2$ are clearly seen and are significantly brighter than the GRB host galaxy.
+23$ are clearly seen and are significantly brighter than the GRB host galaxy.
 The consistency between the deeper FORS2 $z$-band image and the upper limit
-derived for the Gemini image lends credence to the inferred magnitudes, see
-Fig. \ref{photplot}.
+derived for the Gemini image lends credence to the inferred magnitudes presented
+here, see Fig. \ref{photplot}.
 
@@
 
-The large $i-z$ color was mistakenly interpreted as a 4000\,\AA~ break driving
-the galaxy photometric redshift of the earlier works. Even without fixing the
-redshift to the spectroscopic value, using the revised photometry from
-Table~\ref{tab:phot_overview}, the photometric redshift of the galaxy is
-$z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic
-value at the 1~$\sigma$ confidence level.
+Without fixing the redshift to the spectroscopic value, using the revised
+photometry from Table~\ref{tab:phot_overview}, the photometric redshift of the
+galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the
+spectroscopic value at the 1-$\sigma$ confidence level. The large $i-z$ color
+found in previous works was mistakenly interpreted as a 4000\,\AA{} break
+driving the galaxy photometric redshift to a lower, erroneous value.
 
@@
 $N_H$-$M_\star$ relation. Additionally, for lGRBs, $N_H$ correlates with the
-surface luminosity at the explosion site \citep{Lyman2017}. The large offset of
-GRB~111117A relative to the host center derived in \citet{Margutti2012} and
-\citet{Sakamoto2013} is difficult to reconcile with galaxy-scale gas providing
-the X-ray absorption. Along with the absence of dust, the large offset from the
-host center indicates that the high $N_H$ arises because the density in the GRB
-surrounding is high, or because the light from the afterglow traverses a region
-of dense gas. The optical darkness of the GRB is an additional hint of the high
-density in the GRB surroundings or along the line of sight.
+surface luminosity at the explosion site \citep{Lyman2017}.
+
+The large offset of GRB~111117A relative to the host center derived in
+\citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
+galaxy-scale gas providing the X-ray absorption. Along with the low dust content
+of the host, the large offset from the host center indicates that the high $N_H$
+arises because the density in the GRB surroundings is high, or because the light
+from the afterglow traverses a region of dense gas. Assuming a dust-to-metals
+ratio typical for the Milky Way, the X-ray derived $N_H$ corresponds to significant
+extinction along the afterglow sight line ($A_V > 10$) \citep{Predehl1995},
+which is contrasted to the absence of dust found for the SED fit.
+The discrepancy between the extinction derived from the GRB afterglow and the one obtained using galaxy-wide measures is also seen in the range of ratios found for lGRBs \citep{Perley2013a, Covino2013}.
+
+The optical faintness of the
+GRB is an additional hint of the high density in the GRB surroundings or along
+the line of sight.
 
@@
 
-The detection of \lya~is consistent with the SED-inferred absence of dust,
+The detection of \lya{} is consistent with the SED-inferred absence of dust,
 despite the moderate stellar age which would suggest the opposite. The centroid

### 2017-10-04 GRB111117A.tex
Addressing referee

@@
 
-The host lies in the brighter end of the expected sGRB host brightness distribution
-at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
-redshift determination, we find that between 43 and 71 per cent of all sGRB redshifts
-should be missed due to host faintness for hosts at $z\sim2$. The high redshift of
-GRB~111117A is evidence against a lognormal delay-time model for sGRBs through
-the predicted redshift distribution of sGRBs, which is very sensitive to
-high-$z$ sGRBs.
+The host lies in the brighter end of the expected sGRB host brightness
+distribution at $z = 2.211$, and is actively forming stars. Using the host as a
+benchmark for redshift determination, we find that between 43 and 71 per cent of
+all \textit{Swift}-detected sGRB should be missing redshifts due to host
+faintness for bursts at $z\sim2$. The high redshift of GRB~111117A is evidence
+against a lognormal delay-time model for sGRBs through the predicted redshift
+distribution of sGRBs, which is very sensitive to high-$z$ sGRBs.
 
@@
 
-There is mounting evidence that short-duration $\gamma$-ray bursts
-come from the merger of NSs, either with another NS, or a black
+There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs)
+come from the merger of neutron stars (NSs), either with another NS, or a black
 hole, due to their apparent association with kilonovae \citep{Barnes2013a,
-	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
+	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
 supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
@@
 distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
-where the duration T$_{90} < 2$ has been regarded as the dividing line between long
+where the duration T$_{90} < 2$ sec has been regarded as the dividing line between long
 and short GRBs. Additionally, it has been found for lGRBs that there is a
@@
 redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, and because
-most of these measurements come from the associated hosts, it is arguably biased
-towards lower redshifts. The host galaxies of sGRBs are diverse. They are more
-massive and less actively star-forming on average than lGRB hosts
-\citep{Fong2013b}, while in some cases, no host galaxy can be identified above
-detection threshold of deep follow-up observations \citep{Berger2010a,
-	Tunnicliffe2014}. Together with their position within their hosts
-\citep{Fong2013a}, this suggests a progenitor system that can be very long lived
-in comparison to lGRBs, and is tracing with host stellar mass rather than
-star-formation rate (SFR).
+most of these measurements come from the associated hosts, it is potentially
+biased towards lower redshifts due to Malmquist bias. The host galaxies of sGRBs
+are diverse. They are more massive and less actively star-forming on average
+than lGRB hosts \citep{Fong2013b}, while in some cases, no host galaxy can be
+identified above detection threshold of deep follow-up observations
+\citep{Berger2010a, Tunnicliffe2014}. Together with their position within their
+hosts \citep{Fong2013a}, this suggests a progenitor system that can be very long
+lived in comparison to lGRBs, with the host stellar mass affecting the sGRB
+rates more than star-formation rate (SFR) \citep{Berger2014}.
 
@@
 	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
-	\caption{SED fit showning the best-fist SED to the derived photometry.
+	\caption{SED fit showing the best-fit SED to the derived photometry.
 The detection of \lya{}  is predicted from the SED fit and confirmed by the spectroscopic observations.
 Overplotted in grey is the observed spectrum, binned by 6 \AA{} for presentation purposes.
-Slit losses are responsible for discrepancies between the measured photometry and the X-shooter spectrum.
+An attempt to correct for slit losses has been carried out based on the average seeing of the observations.
+The blue, dashed line is the corresponding error spectrum, smoothed for presentation purposes.
 The reason for the spectral gaps at 5500 \AA{} and 10000 \AA{} is from the merging of the arms.}
@@
 arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
-24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across three
-spectroscopic arms. We carried out the bias-correction, flat-fielding, order
-tracing, wavelength calibration, rectification, and flux calibration using the
-VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run in physical
-mode. Because the echelle orders are curved across each detector, a
+24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across
+three spectroscopic arms. We carried out the bias-correction, flat-fielding,
+order tracing, wavelength calibration, rectification, and flux calibration using
+the VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run
+in physical mode. Because the echelle orders are curved across each detector, a
 rectification algorithm is employed which introduces correlations between
@@
 spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
-slit position is shown in Fig.~\ref{fig:spec_setup}. We show the extracted
-spectrum in Fig.~\ref{fig:SED}.
+slit position is shown in Fig.~\ref{fig:spec_setup}. Each of the epochs are
+extracted individually and combined in a weighted fashion where the weights at
+each pixel is chosen as median variance spectrum of the region surrounding that
+pixel, thus avoiding to base the weight on the pixel variance. An attempt to
+correct for slit-losses has been carried out on the combined spectrum based on
+the average seeing of the observations. We show the extracted spectrum in
+Fig.~\ref{fig:SED}.
 
@@
 We determine a redshift of $z = 2.211$ from the simultaneous detection of
-emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$5007,
-and \ha. \hb{} is detected at low significance ($\sim$ 3-$\sigma$). We show
-\oiii$\lambda$5007 in the insert in Fig.~\ref{fig:line}.  \ha{} is only visible in
-the first epoch due to the $K$-band blocking filter used for the remaining
-observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and
-show significant velocity structure along the slit. A drop in the continuum
-bluewards of the \lya{} line further supports the inferred redshift.
+emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959,
+\oiii$\lambda$5007, and \ha. \hb{} and \oiii$\lambda$4959 are both detected at
+low significance ($\sim$ 3-$\sigma$). We show \oiii$\lambda$5007 in the insert
+in Fig.~\ref{fig:line}.  \ha{} is only visible in the first epoch due to the
+$K$-band blocking filter used for the remaining observations. The nebular lines
+exhibit a spatial extent of $\sim$ 1\farc5 and show significant velocity
+structure along the slit. A drop in the continuum bluewards of the \lya{} line
+further supports the inferred redshift. No spectral evolution is observed across
+the epochs, supporting the lack of afterglow contribution.
 
 Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
-host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed at
-21\,000~\AA{} where the night sky is very bright. In addition, several bright
+host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed
+at 21\,000~\AA{} where the night sky is very bright. In addition, several bright
 sky-lines are superposed on the line, making an accurate estimate of the
 \ha-flux difficult. Due to their velocity structure, the lines exhibit clear
-deviations from a Gaussian and given the low S/N of the spectra we do not attempt
-any parametric fits. We instead obtain a limit on the SFR by numerically
-integrating the part of \ha{} free of contamination and obtain $F_{H \alpha} > 4.1
-\times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting
-the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass
-function using \citet{Madau2014}, we derive a limit of $SFR > 7
-M_\odot$~yr$^{-1}$. 
+deviations from a Gaussian and given the low S/N of the spectra we do not
+attempt any parametric fits. We instead obtain a limit on the SFR by numerically
+integrating the part of \ha{} free of contamination and obtain $F_{H \alpha} >
+4.1 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After
+converting the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial
+mass function using \citet{Madau2014}, we derive a limit of $SFR > 7
+M_\odot$~yr$^{-1}$. We additionally obtain the \lya{} line flux by numerically integrating
+the entire \lya{} line complex and obtain $F_{\mathrm{Ly}\alpha} = 2.0 \pm 0.5
+\times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$.
+
 %Additionally, based on the width of \oiii, an aperture
@@
 %\ha-line and excluding the intrinsic scatter of the relation.
-From the SED-fit
-(Sect.~\ref{SED}), and the detection of \lya, the host is constrained to contain
-very little or no dust, although the presence of Lya does not exclude dust.
-Therefore we do not apply a dust-correction to the measured \ha{} flux here.
-\oii{} is close to a region of strong telluric absorption, which is why no SFR is
-inferred from this line.
+
+From the SED-fit (Sect.~\ref{SED}), and the detection of \lya{}, the host is
+constrained to contain very little or no dust, although the presence of \lya{}
+does not exclude dust. Therefore we do not apply a dust-correction to the
+measured \ha{} flux here. \oii{} is close to a region of strong telluric
+absorption, which is why no SFR is inferred from this line.
 
@@
 observations and measured brightnesses is given in
-Table~\ref{tab:phot_overview}. Most data were taken long after the GRB when no
-afterglow contribution was present. Given the faintness of the afterglow (see
-Sect. \ref{xray}, and \citealt{Cucchiara2011, Cenko2011}), we also expect
+Table~\ref{tab:phot_overview}. Most data were taken long after the GRB had faded
+when no afterglow contribution was present. Given the faintness of the afterglow
+(see Sect. \ref{xray}, and \citealt{Cucchiara2011, Cenko2011}), we also expect
 negligible contribution to the earliest epochs, which is confirmed by the
@@
 necessary), sky-subtraction, and stacking of individual images. On the final
-reduced products, seeing-matched aperture photometry was carried out using
-\texttt{Source Extractor} \citep{Bertin1996} where the size of the aperture was
-chosen to be twice the seeing FWHM. (THOMAS! WAS THE APERTURE FIXED BETWEEN
-FILTERS AND IS THIS CORRECT?) Photometric calibration was fixed relative to
-field stars from the SDSS and 2MASS catalogs in the case of $griz$ and
-$JHK_{\mathrm{s}}$ filters, respectively. For the $R$- and $I$-band photometry,
-we used the color transformations of Lupton%
+reduced image products, \texttt{DAOPHOT} \citep{Stetson1987} was used to derive
+the reported photometry, where the size of the aperture was chosen to be
+2\farc0. Because image quality ranges from 0\farc6 - 1\farc7 and the galaxy
+major axis is $\sim$ 1\farc, this ensures that, in all cases, the large majority
+of the flux is collected and low-surface brightness light missed in the aperture
+will not influence the measurement. This method sacrifices some S/N in the best
+seeing cases for more reliable photometry across differing observing conditions.
+
+Photometric calibration was fixed relative to field stars from the SDSS and
+2MASS catalogs in the case of $griz$ and $JHK_{\mathrm{s}}$ filters,
+respectively. For the $R$- and $I$-band photometry, we used the color
+transformations of Lupton%
 \footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. We
@@
 \citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
-$z$-band magnitude that is brighter by 0.8-1.0 mag compared to our value, where
-data taken in other filters are consistent within the errors. Visual inspection
-of the Gemini image shows only a marginal detection of the host, which we report
-here as a $\sim$ 2-$\sigma$ measurement. More conservatively, the 3-$\sigma$
-upper limit for the Gemini image is $z > 24.06$. Objects of magnitude $z \sim
-23$ are clearly seen and are significantly brighter than the GRB host galaxy.
-The consistency between the deeper FORS2 $z$-band image and the upper limit
-derived for the Gemini image lends credence to the inferred magnitudes presented
-here, see Fig. \ref{photplot}.
+$z$-band magnitudes that are brighter by $\sim 1.0$ mag compared to our values,
+where data taken in other filters are consistent within the errors. Visual
+inspection of the Gemini image shows only a marginal detection of the host,
+which we report here as a $\sim$ 2-$\sigma$ measurement. More conservatively,
+the 3-$\sigma$ upper limit for the Gemini image is $z > 24.06$. Objects of
+magnitude $z \sim 23$ are clearly seen and are significantly brighter than the
+GRB host galaxy. The consistency between the deeper FORS2 $z$-band image and the
+upper limit derived for the Gemini image lends credence to the inferred
+magnitudes presented here, see Fig. \ref{fig:photplot}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=9cm]{figures/photometry_mosaic.pdf}
+	\caption{Mosaic showing all obtained imaging.
+The host of GRB~111117A is marked by a red circle with a 2\farc0 radius.
+This is the same size as the aperture used for deriving the photometry.
+Worth noting is the relative depth of the GMOS-N $z$-band and the FORS2 $z$-band.
+For reference, the object located to the south-east of the host and marked with a blue, dashed circle has an extinction corrected magnitude of $23.11 \pm 0.09$ ($23.10 \pm 0.18$) in the FORS2 (GMOS-N) $z$-band.}
+	\label{fig:photplot}
+\end{figure}
 
@@
 "darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
-afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
-interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
-< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
+afterglow limits ($r' > 25.5$, 13.5 hrs post-burst) \citep{Cucchiara2011,
+	Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time
+of the non-detection. We find $\beta_{OX} < 0.79$, consistent with what was
+reported in \citet{Sakamoto2013}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-As pointed out by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is
-securely classified as a sGRB. Because the observed classification indicators,
-$T_{90}$ and hardness ratio, do not depend strongly on redshift
-\citep{Littlejohns2013a}, the updated redshift does not change this designation.
-The intrinsic spectral lag shortens, but since it is already consistent with
-zero, this does not affect the classification.
+Based on the BAT light-curve, $T_{90}=0.46$~s which is shorter than both the
+prototypical 2 s \citep{Kouveliotou1993} and the $\sim$0.7 s suggested to also
+exclude the shorter tail of the \textit{Swift}-observed lGRB population
+\citep{Bromberg2012a}. Additionally, no signs of extended emission was found by
+\citep{Sakamoto2013}. A spectral lag of $0.6 \pm2.4$ ms, consistent with zero,
+is not uniquely associated to sGRBs \citep{Bernardini2014}, but in conjunction
+with the duration and the spectral hardness \citep{Sakamoto2011}, as pointed out
+by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is securely
+classified as a sGRB. Because the observed classification indicators, $T_{90}$
+and hardness ratio, do not depend strongly on redshift \citep{Littlejohns2013a},
+the updated redshift does not change this designation. The intrinsic spectral
+lag shortens, but since it is already consistent with zero, this does not affect
+the classification.
 
@@
 long GRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
-distribution at 10 ks. The sub-samples comprise of 333 long, 19 short GRBs, and
-GRB~111117A. The mean and the 1-$\sigma$ dispersions of the samples are
-L$_{lGRB} = 46.59 \pm 0.87$ and L$_{sGRB} = 44.96 \pm 0.94$. GRB~111117A had a
-luminosity of 44.95 at 10 ks which is very close to the peak of the sGRB
-luminosity distribution at 10 ks, but an outlier from the lGRG distribution,
-further supporting the short classification.
+distribution at 10 ks. The sub-samples comprise of 333 lGRBs, 19 sGRBs, and
+GRB~111117A. The sample of lGRBs are from \citet{Evans2007, Evans2009} and the
+sample of sGRBs are compiled from \citet{Kann2011, Berger2014} and
+\citet{DAvanzo2014a}. The mean and the 1-$\sigma$ dispersions of the samples are
+log(L$_{lGRB}$)$ = 46.59 \pm 0.87$ erg s$^{-1}$ and log(L$_{sGRB}$)$  = 44.96
+\pm 0.94$ erg s$^{-1}$ for the lGRB and sGRB samples respectively. GRB~111117A
+had a log(L) $= 44.95$ erg s$^{-1}$ at 10 ks which is very close to the peak of
+the sGRB luminosity distribution at 10 ks, but an outlier from the lGRG
+distribution, further supporting the short classification.
 
@@
 percent probability of being a sGRB. Compared to the other two sGRB candidates
-at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
-\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
+at high redshift, GRB~060121 ($T_{90}=1.97 \pm 0.06$~s)
+\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
+($17_{-15}^{+14}$ per cent) and GRB~090426 ($T_{90}=1.28 \pm 0.09$~s)
 \citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
@@
 \citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
-mergers.
+mergers. GRB~111117A is also securely classified as a sGRB according to the
+\citet{Horvath2010} classification scheme.
 
@@
 from the afterglow traverses a region of dense gas. Assuming a dust-to-metals
-ratio typical for the Milky Way, the X-ray derived $N_H$ corresponds to significant
-extinction along the afterglow sight line ($A_V > 10$) \citep{Predehl1995},
-which is contrasted to the absence of dust found for the SED fit.
-The discrepancy between the extinction derived from the GRB afterglow and the one obtained using galaxy-wide measures is also seen in the range of ratios found for lGRBs \citep{Perley2013a, Covino2013}.
+ratio typical for the Milky Way, the X-ray derived $N_H$ corresponds to
+significant extinction along the afterglow sight line ($A_V > 10$)
+\citep{Predehl1995}, which is contrasted to the absence of dust found for the
+SED fit and supported by the detection of \lya{}. The discrepancy between the extinction derived from the GRB afterglow
+and the one obtained using galaxy-wide measures is also seen in the range of
+ratios found for lGRBs \citep{Perley2013a, Covino2013}.
 
@@
 
-The detection of \lya{} is consistent with the SED-inferred absence of dust,
-despite the moderate stellar age which would suggest the opposite. The centroid
-of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
-with respect to systemic, which is similar to what is found for long GRB hosts
-\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.
+The simultaneus detection of \lya{} and \ha{} allows us to put constraints of
+the escape fraction of \lya{}, $f_{esc}$(\lya). Using the intrinsic ratio
+between \ha{} and \lya{}, assuming case B recombination
+\citep{Brocklehurst1971}, and the measured fluxes from the spectrum, we find
+$f_{esc}$(\lya)$ < 0.06$. While the $f_{esc}$(\lya) scales with the dust column
+\citep{Hayes2011}, the resonant scattering of \lya{}-photons with neutral
+hydrogen makes the effective path length of \lya{} is longer than \ha{}
+\citep{Atek2009}. This makes $f_{esc}$(\lya) an unreliable proxy for dust
+column, especially at low dust columns \citep{Atek2014} where the geometry and
+dynamics of the \hi{} within the galaxy will affect the \lya{} path the most.
+The $f_{esc}$(\lya) inferred for the host is entirely consistent with what is
+found for field galaxies with similar properties \citep{Oyarzun2017}. The same
+authors also find that \lya{} emitting galaxies have mostly little dust,
+consistent with what inferred from the SED-fit, see Sect. \ref{SED}. The
+centroid of the Ly$\alpha$ emission is found to be redshifted by
+$\sim$~220~km~s$^{-1}$ with respect to systemic, which is similar to what is
+found for long GRB hosts \citep{Milvang-Jensen2012a} and Lyman Break Galaxies
+\citep{Shapley2003a} in which the outflow is attributed to star formation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The redshift distributions of GRBs are powerful informants on both the
+conditions which drive the formation of these events, but also on the potential
+influence these cosmic explosions have on the evolution of the universe. Due to
+elevated brightness of lGRBs compared to sGRBs \citep{Berger2014} and their
+tendency to be associated with the most star-forming and therefore dense regions
+in their hosts \citep{Fruchter2006, Lyman2017}, the afterglows of lGRBs are
+easier to observe than their sGRB counterparts where only a single burst have a
+redshift measurement from the GRB afterglow \citep{Cucchiara2013,
+	DeUgartePostigo2014c}. Correspondingly, the redshift distribution of lGRBs is
+better determined than that of sGRBs \citep[e.g., see][]{Jakobsson2012a,
+	DAvanzo2015, Perley2016d}.
 
@@
 
-Using the complete, flux-limited selection of bursts from \citet{DAvanzo2014a},
-excluding GRB~111117A and the likely non-sGRB GRB~090426, we have a
-statistically homogeneous sample from which we can address the implications of
-the redshift of GRB~111117A. Out of the 14 hosts in the sample, 10 (71 per cent)
-have measured redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the
-host of GRB~111117A  is brighter than 80 per cent of the hosts with measured
+Using the complete, flux-limited selection of \textit{Swift}-detected bursts
+from \citet{DAvanzo2014a}, excluding GRB~111117A and the likely non-sGRB
+GRB~090426, we have a statistically homogeneous sample from which we can address
+the implications of the redshift of GRB~111117A. This sample includes sGRB
+originating in star-forming galaxies, elliptical galaxies, and apparent hostless
+sGRBs. Out of the 14 hosts in the sample, 10 (71 per cent) have measured
+redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the host of
+GRB~111117A  is brighter than 80 per cent of the hosts with measured
 $L_B/L^{\star}_{B}$. Even if we conservatively assume that \textit{all} the
@@
 = 2.211$. The two limits indicate that we would miss between 43 and 71 per cent
-of sGRB hosts at $z \sim 2$ due to host faintness.
+of \textit{Swift}-detected sGRB hosts at $z \sim 2$ due to host faintness.
 
@@
 	Ghirlanda2016}, meaning we could be missing a large fraction of sGRBs.
+
+A critical test to assess whether the power law delay time models can be
+accommodated is to check if the implied sGRB rate at higher redshift can be
+contained in the observed sGRBs without redshifts and the potential number of
+missed bursts due to \textit{Swift}'s sensitivity limit. Of the 65??? definite
+sGRBs observed by \textit{Swift} to date, 10 have redshifts \citep{GRBOX},
+meaning 85 per cent of all sGRBs observed with \textit{Swift} are missing
+redshifts. Additionally, \citet{Behroozi2014} parametrized the \textit{Swift}
+redshift sensitivity and found that mean detection probability for sGRBs at $z
+\sim2$ was only $\sim1$ per cent of the mean detection probability at $z \sim1$,
+assuming that the unknown beaming angle of sGRBs stays constant with time. What
+this means is that at the present, there is almost no limit on the number of
+sGRBs that could be at higher redshifts.
 
@@
 Using the new distance, the X-ray derived $N_H$ towards GRB~111117A is the
-highest within a complete sample of sGRB hosts and is consistent with the
+highest within a complete sample of sGRBs and is consistent with the
 $N_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
@@
 are missed due to the faintness of their hosts, a lognormal delay time model
-will be disfavored. Compared to a sample of sGRB hosts, GRB~111117A is more
-luminous than 80 per cent of a complete sample of sGRB hosts with measured
-luminosities. Assuming a host brightness redshift determination threshold, for
-between 43 and 71 per cent of sGRB hosts, we would be unable to determine a
-redshift should they be at a similar redshift as GRB~111117A. This could
-indicate that potentially, a significant fraction of sGRBs are at high $z$ but
-missed due to host faintness.
+will be disfavored. Compared to a complete sample of \textit{Swift}-detected
+sGRB, the host of GRB~111117A is more luminous than 80 per cent of sGRB hosts
+with measured luminosities. Assuming a host brightness redshift determination
+threshold, for between 43 and 71 per cent of sGRB hosts, we would be unable to
+determine a redshift should they be at a similar redshift as GRB~111117A. This
+could indicate that potentially, a significant fraction of
+\textit{Swift}-detected sGRBs are at high $z$, but missed due to host faintness.
 
@@
 We thank Mathieu Puech for testing the possible contribution from an older stellar population in the SED.
+We thank Peter Laursen for fruitful discussions regarding the \lya escape fraction.
 %	

### 2017-11-24 referee_reply.txt
Referee commnets final

We thank the referee for a very professional and constructive report. We have
done our best to take the suggestions into consideration and the message of the
paper has been strengthened as a consequence.

Kindest regards

Jonatan Selsing

------------------- Referee Report

The authors present a revised redshift for the short GRB 111117A that will have
a significant impact on the redshift distribution of short GRBs, and potentially
the inferred delay times and detectability of short GRB hosts. This makes the
host of GRB 111117A an outlier in several respects, and a benchmark event, which
naturally merits publication.

While the results are inherently very interesting, the paper suffers from lack
of attention to detail, which are summed up in several major and minor comments
that I have listed below. I feel the authors should address these or explain
their reasoning otherwise before publication.

Major comments: --------------

1) VLT spectrum identification of emission lines: The revised redshift of
z=2.211 relies on the detection of several emission lines in the VLT/X-Shooter
spectrum. Because this is the main point of the paper, the reader really needs
to be convinced that this redshift is correct. While it is clear from Figure 2
that [OIII] 5007 is detected, the determination of the redshift relies on
multiple lines. It is also not obvious from the unbinned spectrum shown in Fig.
3 that the lines are actually detected, because the SED fit line is much
thicker. Thus, the following points should be addressed to make a stronger case
for why the authors think this redshift is correct:

-The authors should show the 2D images (as they have done for [OIII]) for all
lines that they identify.

JS: We agree and we have added a panel showing all the detected lines, which
highlights the certainty of the spectroscopic redshift.

-While it is good to show the unbinned spectrum, I suggest either binning the 1D
spectrum to convince the reader that there are actually emission lines present,
or providing plots that zoom in at the locations of the lines. I suggest
adjusting the plot such that the reader can see the actual data (as currently
one can only see the model due to the thickness of the line).

JS: While we agree that it would be nice if the presence of the lines could be shown
in the 1D spectrum, the lines are too narrow and the S/N of the spectrum too low
for the individual lines to stand out clearly in the full 1D-spectrum,
independently of the binning chosen. We believe that the newly added panel of
the lines in the 2D-spectra convincingly demonstrate the robustness of the
inferred redshift and a corresponding panel for the 1D-lines would not add
valuable information.

-What is the uncertainty in the redshift?

JS: An attempt to attach an uncertainty to the redshift has been added, based on
individual redshift measurement using each of the detected lines.

-The 1D error spectrum should be plotted and the positions of the emission lines
in the 1D spectrum should be labeled.

JS: We agree and this has been included in the spectrum.

2) Deduction of star formation rates:

-When calculating the SFR from H-alpha, it is not clear why this is a lower
limit if the authors have corrected H-alpha for the missing flux.

JS: We have clarified this, which was stated incorrectly in the text. When the
deriving the SFR, we did in fact not correct for the missing flux of H-alpha,
but only used the region free from contamination. This also explains why the
inferred SFR is a limit.

-For both H-alpha and OIII, the fits to the line profiles (after
masking/correcting for the sky lines) should be shown.

JS: Because the profiles are not gaussian or particularly well behaved and given the
low S/N of the spectrum, we have not fitted any of the lines. This is now stated
explicitly in the text. Additionally, OIII was never used to infer the SFR and
this has been cleared up in the text.

-The text currently says, "From the integrated H-alpha line, we estimate SFR =
18 +/- 3 Msun/yr." Is this a typo and H-alpha should really be OIII? If that's
the case, the OIII line is very sensitive to the ionization parameter. The
authors should state what ionization parameter they are assuming, and also
reference any caveats in using OIII as a SFR indicator. The authors should also
make clear how the uncertainty in the OIII SFR was derived.

JS: In the original version of the text, we were referring the h-alpha, although the
text was ambiguous. We have now decided to stick to only the lower limit,
because it was unclear how the systematics of the procedure affected the
uncertainty in the reported error.

-Finally, even though weakly detected, it would be interesting to know if the
derived SFR from H-beta is consistent with the results from H-alpha and OIII.

JS: While certainly interesting, this was also attempted to constrain the reddening
through the use of the Balmer decrement, but unfortunately the large uncertainty
on the H-beta flux prevents obtaining useful information.

3) SED fit: The authors demonstrate a best-fit model of an unreddened galaxy
template with various best-fit parameters for the stellar population properties.
What is the goodness-of-fit parameter used? If it is chi^2 or some other
statistic, the reduced chi^2 (or equivalent) should be stated. The range of
models that were considered (in terms of stellar mass, stellar population age,
and star formation rate), should be included. Was a metallicity assumed, and if
so, what metallicity and why? Finally, how do these inferred properties compare
to previous works?

JS: The SED fit is a template fit, and not a parametric fit. Hence, the number of
"degrees of freedom" is ill-defined, and the reduced chi^2 test does not apply
(e.g. see Andrae et al. 2010 - https://arxiv.org/abs/1012.3754). We have updated
and expanded the text to clarify the procedure. We refer to previous work,
where all the details can be found. In particular we use 4 different input
metallicities, which we have now explicitly stated in the text. We have not
compared the inferred values to previous works, since the previous SED fits are
invalidated by the revised redshift.

4) Photometry: The authors mention that the revised z-band measurement is the
primary reason for why the redshift was previously derived incorrectly by two
different groups. It is interesting and worrisome that the two previous works
(Margutti+ and Sakamoto+) both derived the same "wrong" z-band photometry based
on observations from entirely different facilities (Margutti+ with Gemini, 23.04
+/- 0.18 mag vs. GMG, z=22.9 +/- 0.3 mag).

JS: Indeed the photometric redshift provided by the two previous papers relies on
the identification of the 4000 AA break constrained by the overestimated z-band
flux. We did worry about the consistency of the two previous measurements, both
inconsistent with ours. We do not have access to the GMG data, but we did
download and reduce the Gemini data from which we measure the reported
value/limit. Objects of magnitude z ~ 23.2 are clearly seen and are
significantly brighter than the GRB host galaxy, which is only marginally (if at
all) detected. We did contact both Sakamoto and Margutti and relayed our finding
to them, which they acknowledged, but did not rebut. We do not know what
affected the previous measurements, but we are not in position where we can
answer this. For the GMG data, taken at early times, potentially an afterglow
contribution could influence the measurement as noticed by those authors. We
provide all the reduced imaging data used at
https://github.com/jselsing/GRB111117A, from which the significance of the
detection can be visually confirmed.

-The revised Gemini z-band photometry in Table 1 has an uncertainty of 0.47 mag,
a 2.1-sigma detection, which in most cases would be quoted as an upper limit. Is
there actually visible host emission in the Gemini image, and how was photometry
performed on an object with such low significance? Was the center of the
aperture forced to be at a position where the host is well-detected in a
different band? The authors should state what size aperture they used relative
to FWHM as well as the background region they used, not only for the Gemini
z-band photometry but for all of the photometry listed. In fact, if the Gemini
z-band was quoted properly as a 3- or 5-sigma upper limit, would it still be
consistent with the VLT/FORS2 measurement by the authors?

JS: We agree that the methodology was not adequately explained in the paper and we
have expanded the text regarding the photometric methodology, especially
commenting on the discrepancy between the z-band measurements. For the Gemini
image, we find a limiting 3(5)-sigma upper limit of z > 24.06(23.50) which is
consistent with the FORS measurement within the errors and with our 2-sigma
detection in the GMOS-N image. Note that it is not our intent to claim a
detection in the GMOS-N image as we have already a robust detection in the FORS
image. It is however reassuring that an object of consistent magnitude is not
excluded by the Gemini data. We have added a mosaic of all the available bands,
which shows the significance of the detection in the different bands. A faint
object can be glimpsed at the position of the host in the Gemini panel, visually
consistent with the quote significance of the detection.

-The authors state that the z-band photometry is discrepant from previous works
while other filters are in excellent agreement. A careful comparison between
this work's Table 1, Margutti+ and Sakamoto+ demonstrates that there are
discrepancies not only in Gemini z-band, but in several other bands as well. For
instance, the J-band host measurement reported by the authors is 0.4-5 mag
fainter than that quoted by Sakamoto+ (after correcting for AB to Vega
conversions). The R-band magnitude is 0.4 mag fainter than Sakamoto+, the r-band
magnitude is fainter by 0.2 mag, i-band is 0.25 mag fainter than Margutti+, and
so on. The overall trend is that the host photometry quoted by the authors is
systematically fainter than two previous and separate works. This is worrisome
that two groups got the same wrong answer using different data sets, and should
be addressed by the authors in more detail. Is there something different about
the aperture used or calibration system (SDSS vs. another system, used between
the current work and two previous works that could help explain the
systematically fainter fluxes?

JS: First, please note that values presented in the table are not corrected for the
expected Galactic foreground, contrary to previous works (this amounts to 0.12 /
0.07 / 0.05 in the g/r/i bands). According to what stated in the previous
papers, photometry was calibrated in all cases against SDSS and 2MASS for griz
and JHK filters. In the case of R-band only, Sakamoto et al. use the USNO
catalogue which has a poor photometric accuracy where we use the Lupton color
transformation from SDSS. For the J-band, the discrepancy is 0.4 +- 0.3
(Sakamoto) and 0.5 +- 0.4 (Margutti); for the K-band the discrepancy is 0.35 +-
0.3 (Sakamoto) - all consistent within ~ 1 sigma. For the R-band, our value is
~0.4 +- 0.2 mag fainter than Sakamoto (USNO based calibration), while in the
nearly overlapping r-band we are 0.33 +- 0.15 mag *brighter* (Sakamoto) and 0.15
+- 0.15 mag *fainter* (Margutti). For the i-band we are 0.08 +- 0.3 mag
*brighter* (Sakamoto) and 0.18 +- 0.23 *fainter* (Margutti). The overall picture
is that our values are not systematically fainter (nor brighter) than the
published ones. Most differences are at the 1-sigma level, in some cases our
value is intermediate between the published ones. The numerically large
discrepancy in the infrared stems from the especially large uncertainties, which
are properly accounted for in the fitting process. We agree though that our
statement of an "excellent" agreement was too optimistic, so we have rectified
that statement to saying thar the measurements agree within the errors. Given the
arguments above, we do not feel that a point-by-point discussion of these
discrepancies contributes to the clarity of the message in the paper. As stated
above, though, we have added further explanations on our photometric methodology.

5) NH, AV and Ly-alpha: The high NH, lack of AV, and presence of Ly-alpha is an
interesting aspect of this host. However, the lack of reddening seems more like
an assumption than proven by the data. I have a few questions for clarification:

JS: First we remark that the NH is derived from the X-ray spectrum of the afterglow
and is therefore a sight-line-measurement, where the AV and Lya are integrated
measures. This is particularly relevant given the sizeable offset between the
GRB location and the host center. Therefore there does not need to be a direct
relation between these numbers. The host reddening is a parameter that is fit
for in the SED-fit and thus constrained by the photometry. This is now stated in
the expanded description of the SED fitting.

-The high inferred intrinsic NH is interesting but also matches the evolution of
NH with redshift nicely. There are relations between NH and AV - what is the
derived value for AV and is it consistent with no reddening?

JS: While certainly interesting, the proposed comparison is not straightforward.
Several studies have compared the afterglow-derived extinction with the
afterglow-derived hydrogen equivalent column density. Comparison between the
integrated (luminosity weigthed) host extinction and hydrogen column from X-rays
has not been attempted to our knowledge. In the case of long GRBs, Perley et al.
(2013, ApJ 778 128, their Figure 15) have compared the extinction measured along
the afterglow sightline with the average extinction measured in the host from
SED-fitting, finding consistency between the majority of points within a factor
of a few, but with a large number of outliers. To our knowledge, a corresponding
study has not been carried out for short GRB environments. We have added a small
section discussing this, but we believe this point can be left future works.

-How does the inferred NH compare to previous works from Sakamoto and Margutti?

JS: Because of the higher redshift used, the intrinsic NH derived for our work
is correspondingly higher. We do not feel it is necessary to state this directly in the text.

-A galaxy can still have detectable Ly-alpha in the presence of dust as it
depends on the escape fraction. Are the authors able to make any quantitative
statments on the Lyman-alpha escape fraction, and place a limit on the dust
column?

JS: We have added a section describing the contraints on the Lya escape
fraction and the implications for the dust column. Note, we do not explicitely
calculate a dust column based on the escape fraction, because the dust in the
host is constrained by the photometry.

-Overall, how does the assumption of zero reddening affect the conclusions of
the paper? For instance, if a reddened model is assumed in the SED fit, how
drastically does this change the stellar population parameters?

JS: Zero reddening is not an assumption, but something that is constrained by the
photometry. It is a fit parameter in the SED-fit. We have added caveats to the
SED-fit section.

6) Long versus short GRB classification: The authors claim that GRB 111117A was
not a collapsar, despite the host having a relatively high star formation rate
and potentially high-density local environment. The reader should be convinced
that this is actually the case and some suggestions are below.

-The classification from Bromberg+ 2013 does not seem to match the conclusions
from host galaxy properties (see Berger+ 2014; Fong+ 2013), and is not a very
reliable metric for whether or not an event was a collapsar. While the authors
state that the burst is consistent with zero lag, the authors should state basic
information about the burst, such as duration, hardness, limit on extended
emission, etc to make the case that this event was in fact a short hard GRB.

JS: We can appreciate this point, and have added the basic high-energy properties of
the burst, calculated by previous authors, which clearly marks it as a sGRB.

-The mention of the intermediate class of bursts is interesting, but what is GRB
111117A with respect to this class? Are the host properties of 060121 and 090426
different than those of more definite short GRBs?

JS: Also according to the classification of Horvath et al. 2010, GRB111117A is a
short GRB. This is now stated exclicitely in the text.

7) Implications for delay time distribution: The high redshift of GRB 111117A
indeed has some important implications on the detectability of hosts and
potential delay time distribution. This point of discussion merits a bit more
detail and the following are points to consider.

-The authors state that about half (45-60%) of short GRB hosts if they were at
z=2.211 would have been missed based on luminosity arguments. This is a bit
misleading since the sample of short GRB hosts with redshifts is inherently
incomplete due to Swift detection bias as well as difficulty of following every
short GRB sightline. Do the D'Avanzo+ 2014 and Berger 2014 samples include
elliptical host galaxies where the majority are L* or brighter? If the authors
only include emission line hosts, that actually neglects ~30% of the short GRB
population which come from elliptical galaxies.

JS: The D'Avanzo+ 2014 sample is chosen because it is flux-limited, but otherwise
complete, meaning only unbiasing selection criteria has been imposed on it. It
should therefore correctly reflect the full, underlying sGRB population down to
some sGRB brightness level. The Berger 2014 sample is not complete, which is why
no conclusions are drawn based on this sample. And yes, both samples also
consider the sGRBs originating is elliptical host galaxies. These points have
been clarified a bit in the text.

-Long GRBs originate in star-forming galaxies with rates comparable to that
inferred for the host of GRB 111117A, while long GRB host luminosities are
generally much lower than short GRB hosts. In the scenario that we are missing
many short GRB redshifts because they originate at z~2, why are we routinely
able to detect long GRB hosts to high redshifts, even when they have lower
luminosities?

JS: This is because the large majority of lGRB redshifts (~70 %) come from GRB
afterglow spectroscopy. Only a single sGRB redshift has been found from
afterglow spectroscopy (GRB 130603B). We have added a small section to highlight
the difference in redshift determinations of the two classes of bursts.

-Given the observed short GRB redshift distribution (many at z<1.3, and only a
handful at higher redshifts of z>1.5), it still seems very plausible that the
log-normal delay time distribution is correct. However, if the authors are
correct that the redshift of GRB 111117A is in fact close to the peak of the
short GRB rate and the power-law delay time model is preferred, what is the
implied rate at low redshifts of z<1? Is this consistent with the observed short
GRB rate out to z~1 (which should be fairly spectroscopically complete)? Are
there enough Swift short GRBs without redshift measurements that is consistent
with an explanation that the majority of events actually occur at z~2?

JS: This is a very interesting question and definitely something that merits some
consideration. We have added a section to address this.

Minor comments: ----------------

Abstract: -"between 43 and 71 percent of all sGRB redshifts should be missed" --
what do you mean by all sGRB redshifts? all sGRBs that occur in the Universe,
all of those detected by Swift, or all based on the sample of short GRBs with
known hosts? I think it is the latter, but I would definitely clarify in the
abstract and when you get to this section in the text.

JS: We have added that this is for the Swift-detected bursts, both in the abstract and in the text.

Section 1:

--1st sentence: Define sGRB, NS. 
JS: Done.

--For detection of kilonovae, also cite Berger+ 2013. 
JS: Definitely.

--"duration T_90 < 2" - make sure to qualify with 2 sec
JS: Done.

--Here and throughout, there is a tendency to use sGRB/short GRB and lGRB/long
GRB. I suggest using one or the other throughout the text. 
JS: We have streamline this a bit, but have kept the wording "long and short GRBs"

--"The bulk of these localizations have associated galaxies at relatively low redshifts... it is
arguably biased towards lower redshifts" -- This is not necessarily true, as the
low redshifts could be a natural consequence of relatively long delay times.
JS: We have cautioned the wording.

--"is tracing with host stellar mass rather than star-formation rate" --
Previous analysis of host galaxy properties have actually found that the short
GRB rate likely traces both stellar mass and star formation (see Berger 2014 and
references therein) 
JS: We have changed the wording to reflect this.

--"will provide new a promising ways" -- delete "a"
JS: Done

Section 2: --The authors mention there were 4 separate epochs of spectroscopy -
is any spectral evolution apparent? This would at least be consistent with the
assumption that there is negligible afterglow contribution from epoch 1. 
JS: We do not see any spectral evolution. This is now explicitly stated in the text.

--Since the spectra were taken across different nights and under different seeing
conditions, details of flux calibration for the combined spectrum should be
included. 
JS: We have added a description of this.

--Is it surprising that [OIII] 4959 is not detected, given such a
strong detection of [OIII] 5007? 
JS: [OIII] 4959 is detected, but very close to a sky-line. This has now been added to the text.

--"...because the slit is not aligned along the
long axis of the host, this value is an upper limit" - I believe this should be
a lower limit 
JS: Correct! This has been changed and expanded.

--In calculating beta_OX, state the actual limits that were used
and in what band. Are these the deepest limits available? 
JS: To our knowledge, these are the deepest limits available, yes. The limits have now been explicitly stated.

--"Both of these
authors report z-band photometry that is brighter by 0.8 mag to 1.0 mag compared
to our value" -- Are the authors comparing their Gemini z-band to z-band
measurements in the other papers? Or the VLT measurement to the measurements in
other papers? If the former, the previous values are actually brighter by
1.2-1.4 mag; if the latter, the values are discrepant by 0.7 to 0.9 mag. I could
not figure out the quoted values of 0.8 to 1.0 mag.
JS: We have changed this to ~ 1 mag. Both our z-band measurements are fainter than the z-band magnitudes reported by other authors.

Fig. 3: best-fist --> best-fit
JS: Done

Section 3: --"The sub-samples comrprise of 333 long, 19 short..." - the papers
where the data came from should be cited. 
These references were in the caption, but have now been repeated in the text. 
--"had a luminosity of 44.95" --> should appear as a power of 10 and with units 
JS: This has been updated.

--"GRB 111117A has a 96% probability of being a sGRB" 
JS: Done

-- The terms sGRB and non-collapsar seem to be used interchangeably here. From the burst properties it certainly
seems like a sGRB. To be more accurate I would say "non-collapsar"
JS: We have removed the usage of non-collapsar and stuck with sGRB.

--"... it occurred when the universe was younger by 1.8 Gyr compared to any
other non-collapsar GRB ever detected"
JS: Done

-- again, how do you find non-collapsar? There is no way to know for certain and
the Bromberg+ 2013 method has been proven unreliable from host galaxy
properties. If you do want to rely on Bromberg+ 2013, what percent probability
do you consider as a "non-collapsar GRB"?
JS: We have stated a f_NC > 50 per cent requirement of the Bromberg classification to consider a given burst short.

--"This number is 3.2 Gyr for the next-highest spectroscopic redshift" - This
sentence needs a citation and would be helpful to state what burst this is
JS: We have expanded this entire section a bit to be more complete.

--Even though the offset does not change much, it is worth recalculating in this
paper for completeness.
JS: Done

--"Along with the absence of dust" -- This is not yet proven in this paper. I
suggest revising to "If the assumption of no dust is correct..."
JS: This has been changed to "low dust content"

--The scenario in which the density of material is high to explain the high NH
would imply a bright afterglow, although would appear faint at z=2.211. Are the
afterglow limits still consistent with a high-density environment at z=2.211?
The X-ray afterglow is not strongly dependent of the density and because of the optical
darkness, dust extinction locally to the GRB can accomodate a large brightness
interval. We have mentioned this in text.

Fig. 4: A comparison of the light curve for GRB 111117A to the identical figure
in Margutti+ (their Fig 6) looks like the luminosity does not change, even
though it should be significantly higher given the revised redshift. In fact,
the Margutti+ light curve exceeds 10^48 erg s^-1 near the start, while the light
curve in Fig. 4 does not. Is the luminosity calculated correctly, or are they in
different energy bands? -Does this include the Chandra data as well? If so, it
would be good to include that in the caption.

Margutti didn't use the light curve from the Swift Burst Analyser. Her early light curve is ~0.18 dex brighter.
The shape is also a bit different. k-correction: Margutti fixed the photon index
to 2.2. The Burst Analyser provides estimates of the photon index for each
observation. At the time of the first observation, the Burst Analyser gives a
photon index of 0.93. The k-correction for a power-law shaped spectrum is:
(1+z)^(Gamma-2). The difference in these parameters corresponds to a difference
in the k-correction of 0.63 dex. The different distances gives a luminosity
difference of 0.58 dex (differences in the cosmology are taken into account).

All in all, logL(us) - logL(Margutti) = (0.58 - 0.18 - 0.63) dex = -0.23 dex.

Comparing her plot with ours gives difference of ~-0.2 dex, which is identical to the expected difference.

Fig. 5: What are the lines and what is the red curve that appear on the top and
side of the plot? 
JS: The caption has been expanded to better explain the different
components.

Section 5: -"this places a hard upper limit on the initial
separation..." - Is there a citation for the equation used to calculate this?
JS: The equation is in the Postnov reference. The actual calculation is included in
the online repository, with reference to the exact equation.

-This may be outside the scope of this work, but how does
the initial separation change under different mass assumptions (unequal mass
binary, or more/less massive binary)? 
JS: We have included the calculation for different mass assumptions and changed the text accordingly.

-The authors should note that this
discussion assumes that GRB 111117A did not form dynamically in a globular
cluster as suggested by some works (e.g., Lee+ 2010; Church+ 2011)
JS: We do not assume anything about the formation channel, only that the neutron
star binary was already in place at the beginning of the universe. We have for
completeness mentioned the possible dynamical origin of neutron star binaries.

Section 6: -"The high NH is difficult to reconcile with the large projected host
offset" - How large is the offset at z=2.211? -"for between 43 and 71 percent of
sGRB hosts" -- should be revised to "sGRBs with known or well-characterized
hosts"
JS: We have changed the wording.
-------------------

### 2017-11-24 GRB111117A.tex
Referee commnets final

@@
 benchmark for redshift determination, we find that between 43 and 71 per cent of
-all \textit{Swift}-detected sGRB should be missing redshifts due to host
-faintness for bursts at $z\sim2$. The high redshift of GRB~111117A is evidence
-against a lognormal delay-time model for sGRBs through the predicted redshift
-distribution of sGRBs, which is very sensitive to high-$z$ sGRBs.
+all \textit{Swift}-detected sGRBs are too faint at $z\sim2$ to allow for a secure
+redshift determination. The high redshift of GRB~111117A is evidence against a
+lognormal delay-time model for sGRBs through the predicted redshift distribution
+of sGRBs, which is very sensitive to high-$z$ sGRBs.
 
 From the age of the universe at the time of GRB explosion, an initial neutron
-star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
+star (NS) separation of $a_0 < 3.1~R_\odot$ is required in the case where the
 progenitor system is a circular pair of inspiralling NSs. This constraint
@@
 
-There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs)
-come from the merger of neutron stars (NSs), either with another NS, or a black
-hole, due to their apparent association with kilonovae \citep{Barnes2013a,
-	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
-supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
-supports this idea and distinguishes the physical origin of sGRBs from their
-long-duration counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.
+There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs) come
+from the merger of neutron stars (NSs), either with another NS, or a black hole,
+due to their apparent association with kilonovae \citep{Barnes2013a,
+	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. This association was
+recently confirmed by the simultaneous and co-spatial detection of gravitational
+waves (GWs) from binary neutron star merger and a sGRB
+\citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}. The
+absence of associated supernovae in deep searches
+\citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011} supports this idea and
+distinguishes the physical origin of sGRBs from their long-duration
+counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.
 
-The classification of GRBs in two groups, initially comes from the bimodal
-distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
+The classification of GRBs in two groups initially comes from the bimodal
+distribution of burst duration and spectral hardness in the BATSE sample \citep{Kouveliotou1993},
 where the duration T$_{90} < 2$ sec has been regarded as the dividing line between long
-and short GRBs. Additionally, it has been found for lGRBs that there is a
+and short GRBs. Additionally, it has been found for long GRBs (lGRBs) that there is a
 spectral lag in the arrival-time of photons, with the most energetic ones
@@
 classes. For this reason, the distinction between long and short GRBs is
-preferably based on a combination of high-energy properties \citep{Zhang2009,
-	Bromberg2012a, Bromberg2013}.
+preferably based on a combination of the high-energy properties \citep{Zhang2009,
+	Kann2011, Bromberg2012a, Bromberg2013}.
 
@@
 The bulk of these localizations have associated galaxies at relatively low
-redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, and because
+redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, but because
 most of these measurements come from the associated hosts, it is potentially
-biased towards lower redshifts due to Malmquist bias. The host galaxies of sGRBs
-are diverse. They are more massive and less actively star-forming on average
-than lGRB hosts \citep{Fong2013b}, while in some cases, no host galaxy can be
-identified above detection threshold of deep follow-up observations
-\citep{Berger2010a, Tunnicliffe2014}. Together with their position within their
-hosts \citep{Fong2013a}, this suggests a progenitor system that can be very long
-lived in comparison to lGRBs, with the host stellar mass affecting the sGRB
-rates more than star-formation rate (SFR) \citep{Berger2014}.
+biased towards lower redshifts. Additionally, because the \textit{Swift}
+sensitivity to sGRB decreases sharply with redshift \citep{Behroozi2014}, the
+intrinsic redshift distribution of sGRBs is largely unknown at higher redshifts.
+
+The host galaxies of sGRBs are diverse. They are more massive and less actively
+star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases,
+no host galaxy can be identified above the detection threshold of deep follow-up
+observations \citep{Berger2010a, Tunnicliffe2014}. Together with their position
+within their hosts \citep{Fong2013a}, this suggests a progenitor system that can
+be very long lived in comparison to lGRBs, with the host stellar mass affecting
+the sGRB rates more than star-formation rate (SFR) \citep{Berger2014}.
 
@@
 localize NS mergers, which holds the promise for a detection of an associated
-gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
-of a sGRB and a GW will provide new promising ways to constrain the binary
-inclination angle \citep{Arun2014} and measure cosmological distances
-\citep{Nissanke2010}.
+gravitational wave (GW) signal \citep{Ghirlanda2016}. This epochal breakthrough
+occurred recently when the first NS-NS GW event ever was detected by LIGO/Virgo
+(GW 170817) and associated to the weak sGRB 170817A detected by the Fermi and
+Integral satellites \citep{LIGOScientificCollaboration2017a, Goldstein2017,
+	Savchenko2017}. The simultaneous detection of a sGRB and a GW provides new
+promising ways to constrain the binary inclination angle \citep{Arun2014, LIGOScientificCollaboration2017a} and
+measure cosmological distances \citep{Nissanke2010, LIGOScientificCollaboration2017c}.
 
@@
 Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
+%	\label{fig:line}
+%\end{figure}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=18cm]{figures/grid.pdf}
+	\caption{2D-images of the emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha.
+The location of a bright skylines are marked by the blue boxes.
+The location of the emission line is indicated with the red ellipse.
+Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.
+For the upper, left panel containing \lya, the systemic redshoft position of \lya is marked by a yellow, vertical, dashed line.
+The red part of the \oii$\lambda$3727-doublet is affected by atmospheric absorption.}
 	\label{fig:line}
-\end{figure}
+\end{figure*}
 
@@
 ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
-091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
+091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, with 5
 arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
@@
 each pixel is chosen as median variance spectrum of the region surrounding that
-pixel, thus avoiding to base the weight on the pixel variance. An attempt to
-correct for slit-losses has been carried out on the combined spectrum based on
-the average seeing of the observations. We show the extracted spectrum in
-Fig.~\ref{fig:SED}.
+pixel, thus avoiding to base the weight on the pixel variance. Slit-loss
+correction has been applied on the combined spectrum based on the average seeing
+of the observations. We show the extracted spectrum in Fig.~\ref{fig:SED}.
 
@@
 
-We determine a redshift of $z = 2.211$ from the simultaneous detection of
-emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959,
+We determine a redshift of $z = 2.211 \pm 0.001$ from the simultaneous detection
+of emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959,
 \oiii$\lambda$5007, and \ha. \hb{} and \oiii$\lambda$4959 are both detected at
-low significance ($\sim$ 3-$\sigma$). We show \oiii$\lambda$5007 in the insert
-in Fig.~\ref{fig:line}.  \ha{} is only visible in the first epoch due to the
+low significance ($\sim$ 3-$\sigma$). The uncertainty on the redshift is the
+standard deviation of independent measurements of the redshift based on the
+individual lines. We show \oiii$\lambda$5007 in the insert in
+Fig.~\ref{fig:line}.  \ha{} is only visible in the first epoch due to the
 $K$-band blocking filter used for the remaining observations. The nebular lines
@@
 further supports the inferred redshift. No spectral evolution is observed across
-the epochs, supporting the lack of afterglow contribution.
+the epochs indicating that there is negligible GRB afterglow contribution to the
+spectra.
 
@@
 converting the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial
-mass function using \citet{Madau2014}, we derive a limit of $SFR > 7
-M_\odot$~yr$^{-1}$. We additionally obtain the \lya{} line flux by numerically integrating
-the entire \lya{} line complex and obtain $F_{\mathrm{Ly}\alpha} = 2.0 \pm 0.5
-\times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$.
+mass function using the conversion factor from \citet{Madau2014}, we derive a limit of $SFR > 7
+M_\odot$~yr$^{-1}$. We additionally obtain the \lya{} line flux by numerically
+integrating the entire \lya{} line complex and obtain $F_{\mathrm{Ly}\alpha} =
+2.0 \pm 0.5 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$,
+where the error on the flux is found by integrating the associated error
+spectrum over the same spectral region.
 
@@
 The multi-color SED is fit using the \citet{Bruzual2003} single stellar
-population models based on a \citet{Chabrier2003} with initial mass function in
+population models (SSPs) based on a \citet{Chabrier2003} with initial mass function in
 \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the
@@
 
-Because the projected distance does not change significantly between $z = 1.3$ and $z =
-2.211$, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013}
-relating to host offset are unaffected.
+\citet{Margutti2012} find an projected offset between the host nucleus and the
+site of GRB~111117A of 1\farc25 $\pm$ 0\farc20 arcsec and \citet{Sakamoto2013}
+find a corresponding offset of 1\farc0 $\pm$ 0\farc20 arcsec. This corresponds
+to a projected physical offset at $z = 2.221$ of $10.6 \pm 1.7$ kpc and $8.5 \pm
+1.7$ kpc respectively. Because the projected distance does not change
+significantly between $z = 1.3$ and $z = 2.211$, all conclusions of
+\citet{Margutti2012} and \citet{Sakamoto2013} relating to host offset are
+unaffected.
 
@@
 
-Based on the BAT light-curve, $T_{90}=0.46$~s which is shorter than both the
-prototypical 2 s \citep{Kouveliotou1993} and the $\sim$0.7 s suggested to also
+Based on the BAT light-curve, $T_{90}=0.46$~sec which is shorter than both the
+prototypical 2~sec \citep{Kouveliotou1993} and the $<0.8$~sec suggested to also
 exclude the shorter tail of the \textit{Swift}-observed lGRB population
@@
 \citep{Sakamoto2013}. A spectral lag of $0.6 \pm2.4$ ms, consistent with zero,
-is not uniquely associated to sGRBs \citep{Bernardini2014}, but in conjunction
-with the duration and the spectral hardness \citep{Sakamoto2011}, as pointed out
+is not uniquely associated to sGRBs \citep{Bernardini2015}, but in conjunction
+with the duration and the spectral hardness \citep{Sakamoto2011}, as already pointed out
 by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is securely
@@
 and hardness ratio, do not depend strongly on redshift \citep{Littlejohns2013a},
-the updated redshift does not change this designation. The intrinsic spectral
-lag shortens, but since it is already consistent with zero, this does not affect
-the classification.
+the updated redshift does not change this designation.
 
@@
 long GRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
-distribution at 10 ks. The sub-samples comprise of 333 lGRBs, 19 sGRBs, and
+distribution at 10 ks. The sub-samples comprise of 402 lGRBs, 31 sGRBs, and
 GRB~111117A. The sample of lGRBs are from \citet{Evans2007, Evans2009} and the
@@
 \pm 0.94$ erg s$^{-1}$ for the lGRB and sGRB samples respectively. GRB~111117A
-had a log(L) $= 44.95$ erg s$^{-1}$ at 10 ks which is very close to the peak of
+had log(L) $= 44.95$ erg s$^{-1}$ at 10 ks which is very close to the peak of
 the sGRB luminosity distribution at 10 ks, but an outlier from the lGRG
@@
 
-\citet{Bromberg2013} investigated the degree to which the long and short
-population distributions overlap and quantified the certainty in class
-membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
-percent probability of being a sGRB. Compared to the other two sGRB candidates
+The separation of GRBs in two distinct classes based on their high-energy
+observables points to an intrinsically different physical origin. LGRBs are
+typically modeled as collapsars \citep{MacFadyen1999} in which a single, massive
+star undergoes gravitational collapse, where the currently preferred model for
+sGRBs is the merger of two neutron stars \citep{Eichler1989, Nakar2007}.
+\citet{Bromberg2013} investigated the degree to which high-energy observables of
+the long and short GRB populations overlap and quantified the certainty in class
+membership. According to \citet{Bromberg2013}, GRB~111117A has a $96_{-5}^{+3}$
+percent probability of being a sGRB. Compared to two other sGRB candidates
 at high redshift, GRB~060121 ($T_{90}=1.97 \pm 0.06$~s)
@@
 
-The redshift and secure classification of GRB~111117A mean that it occurred
-when the universe was younger by 1.8 Gyr compared to any other non-collapsar GRB
-ever detected. This number is 3.2 Gyr for the next-highest spectroscopic
-redshift. If the merger of NSs is the primary agent for the $r$-process element
-enrichment of the universe \citep{Goriely2011, Ji2016, Komiya2016}, this
-marks the earliest detection of this process.
+A number of other GRBs are claimed to be short and at high ($z > 0.9$) redshift.
+If we consider bursts with a probability of being short, $f_{\mathrm{NC}} > 50$ per cent according to the
+\citet{Bromberg2013} classification scheme as sGRBs, 5 sGRB are found.
+GRB~051210 at $z \sim 1.3$ ($f_{\mathrm{NC}} = 82_{-61}^{+10}$ per cent)
+\cite{Leibler2010},
+GRB~060801 at $z = 1.131$ ($f_{\mathrm{NC}} = 95_{-5}^{+3}$ per cent)
+\cite{Berger2007}
+GRB~070714 at $z = 0.923$ (no $f_{\mathrm{NC}}$ due to extended emission)
+\cite{Graham2009}
+GRB~090510 at $z = 0.903$ ($f_{\mathrm{NC}} = 97_{-29}^{+1}$ per cent)
+\cite{McBreen2010}
+GRB~100117 at $z = 0.915$ ($f_{\mathrm{NC}} = 97_{-3}^{+1}$ per cent)
+\cite{Fong2011} and although simultaneous secure host association and certain
+redshift estimate is difficult, there does seem to be a number of sGRBs at $z
+\sim 1$.
+
+This securely makes GRB~111117A, by far, the highest redshift sGRB detected to
+date. The redshift and classification of GRB~111117A means that it occurred when
+the universe was younger by $\sim$ 3 Gyr compared to any other sGRB ever
+detected. If the merger of NSs is the primary agent for the $r$-process element
+enrichment of the universe \citep{Goriely2011, Ji2016, Komiya2016}, this marks
+the earliest detection of this process.
 
@@
 The thick green line is GRB~111117A.
-Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and subluminous compared to the lGRB population.}
+Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and subluminous compared to the lGRB population.
+The insert shows the respective X-ray luminosity distributions of sGRBs and lGRBs at 10ks, indicated by the vertical dashed line.}
 	\label{fig:sxray_lightcurve}
@@
 GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
-$N_H$ and highest redshift of all sGRBs. The short sample, excluding
+$N_\mathrm{H}$ and highest redshift of all sGRBs. The short sample, excluding
 GRB~111117A, is located at low redshifts ($z < 1$) and is found to populate a
@@
 \citep{DAvanzo2014a}. The inferred hydrogen column density for GRB~111117A seems
-to follow the trend with increasing $N_H$ as a function of redshift as found for
+to follow the trend with increasing $N_\mathrm{H}$ as a function of redshift as found for
 the lGRB afterglows \citep{Campana2010, Starling2013, Arcodia2016}. This is
-related to what is found by \citet{Kopac2012, Margutti2013}, that $N_H$ seems to
+related to what is found by \citet{Kopac2012, Margutti2013}, that $N_\mathrm{H}$ seems to
 be comparable for long and short GRBs when compared at similar redshifts.
 
-The redshift evolution of $N_H$ in the hosts of lGRBs is not reproduced by
-\citet{Buchner2017}, using a different $N_H$ inference methodology. Instead a
-correlation between $N_H$ and host stellar mass is suggested. Assuming that the
-different $N_H$-fitting methodologies yield comparable results, GRB~111117A is
+The redshift evolution of $N_\mathrm{H}$ in the hosts of lGRBs is not reproduced by
+\citet{Buchner2017}, using a different $N_\mathrm{H}$ inference methodology. Instead a
+correlation between $N_\mathrm{H}$ and host stellar mass is suggested. Assuming that the
+different $N_\mathrm{H}$-fitting methodologies yield comparable results, GRB~111117A is
 an outlier from the relation suggested by \citet{Buchner2017} by more than the
 intrinsic scatter, although some lGRB hosts populate a similar region in the
-$N_H$-$M_\star$ relation. Additionally, for lGRBs, $N_H$ correlates with the
+$N_\mathrm{H}$-$M_\star$ relation. Additionally, for lGRBs, $N_\mathrm{H}$ correlates with the
 surface luminosity at the explosion site \citep{Lyman2017}.
@@
 galaxy-scale gas providing the X-ray absorption. Along with the low dust content
-of the host, the large offset from the host center indicates that the high $N_H$
-arises because the density in the GRB surroundings is high, or because the light
-from the afterglow traverses a region of dense gas. Assuming a dust-to-metals
-ratio typical for the Milky Way, the X-ray derived $N_H$ corresponds to
-significant extinction along the afterglow sight line ($A_V > 10$)
-\citep{Predehl1995}, which is contrasted to the absence of dust found for the
-SED fit and supported by the detection of \lya{}. The discrepancy between the extinction derived from the GRB afterglow
-and the one obtained using galaxy-wide measures is also seen in the range of
-ratios found for lGRBs \citep{Perley2013a, Covino2013}.
+of the host, the large offset from the host center indicates that the high
+$N_\mathrm{H}$ arises because the density in the GRB surroundings is high, or
+because the light from the afterglow transverses localized regions of dense gas, similar to what is found by \citep{}.
+Alternatively, a significant contribution to the observed X-ray $N_\mathrm{H}$
+can come from the diffuse intergalactic medium and the intervening systems along
+the line of sight of the GRB \citep{Campana2012, Arcodia2016}, although see \citep{}
 
-The optical faintness of the
-GRB is an additional hint of the high density in the GRB surroundings or along
-the line of sight.
+Even assuming a low dust-to-metals ratio as typically observed in long GRB
+afterglow sightlines \citep{Galama2001, Schady2010, Covino2013}, the
+$N_\mathrm{H}$ value derived from the X-ray spectrum corresponds to significant
+extinction along the afterglow line of sight ($A_V \gtrsim 1$~mag), which is
+contrasted to the absence of dust found for the SED fit and supported by the
+detection of \lya{}. The discrepancy between the extinction derived from the GRB
+afterglow and the one obtained using galaxy-wide measures is also seen in the
+range of values found for lGRBs \citep{Perley2013a, Covino2013}.
+
+The lack of optical detection is also consistent with a high column along the
+GRB line of sight, as dust extinction could contribute to the optical faintness.
+The X-ray flux is in first approximation independent of the surrounding medium
+density \citep{Berger2003, Nysewander2009}, and the X-ray afterglow flux of
+GRB~111117A lies within the expected distribution given its gamma-ray fluence
+\citep{DAvanzo2014a}.
 
@@
 	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations of GRBs.
-The detections are replaced with contours and the limits are shown with arrows.
-Marginalizations over both axes are shown where the limits are shown as transparent bars and detections as solid ones.}
+In red is shown the sample of lGRBs from \citet{Arcodia2016} where detections are replaced with a kernel density estimate of the points and the limits on $N_\mathrm{H}$ are shown with arrows.
+In blue is shown the complete sample of sGRB by \citet{DAvanzo2014a}, where again the limits are indicated by arrows.
+Marginalizations over both axes are shown onthe side and on the plot of the plot where the limits are shown as transparent bars and detections as solid ones.
+The red lines in the marginalization plots are again the kernel density estimates of the \citet{Arcodia2016} sample.}
 	\label{fig:NH_z}
@@
 determine. Additionally, it is natural to expect some evolution in the hosts of
-sGRBs with redshift as illustrated for $N_H$ in Sect.~\ref{restnH}.
+sGRBs with redshift as illustrated for $N_\mathrm{H}$ in Sect.~\ref{restnH}.
 
@@
 \citep{Hayes2011}, the resonant scattering of \lya{}-photons with neutral
-hydrogen makes the effective path length of \lya{} is longer than \ha{}
+hydrogen makes the effective path length of \lya{} longer than \ha{}
 \citep{Atek2009}. This makes $f_{esc}$(\lya) an unreliable proxy for dust
@@
 centroid of the Ly$\alpha$ emission is found to be redshifted by
-$\sim$~220~km~s$^{-1}$ with respect to systemic, which is similar to what is
+$\sim$~$240\pm 90$~km~s$^{-1}$ with respect to systemic, which is similar to what is
 found for long GRB hosts \citep{Milvang-Jensen2012a} and Lyman Break Galaxies
@@
 elevated brightness of lGRBs compared to sGRBs \citep{Berger2014} and their
-tendency to be associated with the most star-forming and therefore dense regions
-in their hosts \citep{Fruchter2006, Lyman2017}, the afterglows of lGRBs are
-easier to observe than their sGRB counterparts where only a single burst have a
-redshift measurement from the GRB afterglow \citep{Cucchiara2013,
+tendency to be associated with the star-forming and therefore dense regions in
+their hosts \citep{Fruchter2006, Lyman2017}, the afterglows of lGRBs are easier
+to estimate redshift for than their sGRB counterparts where only a single burst
+have a redshift measurement from the GRB afterglow \citep{Cucchiara2013,
 	DeUgartePostigo2014c}. Correspondingly, the redshift distribution of lGRBs is
@@
 $z = 0.9$. According to \citet{Wanderman2015}, this preference depends
-critically on the absence of non-collapsar sGRBs at $z \gtrsim 1.2$. The higher
+critically on the absence of sGRBs at $z \gtrsim 1.2$. The higher
 determined redshift of GRB~111117A, and the likely number of additional high-z
@@
 
-A critical test to assess whether the power law delay time models can be
-accommodated is to check if the implied sGRB rate at higher redshift can be
-contained in the observed sGRBs without redshifts and the potential number of
-missed bursts due to \textit{Swift}'s sensitivity limit. Of the 65??? definite
-sGRBs observed by \textit{Swift} to date, 10 have redshifts \citep{GRBOX},
-meaning 85 per cent of all sGRBs observed with \textit{Swift} are missing
-redshifts. Additionally, \citet{Behroozi2014} parametrized the \textit{Swift}
-redshift sensitivity and found that mean detection probability for sGRBs at $z
-\sim2$ was only $\sim1$ per cent of the mean detection probability at $z \sim1$,
-assuming that the unknown beaming angle of sGRBs stays constant with time. What
-this means is that at the present, there is almost no limit on the number of
-sGRBs that could be at higher redshifts.
+A critical test to assess if the power law delay time models can be accommodated
+by the current observation is to check if the implied sGRB rate at higher
+redshift can be contained in the observed sGRBs without redshifts and the
+potential number of missed bursts due to \textit{Swift}'s sensitivity limit. Of
+the  100  sGRBs observed by \textit{Swift}, 20 have secure redshifts, and
+another 7 has a tentative redshift measurement\footnote{This is based on
+	\url{http://www.astro.caltech.edu/grbox/grbox.php}, selecting all short,
+	\textit{Swift}-detected GRBs up to GRB~170428A.}, meaning $> 73$ per cent of all
+sGRBs observed with \textit{Swift} are missing redshifts. More recently,
+\citet{Fong2017} \citep{LIGOScientificCollaboration2017b} compiled a list of 36
+(33) sGRBs with redshift measurements and using this number, the redshift
+incompleteness of sGRBs decrease to 64 per cent. Additional to the potential
+number of high-$z$ event already detected but missing redshifts,
+\citet{Behroozi2014} parametrized the \textit{Swift} redshift sensitivity and
+found that mean detection probability for sGRBs at $z \sim2$ was only $\sim1$
+per cent of the mean detection probability at $z \sim1$, assuming that the
+unknown beaming angle of sGRBs stays constant with time. What this means is that
+at the present, there is almost no limit on the number of sGRBs that could be at
+higher redshifts.
 
@@
 gravitational waves, which in turn is set by the mass of the constituent compact
-objects, the eccentricity of the orbit and the separation of the two \citep{Postnov2014}. If we assume that the
-formation timescale of the first galaxies is short compared to the time since
-the Big Bang \citep{Richard2011}, and if we assume a mass of 1.4~$M_\odot$ for
-each of the NSs in a circular orbit at the time of system formation, this places
-a hard upper limit on the initial separation, of $a_0 < 3.2~R_\odot$.
+objects, the eccentricity of the orbit and the separation of the two
+\citep{Postnov2014}. If we assume that the formation timescale of the first
+galaxies is short compared to the time since the Big Bang \citep{Richard2011},
+and that any binary neutron star formation channel can work sufficiently fast,
+we can, assuming a mass of each of the NSs in a circular orbit at the time of
+system formation, place a hard upper limit on the initial separation, $a_0$.
 
@@
 initial separation constraints. As noted by \citet{Postnov2014}, it takes
-eccentricities > 0.6 to significantly shorten the merger time.
+eccentricities > 0.6 to significantly shorten the merger time. Due to tidal
+interactions between the two NSs, the orbits will also tend to circularize with
+time, lessening the impact of the eccentricity on the constraint.
+
+Additionally, the constraint on initial progenitor separation will change
+depending on the mass assumed for the constituent NS masses, with larger masses
+generally resulting in faster inspiral times and weaker constraints of initial
+separation. We use the NS mass distribution from \citet{Kiziltan2013} to compute
+a grid of initial progenitor separation constraints, given the range of NS
+masses allowed. We show the grid in Fig. \ref{fig:prog_sep}. The double NS
+binary systems has a constituent mass distribution peaked at
+$1.33^{+0.10}_{-0.12}$ M$_\odot$ \citep{Kiziltan2013}, which corresponds to  $a_0 <
+3.1^{+0.2}_{-0.2}~R_\odot$ where the errors are the 68 per cent posterior
+predictive intervals.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=9cm]{figures/prog_sep.pdf}
+	\caption{Constraints on the initial progenitor separation, given binary constituent masses.
+As can be seen, a heavier system will inspiral faster, leading to a weaker constraint on the initial separation of the binary.
+The most probable value, the 68 per cent, and 98 percent posterior predictive intervals of the neutron star binary mass distribution from \citet{Kiziltan2013} are shown in the dashed ellipses, which corresponds to a constraint on maximal initial separation of $a_0 <3.1^{+0.2(+0.4)}_{-0.2(-0.4)}~R_\odot$, respectively.}
+	\label{fig:prog_sep}
+\end{figure}
 
 Using the inferred stellar population age from our SED fit, then we obtain a
-(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
-does not account for the possibility there could be an underlying stellar
-population of older stars from a previous star-formation epoch. To investigate
-the possible impact of the presence of an old stellar population, we followed
-\citet{Papovich2001} and re-fitted the observed SED with the best-fit template
-to which an additional stellar population of old stars was added. For each
-galaxy, this old population was set as the SPSs with same parameters that the
-best-fit SED excepts the age, which was set to the age of the Universe at the
-observed redshift. In principle, this can constrain the maximum contribution of
-old populations within the photometric error bars \citep[see][ for
+(softer) limit on the initial separation of $a_0 < 2.0^{+0.4}_{-0.4}~R_\odot$.
+However, this does not account for the possibility there could be an underlying
+stellar population of older stars from a previous star-formation epoch. To
+investigate the possible impact of the presence of an old stellar population, we
+followed \citet{Papovich2001} and re-fitted the observed SED with the best-fit
+template to which an additional stellar population of old stars was added. For
+each galaxy, this old population was set as the SSPs with same parameters that
+the best-fit SED excepts the age, which was set to the age of the Universe at
+the observed redshift. In principle, this can constrain the maximum contribution
+of old populations within the photometric error bars \citep[see][for
 details]{Papovich2001}. We find a negligible contribution to the stellar mass
 (i.e. variations much smaller than the statistical uncertainty associated with
-the best-fit template). The delay time between formation and explosion is well
-accommodated by the models of \citet{Belczynski2006}, although the longest delay
-times are excluded. This is especially true given the late type nature of the
-host \citep{OShaughnessy2008}.
+the best-fit template).
+
+The delay time between formation and explosion is well accommodated by the
+models of \citet{Belczynski2006}, although the longest delay times are excluded.
+This is especially true given the late type nature of the host
+\citep{OShaughnessy2008}. The same holds, if neutron star binaries are primarily
+formed through dynamical interactions in globular clusters \citep{Lee2010,
+	Church2011}.
 
@@
 We have here provided a revised, spectroscopic redshift measurement for the
-short GRB~111117A based on host galaxy emission lines setting it at $z = 2.211$.
-This value supersedes the previous photometric redshift of \citet{Margutti2012}
-and \citet{Sakamoto2013}. The erroneous redshift estimate of previous authors is
-attributed to a discrepancy in the measured $z$-band magnitude, and highlights
-the importance of deep spectroscopic studies of sGRB hosts at medium resolution.
+short GRB~111117A based on host galaxy emission lines setting it at $z = 2.211
+\pm 0.001$. This value supersedes the previous photometric redshift of
+\citet{Margutti2012} and \citet{Sakamoto2013}. The erroneous redshift estimate
+of previous authors is attributed to a discrepancy in the measured $z$-band
+magnitude, and highlights the importance of deep spectroscopic studies of sGRB
+hosts at medium resolution.
 
-Using the new distance, the X-ray derived $N_H$ towards GRB~111117A is the
+Using the new distance, the X-ray derived $N_\mathrm{H}$ towards GRB~111117A is the
 highest within a complete sample of sGRBs and is consistent with the
-$N_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
+$N_\mathrm{H}-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
 upper end of the sGRB host SFR distribution and no significant amount of dust is
-present. The high $N_H$ is difficult to reconcile with the large projected host
+present. The high $N_\mathrm{H}$ is difficult to reconcile with the large projected host
 offset and the absence of dust. One possible explanation could be, that
@@
 with measured luminosities. Assuming a host brightness redshift determination
-threshold, for between 43 and 71 per cent of sGRB hosts, we would be unable to
+threshold, for between 43 and 71 per cent of the sample hosts, we would be unable to
 determine a redshift should they be at a similar redshift as GRB~111117A. This
@@
 \begin{acknowledgements}
+We thank the anonymous referee for a constructive report. 
+%
 We thank Jens Hjorth and Lise Christensen for useful discussions regarding the interpretation of this event.
 We thank Mathieu Puech for testing the possible contribution from an older stellar population in the SED.
-We thank Peter Laursen for fruitful discussions regarding the \lya escape fraction.
+We thank Peter Laursen for fruitful discussions regarding the \lya~escape fraction.
 %	
@@
 SDV is supported by the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro 
+%
+PDA, SCo, acknowledge support from ASI grant I/004/11/3
 %

### 2017-11-24 authorlist.tex
Referee commnets final

@@
 \and T.~Kr\"{u}hler\inst{2}
-\and D.~Malesani\inst{1,3}
-\and P.~D'Avanzo\inst{4}
-\and S.~Schulze \inst{5, 6, 7}
+\and D.~Malesani\inst{1}
+\and P.~D'Avanzo\inst{3}
+\and S.~Schulze \inst{4, 5, 6}
+\and S.~D.~Vergani\inst{3, 7}
 \and J.~Palmerio\inst{8}
-\and S.~D.~Vergani\inst{9}
-\and J.~Japelj\inst{10}
+\and J.~Japelj\inst{9}
 \and B.~Milvang-Jensen\inst{1}
 \and D.~Watson\inst{1}
-\and P.~Jakobsson\inst{11}
+\and P.~Jakobsson\inst{10}
 %
 \and J.~Bolmer \inst{2}
-\and Z.~Cano\inst{12}
-\and S.~Covino \inst{13}
+\and Z.~Cano\inst{11}
+\and S.~Covino \inst{3}
 %\and L.~B.~Christensen\inst{1}
-\and V.~D'Elia\inst{13, 14}
-\and A.~de~Ugarte~Postigo\inst{12, 1}
+\and V.~D'Elia\inst{12, 13}
+\and A.~de~Ugarte~Postigo\inst{11, 1}
 \and J.~P.~U.~Fynbo\inst{1}
-\and A.~Gomboc\inst{15}
-\and K.~E.~Heintz\inst{11,1}
-\and A.~J.~Levan \inst{16}
-\and S.~Piranomonte \inst{17}
-\and G.~Pugliese \inst{10}
-\and R.~S\'{a}nchez-Ram\'{\i}rez \inst{12, 18}
-\and M.~Sparre\inst{1,19}
-\and N.~R.~Tanvir\inst{20}
-\and C.~C.~Th\"one\inst{12}
-\and K.~Wiersema \inst{20}
+\and A.~Gomboc\inst{14}
+\and K.~E.~Heintz\inst{10,1}
+\and L.~Kaper \inst{9}
+\and A.~J.~Levan \inst{15}
+\and S.~Piranomonte \inst{12}
+\and G.~Pugliese \inst{9}
+\and R.~S\'{a}nchez-Ram\'{\i}rez \inst{11, 17}
+\and M.~Sparre\inst{1,18}
+\and N.~R.~Tanvir\inst{19}
+\and C.~C.~Th\"{o}ne\inst{11}
+\and K.~Wiersema \inst{19}
 }
@@
 %3
-\and DTU Space, National Space Institute, Technical University of Denmark, Elektrovej 327, DK-2800 Lyngby, Denmark
+\and INAF - Osservatorio Astronomico di Brera, via E. Bianchi 46, I-23807, Merate (LC), Italy
 %4
-\and INAF - Osservatorio Astronomico di Brera, via E. Bianchi 46, I-23807, Merate (LC), Italy
+\and Instituto de Astrof\'isica, Facultad de F\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
 %5
-\and Instituto de Astrof\'isica, Facultad de F\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+\and Millennium Institute of Astrophysics, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
 %6
-\and Millennium Institute of Astrophysics, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
 %7
-\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
+\and GEPI, Observatoire de Paris, PSL Research University, CNRS, Place Jules Janssen, 92190 Meudon, France
 %8
@@
 %9
-\and GEPI, Observatoire de Paris, PSL Research University, CNRS, Univ. Paris Diderot, Sorbonne Paris Cité, 5 Place Jules Janssen, 92195 Meudon, France
+\and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands
 %10
-\and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands
+\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
 %11
-\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
+\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain
 %12
-\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain
+\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
 %13
-\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
+\and ASI-Science Data Centre, Via del Politecnico snc, I-00133 Rome, Italy
 %14
-\and ASI-Science Data Centre, Via del Politecnico snc, I-00133 Rome, Italy
+\and Centre for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
 %15
-\and Centre for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
+\and Department of Physics, University of Warwick, Coventry CV4 7AL, UK
 %16
-\and Department of Physics, University of Warwick, Coventry CV4 7AL, UK
+\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate (LC), Italy
 %17
-\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate (LC), Italy
+\and Istituto de Astrofisica e Planetologia Spaziali, INAF, Via Fosso del Cavaliere 100, I-00133 Roma, Italy
 %18
-\and Istituto de Astrofisica e Planetologia Spaziali, INAF, Via Fosso del Cavaliere 100, I-00133 Roma, Italy
+\and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
 %19
-\and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
-%20
-\and Department of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK
-%\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
+\and Department of Physics \& Astronomy and Leicester Institute of Space \& Earth Observation,
+University of Leicester, University Road, Leicester, LE1 7RH, UK
 }

### 2017-11-24 py/NH_z.py

@@ -105,8 +105,10 @@ def main():
     cmap = sns.blend_palette(colors, as_cmap=True)
     g = g.plot_marginals(sns.distplot, hist=False, color=color)
     g = g.plot_joint(sns.kdeplot, cmap=cmap, label="Arcodia et al. 2016", zorder=1)
+    # g = g.plot_joint(pl.scatter, cmap=cmap, label="Arcodia et al. 2016", zorder=1)
     # g = g.plot_marginals(sns.distplot, hist=False, rug=True, kde=False, color=color, rug_kws={"height": 0.7, "lw": 2.0})
     # g = g.plot_joint(pl.scatter, color=color, label="Arcodia et al. 2016")
+    print(z_long_uplim, NH_long_uplim)
     g.x = z_long_uplim
     g.y = NH_long_uplim
     g = g.plot_marginals(sns.distplot, hist=False, rug=True, kde=False, color=color, rug_kws={"height": 0.7, "lw": 2.0, "alpha": 0.5})
@@ -149,7 +151,7 @@ def main():
     g = g.plot_joint(pl.plot, color=sns.color_palette()[2], label="Arcodia et al. 2016")
 
     ax = pl.gca()
-    g.set_axis_labels(r"z", r"log(N$_H$) [cm$^{-2}$]")
+    g.set_axis_labels(r"z", r"log($N_\mathrm{H}$) [cm$^{-2}$]")
     pl.tight_layout()
 
     # Save figure for tex

### 2017-11-24 py/SED_plot.py

@@ -10,8 +10,10 @@ import astropy.units as u
 import astropy.constants as c
 
 # Plotting
+import matplotlib; matplotlib.use('TkAgg')
 import matplotlib.pyplot as pl
 import seaborn; seaborn.set_style('ticks')
+import lineid_plot
 
 import matplotlib as mpl
 from matplotlib.ticker import FormatStrFormatter
@@ -21,7 +23,7 @@ params = {
    'legend.fontsize': 24,
    'xtick.labelsize': 24,
    'ytick.labelsize': 24,
-   'text.usetex': True,
+   'text.usetex': False,
    'figure.figsize': [16, 16/1.618]
    }
 mpl.rcParams.update(params)
@@ -31,9 +33,17 @@ def main():
 
     # Load data
     spectrum = np.genfromtxt("../data/spectroscopy/stitched_spectrum_bin5.dat")
-    wl_spec = spectrum[:, 0] * u.AA
-    flux_spec = spectrum[:, 1] * (u.erg/(u.s * u.cm**2 * u.AA))
-    error_spec = spectrum[:, 2] * (u.erg/(u.s * u.cm**2 * u.AA))
+    wl_spec = spectrum[:, 0] #* u.AA
+    flux_spec = spectrum[:, 1] #* (u.erg/ u.s / u.cm**2 / u.AA)
+    error_spec = spectrum[:, 2] #* (u.erg/(u.s * u.cm**2 * u.AA))
+    f = interp1d(wl_spec, flux_spec, fill_value = np.nan)
+    g = interp1d(wl_spec, error_spec, fill_value = np.nan)
+    x = np.arange(min(wl_spec), max(wl_spec), np.median(np.diff(wl_spec)))# * u.AA
+    flux_spec = f(x) * (u.erg/(u.s * u.cm**2 * u.AA))
+    error_spec = g(x) * (u.erg/(u.s * u.cm**2 * u.AA))
+    wl_spec = x * u.AA
+
+
 
     SED = np.genfromtxt("../data/SED.dat")
     wl_SED = SED[:, 0]
@@ -47,26 +57,37 @@ def main():
     wl_plot_wid = phot[:, 3][magmask]/2.35
 
     AB = flux_spec.to(u.ABmag, u.spectral_density(wl_spec))
+    flux_jan = flux_spec.to(u.Jy, u.spectral_density(wl_spec))*1e6
[…]

### 2017-11-24 py/combine_arms_and_bin.py

@@ -181,14 +181,16 @@ def main():
     VIS_mask = (VIS_wl > 5700) & (VIS_wl < 9600)
     NIR_mask = (NIR_wl > 10500) & (NIR_wl < 20500) #& (NIR_flux > -1e-19)
     waves = [UVB_wl[UVB_mask], VIS_wl[VIS_mask], NIR_wl[NIR_mask]]
-    flux = [UVB_flux[UVB_mask], VIS_flux[VIS_mask], NIR_flux[NIR_mask]]
-    error = [UVB_error[UVB_mask], VIS_error[VIS_mask], NIR_error[NIR_mask]]
+    flux = [UVB_flux[UVB_mask]*1.3, VIS_flux[VIS_mask]*1.1, NIR_flux[NIR_mask]*1.1] # Slitloss correction based on seeing
+    error = [UVB_error[UVB_mask]*1.3, VIS_error[VIS_mask]*1.1, NIR_error[NIR_mask]] # Slitloss correction based on seeing
     bpmap = [UVB_bp[UVB_mask], VIS_bp[VIS_mask], NIR_bp[NIR_mask]]
 
     # Stitch!
     wl, flux, error, bpmap = stitch_XSH_spectra(waves, flux, error, bpmap, scale = scale)
 
-    np.savetxt(data_dir + "stitched_spectrum.dat", zip(wl, flux, error, bpmap), fmt = ['%10.6e', '%10.6e', '%10.6e', '%10.6f'], header=" wl flux error bpmap")
+
+
+    np.savetxt(data_dir + "stitched_spectrum.dat",list(zip(wl, flux, error, bpmap)), fmt = ['%10.6e', '%10.6e', '%10.6e', '%10.6f'], header=" wl flux error bpmap")
 
     hbin = 5
     wl_bin, flux_bin, error_bin, bp_bin = bin_spectrum(wl, flux, error, bpmap.astype("bool"), hbin, weight=True)
@@ -203,24 +205,24 @@ def main():
     # exit()
 
 
-    np.savetxt(data_dir + "stitched_spectrum_bin"+str(hbin)+".dat", zip(wl_bin, flux_bin, error_bin, bp_bin), fmt = ['%10.6e', '%10.6e', '%10.6e', '%10.6f'], header=" wl flux error bpmap")
[…]

### 2017-11-24 py/dust111117.py

@@ -1,0 +1,41 @@
+import numpy as np
+import sys
+sys.path.append('/Users/pela/Dust/')
+import Pei
+import Cardelli
+import Calzetti
+
+RV_SMC, RV_LMC, RV_MW = 2.93, 3.16, 3.08        #RV values
+
+def calcit(fesc,ext):
+  """
+  Usage:
+  ------
+  To calculate AV given a Lya escape fraction of 30%, assuming an SMC
+  extinction, use
+  >>> calcit(.3,'SMC')
+  ---> 0.225
+  """
+  tau  = -np.log(fesc)
+  ALya = 2.5/np.log(10) * tau
+  lam0 = .121567
+  lamV = .5510
+  
+  if ext == 'SMC':
+    RV   = 2.93
+    kLya = Pei.k(lam0,'SMC') 
+  elif ext == 'LMC':
+    RV   = 3.16
+    kLya = Pei.k(lam0,'LMC') 
+  elif ext == 'MW':
+    RV   = 3.08
+    kLya = Pei.k(lam0,'MW') 
+  elif ext == 'Cardelli':
+    RV   = 3.1
+    kLya = Cardelli.k(lam0)
+  elif ext == 'Calzetti':
+    RV   = 4.05
+    kLya = Calzetti.k(lam0)
+  
+  AV = (ALya/kLya) * RV
+  return AV

### 2017-11-24 py/line_flux_ha.py

@@ -46,7 +46,7 @@ def main():
     # Script to get lineflux
     """
     # Get extraction
-    data = np.genfromtxt("../data/NIRext.dat", dtype=None)
+    data = np.genfromtxt("../data/spectroscopy/NIRext.dat", dtype=None)
     # data = np.genfromtxt("../data/NIROB4skysubstdext.dat", dtype=None)
     # print(np.std([18, 31, 38, 20]))
     # exit()
@@ -55,7 +55,7 @@ def main():
     error = data[:, 3]
     error[error > 1e-15] = np.median(error)
     # Load synthetic sky
-    sky_model = fits.open("../data/NIRskytable.fits")
+    sky_model = fits.open("../data/spectroscopy/NIRskytable.fits")
     wl_sky = 10*(sky_model[1].data.field('lam')) # In nm
     # Convolve to observed grid
     f = interpolate.interp1d(wl_sky, convolve(sky_model[1].data.field('flux'), Gaussian1DKernel(stddev=3)), bounds_error=False, fill_value=np.nan)
@@ -70,11 +70,11 @@ def main():
     flux = g(wl)
 
 
-    mask = (wl > convert_air_to_vacuum(21077) - 20) & (wl < convert_air_to_vacuum(21077) + 20)
+    mask = (wl > convert_air_to_vacuum(21076) - 9) & (wl < convert_air_to_vacuum(21076) + 9)
     # mask = (wl > 21020) & (wl < 21150)
-    F_ha = np.trapz(flux[mask])
+    F_ha = np.trapz(flux[mask], x=wl[mask])
     print("Total %0.1e" %F_ha)
-    F_ha_err = np.sqrt(np.trapz((error**2.)[mask]))
+    F_ha_err = np.sqrt(np.trapz((error**2.)[mask], x=wl[mask]))
     print("Total %0.1e +- %0.1e" % (F_ha, F_ha_err))
     # exit()
     # F_ha = 4.1e-17

### 2017-11-24 py/line_flux_lya.py

@@ -1,12 +1,17 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
 
-from __future__ import division, print_function
+
+
+
+# Plotting
+import matplotlib; matplotlib.use('TkAgg')
+import matplotlib.pyplot as pl
+import seaborn; seaborn.set_style('ticks')
+
 
 from astropy.io import fits
 import pandas as pd
-import matplotlib.pyplot as pl
-import seaborn as sns; sns.set_style('ticks')
 import numpy as np
 from scipy import stats, interpolate
 import matplotlib as mpl
@@ -46,62 +51,56 @@ def main():
     # Script to get lineflux
     """
     # Get extraction
-    data = np.genfromtxt("../data/NIRext.dat", dtype=None)
-    # data = np.genfromtxt("../data/NIROB4skysubstdext.dat", dtype=None)
-    # print(np.std([20.2, 36.5, 43.9, 25.6]))
-    # exit()
+    data = np.genfromtxt("../data/spectroscopy/UVBext_lya.dat", dtype=None)
+
     wl = data[:, 1]
+    wl_mask = (wl > 3850) & (wl < 3960)
+
     flux = data[:, 2]
     error = data[:, 3]
+
+    wl, flux, error = wl[wl_mask], flux[wl_mask], error[wl_mask]
+
     error[error > 1e-15] = np.median(error)
-    # Load synthetic sky
-    sky_model = fits.open("../data/NIRskytable.fits")
-    wl_sky = 10*(sky_model[1].data.field('lam')) # In nm
-    # Convolve to observed grid
-    f = interpolate.interp1d(wl_sky, convolve(sky_model[1].data.field('flux'), Gaussian1DKernel(stddev=7)), bounds_error=False, fill_value=np.nan)
-    # f = interpolate.interp1d(wl_sky, sky_model[1].data.field('flux'), bounds_error=False, fill_value=np.nan)
-    flux_sky = f(wl)
-    # flux[(wl > 21055) & (wl < 21068)] = 0
 
-    m = (flux_sky < 9000)
-    g = interpolate.interp1d(wl[m], flux[m], bounds_error=False, fill_value=np.nan)
-    flux = g(wl)
[…]

### 2017-11-24 py/progenitor_seperation.py

@@ -6,11 +6,26 @@ from __future__ import division, print_function
 # Imports
 from astropy import units as u
 from astropy import constants as c
+import numpy as np
+import matplotlib; matplotlib.use('TkAgg')
+import matplotlib.pyplot as pl
+import seaborn as sns; sns.set_style('ticks')
+import matplotlib as mpl
 
+params = {
+   'axes.labelsize': 16,
+   'font.size': 12,
+   'legend.fontsize': 12,
+   'xtick.labelsize': 16,
+   'ytick.labelsize': 16,
+   'text.usetex': False,
+   'figure.figsize': [7.281, 4.5]
+   }
+mpl.rcParams.update(params)
 
 def calc_sep(age, m1, m2):
     """
-    Small calculation of progenitor seperation.
+    Calculation of progenitor seperation.
     age in Myr, progenitor masses in Msun
     returns seperation in units of Rsun
     """
@@ -36,16 +51,51 @@ def main():
     # Calculate initial progenitor seperation from eq 29 from Postnov and Yungelson 2014
     # Time since big bang
     t_constraint = 2970 #Myr
-    # t_constraint = 500 #Myr
+    # t_constraint = 1000 #Myr
     # Neutron star masses
     m1 = 1.4 # Msun
     m2 = 1.4 # Msun
     # Get seperation
-    a0 = calc_sep(t_constraint, m1, m2)
+    ratio = np.arange(0.5, 2, 0.01)
+    masses = np.arange(0.9, 2.1, 0.01)
+    xx, yy = np.meshgrid(masses, masses[::-1])
+    a0 = calc_sep(t_constraint, xx, yy)
 
-    print(a0)
+    color = sns.color_palette()[2]
+    color_rgb = mpl.colors.colorConverter.to_rgb(color)
+    colors = [sns.set_hls_values(color_rgb, l=l) for l in np.linspace(1, 0, 12)]
+    cmap = sns.blend_palette(colors, as_cmap=True)
+    fig, ax = pl.subplots()
+    im = ax.imshow(a0, extent=[min(masses), max(masses), min(masses), max(masses)], cmap="viridis_r")
+    cbar = pl.colorbar(im)
+    cbar.set_label(r'a$_0$ [R$_\odot$]')
 
 
[…]

### 2017-11-25 referee_reply.txt
Last edits before resubmission

@@
 GRB. I suggest using one or the other throughout the text. 
-JS: We have streamline this a bit, but have kept the wording "long and short GRBs"
+JS: We have streamlined this a bit, but have kept the wording "long and short GRBs"
 
@@
 afterglow limits still consistent with a high-density environment at z=2.211?
-The X-ray afterglow is not strongly dependent of the density and because of the optical
+JS: The X-ray afterglow is not strongly dependent of the density and because of the optical
 darkness, dust extinction locally to the GRB can accomodate a large brightness
@@
 
-Margutti didn't use the light curve from the Swift Burst Analyser. Her early light curve is ~0.18 dex brighter.
-The shape is also a bit different. k-correction: Margutti fixed the photon index
-to 2.2. The Burst Analyser provides estimates of the photon index for each
-observation. At the time of the first observation, the Burst Analyser gives a
-photon index of 0.93. The k-correction for a power-law shaped spectrum is:
-(1+z)^(Gamma-2). The difference in these parameters corresponds to a difference
-in the k-correction of 0.63 dex. The different distances gives a luminosity
-difference of 0.58 dex (differences in the cosmology are taken into account).
+Margutti+ didn't use the light curve from the Swift Burst Analyser. Her early light
+curve is ~0.18 dex brighter. The shape is also a bit different. k-correction:
+Margutti+ fixed the photon index to 2.2. The Burst Analyser provides estimates of
+the photon index for each observation. At the time of the first observation, the
+Burst Analyser gives a photon index of 0.93. The k-correction for a power-law
+shaped spectrum is: (1+z)^(Gamma-2). The difference in these parameters
+corresponds to a difference in the k-correction of 0.63 dex. The different
+distances gives a luminosity difference of 0.58 dex (differences in the
+cosmology are taken into account).
 
@@
 
-Comparing her plot with ours gives difference of ~-0.2 dex, which is identical to the expected difference.
+Comparing her plot with ours gives difference of ~-0.2 dex, which is identical
+to the expected difference.
 

### 2017-11-25 GRB111117A.tex
Last edits before resubmission

@@
 complete sample of sGRBs and seems to follow the evolution with redshift as
-traced by the hosts of long GRBs (lGRBs).
+traced by the hosts of long GRBs.
 
 The host lies in the brighter end of the expected sGRB host brightness
-distribution at $z = 2.211$, and is actively forming stars. Using the host as a
-benchmark for redshift determination, we find that between 43 and 71 per cent of
-all \textit{Swift}-detected sGRBs are too faint at $z\sim2$ to allow for a secure
-redshift determination. The high redshift of GRB~111117A is evidence against a
-lognormal delay-time model for sGRBs through the predicted redshift distribution
-of sGRBs, which is very sensitive to high-$z$ sGRBs.
+distribution at $z = 2.211$, and is actively forming stars. Using the observed
+sGRB host luminosity distribution, we find that between 43 and 71 per cent of
+all \textit{Swift}-detected sGRBs have too faint hosts at $z\sim2$ to allow for
+a secure redshift determination. This implies that the measured sGRB redshift
+distribution could be incomplete at high redshift. The large $z$ of
+GRB~111117A is evidence against a lognormal delay-time model for sGRBs through
+the predicted redshift distribution of sGRBs, which is very sensitive to
+high-$z$ sGRBs.
 
@@
 There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs) come
-from the merger of neutron stars (NSs), either with another NS, or a black hole,
+from the merger of a neutron star (NS), either with another NS, or a black hole,
 due to their apparent association with kilonovae \citep{Barnes2013a,
@@
 recently confirmed by the simultaneous and co-spatial detection of gravitational
-waves (GWs) from binary neutron star merger and a sGRB
+waves (GWs) from a binary neutron star merger and a sGRB
 \citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}. The
@@
 distribution of burst duration and spectral hardness in the BATSE sample \citep{Kouveliotou1993},
-where the duration T$_{90} < 2$ sec has been regarded as the dividing line between long
+where the duration $T_{90} < 2$~s has been regarded as the dividing line between long
 and short GRBs. Additionally, it has been found for long GRBs (lGRBs) that there is a
@@
 The bulk of these localizations have associated galaxies at relatively low
-redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, but because
-most of these measurements come from the associated hosts, it is potentially
-biased towards lower redshifts. Additionally, because the \textit{Swift}
-sensitivity to sGRB decreases sharply with redshift \citep{Behroozi2014}, the
-intrinsic redshift distribution of sGRBs is largely unknown at higher redshifts.
+redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}. Unlike lGRBs,
+sGRB afterglows are faint making absorption spectroscopy often ineffective.
+Therefore, most of these measurements come from the associated hosts and is
+potentially biased towards lower redshifts. Additionally, because the
+\textit{Swift} sensitivity to sGRB decreases sharply with redshift
+\citep{Behroozi2014}, the intrinsic redshift distribution of sGRBs is largely
+unknown at higher redshifts.
 
@@
 be very long lived in comparison to lGRBs, with the host stellar mass affecting
-the sGRB rates more than star-formation rate (SFR) \citep{Berger2014}.
+the sGRB rates more than the star-formation rate (SFR) \citep{Berger2014}.
 
 The electromagnetic signals from sGRBs are a promising channel to accurately
-localize NS mergers, which holds the promise for a detection of an associated
-gravitational wave (GW) signal \citep{Ghirlanda2016}. This epochal breakthrough
-occurred recently when the first NS-NS GW event ever was detected by LIGO/Virgo
-(GW 170817) and associated to the weak sGRB 170817A detected by the Fermi and
-Integral satellites \citep{LIGOScientificCollaboration2017a, Goldstein2017,
+localize NS mergers \citep{Ghirlanda2016}. This epochal breakthrough occurred
+recently when the first NS-NS GW event ever was detected by LIGO/Virgo (GW
+170817) and associated to the weak sGRB 170817A detected by the \textit{Fermi}
+and INTEGRAL satellites \citep{LIGOScientificCollaboration2017a, Goldstein2017,
 	Savchenko2017}. The simultaneous detection of a sGRB and a GW provides new
-promising ways to constrain the binary inclination angle \citep{Arun2014, LIGOScientificCollaboration2017a} and
-measure cosmological distances \citep{Nissanke2010, LIGOScientificCollaboration2017c}.
+promising ways to constrain the binary inclination angle \citep{Arun2014,
+	LIGOScientificCollaboration2017a} and measure cosmological distances
+\citep{Nissanke2010, LIGOScientificCollaboration2017c}.
 
-The total lifetimes of NS binaries depends on their orbit, mass, spin, initial
+The total lifetime of NS binaries depends on their orbit, mass, spin, initial
 separations and subsequent inspiral times. The delay time from formation to
@@
 
-We here present a spectrum of the host galaxy of the short
-GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This
-value is significantly higher than the previously estimated redshift based on
-photometric studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest
-frame properties based on this new distance compared to previous analyses and
-revisit the host properties derived from the new solution to the spectral energy distribution (SED) fit.
-While no optical afterglow was detected, the excellent localization from a
-detection of the X-ray afterglow by the \emph{Chandra X-ray Observatory}
-allows us to discuss the positioning and environmental properties of this
-remarkably distant sGRB.
-We use the $\Lambda$CDM cosmology parameters provided by
-\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
-and $\Omega_m = 0.307$. All magnitudes are given in the AB system.
+We here present a spectrum of the host galaxy of the short GRB~111117A
+($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is
+significantly higher than the previously estimated redshift based on photometric
+studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest frame
+properties based on this new distance compared to previous analyses and revisit
+the host properties derived from the new solution to the spectral energy
+distribution (SED) fit. While no optical afterglow was detected, the excellent
+localization from a detection of the X-ray afterglow by the \emph{Chandra} X-ray
+Observatory allows us to discuss the positioning and environmental properties of
+this remarkably distant sGRB. We use the $\Lambda$CDM cosmology parameters
+provided by \citet{Planck2015} in which the universe is flat with $H_0 =
+67.7$\,km\,s$^{-1}$ Mpc$^{-1}$ and $\Omega_\mathrm{m} = 0.307$. All magnitudes
+are given in the AB system.
 
@@
 	\caption{
-	FORS2 R-band imaging of the field of GRB~111117A with the X-shooter slit overlaid. The slit
-	position represents 4 epochs of spectroscopic observations taken at similar
-	position angles. The corresponding photometry
-	is shown in Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as
-	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. 
+	FORS2 $R$-band imaging of the field of GRB~111117A with the X-shooter slit
+overlaid. The slit position represents 4 epochs of spectroscopic observations
+taken at similar position angles. The corresponding photometry is shown in
+Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as derived
+from the \emph{Chandra} observations in \citet{Sakamoto2013}.
 	}
@@
 	\includegraphics[width=18cm]{figures/grid.pdf}
-	\caption{2D-images of the emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha.
-The location of a bright skylines are marked by the blue boxes.
-The location of the emission line is indicated with the red ellipse.
-Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.
-For the upper, left panel containing \lya, the systemic redshoft position of \lya is marked by a yellow, vertical, dashed line.
+	\caption{2D-images of the emission lines corresponding to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha.
+The location of bright skylines are marked by blue boxes.
+The locations of the emission lines are indicated with red ellipses.
+Because the host is observed in nodding-mode, negative images of the emission lines appear on both sides in the spatial direction.
+For the upper, left panel containing \lya, the systemic redshift position of \lya{} is marked by a yellow, vertical, dashed line.
 The red part of the \oii$\lambda$3727-doublet is affected by atmospheric absorption.}
@@
 	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
-	\caption{SED fit showing the best-fit SED to the derived photometry.
+	\caption{Best-fit SED to the derived photometry.
 The detection of \lya{}  is predicted from the SED fit and confirmed by the spectroscopic observations.
 Overplotted in grey is the observed spectrum, binned by 6 \AA{} for presentation purposes.
-An attempt to correct for slit losses has been carried out based on the average seeing of the observations.
+Slit losses has been corrected for based on the average seeing of the observations, as confirmed by the comparison with the photometry.
 The blue, dashed line is the corresponding error spectrum, smoothed for presentation purposes.
@@
 091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, with 5
-arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
+\arcsec~nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
 24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across
@@
 slit position is shown in Fig.~\ref{fig:spec_setup}. Each of the epochs are
-extracted individually and combined in a weighted fashion where the weights at
+extracted individually and combined in a weighted fashion where the weight at
 each pixel is chosen as median variance spectrum of the region surrounding that
@@
 of emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959,
-\oiii$\lambda$5007, and \ha. \hb{} and \oiii$\lambda$4959 are both detected at
-low significance ($\sim$ 3-$\sigma$). The uncertainty on the redshift is the
-standard deviation of independent measurements of the redshift based on the
-individual lines. We show \oiii$\lambda$5007 in the insert in
-Fig.~\ref{fig:line}.  \ha{} is only visible in the first epoch due to the
-$K$-band blocking filter used for the remaining observations. The nebular lines
-exhibit a spatial extent of $\sim$ 1\farc5 and show significant velocity
-structure along the slit. A drop in the continuum bluewards of the \lya{} line
-further supports the inferred redshift. No spectral evolution is observed across
-the epochs indicating that there is negligible GRB afterglow contribution to the
-spectra.
+\oiii$\lambda$5007, and \ha. \oii$\lambda$3727, \hb{}, and \oiii$\lambda$4959
+are detected at low significance ($\sim$ 3-$\sigma$). The uncertainty on the
+redshift is the standard deviation of independent measurements of the redshift
+based on the individual line centroids (excluding \lya). We show cutouts of the
+2D spectrum at the position of all the detected lines in Fig.~\ref{fig:line}. 
+\ha{} is only visible in the first epoch due to the $K$-band blocking filter
+used for the remaining observations. The nebular lines exhibit a spatial extent
+of $\sim$ 1\farc5 and show significant velocity structure along the slit. A drop
+in the continuum bluewards of the \lya{} line further supports the redshift. No
+spectral evolution is observed across the epochs indicating that there is
+negligible GRB afterglow contribution to the first epoch spectrum.
 
-Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
-host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed
-at 21\,000~\AA{} where the night sky is very bright. In addition, several bright
-sky-lines are superposed on the line, making an accurate estimate of the
-\ha-flux difficult. Due to their velocity structure, the lines exhibit clear
+Using the luminosity of \ha, we can infer the SFR of the host
+\citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed at
+around 21\,000~\AA{} where the night sky is very bright. In addition, several
+bright sky-lines are superposed on the line, making an accurate estimate of the
+\ha~flux difficult. Due to their velocity structure, the lines exhibit clear
 deviations from a Gaussian and given the low S/N of the spectra we do not
 attempt any parametric fits. We instead obtain a limit on the SFR by numerically
-integrating the part of \ha{} free of contamination and obtain $F_{H \alpha} >
-4.1 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After
+integrating the part of \ha{} free of contamination and obtain $F_{\mathrm{H,X} \alpha} >
+4.1 \times 10^{-17}~\mathrm{erg}~\mathrm{s}^{-1}~\mathrm{cm}^{-2}$. After
 converting the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial
-mass function using the conversion factor from \citet{Madau2014}, we derive a limit of $SFR > 7
-M_\odot$~yr$^{-1}$. We additionally obtain the \lya{} line flux by numerically
-integrating the entire \lya{} line complex and obtain $F_{\mathrm{Ly}\alpha} =
-2.0 \pm 0.5 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$,
-where the error on the flux is found by integrating the associated error
-spectrum over the same spectral region.
+mass function using the conversion factor from \citet{Madau2014}, we derive a
+limit of $\mathrm{SFR} > 7 M_\odot$~yr$^{-1}$. We additionally obtain the \lya{} line
+flux by numerically integrating the entire \lya{} line complex and obtain
+$F_{\mathrm{Ly}\alpha} = (2.0 \pm 0.5) \times 10^{-17}~\mathrm{erg}~
+\mathrm{s}^{-1}~\mathrm{cm}^{-2}$, where the error on the flux is found by
+integrating the associated error spectrum over the same spectral region.
 
@@
 
-From the SED-fit (Sect.~\ref{SED}), and the detection of \lya{}, the host is
-constrained to contain very little or no dust, although the presence of \lya{}
-does not exclude dust. Therefore we do not apply a dust-correction to the
-measured \ha{} flux here. \oii{} is close to a region of strong telluric
-absorption, which is why no SFR is inferred from this line.
+From the SED-fit (Sect.~\ref{SED}) the host is constrained to contain very
+little or no dust. Consistently, \lya{} is detected although its presence does
+not exclude dust. Therefore we do not apply a dust-correction to the measured
+\ha{} flux here. \oii{} is close to a region of strong telluric absorption,
+which is why no SFR is inferred from this line.
 
-The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
+The total extent of the lines in velocity space is $\sim$ 450 km s$^{-1}$. The line
 profiles shows an asymmetric "double-horned" profile, indicating that we are
@@
 \citep{DeBlok2014}. Based on the physical size along the slit and the velocity
-width of \oiii$\lambda5007$, we infer M$_\text{dyn} \gtrsim 10^{10.8}$
-M$_\odot$. Because we are viewing the host inclined at an angle relative to
+width of \oiii$\lambda5007$, we infer $M_\text{dyn} \gtrsim 10^{10.8}$
+$M_\odot$. Because we are viewing the host inclined at an angle relative to
 edge-on and because the slit is not aligned along the long axis of the host,
@@
 GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
-($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
-faded. These new data are complemented by a re-analysis of some of the imaging
-used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
-(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
+($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters). These new data are
+complemented by a re-analysis of some of the imaging used in
+\citet{Margutti2012} and \citet{Sakamoto2013} that are available to us (GTC
+$gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
 observations and measured brightnesses is given in
@@
 All data were reduced, analyzed and fitted in a similar manner as described in
-detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
-Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
-reduction which includes bias/flat-field correction, de-fringing (where
-necessary), sky-subtraction, and stacking of individual images. On the final
-reduced image products, \texttt{DAOPHOT} \citep{Stetson1987} was used to derive
-the reported photometry, where the size of the aperture was chosen to be
-2\farc0. Because image quality ranges from 0\farc6 - 1\farc7 and the galaxy
-major axis is $\sim$ 1\farc, this ensures that, in all cases, the large majority
-of the flux is collected and low-surface brightness light missed in the aperture
-will not influence the measurement. This method sacrifices some S/N in the best
-seeing cases for more reliable photometry across differing observing conditions.
+detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. We
+use our own \texttt{Python} and IRAF routines to perform a standard reduction
+which includes bias/flat-field correction, de-fringing (where necessary),
+sky-subtraction, and stacking of individual images. On the final reduced image
+products, \texttt{DAOPHOT} \citep{Stetson1987} was used to derive the reported
+photometry, where the size of the aperture was chosen to be 2\farc0. Because
+image quality ranges from 0\farc6 - 1\farc7 and the galaxy major axis is $\sim$
+1\farc, this ensures that, in all cases, the large majority of the flux is
+collected and low-surface brightness light missed in the aperture will not
+influence the measurement. This method sacrifices some S/N in the best seeing
+cases for more reliable photometry across differing observing conditions.
 
@@
 transformations of Lupton%
-\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. We
+\footnote{\url{http://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. We
 convert all magnitudes into the AB system, and correct for a Galactic foreground
@@
 re-analysis of the Gemini data compared to the $z$-band measurements of
-\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
-$z$-band magnitudes that are brighter by $\sim 1.0$ mag compared to our values,
-where data taken in other filters are consistent within the errors. Visual
-inspection of the Gemini image shows only a marginal detection of the host,
-which we report here as a $\sim$ 2-$\sigma$ measurement. More conservatively,
-the 3-$\sigma$ upper limit for the Gemini image is $z > 24.06$. Objects of
-magnitude $z \sim 23$ are clearly seen and are significantly brighter than the
-GRB host galaxy. The consistency between the deeper FORS2 $z$-band image and the
-upper limit derived for the Gemini image lends credence to the inferred
-magnitudes presented here, see Fig. \ref{fig:photplot}.
+\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report $z
+\sim 23$, which is brighter than our measurement by $\sim 1.0$ mag, while data
+taken in other filters are consistent within the errors. Visual inspection of
+the Gemini image shows only a marginal detection of the host, which we report
+here as a $\sim$ 2-$\sigma$ measurement. More conservatively, the 3-$\sigma$
+upper limit for the Gemini image is $z > 24.06$. Objects of magnitude $z \sim
+23$ are clearly seen and are significantly brighter than the GRB host galaxy.
+The consistency between the deeper FORS2 $z$-band image and the upper limit
+derived for the Gemini image lends credence to the inferred magnitudes presented
+here, see Fig. \ref{fig:photplot}.
 
 \begin{figure}
-	\centering
-	\includegraphics[width=9cm]{figures/photometry_mosaic.pdf}
-	\caption{Mosaic showing all obtained imaging.
-The host of GRB~111117A is marked by a red circle with a 2\farc0 radius.
-This is the same size as the aperture used for deriving the photometry.
-Worth noting is the relative depth of the GMOS-N $z$-band and the FORS2 $z$-band.
-For reference, the object located to the south-east of the host and marked with a blue, dashed circle has an extinction corrected magnitude of $23.11 \pm 0.09$ ($23.10 \pm 0.18$) in the FORS2 (GMOS-N) $z$-band.}
+	\centering \includegraphics[width=9cm]{figures/photometry_mosaic.pdf}
+\caption{Mosaic showing all used imaging. The host of GRB~111117A is marked by
+	a red circle with a 2\farc0 radius. This is the same size as the aperture used
+	for deriving the photometry. Each panel is 20\arcsec~in size, north is up, east
+	is left. Worth noting is the relative depth of the GMOS-N and FORS2 $z$-band
+	images. For reference, the object located to the south-east of the host and
+	marked with a blue, dashed circle has an extinction corrected magnitude of
+	$23.11 \pm 0.09$ ($23.10 \pm 0.18$) in the FORS2 (GMOS-N) $z$-band.}
 \label{fig:photplot}
@@
 The multi-color SED is fit using the \citet{Bruzual2003} single stellar
-population models (SSPs) based on a \citet{Chabrier2003} with initial mass function in
-\emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the
-spectroscopic value of $z=2.211$. For the SED fitting, we create a grid of
-templates consisting of $\sim 10^6$  different galaxy templates with four
-metallicities 0.02, 0.2, 0.4, 1.0 $Z_{\odot}$, different ages, star-formation
-histories and degrees of extinction. For every model, we calculate the
-likelihood, and create a probability density function (PDF) for a given
-parameter by marginalizing over the other parameters. We quote the median of the
-PDF as the best fit parameters and the errors are the 16th and 84th percentiles
-of the PDFs \citep[e.g. see][for details on the SED fitting
-procedure]{Schulze2016}.
+population models (SSPs) based on a \citet{Chabrier2003} with initial mass
+function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
+the spectroscopic value of $z=2.211$. For the SED fitting, we create a grid
+consisting of $\sim 10^6$  different galaxy templates with four metallicities
+0.02, 0.2, 0.4, 1.0 $Z_{\odot}$, different ages, star-formation histories and
+degrees of extinction. For every model, we calculate the likelihood, and create
+a probability density function (PDF) for a given parameter by marginalizing over
+the other parameters. We quote the median of the PDF as the best fit parameters
+and the errors are the 16th and 84th percentiles of the PDFs \citep[e.g.
+see][for details on the SED fitting procedure]{Schulze2016}.
 
-The best fit model is an unreddened galaxy
-template, and returns physical parameters of absolute magnitude
-($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$),
-stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
-($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). We show the SED
-fit in Fig.~\ref{fig:SED}.
+The best fit model is an unreddened galaxy template. The inferred physical
+parameters are the absolute magnitude ($M_B=-22.0\pm0.1$\,mag), the stellar mass
+($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), the stellar population age ($\tau =
+0.5_{-0.3}^{+0.5}$ Gyr), and the star-formation rate
+($\mathrm{SFR}_{\mathrm{SED}} = 11_{-4}^{+9}~M_\odot\,\mathrm{yr}^{-1}$). We
+show the SED fit in Fig.~\ref{fig:SED}.
 
@@
 spectroscopic value at the 1-$\sigma$ confidence level. The large $i-z$ color
-found in previous works was mistakenly interpreted as a 4000\,\AA{} break
+found in previous works was mistakenly interpreted as the 4000\,\AA{} break,
 driving the galaxy photometric redshift to a lower, erroneous value.
@@
 model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
-an intrinsic equivalent hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
-\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a solar abundance and a Galactic $N_{\mathrm{H}}$ in
+an intrinsic equivalent hydrogen column density $N_{\mathrm{H,X}}$ of $2.4_{-1.6}^{+2.4}
+\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a solar abundance and a Galactic $N_{\mathrm{H,X}}$ in
 the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.
@@
 A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
-"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
-afterglow limits ($r' > 25.5$, 13.5 hrs post-burst) \citep{Cucchiara2011,
+"darkness"-parameter $\beta_\mathrm{OX} $ \citep{Jakobsson2004}. Using the optical
+afterglow limits ($r' > 25.5$, 13.5 hr after the burst; \citealt{Cucchiara2011,
 	Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time
-of the non-detection. We find $\beta_{OX} < 0.79$, consistent with what was
+of the non-detection. We find $\beta_\mathrm{OX} < 0.79$, consistent with what was
 reported in \citet{Sakamoto2013}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Reinterpretation of the restframe properties}
+\section{Reinterpretation of the rest-frame properties}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\citet{Margutti2012} find an projected offset between the host nucleus and the
-site of GRB~111117A of 1\farc25 $\pm$ 0\farc20 arcsec and \citet{Sakamoto2013}
-find a corresponding offset of 1\farc0 $\pm$ 0\farc20 arcsec. This corresponds
+\citet{Margutti2012} find a projected offset between the host nucleus and the
+GRB site of 1\farc25 $\pm$ 0\farc20 arcsec; \citet{Sakamoto2013}
+find a similar value of 1\farc0 $\pm$ 0\farc20 arcsec. These correspond
 to a projected physical offset at $z = 2.221$ of $10.6 \pm 1.7$ kpc and $8.5 \pm
-1.7$ kpc respectively. Because the projected distance does not change
+1.7$ kpc respectively. Because the angular distance does not change
 significantly between $z = 1.3$ and $z = 2.211$, all conclusions of
@@
 
-Based on the BAT light-curve, $T_{90}=0.46$~sec which is shorter than both the
-prototypical 2~sec \citep{Kouveliotou1993} and the $<0.8$~sec suggested to also
+Based on the BAT light-curve, $T_{90}=0.46$~s which is shorter than both the
+prototypical 2~s \citep{Kouveliotou1993} and the $<0.8$~s suggested to also
 exclude the shorter tail of the \textit{Swift}-observed lGRB population
 \citep{Bromberg2012a}. Additionally, no signs of extended emission was found by
-\citep{Sakamoto2013}. A spectral lag of $0.6 \pm2.4$ ms, consistent with zero,
-is not uniquely associated to sGRBs \citep{Bernardini2015}, but in conjunction
-with the duration and the spectral hardness \citep{Sakamoto2011}, as already pointed out
-by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is securely
-classified as a sGRB. Because the observed classification indicators, $T_{90}$
-and hardness ratio, do not depend strongly on redshift \citep{Littlejohns2013a},
-the updated redshift does not change this designation.
+\citet{Sakamoto2013}. The spectral lag is $0.6 \pm2.4$ ms, consistent with zero.
+As already pointed out by \citet{Margutti2012} and \citet{Sakamoto2013}, this is
+typical of sGRBs, (however see also \citealt{Bernardini2015}). In conjunction
+with the duration and the spectral hardness \citep{Sakamoto2011}, GRB~111117A is
+thus securely classified as a sGRB. Because the observed classification
+indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
+\citep{Littlejohns2013a}, the updated redshift does not change this designation.
 
@@
 \ref{fig:sxray_lightcurve}) and it is sub-luminous compared to the majority of
-long GRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
+lGRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
 distribution at 10 ks. The sub-samples comprise of 402 lGRBs, 31 sGRBs, and
-GRB~111117A. The sample of lGRBs are from \citet{Evans2007, Evans2009} and the
-sample of sGRBs are compiled from \citet{Kann2011, Berger2014} and
+GRB~111117A. The sample of lGRBs is from \citet{Evans2007, Evans2009} and the
+sample of sGRBs is compiled from \citet{Kann2011, Berger2014} and
 \citet{DAvanzo2014a}. The mean and the 1-$\sigma$ dispersions of the samples are
-log(L$_{lGRB}$)$ = 46.59 \pm 0.87$ erg s$^{-1}$ and log(L$_{sGRB}$)$  = 44.96
-\pm 0.94$ erg s$^{-1}$ for the lGRB and sGRB samples respectively. GRB~111117A
-had log(L) $= 44.95$ erg s$^{-1}$ at 10 ks which is very close to the peak of
-the sGRB luminosity distribution at 10 ks, but an outlier from the lGRG
-distribution, further supporting the short classification.
+$\log(L_{\mathrm{lGRB}} / \mathrm{erg~s}^{-1}) = 46.59 \pm 0.87$ and
+$\log(L_{\mathrm{sGRB}} / \mathrm{erg~s}^{-1})  = 44.96 \pm 0.94$ for the lGRB
+and sGRB samples respectively. GRB~111117A had $\log(L / \mathrm{erg~s}^{-1}) =
+44.95$ at 10 ks. This is very close to the peak of the sGRB luminosity
+distribution at 10 ks, but an outlier from the lGRB distribution, further
+supporting the short classification.
 
@@
 observables points to an intrinsically different physical origin. LGRBs are
-typically modeled as collapsars \citep{MacFadyen1999} in which a single, massive
+typically interpreted as collapsars \citep{MacFadyen1999} in which a single, massive
 star undergoes gravitational collapse, where the currently preferred model for
-sGRBs is the merger of two neutron stars \citep{Eichler1989, Nakar2007}.
+sGRBs is the merger of two NSs \citep{Eichler1989, Nakar2007}.
 \citet{Bromberg2013} investigated the degree to which high-energy observables of
@@
 percent probability of being a sGRB. Compared to two other sGRB candidates
-at high redshift, GRB~060121 ($T_{90}=1.97 \pm 0.06$~s)
-\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
-($17_{-15}^{+14}$ per cent) and GRB~090426 ($T_{90}=1.28 \pm 0.09$~s)
-\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
+at high redshift, GRB~060121 ($T_{90}=1.97 \pm 0.06$~s;
+\citealt{DeUgartePostigo2006, Levan2006}) at $1.7 \lesssim z \lesssim 4.5$
+($17_{-15}^{+14}$ per cent) and GRB~090426 ($T_{90}=1.28 \pm 0.09$~s;
+\citealt{Antonelli2009, Levesque2010, Thone2011}) at $z = 2.609$ ($10_{-10}^{+15}$
 per cent), the certainty in class membership for GRB~111117A is much higher.
 
-Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
-intermediate duration bursts. This comes from both events having very soft
+Additionally, \citet{Horvath2010} classify both GRB~060121 and GRB~090426 as
+intermediate-duration bursts. This comes from both events having very soft
 spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
@@
 
-A number of other GRBs are claimed to be short and at high ($z > 0.9$) redshift.
-If we consider bursts with a probability of being short, $f_{\mathrm{NC}} > 50$ per cent according to the
-\citet{Bromberg2013} classification scheme as sGRBs, 5 sGRB are found.
-GRB~051210 at $z \sim 1.3$ ($f_{\mathrm{NC}} = 82_{-61}^{+10}$ per cent)
-\cite{Leibler2010},
-GRB~060801 at $z = 1.131$ ($f_{\mathrm{NC}} = 95_{-5}^{+3}$ per cent)
-\cite{Berger2007}
-GRB~070714 at $z = 0.923$ (no $f_{\mathrm{NC}}$ due to extended emission)
-\cite{Graham2009}
-GRB~090510 at $z = 0.903$ ($f_{\mathrm{NC}} = 97_{-29}^{+1}$ per cent)
-\cite{McBreen2010}
-GRB~100117 at $z = 0.915$ ($f_{\mathrm{NC}} = 97_{-3}^{+1}$ per cent)
-\cite{Fong2011} and although simultaneous secure host association and certain
-redshift estimate is difficult, there does seem to be a number of sGRBs at $z
+A number of other GRBs are claimed to be short and at relatively high ($z \gtrsim 0.9$) redshift.
+If we consider bursts with a probability of being short $f_{\mathrm{NC}} > 50$ per cent, according to the
+\citet{Bromberg2013} classification scheme, 5 sGRB are found:
+GRB~051210 at $z \sim 1.3$ ($f_{\mathrm{NC}} = 82_{-61}^{+10}$ per cent;
+\citealt{Leibler2010}),
+GRB~060801 at $z = 1.131$ ($f_{\mathrm{NC}} = 95_{-5}^{+3}$ per cent;
+\citealt{Berger2007}),
+GRB~070714 at $z = 0.923$ (no $f_{\mathrm{NC}}$ due to extended emission;
+\citealt{Graham2009}),
+GRB~090510 at $z = 0.903$ ($f_{\mathrm{NC}} = 97_{-29}^{+1}$ per cent;
+\citealt{McBreen2010}), and
+GRB~100117 at $z = 0.915$ ($f_{\mathrm{NC}} = 97_{-3}^{+1}$ per cent;
+\citealt{Fong2011}). Although in individual cases a secure host association
+(hence redshift determination) is uncertain, there does seem to be a number of sGRBs at $z
 \sim 1$.
@@
 This securely makes GRB~111117A, by far, the highest redshift sGRB detected to
-date. The redshift and classification of GRB~111117A means that it occurred when
-the universe was younger by $\sim$ 3 Gyr compared to any other sGRB ever
-detected. If the merger of NSs is the primary agent for the $r$-process element
-enrichment of the universe \citep{Goriely2011, Ji2016, Komiya2016}, this marks
-the earliest detection of this process.
+date. The redshift and classification of GRB~111117A imply that it occurred when
+the universe was younger by $\sim$ 3 Gyr compared to any other securely
+classified sGRB ever detected. If the merger of NSs is the primary agent for the
+$r$-process element enrichment of the universe \citep{Goriely2011, Ji2016,
+	Komiya2016}, this marks the earliest detection of this process.
 
@@
 	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
-	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
+	\caption{Rest-frame XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
 The grey shaded region is a compilation of long GRB lightcurves \citep{Evans2007, Evans2009} where the color represents density.
@@
 Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and subluminous compared to the lGRB population.
-The insert shows the respective X-ray luminosity distributions of sGRBs and lGRBs at 10ks, indicated by the vertical dashed line.}
+The insert shows the X-ray luminosity distributions of sGRBs and lGRBs at 10 ks, indicated by the vertical dashed line in the main panel.}
 	\label{fig:sxray_lightcurve}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Restframe $N_\mathrm{H}$} \label{restnH}
+\subsection{Rest-frame $N_\mathrm{H,X}$ } \label{restnH}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We show the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare
-with the distributions of complete samples of both long and short GRBs. The lGRB
-sample is from \citet{Arcodia2016} and the sGRB sample is from
-\citet{DAvanzo2014a}. From the sGRB sample of \citet{DAvanzo2014a} we have
-excluded GRB~090426 which does likely not belong in a short sample as highlighted in
-Sect. \ref{classification}. Both comparison samples consider the largest
-temporal interval of constant hardness ratio to prevent spectral changes that
-can affect the X-ray derived column density. The 17 (5) of the 99 (15) long
-(short) GRBs which do not have redshift have been excluded from our analysis.
+We show the recalculated hydrogen equivalent X-ray derived column density,
+$N_\mathrm{H,X}$, in Fig.~\ref{fig:NH_z} where we compare with the distributions
+of complete samples of both long and short GRBs. The lGRB sample is from
+\citet{Arcodia2016} and the sGRB sample is from \citet{DAvanzo2014a}. From the
+sGRB sample of \citet{DAvanzo2014a} we have excluded GRB~090426 which does
+likely not belong in a short sample as highlighted in Sect.
+\ref{classification}. Both comparison samples infer $N_\mathrm{H,X}$  over the
+largest temporal interval of constant hardness ratio to exclude spectral changes
+that can affect the X-ray derived column density. The 17 (5) of the 99 (15) long
+(short) GRBs which do not have measured redshifts have been excluded from our
+analysis.
 
 GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
-$N_\mathrm{H}$ and highest redshift of all sGRBs. The short sample, excluding
-GRB~111117A, is located at low redshifts ($z < 1$) and is found to populate a
+$N_\mathrm{H,X}$ and highest redshift of all sGRBs. The short sample, excluding
+GRB~111117A, is located at low redshift ($z < 1$) and is found to populate a
 column density environment similar to that of lGRBs at comparable redshifts
 \citep{DAvanzo2014a}. The inferred hydrogen column density for GRB~111117A seems
-to follow the trend with increasing $N_\mathrm{H}$ as a function of redshift as found for
+to follow the trend with increasing $N_\mathrm{H,X}$ as a function of redshift as found for
 the lGRB afterglows \citep{Campana2010, Starling2013, Arcodia2016}. This is
-related to what is found by \citet{Kopac2012, Margutti2013}, that $N_\mathrm{H}$ seems to
+related to what is found by \citet{Kopac2012, Margutti2013}, that $N_\mathrm{H,X}$ seems to
 be comparable for long and short GRBs when compared at similar redshifts.
 
-The redshift evolution of $N_\mathrm{H}$ in the hosts of lGRBs is not reproduced by
-\citet{Buchner2017}, using a different $N_\mathrm{H}$ inference methodology. Instead a
-correlation between $N_\mathrm{H}$ and host stellar mass is suggested. Assuming that the
-different $N_\mathrm{H}$-fitting methodologies yield comparable results, GRB~111117A is
-an outlier from the relation suggested by \citet{Buchner2017} by more than the
-intrinsic scatter, although some lGRB hosts populate a similar region in the
-$N_\mathrm{H}$-$M_\star$ relation. Additionally, for lGRBs, $N_\mathrm{H}$ correlates with the
-surface luminosity at the explosion site \citep{Lyman2017}.
+The redshift evolution of $N_\mathrm{H,X}$ in the hosts of lGRBs is not
+reproduced by \citet{Buchner2017}, using a different $N_\mathrm{H,X}$ inference
+methodology. Instead a correlation between $N_\mathrm{H,X}$ and host stellar
+mass is suggested. Assuming that the different $N_\mathrm{H,X}$-fitting
+methodologies yield comparable results, GRB~111117A has a larger
+$N_\mathrm{H,X}$ compared to the relation suggested by \citet{Buchner2017} by
+more than the intrinsic scatter, although some lGRB hosts populate a similar
+region in the $N_\mathrm{H,X}$-$M_\star$ relation.
+% Additionally, for lGRBs,
+%$N_\mathrm{H,X}$ correlates with the surface luminosity at the explosion site
+%\citep{Lyman2017}.
 
@@
 \citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
-galaxy-scale gas providing the X-ray absorption. Along with the low dust content
-of the host, the large offset from the host center indicates that the high
-$N_\mathrm{H}$ arises because the density in the GRB surroundings is high, or
-because the light from the afterglow transverses localized regions of dense gas, similar to what is found by \citep{}.
-Alternatively, a significant contribution to the observed X-ray $N_\mathrm{H}$
-can come from the diffuse intergalactic medium and the intervening systems along
-the line of sight of the GRB \citep{Campana2012, Arcodia2016}, although see \citep{}
+galaxy-scale gas being the source of the X-ray absorption. Along with the low
+dust content of the host, the large offset from the host center indicates that
+the high $N_\mathrm{H,X}$ arises because the density in the GRB surroundings is
+high, or because the light from the afterglow transverses localized regions of
+dense gas, similar to what is found by \citep{}. Alternatively, a significant
+contribution to the observed X-ray $N_\mathrm{H,X}$ can come from the diffuse
+intergalactic medium and the intervening systems along the line of sight of the
+GRB \citep{Campana2012, Arcodia2016}, although see \citep{}
+
+The large offset of GRB\,111117A relative to the host center derived in
+\citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
+galaxy-scale gas being the source of the X-ray absorption. Along with the low
+dust content of the host, the large offset from the host center indicates that
+the high $N_\mathrm{H,X}$ arises because the density in the GRB surroundings is
+high (or possibly because the light from the afterglow transverses localized
+regions of dense gas), see for example,
+\citep[e.g.][]{Galama1998,Watson2013,Krongold2013}. Alternatively, it has been
+hypothesised that a significant contribution to the observed X-ray
+$N_\mathrm{H,X}$ could come from the diffuse intergalactic medium and the
+intervening systems along the line of sight of the GRB \citep{Campana2012,
+	Arcodia2016}, although see \citep{Watson2013, Krongold2013} for counter
+arguments.
 
@@
 afterglow sightlines \citep{Galama2001, Schady2010, Covino2013}, the
-$N_\mathrm{H}$ value derived from the X-ray spectrum corresponds to significant
-extinction along the afterglow line of sight ($A_V \gtrsim 1$~mag), which is
-contrasted to the absence of dust found for the SED fit and supported by the
-detection of \lya{}. The discrepancy between the extinction derived from the GRB
-afterglow and the one obtained using galaxy-wide measures is also seen in the
-range of values found for lGRBs \citep{Perley2013a, Covino2013}.
+$N_\mathrm{H,X}$ value derived from the X-ray spectrum corresponds to
+significant extinction along the afterglow line of sight ($A_V \gtrsim 1$~mag),
+which is contrasted to the absence of dust found from the SED fit and supported
+by the detection of \lya{}. Such a discrepancy between the extinction derived
+from the GRB afterglow and the one obtained using galaxy-wide measures has also
+been observed occasionally for lGRBs \citep{Perley2013a}. For the one sGRB where
+both parameters were measured (GRB~130603B; \citealt{DeUgartePostigo2014d}),
+they were found to be consistent with $A_V \sim 1$~mag.
 
@@
 GRB line of sight, as dust extinction could contribute to the optical faintness.
-The X-ray flux is in first approximation independent of the surrounding medium
-density \citep{Berger2003, Nysewander2009}, and the X-ray afterglow flux of
-GRB~111117A lies within the expected distribution given its gamma-ray fluence
-\citep{DAvanzo2014a}.
+On the contrary, its X-ray afterglow flux lies within the expected distribution
+given its gamma-ray fluence \citep{DAvanzo2014a}. This is not unexpected, as the
+X-ray flux is independent of the surrounding medium density \citep{Freedman2001,
+	Berger2003, Nysewander2009}.
 
 \begin{figure}
-	\centering
-	\includegraphics[width=9cm]{figures/NH_z.pdf}
-	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations of GRBs.
-In red is shown the sample of lGRBs from \citet{Arcodia2016} where detections are replaced with a kernel density estimate of the points and the limits on $N_\mathrm{H}$ are shown with arrows.
-In blue is shown the complete sample of sGRB by \citet{DAvanzo2014a}, where again the limits are indicated by arrows.
-Marginalizations over both axes are shown onthe side and on the plot of the plot where the limits are shown as transparent bars and detections as solid ones.
-The red lines in the marginalization plots are again the kernel density estimates of the \citet{Arcodia2016} sample.}
-	\label{fig:NH_z}
+	\centering \includegraphics[width=9cm]{figures/NH_z.pdf} \caption{Rest frame,
+	X-ray derived equivalent hydrogen column density of GRB~111117A compared to
+	complete samples of both long and short populations of GRBs. The sample of
+	lGRBs from \citet{Arcodia2016} is shown in red, where detections are also shown
+	with a kernel density estimate of the points and the limits on $N_\mathrm{H,X}$
+	are shown with arrows. The complete sample of sGRB by \citet{DAvanzo2014a} is
+	shown in blue, where again the limits are indicated by arrows. Marginalizations
+	over both axes are shown on the right and on the top of the plot, where the
+	limits are shown as semi-transparent bars and detections as solid ones. The red
+	curves in the marginalization plots are again the kernel density estimates of
+	the \citet{Arcodia2016} sample.} \label{fig:NH_z}
 \end{figure}
-
-% \begin{figure}
-% 	\centering
-% 	\includegraphics[width=9cm]{figures/Eiso_z.pdf}
-% 	\caption{Isotropic equivalent energy for all the representative samples of both long and short GRBs. The sample is from \citet{Turpin2015} and the short is from \citet{DAvanzo2014a}.}
-% 	\label{fig:Eiso_z}
-% \end{figure}
 
@@
 
-From the clear host association, GRB~111117A does not belong to the hostless
-class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
-this is indicative of a population of relatively young stars. As the majority of
-sGRBs \citep{Fong2013b}, the host of GRB~111117A is therefore a late-type galaxy
-and is entirely consistent in terms of stellar mass and stellar age with the
-general population of sGRB hosts ($\left\langle M _* \right\rangle = 10^{10.1}
-M_{\odot}$ and $\left\langle \tau _* \right\rangle = 0.3 $~Gyr)
-\citep{Leibler2010}. Being a late-type host, both the stellar mass and sSFR are
-entirely within the range expected for the hosts of sGRBs \citep{Behroozi2014}.
-Our constraints on the dynamical mass is also well accommodated by the expected sGRB
-host halo mass \citep{Behroozi2014}.
+Because of the secure host association, GRB~111117A does not belong to the
+hostless class of sGRBs \citep{Berger2010a} and because the host exhibits
+emission lines this is indicative of a population of relatively young stars.
+Like the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is
+therefore a late-type galaxy and is entirely consistent in terms of stellar mass
+and stellar age with the general population of sGRB hosts ($\left\langle M _*
+\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
+0.3 $~Gyr; \citealt{Leibler2010}). Being a late-type host, both the stellar mass
+and sSFR are entirely within the range expected for the hosts of sGRBs
+\citep{Behroozi2014}. Our constraint on the dynamical mass is also well
+accommodated by the expected sGRB host halo mass \citep{Behroozi2014}.
 
@@
 lGRBs at a corresponding redshift \citep{Kruhler2015}. Only two hosts in the
-sample of short GRBs compiled in \citet{Berger2014} have a more vigorous star
-formation, meaning that it is in the very upper end of the star formation
+sample of short GRBs compiled by \citet{Berger2014} have a more vigorous star
+formation, placing it is in the very upper end of the star formation
 distribution. The cosmic SFR evolution of the universe likely plays a role due
@@
 determine. Additionally, it is natural to expect some evolution in the hosts of
-sGRBs with redshift as illustrated for $N_\mathrm{H}$ in Sect.~\ref{restnH}.
+sGRBs with redshift as illustrated for $N_\mathrm{H,X}$ in Sect.~\ref{restnH}.
 
 The simultaneus detection of \lya{} and \ha{} allows us to put constraints of
-the escape fraction of \lya{}, $f_{esc}$(\lya). Using the intrinsic ratio
-between \ha{} and \lya{}, assuming case B recombination
+the escape fraction of \lya{}, $f_{\mathrm{esc}}$(\lya). Using the intrinsic
+ratio between \ha{} and \lya{}, assuming case B recombination
 \citep{Brocklehurst1971}, and the measured fluxes from the spectrum, we find
-$f_{esc}$(\lya)$ < 0.06$. While the $f_{esc}$(\lya) scales with the dust column
-\citep{Hayes2011}, the resonant scattering of \lya{}-photons with neutral
-hydrogen makes the effective path length of \lya{} longer than \ha{}
-\citep{Atek2009}. This makes $f_{esc}$(\lya) an unreliable proxy for dust
-column, especially at low dust columns \citep{Atek2014} where the geometry and
-dynamics of the \hi{} within the galaxy will affect the \lya{} path the most.
-The $f_{esc}$(\lya) inferred for the host is entirely consistent with what is
-found for field galaxies with similar properties \citep{Oyarzun2017}. The same
-authors also find that \lya{} emitting galaxies have mostly little dust,
-consistent with what inferred from the SED-fit, see Sect. \ref{SED}. The
-centroid of the Ly$\alpha$ emission is found to be redshifted by
-$\sim$~$240\pm 90$~km~s$^{-1}$ with respect to systemic, which is similar to what is
-found for long GRB hosts \citep{Milvang-Jensen2012a} and Lyman Break Galaxies
-\citep{Shapley2003a} in which the outflow is attributed to star formation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Comparison to long GRBs at z $\sim$ 2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+$f_{\mathrm{esc}}$(\lya)$ < 0.06$. While the $f_{\mathrm{esc}}$(\lya) scales
+with the dust column \citep{Hayes2011}, the resonant scattering of
+\lya{}-photons with neutral hydrogen makes the effective path length of \lya{}
+longer than for \ha{} \citep{Atek2009}. This makes $f_{\mathrm{esc}}$(\lya) an
+unreliable proxy for dust column, especially at low dust columns
+\citep{Atek2014} where the geometry and dynamics of the \hi{} within the galaxy
+will affect the \lya{} path the most. The $f_{\mathrm{esc}}$(\lya) inferred for
+the host is entirely consistent with what is found for field galaxies with
+similar properties \citep{Oyarzun2017}. The same authors also find that \lya{}
+emitting galaxies have mostly little dust, consistent with what inferred from
+the SED fit(see Sect. \ref{SED}). The centroid of the Ly$\alpha$ emission is
+found to be redshifted by $\sim$~$240\pm 90$~km~s$^{-1}$ with respect to
+systemic, which is similar to what is found for long GRB hosts
+\citep{Milvang-Jensen2012a} and Lyman break galaxies \citep{Shapley2003a} in
+which the outflow is attributed to star formation.
 
@@
 
-The redshift distributions of GRBs are powerful informants on both the
-conditions which drive the formation of these events, but also on the potential
-influence these cosmic explosions have on the evolution of the universe. Due to
+The redshift distribution of GRBs is a powerful informant on both the conditions
+which drive the formation of these events, but also on the potential influence
+these cosmic explosions have on the evolution of the universe. Due to the
 elevated brightness of lGRBs compared to sGRBs \citep{Berger2014} and their
 tendency to be associated with the star-forming and therefore dense regions in
-their hosts \citep{Fruchter2006, Lyman2017}, the afterglows of lGRBs are easier
-to estimate redshift for than their sGRB counterparts where only a single burst
-have a redshift measurement from the GRB afterglow \citep{Cucchiara2013,
-	DeUgartePostigo2014c}. Correspondingly, the redshift distribution of lGRBs is
-better determined than that of sGRBs \citep[e.g., see][]{Jakobsson2012a,
+their hosts \citep{Fruchter2006, Lyman2017}, the redshifts of lGRBs are easier
+to measure than for sGRB counterparts, where only a single burst has a redshift
+measurement from the GRB afterglow \citep{Cucchiara2013, DeUgartePostigo2014c}.
+Correspondingly, the redshift distribution of sGRBs is still substantially
+unconstrained compared to that of lGRBs \citep[e.g., see][]{Jakobsson2012a,
 	DAvanzo2015, Perley2016d}.
@@
 A single sGRB at high redshift does little in terms of constraining the redshift
-distribution of sGRBs. In particular, other sGRB hosts could be missed because
-they are intrinsically fainter and thus the high redshift of GRB~111117A is only
-measured due to the brightness of its host. \citet{Berger2014} compiled a
-sample of sGRB host luminosities, normalized by the characteristic galaxy
-luminosity at their respective redshift, $L_B/L^{\star}_{B}$. To convert the
-SED-inferred $M_B$ of GRB~111117A to $L_B/L^{\star}_{B}$, we use the
-characteristic absolute $B$-band magnitude of the Schechter function for blue
-galaxies ($U - V < 0.25$) in the redshift window $2.0 \leq z \leq 2.5$ from
-\citet{Marchesini2007} and obtain $L_B/L^{\star}_{B} = 1.2$.
+distribution of sGRBs. In particular, other sGRB host redshifts could have been
+missed because their hosts are intrinsically fainter and thus the high redshift
+of GRB~111117A is only measured due to the brightness of its host.
+\citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by
+the characteristic galaxy luminosity at their respective redshift,
+$L_B/L^{\star}_{B}$. To convert the SED-inferred $M_B$ of GRB~111117A to
+$L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band magnitude of
+the Schechter function for blue galaxies ($U - V < 0.25$) in the redshift window
+$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and obtain $L_B/L^{\star}_{B}
+= 1.2$.
 
@@
 originating in star-forming galaxies, elliptical galaxies, and apparent hostless
-sGRBs. Out of the 14 hosts in the sample, 10 (71 per cent) have measured
-redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the host of
+sGRBs. Out of the 14 hosts in the sample, 10 (71 per cent) have both measured
+redshifts and $L_B/L^{\star}_{B}$. Compared to the complete sample, the host of
 GRB~111117A  is brighter than 80 per cent of the hosts with measured
@@
 
-If we assume that we are able to obtain emission-line redshifts from hosts, 0.5
-mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed
-60 per cent of the redshifts (6 out of 10 hosts), due too the host being too
-faint, were they at the redshift of GRB~11117A. The corresponding number is
-around 45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
-reflecting the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because
-the average SFR of galaxies hosting lGRBs is higher than for galaxies hosting
-sGRBs, the fraction of missed burst redshifts is likely higher although the
-cosmic SFR evolution could play a role in improving redshift determinability.
+If we assume that we are able to obtain emission-line redshifts from hosts which
+are at most 0.5 mag fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we
+would have missed 60 per cent of the redshifts (6 out of 10 hosts), due too the
+host being too faint, were they at the redshift of GRB~11117A. The corresponding
+number is around 45 per cent (12 out of 26) from the full sample of
+\citet{Berger2014}, reflecting the lower mean $L_B/L^{\star}_{B}$ of the
+complete sample. Because the average SFR of galaxies hosting lGRBs is higher
+than for galaxies hosting sGRBs, the fraction of missed burst redshifts is
+likely higher although the cosmic SFR evolution could play a role in improving
+redshift determinability at high $z$.
 
@@
 time delay models investigated by \citet{Wanderman2015} predict a sGRB rate at
-$z = 2.211$, $\sim$ two orders of magnitude lower compared to the peak rate at
+$z = 2.211$, around two orders of magnitude lower compared to the peak rate at
 $z = 0.9$. According to \citet{Wanderman2015}, this preference depends
-critically on the absence of sGRBs at $z \gtrsim 1.2$. The higher
-determined redshift of GRB~111117A, and the likely number of additional high-z
-sGRB could change the preferred time delay models. The redshift of GRB~111117A,
-on the other hand, is close to the expected peak in sGRB rate calculated using
-the power law delay time models \citep{Behroozi2014, Wanderman2015,
-	Ghirlanda2016}, meaning we could be missing a large fraction of sGRBs.
+critically on the absence of sGRBs at $z \gtrsim 1.2$. The higher determined
+redshift of GRB~111117A, and the likely number of additional high-$z$ sGRB could
+change the preferred time delay models. The redshift of GRB~111117A, on the
+other hand, is close to the expected peak in sGRB rate calculated using the
+power law delay time models \citep{Behroozi2014, Wanderman2015, Ghirlanda2016},
+meaning we would be missing a large fraction of sGRB redshifts.
 
@@
 by the current observation is to check if the implied sGRB rate at higher
-redshift can be contained in the observed sGRBs without redshifts and the
-potential number of missed bursts due to \textit{Swift}'s sensitivity limit. Of
-the  100  sGRBs observed by \textit{Swift}, 20 have secure redshifts, and
-another 7 has a tentative redshift measurement\footnote{This is based on
+redshift does not exceed the number of observed sGRBs without redshifts. Of the
+100  sGRBs observed by \textit{Swift}, 20 have secure redshifts, and another 7
+have a tentative redshift measurement\footnote{This is based on
 	\url{http://www.astro.caltech.edu/grbox/grbox.php}, selecting all short,
-	\textit{Swift}-detected GRBs up to GRB~170428A.}, meaning $> 73$ per cent of all
+	\textit{Swift}-detected GRBs up to GRB~170428A.}, meaning that $> 73$ per cent of all
 sGRBs observed with \textit{Swift} are missing redshifts. More recently,
 \citet{Fong2017} \citep{LIGOScientificCollaboration2017b} compiled a list of 36
-(33) sGRBs with redshift measurements and using this number, the redshift
-incompleteness of sGRBs decrease to 64 per cent. Additional to the potential
+(33) sGRBs with redshift measurements and, using this number, the redshift
+incompleteness of sGRBs decreases to 64 per cent. Additional to the potential
 number of high-$z$ event already detected but missing redshifts,
 \citet{Behroozi2014} parametrized the \textit{Swift} redshift sensitivity and
-found that mean detection probability for sGRBs at $z \sim2$ was only $\sim1$
+found that the mean detection probability for sGRBs at $z \sim2$ was only $\sim1$
 per cent of the mean detection probability at $z \sim1$, assuming that the
@@
 of the orbital decay of the system is set by the energy loss due to
-gravitational waves, which in turn is set by the mass of the constituent compact
-objects, the eccentricity of the orbit and the separation of the two
+GWs, which in turn is set by the mass of the constituent compact
+objects, the eccentricity of the orbit, and the separation of the two
 \citep{Postnov2014}. If we assume that the formation timescale of the first
 galaxies is short compared to the time since the Big Bang \citep{Richard2011},
-and that any binary neutron star formation channel can work sufficiently fast,
+and that any binary NS formation channel can work sufficiently fast,
 we can, assuming a mass of each of the NSs in a circular orbit at the time of
@@
 natal kicks. For more eccentric orbits, the coalescence timescale decreases,
-leading to a decrease in the constraint on the initial separation and larger
-initial separation constraints. As noted by \citet{Postnov2014}, it takes
-eccentricities > 0.6 to significantly shorten the merger time. Due to tidal
-interactions between the two NSs, the orbits will also tend to circularize with
-time, lessening the impact of the eccentricity on the constraint.
+leading to a larger initial separation constraint. As noted by
+\citet{Postnov2014}, it takes eccentricities > 0.6 to significantly shorten the
+merger time. Due to tidal interactions between the two NSs, the orbits will also
+tend to circularize with time, lessening the impact of the eccentricity on the
+constraint.
 
@@
 masses allowed. We show the grid in Fig. \ref{fig:prog_sep}. The double NS
-binary systems has a constituent mass distribution peaked at
-$1.33^{+0.10}_{-0.12}$ M$_\odot$ \citep{Kiziltan2013}, which corresponds to  $a_0 <
+binary systems have a constituent mass distribution peaked at
+$1.33^{+0.10}_{-0.12}~M_\odot$ \citep{Kiziltan2013}, which corresponds to  $a_0 <
 3.1^{+0.2}_{-0.2}~R_\odot$ where the errors are the 68 per cent posterior
@@
 As can be seen, a heavier system will inspiral faster, leading to a weaker constraint on the initial separation of the binary.
-The most probable value, the 68 per cent, and 98 percent posterior predictive intervals of the neutron star binary mass distribution from \citet{Kiziltan2013} are shown in the dashed ellipses, which corresponds to a constraint on maximal initial separation of $a_0 <3.1^{+0.2(+0.4)}_{-0.2(-0.4)}~R_\odot$, respectively.}
+The most probable value, the 68 per cent, and 98 percent posterior predictive intervals of the NS binary mass distribution from \citet{Kiziltan2013} are shown with the dashed ellipses, which corresponds to a constraint on maximal initial separation of $a_0 <3.1^{+0.2(+0.4)}_{-0.2(-0.4)}~R_\odot$, respectively.}
 	\label{fig:prog_sep}
@@
 
-Using the inferred stellar population age from our SED fit, then we obtain a
-(softer) limit on the initial separation of $a_0 < 2.0^{+0.4}_{-0.4}~R_\odot$.
+Using the inferred stellar population age from our SED fit, we obtain a less
+robust limit on the initial separation of $a_0 < 2.0^{+0.4}_{-0.4}~R_\odot$.
 However, this does not account for the possibility there could be an underlying
-stellar population of older stars from a previous star-formation epoch. To
+stellar population of older stars from a previous star-formation episode. To
 investigate the possible impact of the presence of an old stellar population, we
@@
 template to which an additional stellar population of old stars was added. For
-each galaxy, this old population was set as the SSPs with same parameters that
-the best-fit SED excepts the age, which was set to the age of the Universe at
+each template, this old population was set as the SSP with the same parameters as
+the best-fit SED except the age, which was set to the age of the Universe at
 the observed redshift. In principle, this can constrain the maximum contribution
@@
 This is especially true given the late type nature of the host
-\citep{OShaughnessy2008}. The same holds, if neutron star binaries are primarily
-formed through dynamical interactions in globular clusters \citep{Lee2010,
-	Church2011}.
+\citep{OShaughnessy2008}. The same holds, if NS binaries are primarily formed
+through dynamical interactions in globular clusters \citep{Lee2010, Church2011}.
 
@@
 short GRB~111117A based on host galaxy emission lines setting it at $z = 2.211
-\pm 0.001$. This value supersedes the previous photometric redshift of
-\citet{Margutti2012} and \citet{Sakamoto2013}. The erroneous redshift estimate
-of previous authors is attributed to a discrepancy in the measured $z$-band
-magnitude, and highlights the importance of deep spectroscopic studies of sGRB
-hosts at medium resolution.
+\pm 0.001$. This value supersedes the previous photometric redshift estimates of
+\citet{Margutti2012} and \citet{Sakamoto2013}. The erroneous best-fit SED
+redshift of previous authors is attributed to a discrepancy in the measured
+$z$-band magnitude, and highlights the importance of deep spectroscopic studies
+of sGRB hosts at medium resolution.
 
-Using the new distance, the X-ray derived $N_\mathrm{H}$ towards GRB~111117A is the
-highest within a complete sample of sGRBs and is consistent with the
-$N_\mathrm{H}-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
-upper end of the sGRB host SFR distribution and no significant amount of dust is
-present. The high $N_\mathrm{H}$ is difficult to reconcile with the large projected host
-offset and the absence of dust. One possible explanation could be, that
-GRB~111117A is formed through a prompt channel of sGRB formation and originates
-in a star forming region located in the outskirts of the host.
+Using the new distance, the X-ray derived $N_\mathrm{H,X}$ towards GRB~111117A
+is the highest within a complete sample of sGRBs and is consistent with the
+$N_\mathrm{H,X}-z$ evolution traced by the sight lines of lGRBs. The SFR of the
+host is in the upper end of the sGRB host SFR distribution and no significant
+amount of dust is present. The high $N_\mathrm{H,X}$ is at odds with the large
+projected host offset and the absence of dust. One possible explanation could be
+that GRB~111117A is formed through a prompt channel of sGRB formation and
+originates in an (unseen) star forming region located in the outskirts of the
+host, or a localized region of high \hi~density along the line of sight.
 
@@
 accommodated in progenitor models. A lognormal delay time model predicts a very
-low volumetric density of bursts at $z = 2.211$, whereas a power law delay time
-model peaks near GRB~111117A. If more sGRBs are at similarly high redshifts, but
-are missed due to the faintness of their hosts, a lognormal delay time model
-will be disfavored. Compared to a complete sample of \textit{Swift}-detected
-sGRB, the host of GRB~111117A is more luminous than 80 per cent of sGRB hosts
-with measured luminosities. Assuming a host brightness redshift determination
-threshold, for between 43 and 71 per cent of the sample hosts, we would be unable to
-determine a redshift should they be at a similar redshift as GRB~111117A. This
-could indicate that potentially, a significant fraction of
-\textit{Swift}-detected sGRBs are at high $z$, but missed due to host faintness.
+low volumetric density of bursts at $z \sim 2$, whereas a power law delay time
+model peaks near the GRB~111117A redshift. If more sGRBs are at similarly high
+redshifts, but are missed due to the faintness of their hosts and afterglows, a
+lognormal delay time model will be disfavored. Compared to a complete sample of
+\textit{Swift}-detected sGRB, the host of GRB~111117A is more luminous than 80
+per cent of sGRB hosts with measured luminosities. Assuming a host brightness
+redshift determination threshold, for between 43 and 71 per cent of the sample
+hosts, we would be unable to determine a redshift should they be at a similar
+redshift of GRB~111117A. This could indicate that, potentially, a significant
+fraction of \textit{Swift}-detected sGRBs are at high $z$, but with redshifts
+unknown due to host faintness.
 
@@
 time of formation. We find that the maximal separation of two NSs at system
-formation time is $a_0 < 3.2~R_\odot$, which excludes some of the formation
+formation time is $a_0 < 3.1~R_\odot$, which excludes some of the formation
 channels with the longest timescales.
@@
 %
-SDV is supported by the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro 
+SDV is supported by the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro.
 %
-PDA, SCo, acknowledge support from ASI grant I/004/11/3
+PDA, SCo, acknowledge support from ASI grant I/004/11/3.
 %
@@
 NRT and KW acknowledge support from STFC Consolidated
-Grant ST/N000757/1
+Grant ST/N000757/1.
 %
@@
 AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation
-Grant for Researchers and Cultural Creators,and the Spanish Ministry of Economy
-and Competitiveness through project AYA2014-58381-P. Partly based on
-observations made with the Gran Telescopio Canarias (GTC).
+Grant for Researchers and Cultural Creators, and the Spanish Ministry of Economy
+and Competitiveness through project AYA2014-58381-P.
 %
@@
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{TheAstropyCollaboration2013}.
-The analysis and plotting has been achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy and Scipy \citep{VanderWalt2011} along with other community-developed packages.
+The analysis and plotting has been achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy, and Scipy \citep{VanderWalt2011}, along with other community-developed packages.
 %

### 2017-11-25 tables/phot_overview.tex
Last edits before resubmission

@@
 		\hline\hline
-{Obs. Date} &  Exptime & Telescope/Instrument & Filter & Airmass & Image Quality & Host Brightness\tablefootmark{a}  \\ [1.5pt]
+{Observation epoch} &  Exptime & Telescope/instrument & Filter & Airmass & Image quality & Host brightness\tablefootmark{a}  \\ [1.5pt]
         \hline
-{} & {ks} &    & {} & & (arcsec)  & (mag$_{\mathrm{AB}}$)  \\ [1.5pt]
+({UT}) & ({ks}) &    & {} & & (arcsec)  & (AB mag)  \\ [1.5pt]
 		\hline
-2013-08-30T07:43 & 1.45 & VLT/FORS2 & $g$ & 1.55 & 0.99 & $24.08\pm 0.09$ \\
-2011-11-17T20:07 & 0.80 & GTC/OSIRIS & $g$ & 1.15 & 1.67 & $24.13\pm 0.09$ \\
-2011-11-17T20:07 & 1.20 & GTC/OSIRIS & $r$ & 1.11 & 1.50 & $23.93\pm 0.08$ \\        
-2013-07-17T08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
-2011-11-28T21:10 & 3.60 & TNG/DOLORES & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
-2011-11-17T20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   
-2013-08-03T09:23 & 1.35 & VLT/FORS2 & $I$ & 1.54 & 0.93 & $24.22\pm 0.15$ \\           
-2011-11-28T06:14 & 1.80 & Gemini/GMOS-N & $z$ & 1.01 & 0.84 & $24.24\pm 0.47$ \\  
-2013-07-13T09:33 & 1.08 & VLT/FORS2 & $z$ & 1.49 & 0.63 & $23.76\pm 0.21$ \\             
-2013-06-24T09:14 & 1.98 & VLT/HAWK-I & $J$ & 1.70 & 0.63 & $23.13\pm 0.18$ \\        
-2013-06-27T09:21 & 1.68 & VLT/HAWK-I & $H$ & 1.63 & 0.91 & $22.94\pm 0.29$ \\   
-2013-06-28T09:14 & 1.92 & VLT/HAWK-I & $K_\mathrm{s}$ & 1.65 & 0.76 & $23.07\pm 0.32$ \\   
+2013-08-30 07:43 & 1.45 & VLT/FORS2 & $g$ & 1.55 & 0.99 & $24.08\pm 0.09$ \\
+2011-11-17 20:07 & 0.80 & GTC/OSIRIS & $g$ & 1.15 & 1.67 & $24.13\pm 0.09$ \\
+2011-11-17 20:07 & 1.20 & GTC/OSIRIS & $r$ & 1.11 & 1.50 & $23.93\pm 0.08$ \\        
+2013-07-17 08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
+2011-11-28 21:10 & 3.60 & TNG/DOLORES & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
+2011-11-17 20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   
+2013-08-03 09:23 & 1.35 & VLT/FORS2 & $I$ & 1.54 & 0.93 & $24.22\pm 0.15$ \\           
+2011-11-28 06:14 & 1.80 & Gemini/GMOS-N & $z$ & 1.01 & 0.84 & $24.24\pm 0.47$ \\  
+2013-07-13 09:33 & 1.08 & VLT/FORS2 & $z$ & 1.49 & 0.63 & $23.76\pm 0.21$ \\             
+2013-06-24 09:14 & 1.98 & VLT/HAWK-I & $J$ & 1.70 & 0.63 & $23.13\pm 0.18$ \\        
+2013-06-27 09:21 & 1.68 & VLT/HAWK-I & $H$ & 1.63 & 0.91 & $22.94\pm 0.29$ \\   
+2013-06-28 09:14 & 1.92 & VLT/HAWK-I & $K_\mathrm{s}$ & 1.65 & 0.76 & $23.07\pm 0.32$ \\   
 \hline\noalign{\smallskip}

### 2017-11-25 tables/spec_overview.tex
Last edits before resubmission

\begin{table*}[!ht]

	\centering
	\caption{Overview of the spectroscopic observations. ``JH'' in the slit width refers to observations where a $K$-band blocking filter has been used. The seeing is determined from the width of the spectral trace of a telluric standard star, observed close in time to the host observation. The spectral resolution, $R$, is measured from unresolved telluric absorption lines in the spectrum of the telluric standard star. \label{tab:spec_overview}}
	\begin{tabular}{cccccccc}
		\hline\hline
		{Observation epoch} &  \multicolumn{3}{c}{Exposure time (s)} & Slit width & Airmass & Seeing & $R$   \\ [1.5pt]
        \hline
		(UT) & UVB  & VIS & NIR &  (arcsec)   & {} & (arcsec)  & {VIS/NIR}  \\ [1.5pt]
		\hline
		2011-11-19 01:33 & 2 $\times$ 2400 & 2 $\times$ 2400 & 8 $\times$ 600 & 1.0/1.0/0.9 & 1.49 & 0.75 & 11600/6700 \\
        2013-07-15 09:02 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.53 & 0.98 & 9600/8900 \\
        2013-08-03 07:37 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.55 & 0.85 & 11400/11300 \\
        2013-08-03 08:34 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.49 & 0.85 & 11400/11300 \\

		\hline\noalign{\smallskip}

	\end{tabular}

\end{table*}

### 2017-11-25 py/NH_z.py

@@ -5,6 +5,7 @@ from __future__ import division, print_function
 
 from astropy.io import fits
 import pandas as pd
+import matplotlib; matplotlib.use('TkAgg')
 import matplotlib.pyplot as pl
 import seaborn as sns; sns.set_style('ticks')
 import numpy as np
@@ -104,10 +105,10 @@ def main():
     colors = [sns.set_hls_values(color_rgb, l=l) for l in np.linspace(1, 0, 12)]
     cmap = sns.blend_palette(colors, as_cmap=True)
     g = g.plot_marginals(sns.distplot, hist=False, color=color)
-    g = g.plot_joint(sns.kdeplot, cmap=cmap, label="Arcodia et al. 2016", zorder=1)
+    g = g.plot_joint(sns.kdeplot, cmap=cmap, label="Arcodia et al. 2016", zorder=1, lw=0.7, alpha=0.6)
     # g = g.plot_joint(pl.scatter, cmap=cmap, label="Arcodia et al. 2016", zorder=1)
     # g = g.plot_marginals(sns.distplot, hist=False, rug=True, kde=False, color=color, rug_kws={"height": 0.7, "lw": 2.0})
-    # g = g.plot_joint(pl.scatter, color=color, label="Arcodia et al. 2016")
+    g = g.plot_joint(pl.scatter, color=color, marker="h", s=20, lw=1.0, alpha=0.8)
     print(z_long_uplim, NH_long_uplim)
     g.x = z_long_uplim
     g.y = NH_long_uplim
@@ -122,18 +123,6 @@ def main():
 
 
 
-    color = sns.color_palette()[1]
-    g.x = np.array([2.211])
-    g.y = np.array([np.log10(2.4e22)])
-    # g.y = np.array([np.log10(2.4e22+2.4e22)])
-    gyupperr = np.log10(2.4e22)/np.log10(2.4e22) # np.log10(2.4e22) - np.log10(1.6e22)
-    gylowerr = np.log10(1.6e22)/np.log10(2.4e22) #  np.log10(4.8e22) - np.log10(2.4e22)
-    g = g.plot_joint(pl.scatter, marker = "*", s = 150, color=sns.color_palette()[1], label="GRB111117A")
-
-    g = g.plot_marginals(sns.rugplot, color=color, height = 1.0, lw = 4.0)
-    g.x = np.array([2.211, 2.211])
[…]

### 2017-11-25 py/SED_plot.py

@@ -80,8 +80,8 @@ def main():
 
     # ax1 = pl.gca()
     ax1.invert_yaxis()
-    ax1.set_xlabel(r"Wavelength [$\mathrm{\AA}$]")
-    ax1.set_ylabel(r'Brightness [AB$_{mag}$]')
+    ax1.set_xlabel(r"Observed wavelength [$\mathrm{\AA}$]")
+    ax1.set_ylabel(r'Brightness [AB mag]')
     ax1.set_ylim(26.1, 17.5)
 
     # ax2 = ax1.twinx()
@@ -91,7 +91,7 @@ def main():
     print(mn)
     ax2.set_ylim(10**((23.9 - mn)/2.5), 10**((23.9 - mx)/2.5))
     ax2.semilogy()
-    ax2.set_ylabel(r'F$_\nu$ [$\mu$Jy]')
+    ax2.set_ylabel(r'Flux density [$\mu$Jy]')
 
     # pl.ylim(-1 * scale, 10 * scale)
 

### 2017-11-25 py/progenitor_seperation.py

@@ -68,7 +68,7 @@ def main():
     fig, ax = pl.subplots()
     im = ax.imshow(a0, extent=[min(masses), max(masses), min(masses), max(masses)], cmap="viridis_r")
     cbar = pl.colorbar(im)
-    cbar.set_label(r'a$_0$ [R$_\odot$]')
+    cbar.set_label(r'$a_0$ [$R_\odot$]')
 
 
     from matplotlib.patches import Ellipse
@@ -78,8 +78,8 @@ def main():
 
 
 
-    ax.set_xlabel(r"M$_{\mathrm{NS}}$ [M$_\odot$]")
-    ax.set_ylabel(r"M$_{\mathrm{NS}}$ [M$_\odot$]")
+    ax.set_xlabel(r"$M_{\mathrm{NS}}$ [$M_\odot$]")
+    ax.set_ylabel(r"$M_{\mathrm{NS}}$ [$M_\odot$]")
     pl.tight_layout()
     pl.savefig("../figures/prog_sep.pdf")
     pl.show()

### 2017-11-26 referee_reply.txt
Last edits

@@
 --1st sentence: Define sGRB, NS.
+
 JS: Done.
@@
 --For detection of kilonovae, also cite Berger+ 2013.
-JS: Definitely.
+
+JS: Definitely. This was an oversight.
 
 --"duration T_90 < 2" - make sure to qualify with 2 sec
+
 JS: Done.
@@
 GRB. I suggest using one or the other throughout the text.
+
 JS: We have streamlined this a bit, but have kept the wording "long and short GRBs"
@@
 low redshifts could be a natural consequence of relatively long delay times.
+
 JS: We have cautioned the wording.
@@
 references therein)
+
 JS: We have changed the wording to reflect this.
@@
 --"will provide new a promising ways" -- delete "a"
+
 JS: Done
@@
 assumption that there is negligible afterglow contribution from epoch 1.
+
 JS: We do not see any spectral evolution. This is now explicitly stated in the text.
@@
 included.
+
 JS: We have added a description of this.
@@
 strong detection of [OIII] 5007?
+
 JS: [OIII] 4959 is detected, but very close to a sky-line. This has now been added to the text.
@@
 a lower limit
+
 JS: Correct! This has been changed and expanded.
@@
 and in what band. Are these the deepest limits available?
+
 JS: To our knowledge, these are the deepest limits available, yes. The limits have now been explicitly stated.
@@
 not figure out the quoted values of 0.8 to 1.0 mag.
+
 JS: We have changed this to ~ 1 mag. Both our z-band measurements are fainter than the z-band magnitudes reported by other authors.
@@
 Fig. 3: best-fist --> best-fit
+
 JS: Done
@@
 --"had a luminosity of 44.95" --> should appear as a power of 10 and with units
+
 JS: This has been updated.
@@
 --"GRB 111117A has a 96% probability of being a sGRB"
+
 JS: Done
@@
 seems like a sGRB. To be more accurate I would say "non-collapsar"
+
 JS: We have removed the usage of non-collapsar and stuck with sGRB.
@@
 other non-collapsar GRB ever detected"
+
 JS: Done
@@
 do you consider as a "non-collapsar GRB"?
+
 JS: We have stated a f_NC > 50 per cent requirement of the Bromberg classification to consider a given burst short.
@@
 sentence needs a citation and would be helpful to state what burst this is
+
 JS: We have expanded this entire section a bit to be more complete.
@@
 paper for completeness.
+
 JS: Done
@@
 suggest revising to "If the assumption of no dust is correct..."
+
 JS: This has been changed to "low dust content"
@@
 afterglow limits still consistent with a high-density environment at z=2.211?
+
 JS: The X-ray afterglow is not strongly dependent of the density and because of the optical
@@
 
-Margutti+ didn't use the light curve from the Swift Burst Analyser. Her early light
+JS: Margutti+ didn't use the light curve from the Swift Burst Analyser. Her early light
 curve is ~0.18 dex brighter. The shape is also a bit different. k-correction:
@@
 cosmology are taken into account).
-
 All in all, logL(us) - logL(Margutti) = (0.58 - 0.18 - 0.63) dex = -0.23 dex.
-
 Comparing her plot with ours gives difference of ~-0.2 dex, which is identical
@@
 side of the plot?
+
 JS: The caption has been expanded to better explain the different
@@
 separation..." - Is there a citation for the equation used to calculate this?
+
 JS: The equation is in the Postnov reference. The actual calculation is included in
@@
 binary, or more/less massive binary)?
+
 JS: We have included the calculation for different mass assumptions and changed the text accordingly.
@@
 cluster as suggested by some works (e.g., Lee+ 2010; Church+ 2011)
+
 JS: We do not assume anything about the formation channel, only that the neutron
@@
 hosts"
+
 JS: We have changed the wording.

### 2017-11-26 GRB111117A.tex
Last edits

@@
 regions of dense gas), see for example,
-\citep[e.g.][]{Galama1998,Watson2013,Krongold2013}. Alternatively, it has been
+\citep[e.g.][]{Watson2013, Krongold2013}. Alternatively, it has been
 hypothesised that a significant contribution to the observed X-ray
@@
 intervening systems along the line of sight of the GRB \citep{Campana2012,
-	Arcodia2016}, although see \citep{Watson2013, Krongold2013} for counter
-arguments.
+	Arcodia2016}, although see \citep{Watson2013, Krongold2013}.
 
@@
 been observed occasionally for lGRBs \citep{Perley2013a}. For the one sGRB where
-both parameters were measured (GRB~130603B; \citealt{DeUgartePostigo2014d}),
+both parameters were measured (GRB~130603B; \citealt{DeUgartePostigo2014b}),
 they were found to be consistent with $A_V \sim 1$~mag.
@@
 to measure than for sGRB counterparts, where only a single burst has a redshift
-measurement from the GRB afterglow \citep{Cucchiara2013, DeUgartePostigo2014c}.
+measurement from the GRB afterglow \citep{Cucchiara2013, DeUgartePostigo2014b}.
 Correspondingly, the redshift distribution of sGRBs is still substantially

### 2018-08-28 Arxiv/GRB111117A.tex
Published

%                                                                 aa.dem
% AA vers. 8.2, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[rnote]{aa} % for the research notes
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
% according to the author-year natbib style

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
%\documentclass[referee]{aa}

\usepackage{graphicx,url,twoopt,natbib}
\usepackage[varg]{txfonts}           %% A&A font choice

\usepackage{pdfcomment}              %% for popup acronym meanings
\usepackage{acronym}                 %% for popup acronym meanings

\usepackage{url}
\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0.3,0.7}
\hypersetup{colorlinks=true,
	linkcolor=linkcolor, 
	citecolor=linkcolor,
	filecolor=linkcolor, 
	urlcolor=linkcolor}

\usepackage{natbib}

\bibpunct{(}{)}{,}{a}{}{,}    %% natbib cite format used by A&A and ApJ

% Manually specified definitions
\input{definitions.tex}

\usepackage{xcolor}
\newcommand\todo[1]{\textbf{(#1)}}

\begin{document}

\title{The host galaxy of the short GRB~111117A at $z = 2.211$: impact on the short GRB redshift distribution and progenitor channels\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}}

\titlerunning{GRB~111117A}

\input{authorlist.tex}

\date{Received/ accepted}

\authorrunning{Selsing et al.}

\abstract{
It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and
their hosts to measure their redshifts. These measurements, however, are
critical to constrain the nature of sGRB progenitors, their redshift
distribution and the $r$-process element enrichment history of the universe.
Here, we present spectroscopy of the host galaxy of GRB~111117A and measure its
redshift to be $z = 2.211$. This makes GRB~111117A the most distant
high-confidence short duration GRB detected to date. Our spectroscopic redshift
supersedes a lower, previously estimated photometric redshift value for this
burst.

We use the spectroscopic redshift, as well as new imaging data to constrain the
nature of the host galaxy and the physical parameters of the GRB. The rest-frame
X-ray derived hydrogen column density, for example, is the highest compared to a
complete sample of sGRBs and seems to follow the evolution with redshift as
traced by the hosts of long GRBs (lGRBs).

The host lies in the brighter end of the expected sGRB host brightness distribution
at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
redshift determination, we find that between 43 and 71 per cent of all sGRB redshifts
should be missed due to host faintness for hosts at $z\sim2$. The high redshift of
GRB~111117A is evidence against a lognormal delay-time model for sGRBs through
the predicted redshift distribution of sGRBs, which is very sensitive to
high-$z$ sGRBs.

From the age of the universe at the time of GRB explosion, an initial neutron
star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
progenitor system is a circular pair of inspiralling NSs. This constraint
excludes some of the longest sGRB formation channels for this burst.
}

\keywords{gamma-ray burst: individual: GRB~111117A -- gamma-ray burst: general -- galaxies: high-redshift -- binaries: general -- X-rays: bursts -- techniques: imaging spectroscopy}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

There is mounting evidence that short-duration $\gamma$-ray bursts
come from the merger of NSs, either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The classification of GRBs in two groups, initially comes from the bimodal
distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
where the duration T$_{90} < 2$ has been regarded as the dividing line between long
and short GRBs. Additionally, it has been found for lGRBs that there is a
spectral lag in the arrival-time of photons, with the most energetic ones
arriving first. This lag is consistent with zero for sGRBs
\citep{Norris2006}. Because both populations have continuous, overlapping
distributions in their observables and because telescopes observe in differing
bands, it is difficult to impose a single demarcation criterion between the two
classes. For this reason, the distinction between long and short GRBs is
preferably based on a combination of high-energy properties \citep{Zhang2009,
	Bromberg2012a, Bromberg2013}.

The \textit{Swift} satellite \citep{Gehrels2004} greatly improved the
understanding of sGRB progenitors thanks to its quick localization capability.
The bulk of these localizations have associated galaxies at relatively low
redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, and because
most of these measurements come from the associated hosts, it is arguably biased
towards lower redshifts. The host galaxies of sGRBs are diverse. They are more
massive and less actively star-forming on average than lGRB hosts
\citep{Fong2013b}, while in some cases, no host galaxy can be identified above
detection threshold of deep follow-up observations \citep{Berger2010a,
	Tunnicliffe2014}. Together with their position within their hosts
\citep{Fong2013a}, this suggests a progenitor system that can be very long lived
in comparison to lGRBs, and is tracing with host stellar mass rather than
star-formation rate (SFR).

The electromagnetic signals from sGRBs are a promising channel to accurately
localize NS mergers, which holds the promise for a detection of an associated
gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
of a sGRB and a GW will provide new promising ways to constrain the binary
inclination angle \citep{Arun2014} and measure cosmological distances
\citep{Nissanke2010}.

The total lifetimes of NS binaries depends on their orbit, mass, spin, initial
separations and subsequent inspiral times. The delay time from formation to
explosion impacts the timing and distribution of the enrichment of the ISM with
heavy $r$-process elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some
limits can be calculated using host galaxy star-formation history models and
spatial distribution of sGRBs within their hosts \citep[][]{Berger2014}. The
most distant cosmological bursts, however, offer direct, hard limits on the
coalescence time scales.

We here present a spectrum of the host galaxy of the short
GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This
value is significantly higher than the previously estimated redshift based on
photometric studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest
frame properties based on this new distance compared to previous analyses and
revisit the host properties derived from the new solution to the spectral energy distribution (SED) fit.
While no optical afterglow was detected, the excellent localization from a
detection of the X-ray afterglow by the \emph{Chandra X-ray Observatory}
allows us to discuss the positioning and environmental properties of this
remarkably distant sGRB.
We use the $\Lambda$CDM cosmology parameters provided by
\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
and $\Omega_m = 0.307$. All magnitudes are given in the AB system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observations and results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Spectroscopic observations and analysis}

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
	\caption{
	FORS2 R-band imaging of the field of GRB~111117A with the X-shooter slit overlaid. The slit
	position represents 4 epochs of spectroscopic observations taken at similar
	position angles. The corresponding photometry
	is shown in Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as
	derived from the \emph{Chandra} observations in \citet{Sakamoto2013}. 
	}
	\label{fig:spec_setup}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/OIII_img.pdf}
	\caption{2D-image of the \oiii$\lambda$5007 emission line. The location of a bright skyline is marked by the blue box. The location of the emission line is indicated with the red ellipse. Because the host is observed in nodding-mode, negative images of the emission line appear on both sides in the spatial direction.}
	\label{fig:line}
\end{figure}

\begin{figure*}
	\centering
	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
	\caption{SED fit showning the best-fist SED to the derived photometry. The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is the observed spectrum, binned down to 3 \AA~per pixel for presentation purposes. Slit losses are responsible for discrepancies between the measured photometry and the X-shooter spectrum. The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
	\label{fig:SED}
\end{figure*}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA~to
24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run in physical
mode. Because the echelle orders are curved across each detector, a
rectification algorithm is employed which introduces correlations between
neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
maximal resolution. The observations are combined and extracted using scripts
described in Selsing et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
where the full spectral point spread function is modeled across each arm and
used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
slit position is shown in Fig.~\ref{fig:spec_setup}. We show the extracted
spectrum in Fig.~\ref{fig:SED}.

\input{tables/spec_overview.tex}

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$5007,
and \ha. \hb~is detected at low significance ($\sim 3 \sigma$). We show
\oiii$\lambda$5007 in the insert in Fig.~\ref{fig:line}.  \ha~is only visible in
the first epoch due to the $K$-band blocking filter used for the remaining
observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and
show significant velocity structure along the slit. A drop in the continuum
bluewards of the \lya~line further supports the inferred redshift.

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed
at 21\,000~\AA~where the night sky is very bright. In addition, several bright
sky-lines are superposed on the line, making an accurate estimate of the
\ha-flux difficult. We obtain a limit on the SFR by numerically integrating the
part of \ha~free of contamination and correcting for the missing flux using the
line profile and obtain $F_{H \alpha} > 4.1 \times 10^{-17} \mathrm{erg}
\mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting the \citet{Kennicutt1998}
relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$.
Additionally, based on the width of \oiii, an aperture covering the line is
integrated over, where synthetic sky lines \citep{Noll2012, Jones2013} have been
masked and interpolated over. From the integrated \ha-line, we estimate SFR =
$18 \pm 3$ M$_\odot$ yr$^{-1}$. From the SED-fit (Sect.~\ref{SED}), and the
detection of \lya, the host is constrained to contain very little or no dust,
although the presence of Lya does not exclude dust. Therefore we do not apply a
dust-correction to the measured \ha~flux here. \oii~is close to a region of
strong telluric absorption, which is why no SFR is inferred from this line.

The total extent of the lines in velocity space is $\sim$ 450 km/s. The line
profiles shows an asymmetric "double-horned" profile, indicating that we are
seeing a galaxy with a large degree of coherent rotational motion relative to
the line-of-sight. If we assume that we are viewing a spiral galaxy edge-on,
this is a measure of the rotational velocity of the gas. If we assume that the
spectral resolution and the turbulent width of the lines are negligible compared
to the rotational velocity, we can, based on the projected size of the source
and the width of the lines, put a constraint on the dynamical mass of the galaxy
\citep{DeBlok2014}. Based on the physical size along the slit and the velocity
width of \oiii$\lambda5007$, we infer M$_\text{dyn} \gtrsim 10^{10.8}$
M$_\odot$. Because we are viewing the host inclined at an angle relative to
edge-on and because the slit is not aligned along the long axis of the host,
this value is a lower limit.

\subsection{Imaging observations and SED analysis} \label{SED}

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, see
Sect. \ref{xray}, there is likely no afterglow contamination on the measured
photometry.

All data were reduced, analyzed and fitted in a similar manner as described in
detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
reduction which includes bias/flat-field correction, de-fringing (if necessary),
sky-subtraction, and stacking of individual images. The photometry of the host
was calibrated relative to field stars from the SDSS and 2MASS catalogs in the
case of $grizJHK_{\mathrm{s}}$ filters.
For the $R$ and $I$-band photometry, we used the color transformations of
Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system, and correct for a Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$ \citep{Schlegel1998, Schlafly2011}.

\input{tables/phot_overview.tex}

The multi-color SED is fit using the \citet{Bruzual2003} single stellar population
models based on a \citet{Chabrier2003} with initial mass function in
\emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to the
spectroscopic value of $z=2.211$. The best fit model is an unreddened galaxy
template, and returns physical parameters of absolute magnitude
($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$),
stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and star-formation rate
($SFR_{\mathrm{SED}}=11_{-4}^{+9} M_\odot\,\mathrm{yr}^{-1}$). We show the SED
fit in Fig.~\ref{fig:SED}.

Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
re-analysis of the Gemini data to the $z$-band measurements of
\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
$z$-band photometry that is brighter by 0.8~mag to 1.0 mag compared to our
value, where data taken in bluer filters are in excellent agreement. The large
$i-z$ color was mistakenly interpreted as a 4000\,\AA\,break driving the galaxy
photometric redshift of the earlier works. Using the revised photometry from
Table~\ref{tab:phot_overview}, the photometric redshift of the galaxy is
$z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the spectroscopic
value at the 1~$\sigma$ confidence level.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{X-ray temporal and spectral analysis}\label{xray}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. 
The X-ray afterglow light curve can be fit with a single power-law decay with an
index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
analysis using data obtained in photon counting (PC) mode in the widest time epoch where
the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
data) to prevent spectral changes that can affect the X-ray column density
determination \citep{Kopac2012}. 
The obtained spectrum is well described by an absorbed power-law
model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic equivalent hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a solar abundance and a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.

A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reinterpretation of the restframe properties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Because the projected distance does not change significantly between $z = 1.3$ and $z =
2.211$, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013}
relating to host offset are unaffected.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Classification} \label{classification}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As pointed out by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is
securely classified as a sGRB. Because the observed classification indicators,
$T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation.
The intrinsic spectral lag shortens, but since it is already consistent with
zero, this does not affect the classification.

The intrinsic luminosity is shown in the X-ray light curve (Fig.
\ref{fig:sxray_lightcurve}) and it is sub-luminous compared to the majority of
long GRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
distribution at 10 ks. The sub-samples comprise of 333 long, 19 short GRBs, and
GRB~111117A. The mean and the 1-$\sigma$ dispersions of the samples are
L$_{lGRB} = 46.59 \pm 0.87$ and L$_{sGRB} = 44.96 \pm 0.94$. GRB~111117A had a
luminosity of 44.95 at 10 ks which is very close to the peak of the sGRB
luminosity distribution at 10 ks, but an outlier from the lGRG distribution,
further supporting the short classification.

\citet{Bromberg2013} investigated the degree to which the long and short
population distributions overlap and quantified the certainty in class
membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
percent probability of being a sGRB. Compared to the other two sGRB candidates
at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
per cent), the certainty in class membership for GRB~111117A is much higher.

Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
intermediate duration bursts. This comes from both events having very soft
spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
bursts are very clearly related in their properties to lGRBs
\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
mergers.

The redshift and secure classification of GRB~111117A mean that it occurred
when the universe was younger by 1.8 Gyr compared to any other non-collapsar GRB
ever detected. This number is 3.2 Gyr for the next-highest spectroscopic
redshift. If the merger of NSs is the primary agent for the $r$-process element
enrichment of the universe \citep{Goriely2011, Ji2016, Komiya2016}, this
marks the earliest detection of this process.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
	\caption{Restframe XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs. The grey shaded region is a compilation of long GRB lightcurves \citep{Evans2007, Evans2009} where the color represents density. The light blue lines are sGRB lightcurves from bursts with duration of $T_{90} \lesssim 2$ s and those that were classified as short in \citet{Kann2011, Berger2014, DAvanzo2014a}. The thick green line is GRB~111117A. Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and subluminous compared to the lGRB population.}
	\label{fig:sxray_lightcurve}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restframe $N_\mathrm{H}$} \label{restnH}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We show the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare
with the distributions of complete samples of both long and short GRBs. The lGRB
sample is from \citet{Arcodia2016} and the sGRB sample is from
\citet{DAvanzo2014a}. From the sGRB sample of \citet{DAvanzo2014a} we have
excluded GRB~090426 which does likely not belong in a short sample as highlighted in
Sect. \ref{classification}. Both comparison samples consider the largest
temporal interval of constant hardness ratio to prevent spectral changes that
can affect the X-ray derived column density. The 17 (5) of the 99 (15) long
(short) GRBs which do not have redshift have been excluded from our analysis.

GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
$N_H$ and highest redshift of all sGRBs. The short sample, excluding
GRB~111117A, is located at low redshifts ($z < 1$) and is found to populate a
column density environment similar to that of lGRBs at comparable redshifts
\citep{DAvanzo2014a}. The inferred hydrogen column density for GRB~111117A seems
to follow the trend with increasing $N_H$ as a function of redshift as found for
the lGRB afterglows \citep{Campana2010, Starling2013, Arcodia2016}. This is
related to what is found by \citet{Kopac2012, Margutti2013}, that $N_H$ seems to
be comparable for long and short GRBs when compared at similar redshifts.

The redshift evolution of $N_H$ in the hosts of lGRBs is not reproduced by
\citet{Buchner2017}, using a different $N_H$ inference methodology. Instead a
correlation between $N_H$ and host stellar mass is suggested. Assuming that the
different $N_H$-fitting methodologies yield comparable results, GRB~111117A is
an outlier from the relation suggested by \citet{Buchner2017} by more than the
intrinsic scatter, although some lGRB hosts populate a similar region in the
$N_H$-$M_\star$ relation. Additionally, for lGRBs, $N_H$ correlates with the
surface luminosity at the explosion site \citep{Lyman2017}. The large offset of
GRB~111117A relative to the host center derived in \citet{Margutti2012} and
\citet{Sakamoto2013} is difficult to reconcile with galaxy-scale gas providing
the X-ray absorption. Along with the absence of dust, the large offset from the
host center indicates that the high $N_H$ arises because the density in the GRB
surrounding is high, or because the light from the afterglow traverses a region
of dense gas. The optical darkness of the GRB is an additional hint of the high
density in the GRB surroundings or along the line of sight.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/NH_z.pdf}
	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations of GRBs. The detections are replaced with contours and the limits are shown with arrows. Marginalizations over both axes are shown where the limits are shown as transparent bars and detections as solid ones.}
	\label{fig:NH_z}
\end{figure}

% \begin{figure}
% 	\centering
% 	\includegraphics[width=9cm]{figures/Eiso_z.pdf}
% 	\caption{Isotropic equivalent energy for all the representative samples of both long and short GRBs. The sample is from \citet{Turpin2015} and the short is from \citet{DAvanzo2014a}.}
% 	\label{fig:Eiso_z}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Host galaxy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From the clear host association, GRB~111117A does not belong to the hostless
class of GRBs \citep{Berger2010a} and because the host exhibits emission lines
this is indicative of a population of relatively young stars. As the majority of
sGRBs \citep{Fong2013b}, the host of GRB~111117A is therefore a late-type galaxy
and is entirely consistent in terms of stellar mass and stellar age with the
general population of sGRB hosts ($\left\langle M _* \right\rangle = 10^{10.1}
M_{\odot}$ and $\left\langle \tau _* \right\rangle = 0.3 $~Gyr)
\citep{Leibler2010}. Being a late-type host, both the stellar mass and sSFR are
entirely within the range expected for the hosts of sGRBs \citep{Behroozi2014}.
Our constraints on the dynamical mass is also well accommodated by the expected sGRB
host halo mass \citep{Behroozi2014}.

The SFR is $\sim$1 order of magnitude higher than the typical SFR for sGRB host
galaxies \citep{Berger2014} and more similar to the SFR found in the hosts of
lGRBs at a corresponding redshift \citep{Kruhler2015}. Only two hosts in the
sample of short GRBs compiled in \citet{Berger2014} have a more vigorous star
formation, meaning that it is in the very upper end of the star formation
distribution. The cosmic SFR evolution of the universe likely plays a role due
to the proximity of GRB~111117A to the peak of cosmic SFR \citep{Madau2014}.
The high SFR is partly a selection effect, because a less star forming galaxy
would exhibit weaker emission lines, thus making the redshift harder to
determine. Additionally, it is natural to expect some evolution in the hosts of
sGRBs with redshift as illustrated for $N_H$ in Sect.~\ref{restnH}.

The detection of \lya~is consistent with the SED-inferred absence of dust,
despite the moderate stellar age which would suggest the opposite. The centroid
of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
with respect to systemic, which is similar to what is found for long GRB hosts
\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Comparison to long GRBs at z $\sim$ 2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implications for the redshift distribution of sGRBs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A single sGRB at high redshift does little in terms of constraining the redshift
distribution of sGRBs. In particular, other sGRB hosts could be missed because
they are intrinsically fainter and thus the high redshift of GRB~111117A is only
measured due to the brightness of its host. \citet{Berger2014} compiled a
sample of sGRB host luminosities, normalized by the characteristic galaxy
luminosity at their respective redshift, $L_B/L^{\star}_{B}$. To convert the
SED-inferred $M_B$ of GRB~111117A to $L_B/L^{\star}_{B}$, we use the
characteristic absolute $B$-band magnitude of the Schechter function for blue
galaxies ($U - V < 0.25$) in the redshift window $2.0 \leq z \leq 2.5$ from
\citet{Marchesini2007} and obtain $L_B/L^{\star}_{B} = 1.2$.

Using the complete, flux-limited selection of bursts from \citet{DAvanzo2014a},
excluding GRB~111117A and the likely non-sGRB GRB~090426, we have a
statistically homogeneous sample from which we can address the implications of
the redshift of GRB~111117A. Out of the 14 hosts in the sample, 10 (71 per cent)
have measured redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the
host of GRB~111117A  is brighter than 80 per cent of the hosts with measured
$L_B/L^{\star}_{B}$. Even if we conservatively assume that \textit{all} the
hosts missing $L_B/L^{\star}_{B}$ are brighter than the host of GRB~111117A, the
host is still brighter than > 60 per cent of sGRB hosts. For all 26 hosts with
$L_B/L^{\star}_{B}$ from \citet{Berger2014}, the host of GRB~111117A is brighter
than 73 per cent.

If we assume that we are able to obtain emission-line redshifts from hosts, 0.5
mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed
60 per cent of the redshifts (6 out of 10 hosts), due too the host being too
faint, were they at the redshift of GRB~11117A. The corresponding number is
around 45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
reflecting the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because
the average SFR of galaxies hosting lGRBs is higher than for galaxies hosting
sGRBs, the fraction of missed burst redshifts is likely higher although the
cosmic SFR evolution could play a role in improving redshift determinability.

A fraction of the bursts missing redshift are host-less but appear to be
spatially correlated with galaxies that are likely at moderate redshifts
\citep{Tunnicliffe2014}, but should some of the remainder be at high redshift,
the missed fraction will increase. If we assume that \textit{all} the bursts
that are missing redshifts are at high-$z$ and missed due to host faintness, 10
out of 14 hosts in the complete sample (71 per cent) would be missed at $z =
2.211$. This serves as an upper limit on the fraction of missed burst redshifts
at high-$z$. Conversely, if all bursts missing redshift are at low redshift and
missed for other reasons, 6 out of 14 hosts (43 per cent) would be missed at $z
= 2.211$. The two limits indicate that we would miss between 43 and 71 per cent
of sGRB hosts at $z \sim 2$ due to host faintness.

The theoretical redshift distribution of sGRBs depends on the type of delay-time
function used to model the progenitor system. The likelihood preferred lognormal
time delay models investigated by \citet{Wanderman2015} predict a sGRB rate at
$z = 2.211$, $\sim$ two orders of magnitude lower compared to the peak rate at
$z = 0.9$. According to \citet{Wanderman2015}, this preference depends
critically on the absence of non-collapsar sGRBs at $z \gtrsim 1.2$. The higher
determined redshift of GRB~111117A, and the likely number of additional high-z
sGRB could change the preferred time delay models. The redshift of GRB~111117A,
on the other hand, is close to the expected peak in sGRB rate calculated using
the power law delay time models \citep{Behroozi2014, Wanderman2015,
	Ghirlanda2016}, meaning we could be missing a large fraction of sGRBs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Constraints on progenitor separation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

At $z = 2.211$, the age of the universe is 3 Gyr. If the progenitor systems of
sGRBs are the merger of two NSs, this sets a hard upper limit to the coalescence
timescale for such a system. In the absence of other mechanisms, the timescale
of the orbital decay of the system is set by the energy loss due to
gravitational waves, which in turn is set by the mass of the constituent compact
objects, the eccentricity of the orbit and the separation of the two \citep{Postnov2014}. If we assume that the
formation timescale of the first galaxies is short compared to the time since
the Big Bang \citep{Richard2011}, and if we assume a mass of 1.4~$M_\odot$ for
each of the NSs in a circular orbit at the time of system formation, this places
a hard upper limit on the initial separation, of $a_0 < 3.2~R_\odot$.

In practice most NS-NS binaries will be eccentric at formation because of the SN
natal kicks. For more eccentric orbits, the coalescence timescale decreases,
leading to a decrease in the constraint on the initial separation and larger
initial separation constraints. As noted by \citet{Postnov2014}, it takes
eccentricities > 0.6 to significantly shorten the merger time.

Using the inferred stellar population age from our SED fit, then we obtain a
(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
does not account for the possibility there could be an underlying stellar
population of older stars from a previous star-formation epoch. To investigate
the possible impact of the presence of an old stellar population, we followed
\citet{Papovich2001} and re-fitted the observed SED with the best-fit template
to which an additional stellar population of old stars was added. For each
galaxy, this old population was set as the SPSs with same parameters that the
best-fit SED excepts the age, which was set to the age of the Universe at the
observed redshift. In principle, this can constrain the maximum contribution of
old populations within the photometric error bars \citep[see][ for
details]{Papovich2001}. We find a negligible contribution to the stellar mass
(i.e. variations much smaller than the statistical uncertainty associated with
the best-fit template). The delay time between formation and explosion is well
accommodated by the models of \citet{Belczynski2006}, although the longest delay
times are excluded. This is especially true given the late type nature of the
host \citep{OShaughnessy2008}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We have here provided a revised, spectroscopic redshift measurement for the
short GRB~111117A based on host galaxy emission lines setting it at $z = 2.211$.
This value supersedes the previous photometric redshift of \citet{Margutti2012}
and \citet{Sakamoto2013}. The erroneous redshift estimate of previous authors is
attributed to a discrepancy in the measured $z$-band magnitude, and highlights
the importance of deep spectroscopic studies of sGRB hosts at medium resolution.

Using the new distance, the X-ray derived $N_H$ towards GRB~111117A is the
highest within a complete sample of sGRB hosts and is consistent with the
$N_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
upper end of the sGRB host SFR distribution and no significant amount of dust is
present. The high $N_H$ is difficult to reconcile with the large projected host
offset and the absence of dust. One possible explanation could be, that
GRB~111117A is formed through a prompt channel of sGRB formation and originates
in a star forming region located in the outskirts of the host.

Although a single burst carries little leverage in terms of constraining the
redshift distribution of sGRB, the high redshift of GRB~111117A needs to be
accommodated in progenitor models. A lognormal delay time model predicts a very
low volumetric density of bursts at $z = 2.211$, whereas a power law delay time
model peaks near GRB~111117A. If more sGRBs are at similarly high redshifts, but
are missed due to the faintness of their hosts, a lognormal delay time model
will be disfavored. Compared to a sample of sGRB hosts, GRB~111117A is more
luminous than 80 per cent of a complete sample of sGRB hosts with measured
luminosities. Assuming a host brightness redshift determination threshold, for
between 43 and 71 per cent of sGRB hosts, we would be unable to determine a
redshift should they be at a similar redshift as GRB~111117A. This could
indicate that potentially, a significant fraction of sGRBs are at high $z$ but
missed due to host faintness.

Using the age of the universe at the time of explosion allows us to set
constraints on the maximal separation between the engine constituents at the
time of formation. We find that the maximal separation of two NSs at system
formation time is $a_0 < 3.2~R_\odot$, which excludes some of the formation
channels with the longest timescales.

All data, code and calculations related to the paper along with the
paper itself are available at \url{https://github.com/jselsing/GRB111117A}.

\begin{acknowledgements}
We thank Jens Hjorth and Lise Christensen for useful discussions regarding the interpretation of this event. We thank Mathieu Puech for testing the possible contribution from an older stellar population in the SED.
%	
TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
%
SDV is supported by the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro 
%
JJ acknowledges support from NOVA and NWO-FAPESP grant for advanced
instrumentation in astronomy.
%
NRT and KW acknowledge support from STFC Consolidated
Grant ST/N000757/1
%
CT acknowledges support from a Spanish National Research Grant of Excellence
under project AYA 2014-58381-P and funding associated to a Ramón y Cajál
fellowship under grant number RyC-2012-09984.
%
AdUP acknowledges support from a Ramón y Cajal fellowship, a BBVA Foundation
Grant for Researchers and Cultural Creators,and the Spanish Ministry of Economy
and Competitiveness through project AYA2014-58381-P. Partly based on
observations made with the Gran Telescopio Canarias (GTC).
%
ZC acknowledges support from the Spanish research project AYA 2014-58381-P and
support from Juan de la Cierva Incorporaci\'on fellowships IJCI-2014-21669.
%
RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural
Creators and support from ASI (Italian Space Agency) through the Contract n. 2015-046-R.0 and from European Union Horizon 2020 Programme under the AHEAD project (grant agreement n. 654215).
%
This research made use of Astropy, a community-developed core Python package for Astronomy \citep{TheAstropyCollaboration2013}. The analysis and plotting has been achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy and Scipy \citep{VanderWalt2011} along with other community-developed packages.
%
This work made use of observations obtained with the Italian 3.6 m Telescopio Nazionale Galileo (TNG) operated on the island of La Palma by the Fundaci\'on Galileo Galilei of the INAF (Istituto Nazionale di Astrofisica) at the Spanish Observatorio del Roque de los Muchachos of the Instituto de Astrof\'isica de Canarias.
%
Based on data from the GTC Archive at CAB (INTA-CSIC) and on observations obtained at the Gemini Observatory, which is operated by the Association of Universities for Research in Astronomy, Inc., under a cooperative agreement with the NSF on behalf of the Gemini partnership: the National Science Foundation (United States), the National Research Council (Canada), CONICYT (Chile), Ministerio de Ciencia, Tecnología e Innovación Productiva (Argentina), and Ministério da Ciência, Tecnologia e Inovação (Brazil).

\end{acknowledgements}

\bibliographystyle{aa_arxiv}
\bibliography{GRB111117A}

\end{document}

### 2018-08-28 Arxiv/tables/spec_overview.tex
Published

\begin{table*}[!ht]

	\centering
	\caption{Overview of the spectroscopic observations. JH in the slit width refers to observations where a K-band blocking filter has been used. The seeing is determined from the width of the spectral trace of a telluric standard star, taken close in time to the host observation. The spectral resolution, R, is measured from unresolved telluric absorption lines in the spectrum of the telluric standard star. \label{tab:spec_overview}}
	\begin{tabular}{cccccccc}
		\hline\hline
		{Obs. Date} &  \multicolumn{3}{c}{Exposure time (s)} & Slit width & Airmass & Seeing & R   \\ [1.5pt]
        \hline
		{} & UVB  & VIS & NIR &  (arcsec)   & {} & (arcsec)  & {VIS/NIR}  \\ [1.5pt]
		\hline
		2011-11-19T01:33 & 2 $\times$ 2400 & 2 $\times$ 2400 & 8 $\times$ 600 & 1.0/1.0/0.9 & 1.49 & 0.75 & 11600/6700 \\
        2013-07-15T09:02 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.53 & 0.98 & 9600/8900 \\
        2013-08-03T07:37 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.55 & 0.85 & 11400/11300 \\
        2013-08-03T08:34 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.49 & 0.85 & 11400/11300 \\

		\hline\noalign{\smallskip}

	\end{tabular}

\end{table*}

### 2018-08-28 GRB111117A_corr.tex
Published

--- 2017-11-26 GRB111117A.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
+\documentclass[longauth]{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
 %\documentclass[referee]{aa}
@@
 % Manually specified definitions
-\input{definitions.tex}
+\newcommand{\farc}{\hbox{$.\!\!^{\prime\prime}$}} 
+\newcommand{\ergA}{$\rm{erg\,cm^{-2}\,s^{-1}\,\AA^{-1}}$} 
+\newcommand{\erg}{$\rm{erg\,cm^{-2}\,s^{-1}}$} 
+\newcommand{\kms}{$\rm{km\,s^{-1}}\,$}
+
+\newcommand{\griz}{$g' r' i' z'$}
+\newcommand{\JHK}{$JHK_{\rm{s}}$}
+\newcommand{\gK}{$g' r' i' z' JHK_{\rm{s}}$}
+\newcommand{\Msun}{$M_\odot$}
+
+% Elements
+\newcommand{\lya}{Ly$\alpha$} 
+\newcommand{\lyb}{Ly$\beta$} 
+\newcommand{\lyg}{Ly$\gamma$} 
+\newcommand{\hb}{H$\beta$} 
+\newcommand{\ha}{H$\alpha$} 
+\newcommand{\hg}{H$\gamma$} 
+\newcommand{\hi}{\ion{H}{i}} 
+\newcommand{\hei}{[\ion{He}{i}]} 
+\newcommand{\oi}{[\ion{O}{i}]} 
+\newcommand{\sii}{[\ion{S}{ii}]} 
+\newcommand{\siii}{[\ion{S}{iii}]} 
+\newcommand{\oii}{[\ion{O}{ii}]} 
+\newcommand{\oiii}{[\ion{O}{iii}]}
+\newcommand{\nii}{[\ion{N}{ii}]} 
+\newcommand{\nv}{[\ion{N}{v}]} 
+\newcommand{\neiii}{[\ion{Ne}{iii}]} 
+\newcommand{\NIii}{[\ion{Ni}{ii}]} 
+\newcommand{\feii}{\ion{Fe}{ii}} 
+\newcommand{\civ}{\ion{C}{iv}} 
+\newcommand{\cii}{\ion{C}{ii}}
+\newcommand{\mgi}{\ion{Mg}{i}} 
+\newcommand{\mgii}{\ion{Mg}{ii}} 
+\newcommand{\ali}{\ion{Al}{i}} 
+\newcommand{\alii}{\ion{Al}{ii}} 
+\newcommand{\aliii}{\ion{Al}{iii}} 
+\newcommand{\SIi}{\ion{Si}{i}} 
+\newcommand{\SIii}{\ion{Si}{ii}} 
+\newcommand{\SIiii}{\ion{Si}{iii}} 
+\newcommand{\SIiv}{\ion{Si}{iv}} 
+\newcommand{\znii}{\ion{Zn}{ii}} 
+\newcommand{\crii}{\ion{Cr}{ii}} 
+\newcommand{\mnii}{\ion{Mn}{ii}} 
+\newcommand{\tiii}{\ion{Ti}{ii}} 
+\newcommand{\caii}{\ion{Ca}{ii}}
 
@@
 
-\input{authorlist.tex}
+\author{J.~Selsing\inst{1}
+        \and T.~Kr\"{u}hler\inst{2}
+        \and D.~Malesani\inst{1}
+        \and P.~D'Avanzo\inst{3}
+        \and S.~Schulze \inst{4, 5, 6}
+        \and S.~D.~Vergani\inst{3, 7}
+        \and J.~Palmerio\inst{8}
+        \and J.~Japelj\inst{9}
+        \and B.~Milvang-Jensen\inst{1}
+        \and D.~Watson\inst{1}
+        \and P.~Jakobsson\inst{10}
+        %
+        \and J.~Bolmer \inst{2}
+        \and Z.~Cano\inst{11}
+        \and S.~Covino \inst{3}
+        %\and L.~B.~Christensen\inst{1}
+        \and V.~D'Elia\inst{12, 13}
+        \and A.~de~Ugarte~Postigo\inst{11, 1}
+        \and J.~P.~U.~Fynbo\inst{1}
+        \and A.~Gomboc\inst{14}
+        \and K.~E.~Heintz\inst{10,1}
+        \and L.~Kaper \inst{9}
+        \and A.~J.~Levan \inst{15}
+        \and S.~Piranomonte \inst{12}
+        \and G.~Pugliese \inst{9}
+        \and R.~S\'{a}nchez-Ram\'{\i}rez \inst{11, 16}
+        \and M.~Sparre\inst{1,17}
+        \and N.~R.~Tanvir\inst{18}
+        \and C.~C.~Th\"{o}ne\inst{11}
+        \and K.~Wiersema \inst{18}
+}
+
+\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
+        %2
+        \and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
+        %3
+        \and INAF - Osservatorio Astronomico di Brera, via E. Bianchi 46, I-23807, Merate (LC), Italy
+        %4
+        \and Instituto de Astrof\'isica, Facultad de F\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+        %5
+        \and Millennium Institute of Astrophysics, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+        %6
+        \and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
+        %7
+        \and GEPI, Observatoire de Paris, PSL Research University, CNRS, Place Jules Janssen, 92190 Meudon, France
+        %8
+        \and Sorbonne Universités, UPMC Univ. Paris 6 et CNRS, UMR 7095, Institut d’Astrophysique de Paris, 98 bis bd Arago, 75014 Paris, France
+        %9
+        \and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands
+        %10
+        \and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
+        %11
+        \and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain
+        %12
+        \and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
+        %13
+        \and ASI-Science Data Centre, Via del Politecnico snc, I-00133 Rome, Italy
+        %14
+        \and Centre for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
+        %15
+        \and Department of Physics, University of Warwick, Coventry CV4 7AL, UK
+        %16
+        \and Istituto de Astrofisica e Planetologia Spaziali, INAF, Via Fosso del Cavaliere 100, I-00133 Roma, Italy
+        %17
+        \and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
+        %18
+        \and Department of Physics \& Astronomy and Leicester Institute of Space \& Earth Observation,
+        University of Leicester, University Road, Leicester, LE1 7RH, UK
+}
 
@@
 their hosts to measure their redshifts. These measurements, however, are
-critical to constrain the nature of sGRB progenitors, their redshift
-distribution and the $r$-process element enrichment history of the universe.
-Here, we present spectroscopy of the host galaxy of GRB~111117A and measure its
+critical for  constraining the nature of sGRB progenitors, their redshift
+distribution, and the $r$-process element enrichment history of the universe.
+Here we present spectroscopy of the host galaxy of GRB~111117A and measure its
 redshift to be $z = 2.211$. This makes GRB~111117A the most distant
@@
 complete sample of sGRBs and seems to follow the evolution with redshift as
-traced by the hosts of long GRBs.
+traced by the hosts of long GRBs. From the detection of \lya~emission in the
+spectrum, we are able to constrain the escape fraction of \lya~in the host.
 
@@
 distribution at $z = 2.211$, and is actively forming stars. Using the observed
-sGRB host luminosity distribution, we find that between 43 and 71 per cent of
-all \textit{Swift}-detected sGRBs have too faint hosts at $z\sim2$ to allow for
+sGRB host luminosity distribution, we find that between 43\%\ and 71\%\ of
+all \textit{Swift}-detected sGRBs have hosts that are  too faint  at $z\sim2$ to allow for
 a secure redshift determination. This implies that the measured sGRB redshift
-distribution could be incomplete at high redshift. The large $z$ of
+distribution could be incomplete at high redshift. The high $z$ of
 GRB~111117A is evidence against a lognormal delay-time model for sGRBs through
@@
 There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs) come
-from the merger of a neutron star (NS), either with another NS, or a black hole,
+from the merger of a neutron star (NS)  either with another NS or with a black hole,
 due to their apparent association with kilonovae \citep{Barnes2013a,
-	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. This association was
-recently confirmed by the simultaneous and co-spatial detection of gravitational
+        Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. This association has
+recently been confirmed by the simultaneous and co-spatial detection of gravitational
 waves (GWs) from a binary neutron star merger and a sGRB
-\citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}. The
-absence of associated supernovae in deep searches
-\citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011} supports this idea and
-distinguishes the physical origin of sGRBs from their long-duration
-counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.
+\citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}; however,
+to what degree the sGRB associated with GW170817 is a cosmological sGRB remains
+an open question \citep{Lyman2018, Margutti2018}. The absence of associated
+supernovae in deep searches \citep[e.g.,][]{Hjorth2005a,Fox2005,Hjorth2005b,
+        Kann2011} supports this idea and distinguishes the physical origin of sGRBs from
+their long-duration counterparts \citep[but see also][]{Fynbo2006b,
+        Valle2006, Gal-Yam2006}.
 
@@
 redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}. Unlike lGRBs,
-sGRB afterglows are faint making absorption spectroscopy often ineffective.
+sGRB afterglows are faint, making absorption spectroscopy often ineffective.
 Therefore, most of these measurements come from the associated hosts and is
@@
 The host galaxies of sGRBs are diverse. They are more massive and less actively
-star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases,
+ star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases
 no host galaxy can be identified above the detection threshold of deep follow-up
@@
 within their hosts \citep{Fong2013a}, this suggests a progenitor system that can
-be very long lived in comparison to lGRBs, with the host stellar mass affecting
-the sGRB rates more than the star-formation rate (SFR) \citep{Berger2014}.
+be very long-lived in comparison to lGRBs, with the host stellar mass affecting
+the sGRB rates more than the star formation rate (SFR) \citep{Berger2014}.
 
@@
 localize NS mergers \citep{Ghirlanda2016}. This epochal breakthrough occurred
-recently when the first NS-NS GW event ever was detected by LIGO/Virgo (GW
-170817) and associated to the weak sGRB 170817A detected by the \textit{Fermi}
+recently when the first ever NS--NS GW event  was detected by LIGO/Virgo (GW
+170817) and associated with the weak sGRB 170817A detected by the \textit{Fermi}
 and INTEGRAL satellites \citep{LIGOScientificCollaboration2017a, Goldstein2017,
@@
 promising ways to constrain the binary inclination angle \citep{Arun2014,
-	LIGOScientificCollaboration2017a} and measure cosmological distances
+        LIGOScientificCollaboration2017a} and to measure cosmological distances
 \citep{Nissanke2010, LIGOScientificCollaboration2017c}.
@@
 The total lifetime of NS binaries depends on their orbit, mass, spin, initial
-separations and subsequent inspiral times. The delay time from formation to
+separations, and subsequent inspiral times. The delay time from formation to
 explosion impacts the timing and distribution of the enrichment of the ISM with
 heavy $r$-process elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some
-limits can be calculated using host galaxy star-formation history models and
+limits can be calculated using host galaxy star formation history models and
 spatial distribution of sGRBs within their hosts \citep[][]{Berger2014}. The
 most distant cosmological bursts, however, offer direct, hard limits on the
-coalescence time scales.
-
-We here present a spectrum of the host galaxy of the short GRB~111117A
+coalescence timescales.
+
+Here we  present a spectrum of the host galaxy of the short GRB~111117A
 ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is
-significantly higher than the previously estimated redshift based on photometric
+significantly higher than the previously estimated redshift of $z \sim 1.3$ based on photometric
 studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest frame
@@
 
+All data, code, and calculations related to the paper along with the
+paper itself are available \href{https://github.com/jselsing/GRB111117A}{online}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
         FORS2 $R$-band imaging of the field of GRB~111117A with the X-shooter slit
-overlaid. The slit position represents 4 epochs of spectroscopic observations
+overlaid. The slit position represents four epochs of spectroscopic observations
 taken at similar position angles. The corresponding photometry is shown in
-Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as derived
+Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position (R.A., Dec.) (J2000) = (00$^{\rm h}$ 50$^{\rm m}$ 46.264$^{\rm s}$, +23$^{\circ}$ 00$^{\prime}$ 39.98$^{\prime\prime}$) as derived
 from the \emph{Chandra} observations in \citet{Sakamoto2013}.
@@
 %       \centering
-%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
+%%      \includegraphics[width=9cm]{figures/OIII_img.pdf}
 %       \caption{2D-image of the \oiii$\lambda$5007 emission line.
@@
         \includegraphics[width=18cm]{figures/grid.pdf}
-	\caption{2D-images of the emission lines corresponding to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha.
+        \caption{Two-dimensional  images of the emission lines corresponding to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha.
 The location of bright skylines are marked by blue boxes.
 The locations of the emission lines are indicated with red ellipses.
-Because the host is observed in nodding-mode, negative images of the emission lines appear on both sides in the spatial direction.
-For the upper, left panel containing \lya, the systemic redshift position of \lya{} is marked by a yellow, vertical, dashed line.
+Because the host is observed in nodding mode, negative images of the emission lines appear on both sides in the spatial direction.
+For the upper left panel containing \lya, the systemic redshift position of \lya{} is marked by a yellow vertical dashed line.
 The red part of the \oii$\lambda$3727-doublet is affected by atmospheric absorption.}
@@
 \end{figure*}
+
+Spectroscopic observations were carried out using the cross-dispersed echelle
+spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
+burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
+ESO program 088.A-0051 (PI: Fynbo) and again later under ESO program
+091.D-0904 (PI: Hjorth). The observations used a simple ABBA nodding pattern, with 5
+\arcsec~nod throws. The X-shooter covers the wavelength range from 3000~\AA{} to
+24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across
+three spectroscopic arms. We carried out the bias-correction, flat-fielding,
+order tracing, wavelength calibration, rectification, and flux calibration using
+the VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run
+in physical mode. Because the echelle orders are curved across each detector, a
+rectification algorithm is employed that introduces correlations between
+neighboring pixels. We selected a pixel scale of 0.2/0.2/0.6 \AA/pix for the
+UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
+maximum resolution. The observations were combined and extracted using the scripts
+described in \citet{2018arXiv180207727S} and available
+online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
+where the full spectral point spread function was modeled across each arm and
+used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
+spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
+slit position is shown in Fig.~\ref{fig:spec_setup}. Each of the epochs was
+extracted individually and combined in a weighted fashion where the weight at
+each pixel was chosen as median variance spectrum of the region surrounding that
+pixel, thus avoiding basing the weight on the pixel variance. Slit-loss
+correction was applied on the combined spectrum based on the average seeing
+of the observations. We show the extracted spectrum in Fig.~\ref{fig:SED}.
 
@@
 The detection of \lya{}  is predicted from the SED fit and confirmed by the spectroscopic observations.
-Overplotted in grey is the observed spectrum, binned by 6 \AA{} for presentation purposes.
-Slit losses has been corrected for based on the average seeing of the observations, as confirmed by the comparison with the photometry.
-The blue, dashed line is the corresponding error spectrum, smoothed for presentation purposes.
-The reason for the spectral gaps at 5500 \AA{} and 10000 \AA{} is from the merging of the arms.}
+Overplotted in gray is the observed spectrum, binned by 6 \AA{} for presentation purposes.
+Slit losses have been corrected for based on the average seeing of the observations, as confirmed by the comparison with the photometry.
+The blue dashed curve is the corresponding error spectrum, smoothed for presentation purposes.
+The spectral gaps at 5500 \AA{} and 10000 \AA{} are due to the merging of the arms.}
 	\label{fig:SED}
@@
 
-Spectroscopic observations were carried out using the cross-dispersed echelle
-spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
-burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
-ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
-091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, with 5
-\arcsec~nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
-24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across
-three spectroscopic arms. We carried out the bias-correction, flat-fielding,
-order tracing, wavelength calibration, rectification, and flux calibration using
-the VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run
-in physical mode. Because the echelle orders are curved across each detector, a
-rectification algorithm is employed which introduces correlations between
-neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
-UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
-maximal resolution. The observations are combined and extracted using scripts
-described in Selsing et al. 2017 (in prep.) and available
-online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
-where the full spectral point spread function is modeled across each arm and
-used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
-spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
-slit position is shown in Fig.~\ref{fig:spec_setup}. Each of the epochs are
-extracted individually and combined in a weighted fashion where the weight at
-each pixel is chosen as median variance spectrum of the region surrounding that
-pixel, thus avoiding to base the weight on the pixel variance. Slit-loss
-correction has been applied on the combined spectrum based on the average seeing
-of the observations. We show the extracted spectrum in Fig.~\ref{fig:SED}.
-
-\input{tables/spec_overview.tex}
-
 We  determine a redshift of $z = 2.211 \pm 0.001$ from the simultaneous detection
 of emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959,
-\oiii$\lambda$5007, and \ha. \oii$\lambda$3727, \hb{}, and \oiii$\lambda$4959
-are detected at low significance ($\sim$ 3-$\sigma$). The uncertainty on the
+\oiii$\lambda$5007, and \ha. The \oii$\lambda$3727, \hb{}, and \oiii$\lambda$4959
+lines are detected at low significance ($\sim$ 3-$\sigma$). The uncertainty on the
 redshift is the standard deviation of independent measurements of the redshift
 based on the individual line centroids (excluding \lya). We show cutouts of the
-2D spectrum at the position of all the detected lines in Fig.~\ref{fig:line}. 
-\ha{} is only visible in the first epoch due to the $K$-band blocking filter
+2D spectrum at the position of all the detected lines in Fig.~\ref{fig:line}. The 
+\ha{} line is only visible in the first epoch due to the $K$-band blocking filter
 used for the remaining observations. The nebular lines exhibit a spatial extent
@@
 around 21\,000~\AA{} where the night sky is very bright. In addition, several
-bright sky-lines are superposed on the line, making an accurate estimate of the
-\ha~flux difficult. Due to their velocity structure, the lines exhibit clear
+bright sky-lines are superposed on the line, making it difficult to obtain  an accurate estimate of the
+\ha~flux. Due to their velocity structure, the lines exhibit clear
 deviations from a Gaussian and given the low S/N of the spectra we do not
@@
 
-From the SED-fit (Sect.~\ref{SED}) the host is constrained to contain very
+From the SED fit (Sect.~\ref{SED}) the host is constrained to contain very
 little or no dust. Consistently, \lya{} is detected although its presence does
-not exclude dust. Therefore we do not apply a dust-correction to the measured
-\ha{} flux here. \oii{} is close to a region of strong telluric absorption,
-which is why no SFR is inferred from this line.
+not exclude dust. Therefore, we do not apply a dust-correction to the measured
+\ha{} flux here. The  \oii{} line is close to a region of strong telluric absorption,
+which is why no SFR is inferred from it.
 
 The total extent of the lines in velocity space is $\sim$ 450 km s$^{-1}$. The line
-profiles shows an asymmetric "double-horned" profile, indicating that we are
+profiles shows an asymmetric ``double-horned'' profile, indicating that we are
 seeing a galaxy with a large degree of coherent rotational motion relative to
-the line-of-sight. If we assume that we are viewing a spiral galaxy edge-on,
+the line of sight. If we assume that we are viewing a spiral galaxy edge-on,
 this is a measure of the rotational velocity of the gas. If we assume that the
 spectral resolution and the turbulent width of the lines are negligible compared
-to the rotational velocity, we can, based on the projected size of the source
-and the width of the lines, put a constraint on the dynamical mass of the galaxy
+to the rotational velocity, based on the projected size of the source
+and the width of the lines we can put a constraint on the dynamical mass of the galaxy
 \citep{DeBlok2014}. Based on the physical size along the slit and the velocity
@@
 In addition to the spectroscopy presented above, we imaged the field of
-GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
+GRB~111117A in multiple broadband filters using the VLT equipped with FORS2
 ($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters). These new data are
@@
 observations and measured brightnesses is given in
-Table~\ref{tab:phot_overview}. Most data were taken long after the GRB had faded
+Table~\ref{tab:phot_overview}. Most of the data were taken long after the GRB had faded
 when no afterglow contribution was present. Given the faintness of the afterglow
@@
 
-All data were reduced, analyzed and fitted in a similar manner as described in
-detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. We
-use our own \texttt{Python} and IRAF routines to perform a standard reduction
-which includes bias/flat-field correction, de-fringing (where necessary),
+All data were reduced, analyzed, and fitted in a similar manner, as described in
+detail in \citet{Kruhler2011a} and more recently in \citet{Schulze2016}. We
+use our own \texttt{Python} and IRAF routines to perform a standard reduction that  includes bias/flat-field correction, de-fringing (where necessary),
 sky-subtraction, and stacking of individual images. On the final reduced image
@@
 photometry, where the size of the aperture was chosen to be 2\farc0. Because
-image quality ranges from 0\farc6 - 1\farc7 and the galaxy major axis is $\sim$
+image quality ranges from 0\farc6 -- 1\farc7 and the galaxy major axis is $\sim$
 1\farc, this ensures that, in all cases, the large majority of the flux is
-collected and low-surface brightness light missed in the aperture will not
+collected and low surface brightness light missed in the aperture will not
 influence the measurement. This method sacrifices some S/N in the best seeing
@@
 
-Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the
+Noteworthy are the discrepancy of  our new VLT/FORS2 photometry and of the
 re-analysis of the Gemini data compared to the $z$-band measurements of
@@
 
+\begin{table*}[!h]
+
+	\centering
+	\caption{Overview of the spectroscopic observations.
+``JH'' in the slit width column refers to observations where a $K$-band blocking filter has been used.
+The seeing is determined from the width of the spectral trace of a telluric standard star, observed close in time to the host observation.
+The spectral resolution, $R$, is measured from unresolved telluric absorption lines in the spectrum of the telluric standard star.
+\label{tab:spec_overview}}
+	\begin{tabular}{cccccccc}
+		\hline\hline
+		{Observation epoch} &  \multicolumn{3}{c}{Exposure time (s)} & Slit width & Airmass & Seeing & $R$   \\ [1.5pt]
+		\hline
+		(UT) & UVB  & VIS & NIR &  (arcsec)   & {} & (arcsec)  & {VIS/NIR}  \\ [1.5pt]
+		\hline
+		2011-11-19 01:33 & 2 $\times$ 2400 & 2 $\times$ 2400 & 8 $\times$ 600 & 1.0/1.0/0.9 & 1.49 & 0.75 & 11600/6700 \\
+		2013-07-15 09:02 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.53 & 0.98 & 9600/8900 \\
+		2013-08-03 07:37 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.55 & 0.85 & 11400/11300 \\
+		2013-08-03 08:34 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.49 & 0.85 & 11400/11300 \\
+
+		\hline\noalign{\smallskip}
+
+	\end{tabular}
+
+\end{table*}
+
 \begin{figure}
@@
         a red circle with a 2\farc0 radius. This is the same size as the aperture used
-	for deriving the photometry. Each panel is 20\arcsec~in size, north is up, east
+        to derive the photometry. Each panel is 20\arcsec~in size, north is up, east
         is left. Worth noting is the relative depth of the GMOS-N and FORS2 $z$-band
-	images. For reference, the object located to the south-east of the host and
-	marked with a blue, dashed circle has an extinction corrected magnitude of
+        images. For reference, the object located to the southeast of the host and
+        marked with a blue dashed circle has an extinction corrected magnitude of
         $23.11 \pm 0.09$ ($23.10 \pm 0.18$) in the FORS2 (GMOS-N) $z$-band.}
@@
 
-\input{tables/phot_overview.tex}
-
-The multi-color SED is fit using the \citet{Bruzual2003} single stellar
-population models (SSPs) based on a \citet{Chabrier2003} with initial mass
+\begin{table*}[!ht]
+
+        \centering
+        \caption{Overview of the photometric observations. \label{tab:phot_overview}}
+        \begin{tabular}{ccccccc}
+                \hline\hline
+                {Observation epoch} &  Exptime & Telescope/instrument & Filter & Airmass & Image quality & Host brightness\tablefootmark{a}  \\ [1.5pt]
+                \hline
+                ({UT}) & ({ks}) &    & {} & & (arcsec)  & (AB mag)  \\ [1.5pt]
+                \hline
+                2013-08-30 07:43 & 1.45 & VLT/FORS2 & $g$ & 1.55 & 0.99 & $24.08\pm 0.09$ \\
+                2011-11-17 20:07 & 0.80 & GTC/OSIRIS & $g$ & 1.15 & 1.67 & $24.13\pm 0.09$ \\
+                2011-11-17 20:07 & 1.20 & GTC/OSIRIS & $r$ & 1.11 & 1.50 & $23.93\pm 0.08$ \\        
+                2013-07-17 08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
+                2011-11-28 21:10 & 3.60 & TNG/DOLORES & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
+                2011-11-17 20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   
+                2013-08-03 09:23 & 1.35 & VLT/FORS2 & $I$ & 1.54 & 0.93 & $24.22\pm 0.15$ \\           
+                2011-11-28 06:14 & 1.80 & Gemini/GMOS-N & $z$ & 1.01 & 0.84 & $24.24\pm 0.47$ \\  
+                2013-07-13 09:33 & 1.08 & VLT/FORS2 & $z$ & 1.49 & 0.63 & $23.76\pm 0.21$ \\             
+                2013-06-24 09:14 & 1.98 & VLT/HAWK-I & $J$ & 1.70 & 0.63 & $23.13\pm 0.18$ \\        
+                2013-06-27 09:21 & 1.68 & VLT/HAWK-I & $H$ & 1.63 & 0.91 & $22.94\pm 0.29$ \\   
+                2013-06-28 09:14 & 1.92 & VLT/HAWK-I & $K_\mathrm{s}$ & 1.65 & 0.76 & $23.07\pm 0.32$ \\   
+                \hline\noalign{\smallskip}
+
+        \end{tabular}
+
+        \tablefoot{
+                \tablefoottext{a}{All magnitudes are given in the AB system and are not corrected for the expected Galactic foreground extinction corresponding to a reddening of $E_{B-V}=0.027$\,mag.}}
+\end{table*}
+
+The multicolor SED is fit using the \citet{Bruzual2003} single stellar
+population models (SSPs) based on  \citet{Chabrier2003} with initial mass
 function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
@@
 consisting of $\sim 10^6$  different galaxy templates with four metallicities
-0.02, 0.2, 0.4, 1.0 $Z_{\odot}$, different ages, star-formation histories and
+(0.02, 0.2, 0.4, 1.0 $Z_{\odot}$), different ages, star formation histories, and
 degrees of extinction. For every model, we calculate the likelihood, and create
 a probability density function (PDF) for a given parameter by marginalizing over
-the other parameters. We quote the median of the PDF as the best fit parameters
-and the errors are the 16th and 84th percentiles of the PDFs \citep[e.g.
-see][for details on the SED fitting procedure]{Schulze2016}.
-
-The best fit model is an unreddened galaxy template. The inferred physical
+the other parameters. We quote the median of the PDF as the best-fit parameters
+and the errors are the 16th and 84th percentiles of the PDFs \citep[see, e.g.,
+][for details on the SED fitting procedure]{Schulze2016}.
+
+The best-fit model is an unreddened galaxy template. The inferred physical
 parameters are the absolute magnitude ($M_B=-22.0\pm0.1$\,mag), the stellar mass
 ($\log(M_{\star}/M_\odot) = 9.9\pm0.2$), the stellar population age ($\tau =
-0.5_{-0.3}^{+0.5}$ Gyr), and the star-formation rate
+0.5_{-0.3}^{+0.5}$ Gyr), and the star formation rate
 ($\mathrm{SFR}_{\mathrm{SED}} = 11_{-4}^{+9}~M_\odot\,\mathrm{yr}^{-1}$). We
@@
 determination \citep{Kopac2012}. 
-The obtained spectrum is well described by an absorbed power-law
+The obtained spectrum is  described well by an absorbed power-law
 model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
@@
 A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
-"darkness"-parameter $\beta_\mathrm{OX} $ \citep{Jakobsson2004}. Using the optical
+darkness-parameter $\beta_\mathrm{OX} $ \citep{Jakobsson2004}. Using the optical
 afterglow limits ($r' > 25.5$, 13.5 hr after the burst; \citealt{Cucchiara2011,
-	Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time
-of the non-detection. We find $\beta_\mathrm{OX} < 0.79$, consistent with what was
+        Cenko2011}, the X-ray light curve can be interpolated and evaluated at the time
+of the non-detection. We find $\beta_\mathrm{OX} < 0.79$, consistent with the value that  was
 reported in \citet{Sakamoto2013}.
@@
 to a projected physical offset at $z = 2.221$ of $10.6 \pm 1.7$ kpc and $8.5 \pm
-1.7$ kpc respectively. Because the angular distance does not change
+1.7$ kpc, respectively. Because the angular distance does not change
 significantly between $z = 1.3$ and $z = 2.211$, all conclusions of
@@
 
-Based on the BAT light-curve, $T_{90}=0.46$~s which is shorter than both the
+Based on the BAT light curve, $T_{90}=0.46$~s, which is shorter than both the
 prototypical 2~s \citep{Kouveliotou1993} and the $<0.8$~s suggested to also
@@
 As already pointed out by \citet{Margutti2012} and \citet{Sakamoto2013}, this is
-typical of sGRBs, (however see also \citealt{Bernardini2015}). In conjunction
+typical of sGRBs (but see also \citealt{Bernardini2015}). In conjunction
 with the duration and the spectral hardness \citep{Sakamoto2011}, GRB~111117A is
@@
 The intrinsic luminosity is shown in the X-ray light curve (Fig.
-\ref{fig:sxray_lightcurve}) and it is sub-luminous compared to the majority of
+\ref{fig:sxray_lightcurve}) and it is subluminous compared to the majority of
 lGRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
-distribution at 10 ks. The sub-samples comprise of 402 lGRBs, 31 sGRBs, and
+distribution at 10 ks. The subsamples comprise  402 lGRBs, 31 sGRBs, and
 GRB~111117A. The sample of lGRBs is from \citet{Evans2007, Evans2009} and the
-sample of sGRBs is compiled from \citet{Kann2011, Berger2014} and
+sample of sGRBs is compiled from \citet{Kann2011, Berger2014}, and
 \citet{DAvanzo2014a}. The mean and the 1-$\sigma$ dispersions of the samples are
@@
 $\log(L_{\mathrm{sGRB}} / \mathrm{erg~s}^{-1})  = 44.96 \pm 0.94$ for the lGRB
-and sGRB samples respectively. GRB~111117A had $\log(L / \mathrm{erg~s}^{-1}) =
+and sGRB samples,  respectively. GRB~111117A had $\log(L / \mathrm{erg~s}^{-1}) =
 44.95$ at 10 ks. This is very close to the peak of the sGRB luminosity
@@
 The separation of GRBs in two distinct classes based on their high-energy
-observables points to an intrinsically different physical origin. LGRBs are
+observables points to an intrinsically different physical origin. Long GRBs are
 typically interpreted as collapsars \citep{MacFadyen1999} in which a single, massive
@@
 \citet{Bromberg2013} investigated the degree to which high-energy observables of
-the long and short GRB populations overlap and quantified the certainty in class
+the long and short GRB populations overlap, and quantified the certainty in class
 membership. According to \citet{Bromberg2013}, GRB~111117A has a $96_{-5}^{+3}$
-percent probability of being a sGRB. Compared to two other sGRB candidates
+\%\ probability of being a sGRB. Compared to two other sGRB candidates
 at high redshift, GRB~060121 ($T_{90}=1.97 \pm 0.06$~s;
 \citealt{DeUgartePostigo2006, Levan2006}) at $1.7 \lesssim z \lesssim 4.5$
-($17_{-15}^{+14}$ per cent) and GRB~090426 ($T_{90}=1.28 \pm 0.09$~s;
+($17_{-15}^{+14}$ \%) and GRB~090426 ($T_{90}=1.28 \pm 0.09$~s;
 \citealt{Antonelli2009, Levesque2010, Thone2011}) at $z = 2.609$ ($10_{-10}^{+15}$
-per cent), the certainty in class membership for GRB~111117A is much higher.
+\%), the certainty in class membership for GRB~111117A is much higher.
 
 Additionally, \citet{Horvath2010} classify both GRB~060121 and GRB~090426 as
-intermediate-duration bursts. This comes from both events having very soft
-spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
+intermediate-duration bursts because both events have very soft
+spectra,  compared to the hard ones typically seen in sGRBs. Intermediate
 bursts are very clearly related in their properties to  lGRBs
@@
 
-A number of other GRBs are claimed to be short and at relatively high ($z \gtrsim 0.9$) redshift.
-If we consider bursts with a probability of being short $f_{\mathrm{NC}} > 50$ per cent, according to the
-\citet{Bromberg2013} classification scheme, 5 sGRB are found:
-GRB~051210 at $z \sim 1.3$ ($f_{\mathrm{NC}} = 82_{-61}^{+10}$ per cent;
+A number of other GRBs are thought to be short and at relatively high redshift ($z \gtrsim 0.9$).
+If we consider bursts with a probability of being short $f_{\mathrm{NC}} > 50$ \%, according to the
+\citet{Bromberg2013} classification scheme, five sGRB are found:
+GRB~051210 at $z \sim 1.3$ ($f_{\mathrm{NC}} = 82_{-61}^{+10}$ \%;
 \citealt{Leibler2010}),
-GRB~060801 at $z = 1.131$ ($f_{\mathrm{NC}} = 95_{-5}^{+3}$ per cent;
+GRB~060801 at $z = 1.131$ ($f_{\mathrm{NC}} = 95_{-5}^{+3}$ \%;
 \citealt{Berger2007}),
@@
 \citealt{Graham2009}),
-GRB~090510 at $z = 0.903$ ($f_{\mathrm{NC}} = 97_{-29}^{+1}$ per cent;
+GRB~090510 at $z = 0.903$ ($f_{\mathrm{NC}} = 97_{-29}^{+1}$ \%;
 \citealt{McBreen2010}), and
-GRB~100117 at $z = 0.915$ ($f_{\mathrm{NC}} = 97_{-3}^{+1}$ per cent;
+GRB~100117 at $z = 0.915$ ($f_{\mathrm{NC}} = 97_{-3}^{+1}$ \%;
 \citealt{Fong2011}). Although in individual cases a secure host association
@@
 
-This securely makes GRB~111117A, by far, the highest redshift sGRB detected to
+This certainly makes GRB~111117A, by far, the highest redshift sGRB detected to
 date. The redshift and classification of GRB~111117A imply that it occurred when
@@
 $r$-process element enrichment of the universe \citep{Goriely2011, Ji2016,
-	Komiya2016}, this marks the earliest detection of this process.
+        Komiya2016, Safarzadeh2017}, this marks the earliest detection of this process.
 
@@
       \includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
-	\caption{Rest-frame XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
-The grey shaded region is a compilation of long GRB lightcurves \citep{Evans2007, Evans2009} where the color represents density.
-The light blue lines are sGRB lightcurves from bursts with duration of $T_{90} \lesssim 2$ s and those that were classified as short in \citet{Kann2011, Berger2014, DAvanzo2014a}.
+        \caption{Rest-frame XRT light curve of GRB~111117A, compared to the general population of XRT light curves of GRBs.
+The gray shaded region is a compilation of long GRB light curves \citep{Evans2007, Evans2009}, where the color represents density.
+The light blue lines are sGRB light curves from bursts with duration of $T_{90} \lesssim 2$ s and those that were classified as short in \citet{Kann2011, Berger2014, DAvanzo2014a}.
 The thick green line is GRB~111117A.
-Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population and subluminous compared to the lGRB population.
-The insert shows the X-ray luminosity distributions of sGRBs and lGRBs at 10 ks, indicated by the vertical dashed line in the main panel.}
+Despite the remarkably high redshift, the luminosity is comparable to the bulk of the short burst population, and subluminous compared to the lGRB population.
+The inset shows the X-ray luminosity distributions of sGRBs and lGRBs at 10 ks, indicated by the vertical dashed line in the main panel.}
         \label{fig:sxray_lightcurve}
@@
 We show the recalculated hydrogen equivalent X-ray derived column density,
-$N_\mathrm{H,X}$, in Fig.~\ref{fig:NH_z} where we compare with the distributions
+$N_\mathrm{H,X}$, in Fig.~\ref{fig:NH_z}, where we compare it with the distributions
 of complete samples of both long and short GRBs. The lGRB sample is from
 \citet{Arcodia2016} and the sGRB sample is from \citet{DAvanzo2014a}. From the
-sGRB sample of \citet{DAvanzo2014a} we have excluded GRB~090426 which does
-likely not belong in a short sample as highlighted in Sect.
+sGRB sample of \citet{DAvanzo2014a} we have excluded GRB~090426, which 
+likely does not belong in a short sample, as highlighted in Sect.
 \ref{classification}. Both comparison samples infer $N_\mathrm{H,X}$  over the
@@
 that can affect the X-ray derived column density. The 17 (5) of the 99 (15) long
-(short) GRBs which do not have measured redshifts have been excluded from our
+(short) GRBs that do not have measured redshifts have been excluded from our
 analysis.
@@
 the lGRB afterglows \citep{Campana2010, Starling2013, Arcodia2016}. This is
-related to what is found by \citet{Kopac2012, Margutti2013}, that $N_\mathrm{H,X}$ seems to
+related to what has been found by \citet{Kopac2012} and \citet{Margutti2013} that $N_\mathrm{H,X}$ seems to
 be comparable for long and short GRBs when compared at similar redshifts.
@@
 The redshift evolution of $N_\mathrm{H,X}$ in the hosts of lGRBs is not
-reproduced by \citet{Buchner2017}, using a different $N_\mathrm{H,X}$ inference
-methodology. Instead a correlation between $N_\mathrm{H,X}$ and host stellar
+reproduced by \citet{Buchner2017}, who use  a different $N_\mathrm{H,X}$ inference
+methodology. Instead, a correlation between $N_\mathrm{H,X}$ and host stellar
 mass is suggested. Assuming that the different $N_\mathrm{H,X}$-fitting
-methodologies yield comparable results, GRB~111117A has a larger
+methodologies yield comparable results, GRB~111117A has a higher
 $N_\mathrm{H,X}$ compared to the relation suggested by \citet{Buchner2017} by
 more than the intrinsic scatter, although some lGRB hosts populate a similar
-region in the $N_\mathrm{H,X}$-$M_\star$ relation.
+region in the $N_\mathrm{H,X}$--$M_\star$ relation.
 % Additionally, for lGRBs,
@@
 %\citep{Lyman2017}.
-
-The large offset of GRB~111117A relative to the host center derived in
-\citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
-galaxy-scale gas being the source of the X-ray absorption. Along with the low
-dust content of the host, the large offset from the host center indicates that
-the high $N_\mathrm{H,X}$ arises because the density in the GRB surroundings is
-high, or because the light from the afterglow transverses localized regions of
-dense gas, similar to what is found by \citep{}. Alternatively, a significant
-contribution to the observed X-ray $N_\mathrm{H,X}$ can come from the diffuse
-intergalactic medium and the intervening systems along the line of sight of the
-GRB \citep{Campana2012, Arcodia2016}, although see \citep{}
 
@@
 high (or possibly because the light from the afterglow transverses localized
-regions of dense gas), see for example,
-\citep[e.g.][]{Watson2013, Krongold2013}. Alternatively, it has been
-hypothesised that a significant contribution to the observed X-ray
+regions of dense gas)
+\citep[see, e.g.,][]{Watson2013, Krongold2013}. Alternatively, it has been
+hypothesized that a significant contribution to the observed X-ray
 $N_\mathrm{H,X}$ could come from the diffuse intergalactic medium and the
 intervening systems along the line of sight of the GRB \citep{Campana2012,
-	Arcodia2016}, although see \citep{Watson2013, Krongold2013}.
-
-Even assuming a low dust-to-metals ratio as typically observed in long GRB
+        Arcodia2016}  \citep[but see also][]{Watson2013, Krongold2013}.
+
+Even assuming a low dust-to-metals ratio, as typically observed in long GRB
 afterglow sightlines \citep{Galama2001, Schady2010, Covino2013}, the
@@
 significant extinction along the afterglow line of sight ($A_V \gtrsim 1$~mag),
-which is contrasted to the absence of dust found from the SED fit and supported
-by the detection of \lya{}. Such a discrepancy between the extinction derived
-from the GRB afterglow and the one obtained using galaxy-wide measures has also
+which is contrasted with the absence of dust found from the SED fit and supported
+by the detection of \lya{}. This discrepancy between the extinction derived
+from the GRB afterglow and that obtained using galaxy-wide measures has also
 been observed occasionally for lGRBs \citep{Perley2013a}. For the one sGRB where
@@
 \begin{figure}
-	\centering \includegraphics[width=9cm]{figures/NH_z.pdf} \caption{Rest frame,
-	X-ray derived equivalent hydrogen column density of GRB~111117A compared to
-	complete samples of both long and short populations of GRBs. The sample of
-	lGRBs from \citet{Arcodia2016} is shown in red, where detections are also shown
-	with a kernel density estimate of the points and the limits on $N_\mathrm{H,X}$
-	are shown with arrows. The complete sample of sGRB by \citet{DAvanzo2014a} is
-	shown in blue, where again the limits are indicated by arrows. Marginalizations
-	over both axes are shown on the right and on the top of the plot, where the
-	limits are shown as semi-transparent bars and detections as solid ones. The red
-	curves in the marginalization plots are again the kernel density estimates of
-	the \citet{Arcodia2016} sample.} \label{fig:NH_z}
+      \centering \includegraphics[width=9cm]{figures/NH_z.pdf} 
+\caption{Rest frame, X-ray derived equivalent hydrogen column density of
+        GRB~111117A compared to complete samples of both long and short populations of
+        GRBs. The sample of lGRBs from \citet{Arcodia2016} is shown in red, where
+        detections are also shown with a kernel density estimate of the points, and the
+        limits on $N_\mathrm{H,X}$ are shown with arrows. The complete sample of sGRB by
+        \citet{DAvanzo2014a} is shown in blue, where again the limits are indicated by
+        arrows. Marginalizations over both axes are shown on the right and at the top of
+        the plot, where the limits are shown as semi-transparent bars and detections as
+        solid ones. The red curves in the marginalization plots are again the kernel
+        density estimates of the \citet{Arcodia2016} sample.} \label{fig:NH_z}
 \end{figure}
@@
 hostless class of sGRBs \citep{Berger2010a} and because the host exhibits
-emission lines this is indicative of a population of relatively young stars.
+emission lines, this is indicative of a population of relatively young stars.
 Like the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is
@@
 sample of short GRBs compiled by \citet{Berger2014} have a more vigorous star
-formation, placing it is in the very upper end of the star formation
+formation, placing it  at the very upper end of the star formation
 distribution. The cosmic SFR evolution of the universe likely plays a role due
 to the proximity of GRB~111117A to the peak of cosmic SFR \citep{Madau2014}.
-The high SFR is partly a selection effect, because a less star forming galaxy
+The high SFR is partly a selection effect;  a less star-forming galaxy
 would exhibit weaker emission lines, thus making the redshift harder to
@@
 
-The simultaneus detection of \lya{} and \ha{} allows us to put constraints of
+The simultaneous detection of \lya{} and \ha{} allows us to put constraints of
 the escape fraction of \lya{}, $f_{\mathrm{esc}}$(\lya). Using the intrinsic
@@
 similar properties \citep{Oyarzun2017}. The same authors also find that \lya{}
-emitting galaxies have mostly little dust, consistent with what inferred from
-the SED fit(see Sect. \ref{SED}). The centroid of the Ly$\alpha$ emission is
+emitting galaxies mostly have  little dust, consistent with what is inferred from
+the SED fit (see Sect. \ref{SED}). The centroid of the Ly$\alpha$ emission is
 found to be redshifted by $\sim$~$240\pm 90$~km~s$^{-1}$ with respect to
 systemic, which is similar to what is found for long GRB hosts
-\citep{Milvang-Jensen2012a} and Lyman break galaxies \citep{Shapley2003a} in
-which the outflow is attributed to star formation.
+\citep{Milvang-Jensen2012a} and Lyman break galaxies \citep{Shapley2003a} where the outflow is attributed to star formation.
 
@@
 
-The redshift distribution of GRBs is a powerful informant on both the conditions
-which drive the formation of these events, but also on the potential influence
+The redshift distribution of GRBs provides valuable information not only on the conditions
+that drive the formation of these events, but also on the potential influence
 these cosmic explosions have on the evolution of the universe. Due to the
@@
 their hosts \citep{Fruchter2006, Lyman2017}, the redshifts of lGRBs are easier
-to measure than for sGRB counterparts, where only a single burst has a redshift
+to measure than those of their sGRB counterparts, where only a single burst has a redshift
 measurement from the GRB afterglow \citep{Cucchiara2013, DeUgartePostigo2014b}.
 Correspondingly, the redshift distribution of sGRBs is still substantially
-unconstrained compared to that of lGRBs \citep[e.g., see][]{Jakobsson2012a,
+unconstrained compared to that of lGRBs \citep[see, e.g., ][]{Jakobsson2012a,
         DAvanzo2015, Perley2016d}.
@@
 GRB~090426, we have a statistically homogeneous sample from which we can address
-the implications of the redshift of GRB~111117A. This sample includes sGRB
+the implications of the redshift of GRB~111117A. This sample includes sGRBs
 originating in star-forming galaxies, elliptical galaxies, and apparent hostless
-sGRBs. Out of the 14 hosts in the sample, 10 (71 per cent) have both measured
-redshifts and $L_B/L^{\star}_{B}$. Compared to the complete sample, the host of
-GRB~111117A  is brighter than 80 per cent of the hosts with measured
-$L_B/L^{\star}_{B}$. Even if we conservatively assume that \textit{all} the
-hosts missing $L_B/L^{\star}_{B}$ are brighter than the host of GRB~111117A, the
-host is still brighter than > 60 per cent of sGRB hosts. For all 26 hosts with
-$L_B/L^{\star}_{B}$ from \citet{Berger2014}, the host of GRB~111117A is brighter
-than 73 per cent.
+sGRBs. Of the 14 hosts in the sample, 10 (71 \%)  have both measured redshifts
+and $L_B/L^{\star}_{B}$. Compared to the complete sample, the host of
+GRB~111117A  is brighter than 80 \%\ of the hosts with measured
+$L_B/L^{\star}_{B}$. Even if we conservatively assume that {all} the hosts
+missing $L_B/L^{\star}_{B}$ are brighter than the host of GRB~111117A, the host
+is still brighter than > 60 \%\ of sGRB hosts. The host of GRB~111117A is
+brighter than 73 \%\ of all 26 hosts with $L_B/L^{\star}_{B}$ from
+\citet{Berger2014}.
 
 If we assume that we are able to obtain emission-line redshifts from hosts which
-are at most 0.5 mag fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we
-would have missed 60 per cent of the redshifts (6 out of 10 hosts), due too the
-host being too faint, were they at the redshift of GRB~11117A. The corresponding
-number is around 45 per cent (12 out of 26) from the full sample of
+are at most 0.5 mag fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, and if they were at the redshift of GRB~11117A,  we
+would miss 60 \%\ of the redshifts (6 out of 10 hosts) because  the
+host is too faint. The corresponding
+number is around 45 \%\ (12 out of 26) of the full sample of
 \citet{Berger2014}, reflecting the lower mean $L_B/L^{\star}_{B}$ of the
@@
 than for galaxies hosting sGRBs, the fraction of missed burst redshifts is
-likely higher although the cosmic SFR evolution could play a role in improving
+likely higher, although the cosmic SFR evolution could play a role in improving
 redshift determinability at high $z$.
 
-A fraction of the bursts missing redshift are host-less but appear to be
+A fraction of the bursts missing redshift are host-less, but appear to be
 spatially correlated with galaxies that are likely at moderate redshifts
-\citep{Tunnicliffe2014}, but should some of the remainder be at high redshift,
-the missed fraction will increase. If we assume that \textit{all} the bursts
-that are missing redshifts are at high-$z$ and missed due to host faintness, 10
-out of 14 hosts in the complete sample (71 per cent) would be missed at $z =
+\citep{Tunnicliffe2014}; should some of the remainder be at high redshift,
+the missed fraction would increase. If we assume that {all} the bursts
+that are missing redshifts are at high $z$ and are missed due to host faintness, 10
+out of 14 hosts in the complete sample (71 \%) would be missed at $z =
 2.211$. This serves as an upper limit on the fraction of missed burst redshifts
-at high-$z$. Conversely, if all bursts missing redshift are at low redshift and
-missed for other reasons, 6 out of 14 hosts (43 per cent) would be missed at $z
-= 2.211$. The two limits indicate that we would miss between 43 and 71 per cent
+at high $z$. Conversely, if all bursts missing redshift were at low redshift and
+missed for other reasons, 6 out of 14 hosts (43 \%) would be missed at $z
+= 2.211$. The two limits indicate that we would miss between 43 \%\ and 71 \%
 of \textit{Swift}-detected sGRB hosts at $z \sim 2$ due to host faintness.
@@
 time delay models investigated by \citet{Wanderman2015} predict a sGRB rate at
-$z = 2.211$, around two orders of magnitude lower compared to the peak rate at
+$z = 2.211$, around two orders of magnitude lower than the peak rate at
 $z = 0.9$. According to \citet{Wanderman2015}, this preference depends
@@
 other hand, is close to the expected peak in sGRB rate calculated using the
-power law delay time models \citep{Behroozi2014, Wanderman2015, Ghirlanda2016},
+power-law delay time models \citep{Behroozi2014, Wanderman2015, Ghirlanda2016},
 meaning we would be missing a large fraction of sGRB redshifts.
 
-A critical test to assess if the power law delay time models can be accommodated
+A critical test to assess whether the power-law delay  time models can be accommodated
 by the current observation is to check if the implied sGRB rate at higher
@@
 have a tentative redshift measurement\footnote{This is based on
-	\url{http://www.astro.caltech.edu/grbox/grbox.php}, selecting all short,
-	\textit{Swift}-detected GRBs up to GRB~170428A.}, meaning that $> 73$ per cent of all
-sGRBs observed with \textit{Swift} are missing redshifts. More recently,
-\citet{Fong2017} \citep{LIGOScientificCollaboration2017b} compiled a list of 36
+        \url{http://www.astro.caltech.edu/grbox/grbox.php}, selecting all 
+        \textit{Swift}-detected sGRBs up to GRB~170428A.}, meaning that $> 73$ \%
+of all sGRBs observed with \textit{Swift} are missing redshifts. More recently,
+\citet{Fong2017} and \citet{LIGOScientificCollaboration2017b} compiled a list of 36
 (33) sGRBs with redshift measurements and, using this number, the redshift
-incompleteness of sGRBs decreases to 64 per cent. Additional to the potential
-number of high-$z$ event already detected but missing redshifts,
+incompleteness of sGRBs decreases to 64 \%. In addition to the potential
+number of high-$z$ events already detected but missing redshifts,
 \citet{Behroozi2014} parametrized the \textit{Swift} redshift sensitivity and
-found that the mean detection probability for sGRBs at $z \sim2$ was only $\sim1$
-per cent of the mean detection probability at $z \sim1$, assuming that the
-unknown beaming angle of sGRBs stays constant with time. What this means is that
-at the present, there is almost no limit on the number of sGRBs that could be at
-higher redshifts.
+found that the mean detection probability for sGRBs at $z \sim2$ was only
+$\sim1$ \%\ of the mean detection probability at $z \sim1$, assuming that
+the unknown beaming angle of sGRBs stays constant with time. What this means is
+that at  present, there is almost no limit on the number of sGRBs that could
+be at redshifts $z > 2$.
 
@@
 and that any binary NS formation channel can work sufficiently fast,
-we can, assuming a mass of each of the NSs in a circular orbit at the time of
-system formation, place a hard upper limit on the initial separation, $a_0$.
-
-In practice most NS-NS binaries will be eccentric at formation because of the SN
+we can--assuming a mass of each of the NSs in a circular orbit at the time of
+system formation--place a hard upper limit on the initial separation, $a_0$.
+
+In practice, most NS--NS binaries will be eccentric at formation because of the SN
 natal kicks. For more eccentric orbits, the coalescence timescale decreases,
@@
 Additionally, the constraint on initial progenitor separation will change
-depending on the mass assumed for the constituent NS masses, with larger masses
+depending on the mass assumed for the constituent NS masses, with higher masses
 generally resulting in faster inspiral times and weaker constraints of initial
@@
 $1.33^{+0.10}_{-0.12}~M_\odot$ \citep{Kiziltan2013}, which corresponds to  $a_0 <
-3.1^{+0.2}_{-0.2}~R_\odot$ where the errors are the 68 per cent posterior
+3.1^{+0.2}_{-0.2}~R_\odot$ where the errors are the 68 \%\ posterior
 predictive intervals.
@@
 As can be seen, a heavier system will inspiral faster, leading to a weaker constraint on the initial separation of the binary.
-The most probable value, the 68 per cent, and 98 percent posterior predictive intervals of the NS binary mass distribution from \citet{Kiziltan2013} are shown with the dashed ellipses, which corresponds to a constraint on maximal initial separation of $a_0 <3.1^{+0.2(+0.4)}_{-0.2(-0.4)}~R_\odot$, respectively.}
+The most probable value, the 68 \% and 98 \%\ posterior predictive intervals of the NS binary mass distribution from \citet{Kiziltan2013}, are shown with the dashed ellipses, which corresponds to a constraint on a maximum initial separation of $a_0 <3.1^{+0.2(+0.4)}_{-0.2(-0.4)}~R_\odot$, respectively.}
         \label{fig:prog_sep}
@@
 robust limit on the initial separation of $a_0 < 2.0^{+0.4}_{-0.4}~R_\odot$.
-However, this does not account for the possibility there could be an underlying
-stellar population of older stars from a previous star-formation episode. To
+However, this does not account for the possibility that there could be an underlying
+stellar population of older stars from a previous star formation episode. To
 investigate the possible impact of the presence of an old stellar population, we
@@
 the observed redshift. In principle, this can constrain the maximum contribution
-of old populations within the photometric error bars \citep[see][for
-details]{Papovich2001}. We find a negligible contribution to the stellar mass
-(i.e. variations much smaller than the statistical uncertainty associated with
+of old populations within the photometric error bars \citep[for details, see][]{Papovich2001}. We find a negligible contribution to the stellar mass
+(i.e., variations much smaller than the statistical uncertainty associated with
 the best-fit template).
@@
 The delay time between formation and explosion is well accommodated by the
-models of \citet{Belczynski2006}, although the longest delay times are excluded.
-This is especially true given the late type nature of the host
-\citep{OShaughnessy2008}. The same holds, if NS binaries are primarily formed
-through dynamical interactions in globular clusters \citep{Lee2010, Church2011}.
+models of \citet{Belczynski2006} where a range of binary NS systems and their
+decays are followed in different host potentials; however, some of the longest delay
+times are excluded. Given the late-type nature of the host, only a small
+subset of the longest population synthesis models violates the delay time
+constraint derived here \citep{OShaughnessy2008}. The same holds if NS binaries
+are primarily formed through dynamical interactions in globular clusters
+\citep{Lee2010, Church2011}.
 
@@
 
-We have here provided a revised, spectroscopic redshift measurement for the
-short GRB~111117A based on host galaxy emission lines setting it at $z = 2.211
+We have  provided here a revised, spectroscopic redshift measurement for the
+short GRB~111117A based on host galaxy emission lines, setting it at $z = 2.211
 \pm 0.001$. This value supersedes the previous photometric redshift estimates of
@@
 that GRB~111117A is formed through a prompt channel of sGRB formation and
-originates in an (unseen) star forming region located in the outskirts of the
+originates in an unseen star-forming region located in the outskirts of the
 host, or a localized region of high \hi~density along the line of sight.
@@
 accommodated in progenitor models. A lognormal delay time model predicts a very
-low volumetric density of bursts at $z \sim 2$, whereas a power law delay time
+low volumetric density of bursts at $z \sim 2$, whereas a power-law delay time
 model peaks near the GRB~111117A redshift. If more sGRBs are at similarly high
-redshifts, but are missed due to the faintness of their hosts and afterglows, a
+redshifts, but are missed because of the faintness of their hosts and afterglows, a
 lognormal delay time model will be disfavored. Compared to a complete sample of
 \textit{Swift}-detected sGRB, the host of GRB~111117A is more luminous than 80
-per cent of sGRB hosts with measured luminosities. Assuming a host brightness
-redshift determination threshold, for between 43 and 71 per cent of the sample
-hosts, we would be unable to determine a redshift should they be at a similar
-redshift of GRB~111117A. This could indicate that, potentially, a significant
+\%\ of sGRB hosts with measured luminosities. Assuming a host brightness
+redshift determination threshold for  43 -- 71 \%\ of the sample
+hosts, we would be unable to determine a redshift if  they were  at a similar
+redshift to that of  GRB~111117A. This could indicate that, potentially, a significant
 fraction of \textit{Swift}-detected sGRBs are at high $z$, but with redshifts
@@
 Using the age of the universe at the time of explosion allows us to set
-constraints on the maximal separation between the engine constituents at the
-time of formation. We find that the maximal separation of two NSs at system
+constraints on the maximum separation between the engine constituents at the
+time of formation. We find that the maximum separation of two NSs at system
 formation time is $a_0 < 3.1~R_\odot$, which excludes some of the formation
@@
 
-All data, code and calculations related to the paper along with the
-paper itself are available at \url{https://github.com/jselsing/GRB111117A}.
-
 \begin{acknowledgements}
-We thank the anonymous referee for a constructive report. 
+We thank the anonymous referee for the constructive report. 
 %
@@
 %
-PDA, SCo, acknowledge support from ASI grant I/004/11/3.
-%
-JJ acknowledges support from NOVA and NWO-FAPESP grant for advanced
+PDA and SCo acknowledge support from ASI grant I/004/11/3.
+%
+JJ acknowledges support from NOVA and a NWO-FAPESP grant for advanced
 instrumentation in astronomy.
@@
 CT acknowledges support from a Spanish National Research Grant of Excellence
-under project AYA 2014-58381-P and funding associated to a Ramón y Cajál
+under project AYA 2014-58381-P and funding associated with a Ramón y Cajál
 fellowship under grant number RyC-2012-09984.
@@
 RSR acknowledges AdUP's BBVA Foundation Grant for Researchers and Cultural
-Creators and support from ASI (Italian Space Agency) through the Contract n. 2015-046-R.0 and from European Union Horizon 2020 Programme under the AHEAD project (grant agreement n. 654215).
+Creators and support from the Italian Space Agency (ASI) through  Contract n. 2015-046-R.0 and from the European Union Horizon 2020 Programme under the AHEAD project (grant agreement n. 654215).
 %
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{TheAstropyCollaboration2013}.
-The analysis and plotting has been achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy, and Scipy \citep{VanderWalt2011}, along with other community-developed packages.
+The analysis and plotting was achieved using the Python-based packages Matplotlib \citep{Hunter2007}, Numpy, and Scipy \citep{VanderWalt2011}, along with other community-developed packages.
 %
@@
 
-\bibliographystyle{mnras}
+\bibliographystyle{aa}
 \bibliography{GRB111117A}

### 2018-08-28 GRB111117A_old.tex
Published

--- 2018-08-28 Arxiv/GRB111117A.tex
@@
 
-\title{The host galaxy of the short GRB~111117A at $z = 2.211$: impact on the short GRB redshift distribution and progenitor channels\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}}
+\title{The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}}
 
@@
 The classification of GRBs in two groups, initially comes from the bimodal
-distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
+distribution of burst duration and spectral hardness \citet{Kouveliotou1993},
 where the duration T$_{90} < 2$ has been regarded as the dividing line between long
@@
 gravitational wave (GW) signal \citep{Ghirlanda2016}. The simultaneous detection
-of a sGRB and a GW will provide new promising ways to constrain the binary
+of a sGRB and a GW will provide new a promising ways to constrain the binary
 inclination angle \citep{Arun2014} and measure cosmological distances
@@
 The detection of \lya~ is predicted from the SED fit and confirmed by the spectroscopic observations.
-Overplotted in grey is the observed spectrum, binned down to 3 \AA~per pixel for presentation purposes.
-Slit losses are responsible for discrepancies between the measured photometry and the X-shooter spectrum.
+Overplotted in grey is the observed spectrum.
 The reason for the spectral gaps at 5500 \AA~and 10000 \AA~is from the merging of the arms.}
@@
 tracing, wavelength calibration, rectification, and flux calibration using the
-VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run in physical
+VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010} run in physical
 mode. Because the echelle orders are curved across each detector, a
@@
 described in Selsing et al. 2017 (in prep.) and available
-online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
-where the full spectral point spread function is modeled across each arm and
-used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
-spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
-slit position is shown in Fig.~\ref{fig:spec_setup}. We show the extracted
-spectrum in Fig.~\ref{fig:SED}.
+online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
+signal-to-noise ratio of the continuum in the near-infrared arm is too low to
+use the optimal extraction algorithm \citep{Horne1986} and therefore the
+extraction is carried out with a simple aperture covering the entire host, where
+we sum the flux in the aperture and derive the error by summing the
+corresponding pixel errors in quadrature. An overview of the spectroscopic
+observations is given in Table~\ref{tab:spec_overview}, and the slit position is
+shown in Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in
+Fig.~\ref{fig:SED}.
 
@@
 edge-on and because the slit is not aligned along the long axis of the host,
-this value is a lower limit.
+this value is an upper limit.
+
+A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
+"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
+afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
+interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
+< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
 
@@
 observations and measured brightnesses is given in
-Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, see
-Sect. \ref{xray}, there is likely no afterglow contamination on the measured
-photometry.
+Table~\ref{tab:phot_overview}. Due to the optical darkness of this burst, there
+is likely no afterglow contamination on the measured photometry.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{X-ray temporal and spectral analysis}\label{xray}
+\subsection{X-ray temporal and spectral analysis}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
-"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
-afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
-interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
-< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 surrounding is high, or because the light from the afterglow traverses a region
-of dense gas. The optical darkness of the GRB is an additional hint of the high
-density in the GRB surroundings or along the line of sight.
+of dense gas.
 
@@
 	\caption{Rest frame, X-ray derived equivalent hydrogen column densities for GRB~111117A compared to complete samples of both long and short populations of GRBs.
-The detections are replaced with contours and the limits are shown with arrows.
-Marginalizations over both axes are shown where the limits are shown as transparent bars and detections as solid ones.}
+The detections are replaced with contours and the limits are shown with arrows.}
 	\label{fig:NH_z}
@@
 If we assume that we are able to obtain emission-line redshifts from hosts, 0.5
-mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed
-60 per cent of the redshifts (6 out of 10 hosts), due too the host being too
-faint, were they at the redshift of GRB~11117A. The corresponding number is
-around 45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
+mags fainter \citep[$R < 24.5$~mag;][]{Kruhler2012}, then we would have missed 60
+per cent of the redshifts (6 out of 10 hosts), were they at the redshift of
+GRB~111117A due to the hosts being too faint. The corresponding number is around
+45 per cent (12 out of 26) from the full sample of \citet{Berger2014},
 reflecting the lower mean $L_B/L^{\star}_{B}$ of the complete sample. Because
@@
 on the other hand, is close to the expected peak in sGRB rate calculated using
-the power law delay time models \citep{Behroozi2014, Wanderman2015,
-	Ghirlanda2016}, meaning we could be missing a large fraction of sGRBs.
+the power law delay time models \citep{Behroozi2014, Wanderman2015}.
 
@@
 does not account for the possibility there could be an underlying stellar
-population of older stars from a previous star-formation epoch. To investigate
-the possible impact of the presence of an old stellar population, we followed
-\citet{Papovich2001} and re-fitted the observed SED with the best-fit template
-to which an additional stellar population of old stars was added. For each
-galaxy, this old population was set as the SPSs with same parameters that the
-best-fit SED excepts the age, which was set to the age of the Universe at the
-observed redshift. In principle, this can constrain the maximum contribution of
-old populations within the photometric error bars \citep[see][ for
-details]{Papovich2001}. We find a negligible contribution to the stellar mass
-(i.e. variations much smaller than the statistical uncertainty associated with
-the best-fit template). The delay time between formation and explosion is well
-accommodated by the models of \citet{Belczynski2006}, although the longest delay
-times are excluded. This is especially true given the late type nature of the
-host \citep{OShaughnessy2008}.
+population of older stars from a previous star-formation epoch. The delay time
+between formation and explosion is well accommodated by the models of
+\citet{Belczynski2006}, although the longest delay times are excluded.
+This is especially true given the late type nature of the host
+\citep{OShaughnessy2008}.
 
@@
 $N_H-z$ evolution traced by the hosts of lGRBs. The SFR of the host is in the
-upper end of the sGRB host SFR distribution and no significant amount of dust is
-present. The high $N_H$ is difficult to reconcile with the large projected host
-offset and the absence of dust. One possible explanation could be, that
-GRB~111117A is formed through a prompt channel of sGRB formation and originates
-in a star forming region located in the outskirts of the host.
+upper end of the sGRB host SFR distribution and no dust is present. The high
+$N_H$ is difficult to reconcile with the large projected host offset and the
+absence of dust. One possible explanation could be, that GRB~111117A is formed
+through a prompt channel of sGRB formation and originates in a star forming
+region located in the outskirts of the host.
 
@@
 We thank Jens Hjorth and Lise Christensen for useful discussions regarding the interpretation of this event. 
-We thank Mathieu Puech for testing the possible contribution from an older stellar population in the SED.
 %	
 TK acknowledges support through the Sofja Kovalevskaja Award to P. Schady.
-%
-SDV is supported by the French National Research Agency (ANR) under contract ANR-16-CE31-0003 BEaPro 
 % 
@@
 
-\bibliographystyle{aa_arxiv}
+\bibliographystyle{mnras}
 \bibliography{GRB111117A}

### 2018-08-28 GRB111117A_old2.tex
Published

--- 2017-11-26 GRB111117A.tex
@@
 
-The large offset of GRB~111117A relative to the host center derived in
-\citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
-galaxy-scale gas being the source of the X-ray absorption. Along with the low
-dust content of the host, the large offset from the host center indicates that
-the high $N_\mathrm{H,X}$ arises because the density in the GRB surroundings is
-high, or because the light from the afterglow transverses localized regions of
-dense gas, similar to what is found by \citep{}. Alternatively, a significant
-contribution to the observed X-ray $N_\mathrm{H,X}$ can come from the diffuse
-intergalactic medium and the intervening systems along the line of sight of the
-GRB \citep{Campana2012, Arcodia2016}, although see \citep{}
-
 The large offset of GRB\,111117A relative to the host center derived in

### 2018-08-28 submission/GRB111117A.tex
Published

--- 2018-08-28 GRB111117A_old2.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
-%\documentclass[referee]{aa}
+%\documentclass{aa}    %% Astronomy & Astrophysics style class aa.cls v8.2
+\documentclass[referee]{aa}
 
@@
 \usepackage{color,hyperref}
-\definecolor{linkcolor}{rgb}{0,0.3,0.7}
-\hypersetup{colorlinks=true,
-	linkcolor=linkcolor, 
-	citecolor=linkcolor,
-	filecolor=linkcolor, 
-	urlcolor=linkcolor}
+%\definecolor{linkcolor}{rgb}{0,0.3,0.7}
+%\hypersetup{colorlinks=true,
+%	linkcolor=linkcolor, 
+%	citecolor=linkcolor,
+%	filecolor=linkcolor, 
+%	urlcolor=linkcolor}
 
@@
 
-\input{authorlist.tex}
+\author{J.~Selsing\inst{1}
+	\and T.~Kr\"{u}hler\inst{2}
+	\and D.~Malesani\inst{1}
+	\and P.~D'Avanzo\inst{3}
+	\and S.~Schulze \inst{4, 5, 6}
+	\and S.~D.~Vergani\inst{3, 7}
+	\and J.~Palmerio\inst{8}
+	\and J.~Japelj\inst{9}
+	\and B.~Milvang-Jensen\inst{1}
+	\and D.~Watson\inst{1}
+	\and P.~Jakobsson\inst{10}
+	%
+	\and J.~Bolmer \inst{2}
+	\and Z.~Cano\inst{11}
+	\and S.~Covino \inst{3}
+	%\and L.~B.~Christensen\inst{1}
+	\and V.~D'Elia\inst{12, 13}
+	\and A.~de~Ugarte~Postigo\inst{11, 1}
+	\and J.~P.~U.~Fynbo\inst{1}
+	\and A.~Gomboc\inst{14}
+	\and K.~E.~Heintz\inst{10,1}
+	\and L.~Kaper \inst{9}
+	\and A.~J.~Levan \inst{15}
+	\and S.~Piranomonte \inst{12}
+	\and G.~Pugliese \inst{9}
+	\and R.~S\'{a}nchez-Ram\'{\i}rez \inst{11, 16}
+	\and M.~Sparre\inst{1,17}
+	\and N.~R.~Tanvir\inst{18}
+	\and C.~C.~Th\"{o}ne\inst{11}
+	\and K.~Wiersema \inst{18}
+}
+
+\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
+	%2
+	\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
+	%3
+	\and INAF - Osservatorio Astronomico di Brera, via E. Bianchi 46, I-23807, Merate (LC), Italy
+	%4
+	\and Instituto de Astrof\'isica, Facultad de F\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+	%5
+	\and Millennium Institute of Astrophysics, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
+	%6
+	\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
+	%7
+	\and GEPI, Observatoire de Paris, PSL Research University, CNRS, Place Jules Janssen, 92190 Meudon, France
+	%8
+	\and Sorbonne Universités, UPMC Univ. Paris 6 et CNRS, UMR 7095, Institut d’Astrophysique de Paris, 98 bis bd Arago, 75014 Paris, France
+	%9
+	\and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands
+	%10
+	\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
+	%11
+	\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain
+	%12
+	\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy
+	%13
+	\and ASI-Science Data Centre, Via del Politecnico snc, I-00133 Rome, Italy
+	%14
+	\and Centre for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
+	%15
+	\and Department of Physics, University of Warwick, Coventry CV4 7AL, UK
+	%16
+	\and Istituto de Astrofisica e Planetologia Spaziali, INAF, Via Fosso del Cavaliere 100, I-00133 Roma, Italy
+	%17
+	\and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
+	%18
+	\and Department of Physics \& Astronomy and Leicester Institute of Space \& Earth Observation,
+	University of Leicester, University Road, Leicester, LE1 7RH, UK
+}
 
@@
 waves (GWs) from a binary neutron star merger and a sGRB
-\citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}. The
-absence of associated supernovae in deep searches
-\citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011} supports this idea and
-distinguishes the physical origin of sGRBs from their long-duration
-counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.
+\citep{LIGOScientificCollaboration2017a, Goldstein2017, Savchenko2017}, however,
+to what degree the sGRB associated with GW170817 is a cosmological sGRB remains
+an open question \citep{Lyman2018, Margutti2018}. The absence of associated
+supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b,
+	Kann2011} supports this idea and distinguishes the physical origin of sGRBs from
+their long-duration counterparts, \citep[albeit see also][]{Fynbo2006b,
+	Valle2006, Gal-Yam2006}.
 
@@
 ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is
-significantly higher than the previously estimated redshift based on photometric
+significantly higher than the previously estimated redshift of $z \sim 1.3$ based on photometric
 studies \citep{Margutti2012,Sakamoto2013}. We present the GRB's rest frame
@@
 	\centering
-	\includegraphics[width=9cm]{figures/GRB111117A_spec_obs.pdf}
+%	\includegraphics[width=9cm]{GRB111117A_spec_obs.pdf}
 	\caption{
@@
 taken at similar position angles. The corresponding photometry is shown in
-Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position as derived
+Fig.~\ref{fig:SED}. The blue asterisk indicates the GRB position (R.A., Dec.) (J2000) = (00$^{\rm h}$ 50$^{\rm m}$ 46.264$^{\rm s}$, +23$^{\circ}$ 00$^{\prime}$ 39.98$^{\prime\prime}$) as derived
 from the \emph{Chandra} observations in \citet{Sakamoto2013}.
@@
 %	\centering
-%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
+%%	\includegraphics[width=9cm]{figures/OIII_img.pdf}
 %	\caption{2D-image of the \oiii$\lambda$5007 emission line.
@@
 	\centering
-	\includegraphics[width=18cm]{figures/grid.pdf}
+%	\includegraphics[width=18cm]{grid.pdf}
 	\caption{2D-images of the emission lines corresponding to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959, \oiii$\lambda$5007, and \ha.
@@
 	\centering
-	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
+%	\includegraphics[width=16cm]{SEDspecphot.pdf}
 	\caption{Best-fit SED to the derived photometry.
@@
 
-\input{tables/spec_overview.tex}
+\begin{table*}[!ht]
+
+	\centering
+	\caption{Overview of the spectroscopic observations.
+``JH'' in the slit width refers to observations where a $K$-band blocking filter has been used.
+The seeing is determined from the width of the spectral trace of a telluric standard star, observed close in time to the host observation.
+The spectral resolution, $R$, is measured from unresolved telluric absorption lines in the spectrum of the telluric standard star.
+\label{tab:spec_overview}}
+	\begin{tabular}{cccccccc}
+		\hline\hline
+		{Observation epoch} &  \multicolumn{3}{c}{Exposure time (s)} & Slit width & Airmass & Seeing & $R$   \\ [1.5pt]
+		\hline
+		(UT) & UVB  & VIS & NIR &  (arcsec)   & {} & (arcsec)  & {VIS/NIR}  \\ [1.5pt]
+		\hline
+		2011-11-19 01:33 & 2 $\times$ 2400 & 2 $\times$ 2400 & 8 $\times$ 600 & 1.0/1.0/0.9 & 1.49 & 0.75 & 11600/6700 \\
+		2013-07-15 09:02 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.53 & 0.98 & 9600/8900 \\
+		2013-08-03 07:37 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.55 & 0.85 & 11400/11300 \\
+		2013-08-03 08:34 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.49 & 0.85 & 11400/11300 \\
+
+		\hline\noalign{\smallskip}
+
+	\end{tabular}
+
+\end{table*}
 
@@
 \begin{figure}
-	\centering \includegraphics[width=9cm]{figures/photometry_mosaic.pdf}
+%%	\centering \includegraphics[width=9cm]{photometry_mosaic.pdf}
 \caption{Mosaic showing all used imaging. The host of GRB~111117A is marked by
@@
 
-\input{tables/phot_overview.tex}
+\begin{table*}[!ht]
+
+	\centering
+	\caption{Overview of the photometric observations. \label{tab:phot_overview}}
+	\begin{tabular}{ccccccc}
+		\hline\hline
+		{Observation epoch} &  Exptime & Telescope/instrument & Filter & Airmass & Image quality & Host brightness\tablefootmark{a}  \\ [1.5pt]
+		\hline
+		({UT}) & ({ks}) &    & {} & & (arcsec)  & (AB mag)  \\ [1.5pt]
+		\hline
+		2013-08-30 07:43 & 1.45 & VLT/FORS2 & $g$ & 1.55 & 0.99 & $24.08\pm 0.09$ \\
+		2011-11-17 20:07 & 0.80 & GTC/OSIRIS & $g$ & 1.15 & 1.67 & $24.13\pm 0.09$ \\
+		2011-11-17 20:07 & 1.20 & GTC/OSIRIS & $r$ & 1.11 & 1.50 & $23.93\pm 0.08$ \\        
+		2013-07-17 08:37 & 1.45 & VLT/FORS2 & $R$ & 1.56 & 0.74 & $23.95\pm 0.06$ \\   
+		2011-11-28 21:10 & 3.60 & TNG/DOLORES & $R$ & 1.01 & 1.08 & $23.96\pm 0.13$ \\           
+		2011-11-17 20:07 & 0.36 & GTC/OSIRIS & $i$ & 1.08 & 1.50 & $23.89\pm 0.23$ \\   
+		2013-08-03 09:23 & 1.35 & VLT/FORS2 & $I$ & 1.54 & 0.93 & $24.22\pm 0.15$ \\           
+		2011-11-28 06:14 & 1.80 & Gemini/GMOS-N & $z$ & 1.01 & 0.84 & $24.24\pm 0.47$ \\  
+		2013-07-13 09:33 & 1.08 & VLT/FORS2 & $z$ & 1.49 & 0.63 & $23.76\pm 0.21$ \\             
+		2013-06-24 09:14 & 1.98 & VLT/HAWK-I & $J$ & 1.70 & 0.63 & $23.13\pm 0.18$ \\        
+		2013-06-27 09:21 & 1.68 & VLT/HAWK-I & $H$ & 1.63 & 0.91 & $22.94\pm 0.29$ \\   
+		2013-06-28 09:14 & 1.92 & VLT/HAWK-I & $K_\mathrm{s}$ & 1.65 & 0.76 & $23.07\pm 0.32$ \\   
+		\hline\noalign{\smallskip}
+
+	\end{tabular}
+
+	\tablefoot{
+		\tablefoottext{a}{All magnitudes are given in the AB system and are not corrected for the expected Galactic foreground extinction corresponding to a reddening of $E_{B-V}=0.027$\,mag.}}
+\end{table*}
 
@@
 $r$-process element enrichment of the universe \citep{Goriely2011, Ji2016,
-	Komiya2016}, this marks the earliest detection of this process.
+	Komiya2016, Safarzadeh2017}, this marks the earliest detection of this process.
 
@@
 	\centering
-	\includegraphics[width=9cm]{figures/XLC_111117A_rest.pdf}
+%%	\includegraphics[width=9cm]{XLC_111117A_rest.pdf}
 	\caption{Rest-frame XRT lightcurve of GRB~111117A, compared to the general population of XRT lightcurves of GRBs.
@@
 \begin{figure}
-	\centering \includegraphics[width=9cm]{figures/NH_z.pdf} \caption{Rest frame,
-	X-ray derived equivalent hydrogen column density of GRB~111117A compared to
-	complete samples of both long and short populations of GRBs. The sample of
-	lGRBs from \citet{Arcodia2016} is shown in red, where detections are also shown
-	with a kernel density estimate of the points and the limits on $N_\mathrm{H,X}$
-	are shown with arrows. The complete sample of sGRB by \citet{DAvanzo2014a} is
-	shown in blue, where again the limits are indicated by arrows. Marginalizations
-	over both axes are shown on the right and on the top of the plot, where the
-	limits are shown as semi-transparent bars and detections as solid ones. The red
-	curves in the marginalization plots are again the kernel density estimates of
-	the \citet{Arcodia2016} sample.} \label{fig:NH_z}
+%%	\centering \includegraphics[width=9cm]{NH_z.pdf} 
+\caption{Rest frame, X-ray derived equivalent hydrogen column density of
+	GRB~111117A compared to complete samples of both long and short populations of
+	GRBs. The sample of lGRBs from \citet{Arcodia2016} is shown in red, where
+	detections are also shown with a kernel density estimate of the points and the
+	limits on $N_\mathrm{H,X}$ are shown with arrows. The complete sample of sGRB by
+	\citet{DAvanzo2014a} is shown in blue, where again the limits are indicated by
+	arrows. Marginalizations over both axes are shown on the right and on the top of
+	the plot, where the limits are shown as semi-transparent bars and detections as
+	solid ones. The red curves in the marginalization plots are again the kernel
+	density estimates of the \citet{Arcodia2016} sample.} \label{fig:NH_z}
 \end{figure}
@@
 	\url{http://www.astro.caltech.edu/grbox/grbox.php}, selecting all short,
-	\textit{Swift}-detected GRBs up to GRB~170428A.}, meaning that $> 73$ per cent of all
-sGRBs observed with \textit{Swift} are missing redshifts. More recently,
+	\textit{Swift}-detected GRBs up to GRB~170428A.}, meaning that $> 73$ per cent
+of all sGRBs observed with \textit{Swift} are missing redshifts. More recently,
 \citet{Fong2017} \citep{LIGOScientificCollaboration2017b} compiled a list of 36
@@
 \citet{Behroozi2014} parametrized the \textit{Swift} redshift sensitivity and
-found that the mean detection probability for sGRBs at $z \sim2$ was only $\sim1$
-per cent of the mean detection probability at $z \sim1$, assuming that the
-unknown beaming angle of sGRBs stays constant with time. What this means is that
-at the present, there is almost no limit on the number of sGRBs that could be at
-higher redshifts.
+found that the mean detection probability for sGRBs at $z \sim2$ was only
+$\sim1$ per cent of the mean detection probability at $z \sim1$, assuming that
+the unknown beaming angle of sGRBs stays constant with time. What this means is
+that at the present, there is almost no limit on the number of sGRBs that could
+be at redshifts $z > 2$.
 
@@
 	\centering
-	\includegraphics[width=9cm]{figures/prog_sep.pdf}
+%%	\includegraphics[width=9cm]{prog_sep.pdf}
 	\caption{Constraints on the initial progenitor separation, given binary constituent masses.
@@
 The delay time between formation and explosion is well accommodated by the
-models of \citet{Belczynski2006}, although the longest delay times are excluded.
-This is especially true given the late type nature of the host
-\citep{OShaughnessy2008}. The same holds, if NS binaries are primarily formed
-through dynamical interactions in globular clusters \citep{Lee2010, Church2011}.
+models of \citet{Belczynski2006}, where a range of binary NS systems and their
+decays are followed in different host potentials.  Some of the longest delay
+times are, however excluded. Given late type nature of the host, only a small
+subset of the longest population synthesis models violates the delay time
+constraint derived here \citep{OShaughnessy2008}. The same holds, if NS binaries
+are primarily formed through dynamical interactions in globular clusters
+\citep{Lee2010, Church2011}.
 
